### 2018-07-03 paper/sp.tex
start of paper, abstract

%% This file is part of the spectroscopic_parallax project
%% Copyright 2018 the authors.

\documentclass[modern]{aastex62}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{TBD}
%\revised{TBD}
%\accepted{TBD}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
%\submitjournal{AAS Journals}

\shorttitle{spectroscopic distance estimates}
\shortauthors{eilers, hogg, rix, ness}
\watermark{kittens}

% project macros
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\wise}{\project{\acronym{WISE}}}
\newcommand{\zmass}{\project{\acronym{2MASS}}}
\newcommand{\sdssv}{\project{\acronym{SDSS-V}}}

% math macros
\newcommand{\logg}{\log g}

\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing

\title{Accurate spectroscopic estimates of luminosity and distance for luminous red-giant stars}

\correspondingauthor{Anna-Christina Eilers}
\email{eilers@mpia.de}

\author[0000-0003-2895-6218]{Anna-Christina Eilers}
\affiliation{Max-Planck-Institut f\"ur Astronomie}

\author[0000-0003-2866-9403]{David W. Hogg}
\affiliation{Max-Planck-Institut f\"ur Astronomie}
\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University}
\affiliation{Center for Data Science, New York University}
\affiliation{Flatiron Institute}

\author{Hans-Walter Rix}
\affiliation{Max-Planck-Institut f\"ur Astronomie}

\author{Melissa Ness}
\affiliation{Flatiron Institute}
\affiliation{Department of Astronomy, Columbia University}

\begin{abstract}\noindent
With contemporary infrared spectroscopic surveys like \apogee,
red-giant stars can be observed to distances and extinctions
at which geometric parallaxes are not precisely measured by \gaia.
Here we use the \apogee--\gaia\ overlap to train a purely linear model of
\gaia, \zmass, and \wise\ photometry and
continuum-normalized \apogee\ spectroscopy that predicts inverse-square-root luminosity
for red stars with $0<\logg<2.2$.
The inverse-square-root is used to make the noise model close to Gaussian,
so that the training can be performed with full use of uncertainties and
without any cuts on parallax or parallax signal-to-noise ratio;
the model includes an L1 regularization that zeros out the contributions of
most spectral pixels to the parallax estimates; and the training was performed
with leave-out subsamples so no star's astrometry is used even indirectly in its
spectroscopic parallax estimate.
We re-label each star in the sample
with a new spectroscopic parallax; this new parallax has an empirical uncertainty of
10\,percent, which is more precise than the \gaia\ parallax
for the vast majority of targets.
Validation with globular and open clusters shows that the spectroscopic parallax
estimates are both accurate and precise.
These distance and luminosity estimates open up great opportunities for
mapping the Milky Way disk with the next
generation of spectroscopic surveys, and especially \sdssv.
\end{abstract}

\keywords{%
methods:~statistical
 ---
techniques:~spectroscopic
 ---
catalogs
 ---
surveys
 ---
parallaxes
 ---
stars:~distances
 ---
Galaxy:~disk
 ---
infrared:~stars
}

\section{Introduction} \label{sec:intro}

foo

bar

\section{Data}

foo

bar

\section{Method}

\section{Validation}

\section{Results and Discussion}

\acknowledgements
This project was developed in part at the
2018 NYC Gaia Sprint, hosted by the Center for Computational Astrophysics of
the Flatiron Institute in New York City.

This work has made use of data from the European Space Agency (ESA) mission
\gaia\ (\url{https://www.cosmos.esa.int/gaia}), processed by the \gaia\ Data
Processing and Analysis Consortium (\acronym{DPAC},
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the
\acronym{DPAC}
has been provided by national institutions, in particular the institutions
participating in the \gaia\ Multilateral Agreement.

\begin{thebibliography}{}

\bibitem[Astropy Collaboration et al.(2013)]{2013A&A...558A..33A} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, \aap, 558, A33 
\bibitem[Bertin \& Arnouts(1996)]{1996A&AS..117..393B} Bertin, E., \& Arnouts, S.\ 1996, \aaps, 117, 393 
\bibitem[Corrales(2015)]{2015ApJ...805...23C} Corrales, L.\ 2015, \apj, 805, 23
\bibitem[Ferland et al.(2013)]{2013RMxAA..49..137F} Ferland, G.~J., Porter, R.~L., van Hoof, P.~A.~M., et al.\ 2013, \rmxaa, 49, 137
\bibitem[Hanisch \& Biemesderfer(1989)]{1989BAAS...21..780H} Hanisch, R.~J., \& Biemesderfer, C.~D.\ 1989, \baas, 21, 780 
\bibitem[Lamport(1994)]{lamport94} Lamport, L. 1994, LaTeX: A Document Preparation System, 2nd Edition (Boston, Addison-Wesley Professional)
\bibitem[Schwarz et al.(2011)]{2011ApJS..197...31S} Schwarz, G.~J., Ness, J.-U., Osborne, J.~P., et al.\ 2011, \apjs, 197, 31  
\bibitem[Vogt et al.(2014)]{2014ApJ...793..127V} Vogt, F.~P.~A., Dopita, M.~A., Kewley, L.~J., et al.\ 2014, \apj, 793, 127

\end{thebibliography}

%% This command is needed to show the entire author+affilation list when
%% the collaboration and author truncation commands are used.  It has to
%% go at the end of the manuscript.
%\allauthors

%% Include this line if you are using the \added, \replaced, \deleted
%% commands to see a summary list of all changes at the end of the article.
%\listofchanges

\end{document}

% End of file `sample62.tex'.

### 2018-07-03 paper/sp.tex

@@
 \shortauthors{eilers, hogg, rix, ness}
-\watermark{kittens}
+%\watermark{kittens}
 
@@
 
-\title{Accurate spectroscopic estimates of luminosity and distance for luminous red-giant stars}
+\title{Accurate spectroscopic estimates of luminosity and distance\\
+       for luminous red-giant stars}
 
@@
 Here we use the \apogee--\gaia\ overlap to train a purely linear model of
-\gaia, \zmass, and \wise\ photometry and
-continuum-normalized \apogee\ spectroscopy that predicts inverse-square-root luminosity
-for red stars with $0<\logg<2.2$.
+continuum-normalized \apogee\ spectroscopy
+(plus \gaia, \zmass, and \wise\ photometry)
+that predicts inverse-square-root luminosity for red stars with $0<\logg<2.2$.
 The inverse-square-root is used to make the noise model close to Gaussian,
@@
 spectroscopic parallax estimate.
+The model is flexible enough to correct for extinction by dust without any
+input of explicit extinction or reddening estimates.
 We re-label each star in the sample
@@
 for the vast majority of targets.
-Validation with globular and open clusters shows that the spectroscopic parallax
-estimates are both accurate and precise.
+Validation with globular and open clusters shows that the spectroscopic parallaxes
+are both accurate and precise.
 These distance and luminosity estimates open up great opportunities for
@@
 This project was developed in part at the
-2018 NYC Gaia Sprint, hosted by the Center for Computational Astrophysics of
+2018 \acronym{NYC} Gaia Sprint, hosted by the Center for Computational Astrophysics of
 the Flatiron Institute in New York City.

### 2018-07-03 infer_parallax.py
test

@@ -29,7 +29,7 @@ fsize = 14
 # -------------------------------------------------------------------------------
 
 # make plots?
-prediction = True
+prediction = False
 
 print('loading labels...')
 hdu = fits.open('data/training_labels_parent.fits')
@@ -398,6 +398,95 @@ if not prediction:
     plt.title(r'$N = {0},\,\lambda = {1}$'.format(len(labels), lam), fontsize = fsize)
     plt.savefig('plots/optimization_results_0_{0}.pdf'.format(name))
     f.close()
+    
+    
+    fig, ax = plt.subplots(1, 3, figsize = (17, 5))
+    for i, sam in enumerate(list(samples)):
+        
+        sam_i_str = samples_str[i]                        
+        dy = (labels['spec_parallax'][sam] - labels['parallax'][sam]) / labels['parallax'][sam]
+        s = 0.5 * (np.percentile(dy, 84) - np.percentile(dy, 16))
+        print('1 sigma inferred parallax for {0} sample: {1}, {2}'.format(sam_i_str, 0.5 * (np.percentile(dy, 84) - np.percentile(dy, 16)), 0.25 * (np.percentile(dy, 97.5) - np.percentile(dy, 2.5))))
+    
+        sc = ax[i].scatter(labels['parallax'][sam], labels['spec_parallax'][sam], c = labels['LOGG'][sam], cmap = 'viridis_r', s = 10, vmin = 0, vmax = 2.2, label = r'$1\sigma={}$'.format(round(s, 3)), rasterized = True)
+        if i == 0:
+            cb = fig.colorbar(sc)
+            cb.set_label(r'$\log g$', fontsize = fsize)
+        ax[i].set_title(r'{} sample'.format(sam_i_str), fontsize = fsize)
+        if i == 2:
+            ax[i].set_title(r'$\varpi/\sigma_{\varpi} \geq 20$', fontsize = fsize)
+        ax[i].plot([-100, 100], [-100, 100], linestyle = '--', color = 'k')
+        ax[i].set_ylim(-0.5, 2)
+        ax[i].set_xlim(-0.5, 2)
+        ax[i].legend(frameon = True, fontsize = fsize)
+        if i == 0:
+            ax[i].tick_params(axis=u'both', direction='in', which='both')
+        else:
+            ax[i].tick_params(axis=u'both', direction='in', which='both', labelleft = False)            
+        ax[i].set_xlabel('Gaia parallax', fontsize = fsize)
[…]

### 2018-07-03 map_making.py

@@ -46,15 +46,6 @@ labels.rename_column('ra_1', 'ra')
 labels.rename_column('dec_1', 'dec')
 
 # -------------------------------------------------------------------------------
-# new cross match with Melissa's ages! (or Sanders+ 2018)
-# -------------------------------------------------------------------------------           
-
-hdu = fits.open('data/labels_match_ages_melissa.fits')
-melissa_labels = hdu[1].data
-melissa_labels = Table(melissa_labels)
-melissa_labels.rename_column('col2', 'ages')
-
-# -------------------------------------------------------------------------------
 # calculate cartesian coordinates
 # -------------------------------------------------------------------------------           
 
@@ -164,9 +155,9 @@ plt.savefig('plots/xy_vy.pdf', bbox_inches = 'tight', dpi = 120)
 plt.close()
 
 fig, ax = plt.subplots(1, 1, figsize = (10, 10))
-sc = plt.scatter(XS[vz_cut, 0], XS[vz_cut, 1], c = labels['parallax_error'][vz_cut], cmap = 'RdBu', vmin = 0, vmax = .2, s = 10, rasterized = True)
+sc = plt.scatter(XS[vz_cut, 0], XS[vz_cut, 1], c = labels['parallax_error'][vz_cut]**2 / (0.1)**2, cmap = 'RdBu', vmin = 0, vmax = 2, s = 10, rasterized = True)
 cb = plt.colorbar(sc, shrink = 0.82)
-cb.set_label(label = r'$\sigma_{\varpi}$', fontsize = fsize)
+cb.set_label(label = r'$\sigma^2_{\varpi,\,\rm Gaia}/\sigma^2_{\varpi,\,\rm inferred}$', fontsize = fsize)
 plt.xlim(Xlimits[0])
 plt.ylim(Xlimits[1])
 plt.tick_params(axis=u'both', direction='in', which='both')

### 2018-07-03 test_distances.py

@@ -1,0 +1,331 @@
+#!/usr/bin/env python2
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jul  2 11:10:48 2018
+
+@author: eilers
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import scipy.optimize as op
+import time
+import pickle
+from astropy.table import Column, Table, join, vstack, hstack
+import sys
+from astropy.io import fits
+from sklearn.decomposition import PCA
+from astropy import units as u
+from astropy.coordinates import SkyCoord
+import astropy.coordinates as coord
+from mpl_toolkits.mplot3d import Axes3D
+import corner
+
+
+# -------------------------------------------------------------------------------
+# plotting settings
+# -------------------------------------------------------------------------------
+
+matplotlib.rcParams['ytick.labelsize'] = 14
+matplotlib.rcParams['xtick.labelsize'] = 14
+fsize = 14
+
+# -------------------------------------------------------------------------------
+# open inferred labels
+# -------------------------------------------------------------------------------
+
+N = 45787
+Kfold = 4
+lam = 100
+name = 'N{0}_lam{1}_K{2}'.format(N, lam, Kfold)
+
+print('loading new labels...')   
+labels = Table.read('data/training_labels_new_{}_2.fits'.format(name), format = 'fits')    
+labels.rename_column('ra_1', 'ra')
+labels.rename_column('dec_1', 'dec')
+
+# -------------------------------------------------------------------------------
+# calculate cartesian coordinates
+# -------------------------------------------------------------------------------           
+
+spec_par = labels['spec_parallax'] * u.mas
+distance = spec_par.to(u.parsec, equivalencies = u.parallax())
+
[…]

### 2018-07-11 paper/sp.tex
equations

@@
 \documentclass[modern]{aastex62}
+\usepackage{amsmath}
 
@@
 \newcommand{\logg}{\log g}
+\DeclareMathOperator*{\argmin}{argmin}
 
@@
 
-foo
+The promise of mapping the Milky Way disk with infrared spectroscopy
+and red-giant stars.
 
-bar
+The information \gaia\ brings to this.
+
+The history and value (and disadvantages)
+of data-driven models for these kinds of problems.
+
+The dangers of aggressive machine-learning methods. Why will we just
+look at linear models? The disadvantages of this approach.
+
+The dangers of cutting on \gaia\ properties. Why we will not cut on
+\gaia\ parallaxes. The disadvantages of this approach.
 
@@
 
-foo
+The \gaia\ Catalog describes a likelihood function!
 
-bar
+There is a parallax offset.
+
+We are only looking at \apogee\ luminous giants (selected spectroscopically).
+
+We use photometry from \gaia, \zmass, \wise.
 
 \section{Method}
+
+\begin{eqnarray}
+Q_n &\equiv& 0.01\,\varpi_n\,10^{0.2\,K_n}
+\\
+\sigma_n &\equiv& \sigma_{\varpi n}\,10^{0.2\,K_n}
+\\
+Q_n &=& \theta_Q\cdot x_n + \mbox{noise}
+\\
+\chi^2 &\equiv& \sum_{n=1}^N \frac{[Q_n - \theta_Q\cdot x_n]^2}{\sigma_n^2}
+\\
+\hat{\theta}_Q &\leftarrow& \argmin_{\theta_Q}\left[\frac{1}{2}\,\chi^2 + \lambda_Q\,||\theta_Q||_1^1\right]
+\quad ,
+\end{eqnarray}
+where...
+
+\begin{eqnarray}
+Q_n &=& \exp(\theta_M\cdot x_n) + \mbox{noise}
+\\
+M_{Kn} &\approx& \frac{1}{0.2\,\ln 10}\,\theta_M\cdot x_n
+\\
+\chi^2 &\equiv& \sum_{n=1}^N \frac{[Q_n - \exp(\theta_M\cdot x_n)]^2}{\sigma_n^2}
+\\
+\hat{\theta}_M &\leftarrow& \argmin_{\theta_M}\left[\frac{1}{2}\,\chi^2 + \lambda_M\,||\theta_M||_1^1\right]
+\quad ,
+\end{eqnarray}
+where...
 

### 2018-07-11 infer_parallax.py
small changes

@@ -35,6 +35,9 @@ print('loading labels...')
 hdu = fits.open('data/training_labels_parent.fits')
 labels = Table(hdu[1].data)
 
+offset = 0.029 # mas as per Lindegren et al. 2018
+labels['parallax'] += offset
+
 if prediction: 
     
     print('loading spectra...')
@@ -110,7 +113,7 @@ def check_H_func(x, y, A, lam, ivar):
 
 Kfold = 2
 lam = 100                       # hyperparameter -- needs to be tuned!
-name = 'N{0}_lam{1}_K{2}'.format(len(labels), lam, Kfold)
+name = 'N{0}_lam{1}_K{2}_offset'.format(len(labels), lam, Kfold)
 
 if prediction:        
     
@@ -245,7 +248,7 @@ if not prediction:
         ax[i].set_xlabel('Gaia parallax', fontsize = fsize)
     ax[0].set_ylabel('inferred parallax', fontsize = fsize)
     plt.subplots_adjust(wspace = 0.08)
-    plt.savefig('plots/parallax_inferred_{0}_vis.pdf'.format(name), dpi = 120)
+    plt.savefig('plots/parallax/parallax_inferred_{0}_vis.pdf'.format(name), dpi = 120)
     plt.close()
     
     fig, ax = plt.subplots(1, 3, figsize = (17, 5))
@@ -273,7 +276,7 @@ if not prediction:
         ax.set_xlabel('Gaia parallax', fontsize = fsize)
     cb = fig.colorbar(sc)
     cb.set_label(r'visibility periods used', fontsize = fsize)
-    plt.savefig('plots/parallax_inferred_{0}_vis_density.pdf'.format(name))
+    plt.savefig('plots/parallax/parallax_inferred_{0}_vis_density.pdf'.format(name))
     plt.close()
     
     fig, ax = plt.subplots(1, 3, figsize = (17, 5))
@@ -301,7 +304,7 @@ if not prediction:
         ax.set_xlabel('Gaia parallax', fontsize = fsize)
     cb = fig.colorbar(sc)
     cb.set_label(r'$\sigma_{\varpi}$', fontsize = fsize)
-    plt.savefig('plots/parallax_inferred_{0}_varpi_density.pdf'.format(name))
+    plt.savefig('plots/parallax/parallax_inferred_{0}_varpi_density.pdf'.format(name))
     plt.close()
     
     fig, ax = plt.subplots(1, 3, figsize = (17, 5))
[…]

### 2018-07-11 map_making.py

@@ -21,7 +21,8 @@ from astropy.coordinates import SkyCoord
 import astropy.coordinates as coord
 from mpl_toolkits.mplot3d import Axes3D
 import corner
-
+from scipy.stats import binned_statistic_2d
+import astropy
 
 # -------------------------------------------------------------------------------
 # plotting settings
@@ -36,7 +37,7 @@ fsize = 14
 # -------------------------------------------------------------------------------
 
 N = 45787
-Kfold = 4
+Kfold = 2
 lam = 100
 name = 'N{0}_lam{1}_K{2}'.format(N, lam, Kfold)
 
@@ -77,6 +78,9 @@ vX_GC_sun_kms = -U_LSR_kms           # = -U              [km/s]
 vY_GC_sun_kms =  V_LSR_kms+vcirc_kms # = V+v_circ(R_Sun) [km/s]
 vZ_GC_sun_kms =  W_LSR_kms           # = W               [km/s]
 
+# keep proper motion of Sgr A* constant! 
+vY_GC_sun_kms = X_GC_sun_kpc * vY_GC_sun_kms / 8.
+
 gc = coord.Galactocentric(galcen_distance = X_GC_sun_kpc*u.kpc,
                           galcen_v_sun = coord.CartesianDifferential([-vX_GC_sun_kms, vY_GC_sun_kms, vZ_GC_sun_kms] * u.km/u.s),
                           z_sun = Z_GC_sun_kpc*u.kpc)
@@ -107,7 +111,7 @@ fig.savefig('plots/corner.pdf')
 plt.close()
 
 # -------------------------------------------------------------------------------
-# maps
+# rings
 # -------------------------------------------------------------------------------           
 
 def overplot_ring(r):
@@ -123,8 +127,120 @@ def overplot_rings():
         overplot_ring(r)
     return
 
+# -------------------------------------------------------------------------------
+# theoretical rotation curves
+# -------------------------------------------------------------------------------           
+
[…]

### 2018-07-11 test_distances.py

@@ -36,9 +36,9 @@ fsize = 14
 # -------------------------------------------------------------------------------
 
 N = 45787
-Kfold = 4
+Kfold = 2
 lam = 100
-name = 'N{0}_lam{1}_K{2}'.format(N, lam, Kfold)
+name = 'N{0}_lam{1}_K{2}_offset'.format(N, lam, Kfold)
 
 print('loading new labels...')   
 labels = Table.read('data/training_labels_new_{}_2.fits'.format(name), format = 'fits')    
@@ -118,7 +118,7 @@ clus = 'ngc6791'
 lat = 'NGC 6791'
 ra_clus_hms = '19:20:53.0' 
 dec_clus_dms = '+37:46:18'
-distance = 5.853 
+distance = 4.078
 feh_clus = '+0.47\pm 0.07'
 t.add_row([clus, ra_clus_hms, dec_clus_dms, feh_clus, distance, lat])
 
@@ -317,15 +317,63 @@ for i, clus in enumerate(cluster_list):
     ax[c, r].set_title(r'{0}, $\rm [Fe/H] = {1}$'.format(t['lat_name'][k][0], t['FE_H'][k][0]), fontsize = 14)
     ax[c, r].set_xlabel(r'$\varpi$', fontsize = 14)
     ax[c, r].set_xlim(-0.1, 1.)
-
-plt.savefig('plots/test_copen_clusters.pdf')
+plt.savefig('plots/open_clusters/test_open_clusters_{}.pdf'.format(name))
 
 # -------------------------------------------------------------------------------
 # take Sagittarius region...
 # -------------------------------------------------------------------------------
 
+hdu = fits.open('data/Sgr_Candidate_in_Gaia.fits')
+data_sgr = hdu[1].data
 
+indices = np.arange(len(labels))     
+xx = []              
+for i in data_sgr['APOGEE_ID']:
+    foo = labels['APOGEE_ID'] == i
+    if np.sum(foo) == 1:
+        xx.append(indices[foo][0])
+xx = np.array(xx)
 
+labels_sgr = labels[xx]
 
+XS_sgr = np.vstack([labels_sgr['ra'], labels_sgr['dec'], labels_sgr['spec_parallax']]).T
+Xlabels_sgr = ['RA', 'DEC', r'$\varpi$']
+fig = corner.corner(XS_sgr, labels = Xlabels_sgr, plot_datapoints = True)
[…]

### 2018-07-11 infer_magnitudes.py
magnitudes

@@ -59,10 +59,10 @@ if prediction:
 
 
 # -------------------------------------------------------------------------------
-# add absolute Q magnitudes
+# add absolute K magnitudes
 # -------------------------------------------------------------------------------
 
-    print('calculating Q...')
+    print('calculating Q_K...')
     m_K = labels['K']
     
     Q_factor = 10**(0.2 * m_K) / 100.                 # assumes parallaxes is in mas
@@ -91,14 +91,16 @@ if prediction:
 # linear algebra
 # -------------------------------------------------------------------------------
 
-def H_func(x, y, A, lam, ivar):    
-    H = 0.5 * np.dot((y - np.dot(A, x)).T, ivar * (y - np.dot(A, x))) + lam * np.sum(np.abs(x))
-    dHdx = -1. * np.dot(A.T, ivar * (y - np.dot(A, x))) + lam * np.sign(x)
+def H_func(x, y, A, lam, ivar):   
+    y_model = np.exp(np.dot(A, x))
+    dy = y - y_model
+    H = 0.5 * np.dot(dy.T, ivar * dy) + lam * np.sum(np.abs(x))
+    dHdx = -1. * np.dot(A.T * y_model[None, :], ivar * dy) + lam * np.sign(x)
     return H, dHdx
 
 def check_H_func(x, y, A, lam, ivar):
     H0, dHdx0 = H_func(x, y, A, lam, ivar)
-    dx = 0.001 # magic
+    dx = 0.0001 # magic
     for i in range(len(x)):
         x1 = 1. * x
         x1[i] += dx
@@ -112,7 +114,7 @@ def check_H_func(x, y, A, lam, ivar):
 # -------------------------------------------------------------------------------
 
 Kfold = 2
-lam = 30                       # hyperparameter -- needs to be tuned!
+lam = 100                       # hyperparameter -- needs to be tuned!
 name = 'N{0}_lam{1}_K{2}_offset'.format(len(labels), lam, Kfold)
 
 if prediction:        
@@ -179,7 +181,7 @@ if prediction:
         print(res)                       
                                
         # prediction
-        y_pred = np.dot(A_all[valid, :], res.x) 
+        y_pred = np.exp(np.dot(A_all[valid, :], res.x))
         y_pred_all[valid] = y_pred
                                            
         f = open('optimization/opt_results_{0}_{1}.pickle'.format(k, name), 'wb')

### 2018-07-12 infer_magnitudes.py
rotation

@@ -114,8 +114,8 @@ def check_H_func(x, y, A, lam, ivar):
 # -------------------------------------------------------------------------------
 
 Kfold = 2
-lam = 100                       # hyperparameter -- needs to be tuned!
-name = 'N{0}_lam{1}_K{2}_offset'.format(len(labels), lam, Kfold)
+lam = 3                      # hyperparameter -- needs to be tuned!
+name = 'N{0}_lam{1}_K{2}_mag_offset'.format(len(labels), lam, Kfold)
 
 if prediction:        
     

### 2018-07-12 test_distances.py

@@ -38,7 +38,7 @@ fsize = 14
 N = 45787
 Kfold = 2
 lam = 100
-name = 'N{0}_lam{1}_K{2}_offset'.format(N, lam, Kfold)
+name = 'N{0}_lam{1}_K{2}_mag_offset'.format(N, lam, Kfold)
 
 print('loading new labels...')   
 labels = Table.read('data/training_labels_new_{}_2.fits'.format(name), format = 'fits')    

### 2018-07-12 rotation_curve_new.py
rotation

@@ -1,0 +1,192 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Jul  6 09:43:21 2018
+
+@author: eilers
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import scipy.optimize as op
+import time
+import pickle
+from astropy.table import Column, Table, join, vstack, hstack
+import sys
+from astropy.io import fits
+from sklearn.decomposition import PCA
+from astropy import units as u
+from astropy.coordinates import SkyCoord
+import astropy.coordinates as coord
+from mpl_toolkits.mplot3d import Axes3D
+import corner
+from scipy.stats import binned_statistic_2d
+from plotting_helpers import histcont
+
+# -------------------------------------------------------------------------------
+# plotting settings
+# -------------------------------------------------------------------------------
+
+matplotlib.rcParams['ytick.labelsize'] = 14
+matplotlib.rcParams['xtick.labelsize'] = 14
+fsize = 14
+
+# -------------------------------------------------------------------------------
+# open inferred labels
+# -------------------------------------------------------------------------------
+
+N = 45787
+Kfold = 2
+lam = 100
+name = 'N{0}_lam{1}_K{2}_mag_offset'.format(N, lam, Kfold)
+
+print('loading new labels...')   
+labels = Table.read('data/training_labels_new_{}_2.fits'.format(name), format = 'fits')    
+labels.rename_column('ra_1', 'ra')
+labels.rename_column('dec_1', 'dec')
+
+# -------------------------------------------------------------------------------
+# calculate cartesian coordinates
+# -------------------------------------------------------------------------------           
+
+spec_par = labels['spec_parallax'] * u.mas
+distance = spec_par.to(u.parsec, equivalencies = u.parallax())
+
+cs = coord.ICRS(ra = labels['ra'] * u.degree, 
+                dec = labels['dec'] * u.degree, 
[…]

### 2018-07-22 paper/sp.tex
update abstract to logs

@@
 (plus \gaia, \zmass, and \wise\ photometry)
-that predicts inverse-square-root luminosity for red stars with $0<\logg<2.2$.
-The inverse-square-root is used to make the noise model close to Gaussian,
+that predicts absolute magnitude (logarithmic luminosity) for red stars with $0<\logg<2.2$.
+The training makes use of a justified likelihood function for the \gaia\ parallaxes,
 so that the training can be performed with full use of uncertainties and
-without any cuts on parallax or parallax signal-to-noise ratio;
-the model includes an L1 regularization that zeros out the contributions of
-most spectral pixels to the parallax estimates; and the training was performed
-with leave-out subsamples so no star's astrometry is used even indirectly in its
+without any cuts on parallax or parallax signal-to-noise ratio.
+The model includes an L1 regularization that zeros out the contributions of
+most spectral pixels to the parallax estimates.
+The training is performed
+with leave-out subsamples such that no star's astrometry is used even indirectly in its
 spectroscopic parallax estimate.
-The model is flexible enough to correct for extinction by dust without any
+The model is flexible enough---and has enough photometric information---to
+correct for extinction by dust without any
 input of explicit extinction or reddening estimates.
 We re-label each star in the sample
-with a new spectroscopic parallax; this new parallax has an empirical uncertainty of
-10\,percent, which is more precise than the \gaia\ parallax
+with a new spectroscopic parallax; this new parallax has an empirical (1-sigma)
+uncertainty of 9~percent, which is more precise than the \gaia\ parallax
 for the vast majority of targets.

### 2018-07-22 paper/sp.tex
document tweaks; assumptions

@@
 
-\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing
+% tweak the modern style -- trust in Hogg
+\setlength{\parindent}{1.0\baselineskip}
+\addtolength{\textheight}{0.5in}
+\addtolength{\topmargin}{-0.3in}
 
-\title{Accurate spectroscopic estimates of luminosity and distance\\
-       for luminous red-giant stars}
+\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust in Hogg
 
-\correspondingauthor{Anna-Christina Eilers}
-\email{eilers@mpia.de}
+\title{\textbf{%
+Accurate spectroscopic estimates of luminosity and distance\\
+for luminous red-giant stars%
+}}
 
@@
 surveys
- ---
-parallaxes
+% ---
+%parallaxes
  ---
@@
 
-\section{Data}
+\section{Assumptions of the method}
 
-The \gaia\ Catalog describes a likelihood function!
+Our position is that a methodological technique is correct inasmuch as
+it delivers correct or best results \emph{under its particular assumptions}.
+In that spirit, we present here the assumptions of the method
+explicitly.
+We will return to these assumptions in the Discussion section below,
+to address the costs and benefits of each in more depth.
+\begin{description}
+\item[features] bar
 
-There is a parallax offset.
+\item[good features] bar
 
-We are only looking at \apogee\ luminous giants (selected spectroscopically).
+\item[representativeness] bar
 
-We use photometry from \gaia, \zmass, \wise.
+\item[sparsity] bar
+
+\item[linearity] bar
+
+\item[likelihood] bar
+\end{description}
 
@@
 
-\section{Validation}
+\section{Data}
 
-\section{Results and Discussion}
+The \gaia\ Catalog describes a likelihood function!
+
+There is a parallax offset.
+
+We are only looking at \apogee\ luminous giants (selected spectroscopically).
+
+We use photometry from \gaia, \zmass, \wise.
+
+\section{Results and validation}
+
+\section{Discussion}
 

### 2018-07-22 paper/sp.tex
first shot at assumptions

@@
 \begin{description}
-\item[features] bar
+\item[features] Perhaps our most fundamental assumption is that the intrinsic
+luminosity of a star can be predicted from the features we provide, which are
+the full set of (pseudo-continuum-normalized) pixels from the \apogee\ spectrum,
+plus the $G$ and $G_{BP}-G_{RP}$ photometry from \gaia,
+plus the $J$, $H$, and $K_s$ photometry from \zmass,
+plus the $W_1$ and $W_2$ photometry from \wise.
+That is, we assume that these spectrophotometric \emph{features} are sufficient
+to predict the intrinsic luminosity in the face of variations in stellar
+age, evolutionary phase, composition and other parameters, and also interstellar
+extinction.
 
-\item[good features] bar
+\item[good features] We assume that the spectrophotometric features are known for
+each star with such high fidelity (both precision and accuracy) that we do not
+need to account for errors or uncertainties or biases in the features.
+That is, we assume that the features are substantially higher in signal-to-noise than the
+quantities we are trying to predict.
+This is true for most features for most stars, but it does not hole universally.
 
-\item[representativeness] bar
+\item[representativeness] We assume that the training set constructed from the overlap
+of \gaia\ and \apogee\ data sets constitutes a representative sample of stars,
+sufficiently representative to train the model for all other stars.
+Although this assumption is not terrible, it has a weakness:
+The stars at greatest distances and greatest local (angular) stellar crowding have
+the least precise \gaia\ parallax measurements, and therefore will effectively get
+less weight in the fits performed to train the model.
 
-\item[sparsity] bar
+\item[sparsity] We expect that only a small subset of the full complement of
+\apogee\ spectral pixels will be relevant to the prediction of stellar luminosity.
+That is, we expect that many of the pixels will or ought to get no weight in the
+final model that we use to predict luminosities and distances.
 
-\item[linearity] bar
+\item[linearity] Perhaps the most restrictive assumption of this work is that
+the stellar logarithmic luminosity (or, equivalently, the absolute magnitude) in
+the \zmass\ $K_s$~band can be predicted as a \emph{completely linear function} of
+the chosen features. We are only assuming this linearity in a small range of stellar
+surface gravity $\logg$, but this assumption is strong, and limits strongly the
+flexibility or capacity of the model.
 
-\item[likelihood] bar
+\item[likelihood] We assume that the \gaia\ parallaxes are generated by a particular
+stochastic process, in which the difference between the \gaia-Catalog parallax (plus
+a small \gaia-recommended offset) and the true parallax is effectively drawn from a
+Gaussian with a width set by the \gaia-Catalog uncertainty on the parallax.
+This is the standard assumption in all properly probabilistic \gaia-Mission inferences
+to date, but it subsumes a number of related assumptions, like that the \gaia\ noise
+model is correct, that the stars are only covariant at negligible levels, and that
+there are no significant outliers or large-scale systematics.
 \end{description}

### 2018-07-24 paper/sp.tex
method section

@@
 That is, we assume that the features are substantially higher in signal-to-noise than the
-quantities we are trying to predict.
-This is true for most features for most stars, but it does not hole universally.
+quantities we are trying to predict; in particular we are assuming that the photometry
+and the spectroscopy is better measured than the astrometry.
+This is true for most features for most stars, but it does not hold universally.
 
@@
 flexibility or capacity of the model.
+We make it to ensure that our method is easy to optimize, and the results are easy
+to interpret.
 
@@
 
+Under the assumptions stated in the previous Section, there are not many choices
+for building a model of the stars that is justifiable probabilistically.
+Here we lay out the model and methodology, and apply it in the following Section.
+
+Since the photometry is (by the ``good features'' assumption above)
+better measured than the parallaxes, we
+presume that the parallax uncertainties dominate
+the uncertainties in the luminosity measurements.
+For each star $n$ in our training set, we define an inverse-square-root
+luminosity $Q_n$ estimate that will be subject to Gaussian noise under these assumptions.
 \begin{eqnarray}
-Q_n &\equiv& 0.01\,\varpi_n\,10^{0.2\,K_n}
+Q_n &\equiv& \varpi_n\,10^{0.2\,(K_n - 10)}
 \\
-\sigma_n &\equiv& \sigma_{\varpi n}\,10^{0.2\,K_n}
+\sigma_n &\equiv& \sigma_{\varpi n}\,10^{0.2\,(K_n - 10)}
+\quad,
+\end{eqnarray}
+where
+$\varpi_n$ is the \gaia\ Catalog entry for the parallax for star $n$,
+$K_n$ is the \zmass\ Catalog $K_s$-band magnitude for star $n$,
+the $Q_n$ have Gaussian errors of root-variance $\sigma_n$,
+and $\sigma_{\varpi n}$ is the \gaia\ Catalog entry for the uncertainty on parallax
+for star $n$.
+In what follows, the $Q_n$ are the training data.
+Naively, the $Q_n$ are related to the absolute magnituges $M_{Kn}$ for the stars by
+\begin{eqnarray}
+M_{Kn} &\approx& \frac{1}{0.2\,\ln 10}\,\ln Q_n
+\quad.
+\end{eqnarray}
+These would be the absolute magnitudes of the stars under the
+``the distance is the inverse of the
+measured parallax'' assumption, although we never, in this work, ever make that
+assumption, by God.
+We construct the $Q_n$ because they are related directly to the absolute magnitudes,
+but they (under assumptions) are subject to known Gaussian noise.
+
+The log-likelihood function can be expressed heuristically as
+\begin{eqnarray}
+Q_n &=& \exp(\theta\cdot x_n) + \mbox{noise}
 \\
-Q_n &=& \theta_Q\cdot x_n + \mbox{noise}
-\\
-\chi^2 &\equiv& \sum_{n=1}^N \frac{[Q_n - \theta_Q\cdot x_n]^2}{\sigma_n^2}
-\\
-\hat{\theta}_Q &\leftarrow& \argmin_{\theta_Q}\left[\frac{1}{2}\,\chi^2 + \lambda_Q\,||\theta_Q||_1^1\right]
+\chi^2 &\equiv& \sum_{n=1}^N \frac{[Q_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
 \quad ,
 \end{eqnarray}
-where...
+where
+the model is that the $\ln Q_n$ can be expressed as a linear combination of the components
+of some $D$-dimensional feature vector $x_n$,
+$\theta$ is a $D$-vector of linear coefficients,
+and $\chi^2$ is the negative-log-likelihood
+under the assumption of known Gaussian noise and
+that there are $N$ independently measured stars $n$.
+This feature vector $x_n$ contains photometry in a few bands, and all 8000-ish pixels
+in the pseudo-continuum-normalized \apogee\ spectrum, so $D$ is on the order of 8000.
 
+In addition, we assume that many entries in the $D$-vector will be zero
+(the ``sparsity'' assumption).
+In order to encourage this, we optimize not $\chi^2$ but a regularized objective function
 \begin{eqnarray}
-Q_n &=& \exp(\theta_M\cdot x_n) + \mbox{noise}
-\\
-M_{Kn} &\approx& \frac{1}{0.2\,\ln 10}\,\theta_M\cdot x_n
-\\
-\chi^2 &\equiv& \sum_{n=1}^N \frac{[Q_n - \exp(\theta_M\cdot x_n)]^2}{\sigma_n^2}
-\\
-\hat{\theta}_M &\leftarrow& \argmin_{\theta_M}\left[\frac{1}{2}\,\chi^2 + \lambda_M\,||\theta_M||_1^1\right]
+\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2 + ||\lambda\cdot\theta||_1^1\right]
 \quad ,
 \end{eqnarray}
-where...
+where
+$\lambda$ is a $D$-vector of regularization parameters (such that we can regularize
+different parts of the $x_n$ vectors differently),
+and $||x||_1^1$ is the L1-norm or sum of absolute values of the components of $x$.
+This kind of regularization adds a convex term to the optimization and leads to
+sparse optima.
+We will set the value $\lambda$ of the $D$-vector of regularization strengths by a
+very coarse cross-validation.
 

### 2018-07-24 paper/sp.tex
todo list

@@
-%% This file is part of the spectroscopic_parallax project
+w%% This file is part of the spectroscopic_parallax project
 %% Copyright 2018 the authors.
+
+% To-do items
+% -----------
+% - draft introduction
+% - draft data+results section
+% - draft discussion
+% - write list of things we never do!
+% - make sure we address the cross-validation point
+% - make sure we address the sensibleness and generality of the linear model for the photometric space
 
@@
 stochastic process, in which the difference between the \gaia-Catalog parallax (plus
-a small \gaia-recommended offset) and the true parallax is effectively drawn from a
+a small \gaia-recommended offset HOGG CITE) and the true parallax is effectively drawn from a
 Gaussian with a width set by the \gaia-Catalog uncertainty on the parallax.
@@
 \end{description}
+
+HOGG: LIST of things we don't do, like don't cut the sample on
+astrometric quantities, and why not!! And like don't assume distance
+is inverse-measured-parallax.
 
@@
 \\
-\chi^2 &\equiv& \sum_{n=1}^N \frac{[Q_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
+\chi^2(\theta) &\equiv& \sum_{n=1}^N \frac{[Q_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
 \quad ,
@@
 $\theta$ is a $D$-vector of linear coefficients,
-and $\chi^2$ is the negative-log-likelihood
+and $\chi^2(\theta)$ is the negative-log-likelihood for the parameters
 under the assumption of known Gaussian noise and
@@
 \begin{eqnarray}
-\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2 + ||\lambda\cdot\theta||_1^1\right]
+\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi(\theta)^2 + ||\lambda\cdot\theta||_1^1\right]
 \quad ,

### 2018-07-24 paper/sp.tex

@@
-w%% This file is part of the spectroscopic_parallax project
+%% This file is part of the spectroscopic_parallax project
 %% Copyright 2018 the authors.
@@
 \begin{eqnarray}
-M_{Kn} &\approx& \frac{1}{0.2\,\ln 10}\,\ln Q_n
+M_{Kn} &\approx& 5\,\log_{10} Q_n
 \quad.
@@
 $\theta$ is a $D$-vector of linear coefficients,
-and $\chi^2(\theta)$ is the negative-log-likelihood for the parameters
+and $\chi^2(\theta)$ is the negative-log-likelihood for the parameters $\theta$
 under the assumption of known Gaussian noise and
@@
 \begin{eqnarray}
-\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi(\theta)^2 + ||\lambda\cdot\theta||_1^1\right]
+\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2(\theta) + ||\lambda\cdot\theta||_1^1\right]
 \quad ,
@@
 very coarse cross-validation.
+
+Once the model is optimized, the prediction for a new stellar absolute magnitude
+$M_{Km}$ for a star $m$ in the validation or test set is
+\begin{eqnarray}
+Q_m &=& \exp(\theta\cdot x_m)
+\\
+M_{Km} &=& 5\,\log_{10} Q_m
+\quad ,
+\end{eqnarray}
+where $x_m$ is the feature $D$-vector for star $m$.
+Thus the model can be trained on a training set of stars and applied to make
+predictions for a validation set or a test set.
 
@@
 \section{Discussion}
+
+What did we do and what's cool?
+
+Why did we make the assumptions we made and what did they gain us and cost us?
+
+What is this useful for and how do you use them safely?
+
+(Mention independence of every result from the \gaia\ astrometry.)
+
+Come back to SDSS-V and show the rotation curve of the Galaxy.
 

### 2018-07-24 infer_magnitudes.py
new

@@ -29,7 +29,7 @@ fsize = 14
 # -------------------------------------------------------------------------------
 
 # make plots?
-prediction = True
+prediction = False
 
 print('loading labels...')
 hdu = fits.open('data/training_labels_parent.fits')

### 2018-07-24 normalize_all_spectra.py

@@ -120,8 +120,7 @@ def NormalizeData(dataall):
         
     Nlambda, Nstar, foo = dataall.shape
     
-    pixlist = np.loadtxt('data/pixtest8_dr13.txt', usecols = (0,), unpack = 1)
-    pixlist = map(int, pixlist)
+    pixlist = np.loadtxt('data/pixtest8_dr13.txt', usecols = (0,), unpack = 1).astype(int)
     LARGE  = 3.0                          # magic LARGE sigma value
    
     continuum = np.zeros((Nlambda, Nstar))

### 2018-07-24 parent_sample_dr15.py

@@ -1,0 +1,174 @@
+#!/usr/bin/env python2
+# -*- coding: utf-8 -*-
+"""
+Created on Tue Jun  5 15:15:21 2018
+
+@author: eilers
+"""
+
+import numpy as np
+import pickle
+from astropy.table import Column, Table, join, vstack, hstack
+from astropy.io import fits
+import os.path
+import subprocess
+
+from normalize_all_spectra import LoadAndNormalizeData, NormalizeData
+
+# -------------------------------------------------------------------------------
+# load catalogs
+# -------------------------------------------------------------------------------
+
+print("opening Gaia DR2 and APOGEE cross match catalogue. ")
+
+tbl1 = Table.read('data/gaiadr2_allStar-t9-l31c-58158.fits', 1)
+tbl1.rename_column('apstar_id', 'APSTAR_ID')
+tbl2 = Table.read('data/allStar-t9-l31c-58158.fits', 1)
+training_lab = join(tbl1, tbl2, keys='APSTAR_ID')
+
+print("Gaia DR2 and APOGEE cross match catalogue: {} entries. ".format(len(training_lab))) 
+
+# -------------------------------------------------------------------------------
+# cut in logg... take only RGB stars!
+# -------------------------------------------------------------------------------
+
+parent_logg_cut = 2.2
+cut = np.logical_and(training_lab['LOGG'] <= parent_logg_cut, training_lab['LOGG'] > 0.)
+training_lab = training_lab[cut]              
+print('logg <= {0} cut: {1}'.format(parent_logg_cut, len(training_lab)))
+
+# -------------------------------------------------------------------------------
+# cut in Q_K (where training_labels['K'] was missing! (i.e. -9999))
+# ------------------------------------------------------------------------------- 
+
+cut = np.where(training_lab['K'] > 0)
+training_lab = training_lab[cut]              
+print('remove missing K: {}'.format(len(training_lab)))
+
+# -------------------------------------------------------------------------------
+# remove missing data
+# ------------------------------------------------------------------------------- 
+
+cut = np.where(np.isfinite(training_lab['bp_rp']))
+training_lab = training_lab[cut]              
+print('remove bad bp-rp: {}'.format(len(training_lab)))
+
+# -------------------------------------------------------------------------------
+# add WISE catalog -- NEEDS TO BE RE-DONE!!
+# -------------------------------------------------------------------------------
+
+print('match to WISE...')
+hdu = fits.open('data/gaia_apogeedr15_wise.fits')
+wise_data = hdu[1].data
[…]

### 2018-07-24 rotation_curve.py

@@ -1,0 +1,218 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Jul  6 09:43:21 2018
+
+@author: eilers
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import scipy.optimize as op
+import time
+import pickle
+from astropy.table import Column, Table, join, vstack, hstack
+import sys
+from astropy.io import fits
+from sklearn.decomposition import PCA
+from astropy import units as u
+from astropy.coordinates import SkyCoord
+import astropy.coordinates as coord
+from mpl_toolkits.mplot3d import Axes3D
+import corner
+from scipy.stats import binned_statistic_2d
+from plotting_helpers import histcont
+
+# -------------------------------------------------------------------------------
+# plotting settings
+# -------------------------------------------------------------------------------
+
+matplotlib.rcParams['ytick.labelsize'] = 14
+matplotlib.rcParams['xtick.labelsize'] = 14
+fsize = 14
+
+# -------------------------------------------------------------------------------
+# open inferred labels
+# -------------------------------------------------------------------------------
+
+N = 45787
+Kfold = 2
+lam = 100
+name = 'N{0}_lam{1}_K{2}'.format(N, lam, Kfold)
+
+print('loading new labels...')   
+labels = Table.read('data/training_labels_new_{}_2.fits'.format(name), format = 'fits')    
+labels.rename_column('ra_1', 'ra')
+labels.rename_column('dec_1', 'dec')
+
+# take only stars in mid-plane 
+vz_cut = (abs(labels['b']) < 2) 
+labels = labels[vz_cut]
+N = len(labels)
+
+# -------------------------------------------------------------------------------
+# re-sampling of each star 
+# -------------------------------------------------------------------------------           
+
+N_samples = 10
+
+# uncertainties in RA, DEC, parallax, proper motions in RA and DEC, position of the sun, velocity of the sun
+# fix position of the sun at the moment! 
[…]

### 2018-07-24 rotation_curve_cylindrical.py

@@ -1,0 +1,295 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Jul  6 09:43:21 2018
+
+@author: eilers
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import scipy.optimize as op
+import time
+import pickle
+from astropy.table import Column, Table, join, vstack, hstack
+import sys
+from astropy.io import fits
+from sklearn.decomposition import PCA
+from astropy import units as u
+from astropy.coordinates import SkyCoord, Galactocentric
+import astropy.coordinates as coord
+from mpl_toolkits.mplot3d import Axes3D
+import corner
+from scipy.stats import binned_statistic_2d
+from plotting_helpers import histcont
+
+
+# -------------------------------------------------------------------------------
+# plotting settings
+# -------------------------------------------------------------------------------
+
+matplotlib.rcParams['ytick.labelsize'] = 14
+matplotlib.rcParams['xtick.labelsize'] = 14
+fsize = 14
+
+def overplot_ring(r):
+    tiny = 1e-4
+    thetas = np.arange(0., 2*np.pi + tiny, 0.001 * np.pi)
+    xs = r * np.cos(thetas)
+    ys = r * np.sin(thetas)
+    plt.plot(xs, ys, "k-", alpha=0.2, lw=1, zorder = -np.inf)
+    plt.scatter(0, 0, s = 10, color = 'k', alpha=0.2)
+    return
+
+def overplot_rings():
+    for r in [5, 10, 15, 20, 25, 30]:
+        overplot_ring(r)
+    return
+
+# -------------------------------------------------------------------------------
+# open inferred labels
+# -------------------------------------------------------------------------------
+
+N = 45787
+Kfold = 2
+lam = 30
+name = 'N{0}_lam{1}_K{2}_mag_allcolors_offset_lams'.format(N, lam, Kfold)
+
+print('loading new labels...')   
[…]

### 2018-07-24 rotation_curve_uncertainties.py

@@ -1,0 +1,462 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Jul  6 09:43:21 2018
+
+@author: eilers
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import scipy.optimize as op
+import time
+import pickle
+from astropy.table import Column, Table, join, vstack, hstack
+import sys
+from astropy.io import fits
+from sklearn.decomposition import PCA
+from astropy import units as u
+from astropy.coordinates import SkyCoord
+import astropy.coordinates as coord
+from mpl_toolkits.mplot3d import Axes3D
+import corner
+from scipy.stats import binned_statistic_2d
+from plotting_helpers import histcont
+
+# -------------------------------------------------------------------------------
+# plotting settings
+# -------------------------------------------------------------------------------
+
+matplotlib.rcParams['ytick.labelsize'] = 14
+matplotlib.rcParams['xtick.labelsize'] = 14
+fsize = 14
+
+# -------------------------------------------------------------------------------
+# open inferred labels
+# -------------------------------------------------------------------------------
+
+N = 45787
+Kfold = 2
+lam = 30
+name = 'N{0}_lam{1}_K{2}_mag_allcolors_offset'.format(N, lam, Kfold)
+
+print('loading new labels...')   
+labels = Table.read('data/training_labels_new_{}_2.fits'.format(name), format = 'fits')    
+labels.rename_column('ra_1', 'ra')
+labels.rename_column('dec_1', 'dec')
+
+# -------------------------------------------------------------------------------
+# position of the sun
+# -------------------------------------------------------------------------------           
+
+# Galactocentric position of the Sun:
+X_GC_sun_kpc = 8.3    #[kpc]
[…]

### 2018-07-25 paper/sp.tex
outline, acks, words

@@
 % - draft discussion
-% - write list of things we never do!
 % - make sure we address the cross-validation point
 % - make sure we address the sensibleness and generality of the linear model for the photometric space
+% - address convexity of the model...?
 
@@
 \affiliation{Max-Planck-Institut f\"ur Astronomie}
-
-\author{Melissa Ness}
-\affiliation{Flatiron Institute}
-\affiliation{Department of Astronomy, Columbia University}
 
@@
 
-HOGG: LIST of things we don't do, like don't cut the sample on
-astrometric quantities, and why not!! And like don't assume distance
-is inverse-measured-parallax.
+In addition to making the above assumptions, we also avoid various practices
+with the data that are tempting but lead to strong biases.
+For example, we never cut on \gaia\ parallax or parallax signal-to-noise.
+The common practices of cutting to parallaxes that are good to 20~percent,
+or Catalog parallaxes that are positive, or parallaxes that are smaller or
+larger than something, are all practices that will bias results on stellar
+collections.
+That is, if you cut on parallax or parallax signal-to-noise and you subsequently
+take an average of parallaxes for some population, or perform a regression (as we
+do here), the results of that average or regression will be strongly biased.
+We never cut on parallax or parallax signal-to-noise.
+By using a justifiable likelihood function (the ``likelihood'' assumption above),
+we can use all of the \gaia\ parallaxes without the low signal-to-noise and
+negative parallaxes causing trouble for our regressions, and without the biases
+that enter when cuts are made.
+
+Along similar lines, we never assume that the distance is the inverse of the
+measured parallax.
+In what follows, a star's distance is a latent property of the star, which generates
+the \gaia\ Catalog parallax through a noisy process (again, this is
+the ``likelihood'' assumption above).
+We never take the inverse or the logarithm of the measured parallax at any time.
+This is related to the no-cuts point above:
+If you take the inverse or logarithm of the parallax, you can't operate safely
+on the negative and low signal-to-noise parallaxes, which in turn will require
+making cuts on the sample, which will in turn bias the results.
+
+Finally, we never use Lutz--Kelker corrections (HOGG CITE) or distance
+posteriors (HOGG CITE). These both involve (implicit or explicit) priors on 
+distance.
+When multiple stars are combined as we combine them here (below),
+use of distance posteriors instead of parallax likelihoods is not just
+unjustified, but it also leads to an
+effective raising of the distance prior to an enormous power.
+That is, the effective prior on a $N$-star inference performed naively
+with distance posteriors (made with a weak prior) can end up
+bringing into the inference an \emph{exceedingly
+strong prior}.
+Therefore we don't use any prior-contaminated parallax or distance
+inputs to the inference.
 
@@
 
-Under the assumptions stated in the previous Section, there are not many choices
-for building a model of the stars that is justifiable probabilistically.
-Here we lay out the model and methodology, and apply it in the following Section.
+There are not many choices
+for building a model of the stars that is both justifiable probabilistically
+and consistent with the assumptions stated in the previous Section.
+Here we lay out the model and methodology.
+We apply the method to real data in the following Sections.
 
@@
 where
-$\varpi_n$ is the \gaia\ Catalog entry for the parallax for star $n$,
+$\varpi_n$ is the \gaia\ Catalog entry for the parallax (in mas) for star $n$,
 $K_n$ is the \zmass\ Catalog $K_s$-band magnitude for star $n$,
@@
 In what follows, the $Q_n$ are the training data.
-Naively, the $Q_n$ are related to the absolute magnituges $M_{Kn}$ for the stars by
+Naively, the $Q_n$ are related to the absolute magnitudes $M_{Kn}$ for the stars by
 \begin{eqnarray}
@@
 
-The \gaia\ Catalog describes a likelihood function!
+We take the \apogee\ data from what subset?
 
-There is a parallax offset.
+We cut on $\logg$ how and why? Comment on the trustworthiness of the $\logg$ values.
 
-We are only looking at \apogee\ luminous giants (selected spectroscopically).
+We take the \gaia\ data and apply the parallax offset.
 
-We use photometry from \gaia, \zmass, \wise.
+We match \apogee\ to \gaia\ how?
+
+We match the catalog to \zmass\ and \wise\ how?
+
+Figures that show off the data.
 
 \section{Results and validation}
+
+We construct the feature vectors $x_n$ how and what is their dimensionality $D$?
+
+Brief comment on the flexibility of the photometric model.
+
+We do 2-fold train-and-test to make results independent of \gaia.
+
+We set the regularization parameters how? And why?
+
+We compute our precision how? Figures.
 
@@
 Why did we make the assumptions we made and what did they gain us and cost us?
+
+(In addition: what did we assume about the \apogee\ data: Good $\logg$ values.
 
@@
 \acknowledgements
+It is a pleasure to thank
+  Daniel Michalik (Lund), and
+  Melissa Ness (Columbia)
+for help with this project.
 This project was developed in part at the

### 2018-07-25 paper/sp.tex
jones

@@
 % - make sure we address the sensibleness and generality of the linear model for the photometric space
-% - address convexity of the model...?
+% - address the non-convexity of the model (see twitter convo with @andy_l_jones).
 
@@
 
-\shorttitle{spectroscopic distance estimates}
-\shortauthors{eilers, hogg, rix, ness}
+\shorttitle{spectrophotometric distance estimates}
+\shortauthors{eilers, hogg, rix}
 %\watermark{kittens}
@@
 \title{\textbf{%
-Accurate spectroscopic estimates of luminosity and distance\\
+Accurate spectrophotometric estimates of luminosity and distance\\
 for luminous red-giant stars%
@@
 It is a pleasure to thank
-  Daniel Michalik (Lund), and
+  Andy L. Jones (HOGG),
+  Daniel Michalik (\acronym{ESTEC}), and
   Melissa Ness (Columbia)
@@
 \begin{thebibliography}{}
-
 \bibitem[Astropy Collaboration et al.(2013)]{2013A&A...558A..33A} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, \aap, 558, A33 
@@
 \bibitem[Vogt et al.(2014)]{2014ApJ...793..127V} Vogt, F.~P.~A., Dopita, M.~A., Kewley, L.~J., et al.\ 2014, \apj, 793, 127  
-
 \end{thebibliography}

### 2018-07-26 paper/gc.tex
start paper on the Galactic Center

--- 2018-07-24 paper/sp.tex
@@
 % -----------
-% - draft introduction
-% - draft data+results section
-% - draft discussion
-% - write list of things we never do!
-% - make sure we address the cross-validation point
-% - make sure we address the sensibleness and generality of the linear model for the photometric space
+% -
 
@@
 
-\shorttitle{spectroscopic distance estimates}
-\shortauthors{eilers, hogg, rix, ness}
+\shorttitle{kinematic distance to the galactic center}
+\shortauthors{hogg, eilers, rix}
 %\watermark{kittens}
@@
 \newcommand{\zmass}{\project{\acronym{2MASS}}}
-\newcommand{\sdssv}{\project{\acronym{SDSS-V}}}
+
+% jargon macros
+\newcommand{\sagastar}{Sag\,A$^{\ast}$}
 
@@
 \title{\textbf{%
-Accurate spectroscopic estimates of luminosity and distance\\
-for luminous red-giant stars%
+A purely kinematic distance to the Galactic Center
 }}
+
+\author[0000-0003-2866-9403]{David W. Hogg}
+\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University}
+\affiliation{Center for Data Science, New York University}
+\affiliation{Max-Planck-Institut f\"ur Astronomie}
+\affiliation{Flatiron Institute}
 
@@
 
-\author[0000-0003-2866-9403]{David W. Hogg}
-\affiliation{Max-Planck-Institut f\"ur Astronomie}
-\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University}
-\affiliation{Center for Data Science, New York University}
-\affiliation{Flatiron Institute}
-
-\author{Hans-Walter Rix}
+\author[0000-0003-4996-9069]{Hans-Walter Rix}
 \affiliation{Max-Planck-Institut f\"ur Astronomie}
 
-\author{Melissa Ness}
-\affiliation{Flatiron Institute}
-\affiliation{Department of Astronomy, Columbia University}
-
 \begin{abstract}\noindent
-With contemporary infrared spectroscopic surveys like \apogee,
-red-giant stars can be observed to distances and extinctions
-at which geometric parallaxes are not precisely measured by \gaia.
-Here we use the \apogee--\gaia\ overlap to train a purely linear model of
-continuum-normalized \apogee\ spectroscopy
-(plus \gaia, \zmass, and \wise\ photometry)
-that predicts absolute magnitude (logarithmic luminosity) for red stars with $0<\logg<2.2$.
-The training makes use of a justified likelihood function for the \gaia\ parallaxes,
-so that the training can be performed with full use of uncertainties and
-without any cuts on parallax or parallax signal-to-noise ratio.
-The model includes an L1 regularization that zeros out the contributions of
-most spectral pixels to the parallax estimates.
-The training is performed
-with leave-out subsamples such that no star's astrometry is used even indirectly in its
-spectroscopic parallax estimate.
-The model is flexible enough---and has enough photometric information---to
-correct for extinction by dust without any
-input of explicit extinction or reddening estimates.
-We re-label each star in the sample
-with a new spectroscopic parallax; this new parallax has an empirical (1-sigma)
-uncertainty of 9~percent, which is more precise than the \gaia\ parallax
-for the vast majority of targets.
-Validation with globular and open clusters shows that the spectroscopic parallaxes
-are both accurate and precise.
-These distance and luminosity estimates open up great opportunities for
-mapping the Milky Way disk with the next
-generation of spectroscopic surveys, and especially \sdssv.
+% Context
+The distance from the Sun to the Galactic Center is a critical parameter
+that sets the length scale for many Galactic measurements and phenomena.
+Radio observations have delivered an extremely precise proper motion
+of the Center (or really \sagastar),
+but not a distance with comparable precision.
+Meanwhile, the \gaia\ Mission is delivering distances and proper motions for
+billions of stars, some of which extend to $>10$~kpc.
+% Aims
+Here we produce a purely kinematic measurement of the position of the
+Galactic Center.
+% Methods
+We use the \gaia\ data to calibrate accurate but simple data-driven
+spectrophotometric distance estimates to luminous red-giant stars.
+We compare the mean proper motions of these stars as a function of distance
+and azimuth in the Milky-Way disk to the azimuth and proper motion of
+\sagastar.
+% Results
+We find a kinematic distance to the Galactic Center of $XXX\pm YYY$~kpc.
+This distance measurement is purely kinematic and geometric;
+it involves no physical or dynamical model of either stars or the Milky Way.
 \end{abstract}
@@
 \keywords{%
-methods:~statistical
+Galaxy:~center
  ---
-techniques:~spectroscopic
+Galaxy:~fundamental~parameters
  ---
-catalogs
+Galaxy:~kinematics~and~dynamics
  ---
-surveys
+infrared:~stars
+ ---
+%methods:~data~analysis
 % ---
-%parallaxes
+proper~motions
  ---
@@
  ---
-Galaxy:~disk
- ---
-infrared:~stars
+stars:~kinematics~and~dynamics
+% ---
+%techniques:~spectroscopic
 }
@@
 \section{Introduction} \label{sec:intro}
-
-The promise of mapping the Milky Way disk with infrared spectroscopy
-and red-giant stars.
-
-The information \gaia\ brings to this.
-
-The history and value (and disadvantages)
-of data-driven models for these kinds of problems.
-
-The dangers of aggressive machine-learning methods. Why will we just
-look at linear models? The disadvantages of this approach.
-
-The dangers of cutting on \gaia\ properties. Why we will not cut on
-\gaia\ parallaxes. The disadvantages of this approach.
 
@@
 \begin{description}
-\item[features] Perhaps our most fundamental assumption is that the intrinsic
-luminosity of a star can be predicted from the features we provide, which are
-the full set of (pseudo-continuum-normalized) pixels from the \apogee\ spectrum,
-plus the $G$ and $G_{BP}-G_{RP}$ photometry from \gaia,
-plus the $J$, $H$, and $K_s$ photometry from \zmass,
-plus the $W_1$ and $W_2$ photometry from \wise.
-That is, we assume that these spectrophotometric \emph{features} are sufficient
-to predict the intrinsic luminosity in the face of variations in stellar
-age, evolutionary phase, composition and other parameters, and also interstellar
-extinction.
+\item[features] Wassuh!
 
-\item[good features] We assume that the spectrophotometric features are known for
-each star with such high fidelity (both precision and accuracy) that we do not
-need to account for errors or uncertainties or biases in the features.
-That is, we assume that the features are substantially higher in signal-to-noise than the
-quantities we are trying to predict; in particular we are assuming that the photometry
-and the spectroscopy is better measured than the astrometry.
-This is true for most features for most stars, but it does not hold universally.
-
-\item[representativeness] We assume that the training set constructed from the overlap
-of \gaia\ and \apogee\ data sets constitutes a representative sample of stars,
-sufficiently representative to train the model for all other stars.
-Although this assumption is not terrible, it has a weakness:
-The stars at greatest distances and greatest local (angular) stellar crowding have
-the least precise \gaia\ parallax measurements, and therefore will effectively get
-less weight in the fits performed to train the model.
-
-\item[sparsity] We expect that only a small subset of the full complement of
-\apogee\ spectral pixels will be relevant to the prediction of stellar luminosity.
-That is, we expect that many of the pixels will or ought to get no weight in the
-final model that we use to predict luminosities and distances.
-
-\item[linearity] Perhaps the most restrictive assumption of this work is that
-the stellar logarithmic luminosity (or, equivalently, the absolute magnitude) in
-the \zmass\ $K_s$~band can be predicted as a \emph{completely linear function} of
-the chosen features. We are only assuming this linearity in a small range of stellar
-surface gravity $\logg$, but this assumption is strong, and limits strongly the
-flexibility or capacity of the model.
-We make it to ensure that our method is easy to optimize, and the results are easy
-to interpret.
-
-\item[likelihood] We assume that the \gaia\ parallaxes are generated by a particular
-stochastic process, in which the difference between the \gaia-Catalog parallax (plus
-a small \gaia-recommended offset HOGG CITE) and the true parallax is effectively drawn from a
-Gaussian with a width set by the \gaia-Catalog uncertainty on the parallax.
-This is the standard assumption in all properly probabilistic \gaia-Mission inferences
-to date, but it subsumes a number of related assumptions, like that the \gaia\ noise
-model is correct, that the stars are only covariant at negligible levels, and that
-there are no significant outliers or large-scale systematics.
 \end{description}
-
-HOGG: LIST of things we don't do, like don't cut the sample on
-astrometric quantities, and why not!! And like don't assume distance
-is inverse-measured-parallax.
 
@@
 
-Under the assumptions stated in the previous Section, there are not many choices
-for building a model of the stars that is justifiable probabilistically.
-Here we lay out the model and methodology, and apply it in the following Section.
-
-Since the photometry is (by the ``good features'' assumption above)
-better measured than the parallaxes, we
-presume that the parallax uncertainties dominate
-the uncertainties in the luminosity measurements.
-For each star $n$ in our training set, we define an inverse-square-root
-luminosity $Q_n$ estimate that will be subject to Gaussian noise under these assumptions.
-\begin{eqnarray}
-Q_n &\equiv& \varpi_n\,10^{0.2\,(K_n - 10)}
-\\
-\sigma_n &\equiv& \sigma_{\varpi n}\,10^{0.2\,(K_n - 10)}
-\quad,
-\end{eqnarray}
-where
-$\varpi_n$ is the \gaia\ Catalog entry for the parallax for star $n$,
-$K_n$ is the \zmass\ Catalog $K_s$-band magnitude for star $n$,
-the $Q_n$ have Gaussian errors of root-variance $\sigma_n$,
-and $\sigma_{\varpi n}$ is the \gaia\ Catalog entry for the uncertainty on parallax
-for star $n$.
-In what follows, the $Q_n$ are the training data.
-Naively, the $Q_n$ are related to the absolute magnituges $M_{Kn}$ for the stars by
-\begin{eqnarray}
-M_{Kn} &\approx& 5\,\log_{10} Q_n
-\quad.
-\end{eqnarray}
-These would be the absolute magnitudes of the stars under the
-``the distance is the inverse of the
-measured parallax'' assumption, although we never, in this work, ever make that
-assumption, by God.
-We construct the $Q_n$ because they are related directly to the absolute magnitudes,
-but they (under assumptions) are subject to known Gaussian noise.
-
-The log-likelihood function can be expressed heuristically as
-\begin{eqnarray}
-Q_n &=& \exp(\theta\cdot x_n) + \mbox{noise}
-\\
-\chi^2(\theta) &\equiv& \sum_{n=1}^N \frac{[Q_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
-\quad ,
-\end{eqnarray}
-where
-the model is that the $\ln Q_n$ can be expressed as a linear combination of the components
-of some $D$-dimensional feature vector $x_n$,
-$\theta$ is a $D$-vector of linear coefficients,
-and $\chi^2(\theta)$ is the negative-log-likelihood for the parameters $\theta$
-under the assumption of known Gaussian noise and
-that there are $N$ independently measured stars $n$.
-This feature vector $x_n$ contains photometry in a few bands, and all 8000-ish pixels
-in the pseudo-continuum-normalized \apogee\ spectrum, so $D$ is on the order of 8000.
-
-In addition, we assume that many entries in the $D$-vector will be zero
-(the ``sparsity'' assumption).
-In order to encourage this, we optimize not $\chi^2$ but a regularized objective function
-\begin{eqnarray}
-\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2(\theta) + ||\lambda\cdot\theta||_1^1\right]
-\quad ,
-\end{eqnarray}
-where
-$\lambda$ is a $D$-vector of regularization parameters (such that we can regularize
-different parts of the $x_n$ vectors differently),
-and $||x||_1^1$ is the L1-norm or sum of absolute values of the components of $x$.
-This kind of regularization adds a convex term to the optimization and leads to
-sparse optima.
-We will set the value $\lambda$ of the $D$-vector of regularization strengths by a
-very coarse cross-validation.
-
-Once the model is optimized, the prediction for a new stellar absolute magnitude
-$M_{Km}$ for a star $m$ in the validation or test set is
-\begin{eqnarray}
-Q_m &=& \exp(\theta\cdot x_m)
-\\
-M_{Km} &=& 5\,\log_{10} Q_m
-\quad ,
-\end{eqnarray}
-where $x_m$ is the feature $D$-vector for star $m$.
-Thus the model can be trained on a training set of stars and applied to make
-predictions for a validation set or a test set.
-
 \section{Data}
-
-The \gaia\ Catalog describes a likelihood function!
-
-There is a parallax offset.
-
-We are only looking at \apogee\ luminous giants (selected spectroscopically).
-
-We use photometry from \gaia, \zmass, \wise.
 
@@
 
-What did we do and what's cool?
-
-Why did we make the assumptions we made and what did they gain us and cost us?
-
-What is this useful for and how do you use them safely?
-
-(Mention independence of every result from the \gaia\ astrometry.)
-
-Come back to SDSS-V and show the rotation curve of the Galaxy.
-
 \acknowledgements
-This project was developed in part at the
-2018 \acronym{NYC} Gaia Sprint, hosted by the Center for Computational Astrophysics of
-the Flatiron Institute in New York City.
+It is a pleasure to thank
+  WHOEVER
+for help with this project.
 
@@
 \bibitem[Vogt et al.(2014)]{2014ApJ...793..127V} Vogt, F.~P.~A., Dopita, M.~A., Kewley, L.~J., et al.\ 2014, \apj, 793, 127  
-
 \end{thebibliography}
 
-%% This command is needed to show the entire author+affilation list when
-%% the collaboration and author truncation commands are used.  It has to
-%% go at the end of the manuscript.
-%\allauthors
-
-%% Include this line if you are using the \added, \replaced, \deleted
-%% commands to see a summary list of all changes at the end of the article.
-%\listofchanges
-
 \end{document}
-
-% End of file `sample62.tex'.

### 2018-07-26 paper/sp.tex
outline tweaks

@@
 
-\author{Hans-Walter Rix}
+\author[0000-0003-4996-9069]{Hans-Walter Rix}
 \affiliation{Max-Planck-Institut f\"ur Astronomie}
@@
 
-We take the \gaia\ data and apply the parallax offset.
+HOGG: We take the \gaia\ data and apply the parallax offset.
 
@@
 
+We construct the feature vectors $x_n$ how and what is their dimensionality $D$?
+
 \section{Results and validation}
 
-We construct the feature vectors $x_n$ how and what is their dimensionality $D$?
+HOGG: We do 2-fold train-and-test to make results independent of \gaia.
 
-Brief comment on the flexibility of the photometric model.
-
-We do 2-fold train-and-test to make results independent of \gaia.
+We compute our precision how? Figures.
 
@@
 
-We compute our precision how? Figures.
+Brief comment on the flexibility of the photometric model and interpretation of the photometric coefficients. Maybe plot some photometric properties vs metallicity or dust?
 

### 2018-07-27 paper/sp.tex
some intro text

@@
 
-The promise of mapping the Milky Way disk with infrared spectroscopy
-and red-giant stars.
+If we want to make precise kinematic and element-abundance maps
+of the Milky Way disk out to large heliocentric radii,
+and especially through the extinction to the other side of the Galactic Center, we will
+need to make use of luminous red giants, and we will need to observe in
+the infrared.
+These arguments motivate the \apogee\ (CITE), \apogee-2 (CITE), and \sdssv\ (CITE)
+projects, which take spectra in the infrared, and deliver detailed abundances
+along the entire red-giant branch up to the tip.
+The maps made with these data will reveal the dynamics of the Milky Way and its disk,
+and show us how and where stars form, and how they migrate around (and out of) the
+disk with time.
 
-The information \gaia\ brings to this.
+As we operate these spectroscopic surveys,
+the \gaia\ Mission has been revolutionizing our view of the disk.
+It delivers good proper motions everywhere,
+and extremely valuable parallax information
+locally.
+However, it does not deliver precise geometric parallaxes over a large fraction of
+the disk, and certainly not in the dusty and crowded regions.
+For these reasons, the value of \gaia\ in these infrared sky projects is not
+to deliver distance information directly, but rather to calibrate stellar models,
+which then deliver distance information through relationships between stellar
+luminosities and spectral characteristics.
+
+In this contribution we develop data-driven techniques to predict
+stellar luminosities and distances with spectroscopy and photometry...
 

### 2018-07-27 paper/sp.tex
author order; tweaks

@@
 
+\author[0000-0003-2866-9403]{David W. Hogg}
+\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University}
+\affiliation{Center for Data Science, New York University}
+\affiliation{Max-Planck-Institut f\"ur Astronomie}
+\affiliation{Flatiron Institute}
+
 \author[0000-0003-2895-6218]{Anna-Christina Eilers}
 \affiliation{Max-Planck-Institut f\"ur Astronomie}
-
-\author[0000-0003-2866-9403]{David W. Hogg}
-\affiliation{Max-Planck-Institut f\"ur Astronomie}
-\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University}
-\affiliation{Center for Data Science, New York University}
-\affiliation{Flatiron Institute}
 
@@
 the disk, and certainly not in the dusty and crowded regions.
-For these reasons, the value of \gaia\ in these infrared sky projects is not
+For these reasons, the value of \gaia\ in these infrared spectroscopy projects is not
 to deliver distance information directly, but rather to calibrate stellar models,

### 2018-07-28 paper/gc.tex
sag -> sgr
no hyphen
note to self

@@
 % jargon macros
-\newcommand{\sagastar}{Sag\,A$^{\ast}$}
+\newcommand{\sagastar}{Sgr\,A$^{\ast}$}
 
@@
 We compare the mean proper motions of these stars as a function of distance
-and azimuth in the Milky-Way disk to the azimuth and proper motion of
+and azimuth in the Milky Way disk to the azimuth and proper motion of
 \sagastar.
@@
 \section{Introduction} \label{sec:intro}
+
+What do we know about the Galactic Center distance? And cite the GRAVITY result.
 

### 2018-07-29 paper/sp.tex
many words in the data section

@@
 
-% project macros
+% text macros
+\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
+\newcommand{\foreign}[1]{\textsl{#1}}
 \newcommand{\acronym}[1]{{\small{#1}}}
@@
 
-We take the \apogee\ data from what subset?
-
+ACE: We take the \apogee\ data from what subset?
 We cut on $\logg$ how and why? Comment on the trustworthiness of the $\logg$ values.
 
-HOGG: We take the \gaia\ data and apply the parallax offset.
+ACE: We match the \apogee\ stars making the $\logg$ cut also to
+the \zmass\ ACE WHICH CATALOG by ACE WHAT CRITERION?
+This gives, for each matching star $n$,
+near-infrared photometry $J_n$, $H_n$, and $K_n$.
+In detail we use from the catalog the entries ACE WHAT SPECIFIC CATALOG ENTRIES.
 
-We match \apogee\ to \gaia\ how?
+ACE: We match the \apogee+\zmass\ stars to he \wise\ ACE WHICH CATALOG by ACE WHAT CRITERION?
+This gives, fore each matching star $n$,
+mid-infrared photometry $W_{1n}$ and $W_{2n}$ at ACE WHAT WAVELENGTHS?
+In detail we use the ACE WHAT SPECIFIC CATALOG ENTRIES?
 
-We match the catalog to \zmass\ and \wise\ how?
+We match this full-match catalog to the \gaia\ \acronym{DR2} (HOGG CITE) by what criteria?
 
-Figures that show off the data.
+For each star $n$, we take from \gaia\ DR2 we the parallax $\varpi_n$
+and its uncertainty $\sigma_{\varpi n}$, which will be used in
+constructing the inverse-root-luminosity $Q_n$ and associated
+uncertainty $\sigma_n$.
+Importantly, to each \gaia\ parallax measurement we add a positive
+offset of $0.029$\,mas to adjust for the under-estimation of
+parallaxes reported by the \gaia\ Collaboration (HOGG CITE LINDEGREN).
+We also take from the \gaia\ data the photometric data $G_n$
+(\code{phot_g_mean_mag}) and $G_{Bn}-G_{Rn}$ (\code{bp_rp}), which will
+become part of the feature vector $x_n$.
+We do not use the entire \gaia\ \acronym{DR2} Catalog but instead the subset
+that meets ACE WHAT ADDITIONAL PHOTOMETRIC CUTS?
 
-We construct the feature vectors $x_n$ how and what is their dimensionality $D$?
+We need complete feature-vector information for all stars.  For this
+reason, we define the Parent Sample to be the set of all stars that
+meet the \apogee\ and \gaia\ cuts and also
+have the complete set of photometry: $G_n$, $G_{Bn}-G_{Rn}$, $J_n$,
+$H_n$, $K_n$, $W_{1n}$, and $W_{2n}$ and also an apogee spectrum.
+This Parent Sample contains ACE HOW MANY STARS? and is shown in HOGG WHAT FIGURES?
+
+Every Parent Sample object gets, in addition, a randomly assigned binary
+label (A or B).
+This is used for two-fold validation and jackknife.
+In short, we will use the model trained on the A data to assign luminosities
+and distances to the B data and \foreign{vice versa}.
+The Parent Sample A has HOGG entries and Parent Sample B has HOGG entries.
+
+For training, we need A and B Training Sets.
+We define the Training Set stars to be Parent Sample stars that also
+have a measured \gaia\ parallax.
+The Training-Set stars, in addition to meeting all Parent-Sample cuts,
+also must meet ACE WHAT ADDITIONAL CUTS to ensure that the parallax
+measurements are good.
+The training sets are shown in HOGG WHAT FIGURES?
+
+But---as we have emphasized above---we do not cut ever on parallax or
+parallax error or their ratio. For this reason, most of the training set
+do not have significantly measured parallaxes and HOGG percent even have
+negative parallaxes!
+But we will perform our training such that it will be unbiased in these
+circumstances.
+Indeed it would be necessarily biased if ever we did cut on parallax or
+parallax signal-to-noise.
+
+For every star $n$ in the full Parent Sample we construct the feature
+vector $x_n$ as HOGG PUT EQUATION HERE.
+The feature vectors live in a $D$-dimensional space where $D=$HOGG.
+For every star $n$ in either of the Training Sets we also construct a
+regression datum $Q_n$ and uncertainty $\sigma_n$ according to HOGG WHAT EQUATION ABOVE?
 

### 2018-07-30 rotation_curve_uncertainties.py
new

@@ -137,7 +137,7 @@ for i in range(N):
     vz = cyl.differentials['s'].d_z.to(u.km/u.s)
     rho = cyl.rho.to(u.kpc)
     phi = cyl.phi
-    z = cyl.z    
+    z = cyl.z.to(u.kpc)
     XS_cyl = np.vstack([rho, phi, z, vr, vphi, vz]).T.value
     
     galcen_true = cs_true.transform_to(gc)
@@ -175,8 +175,8 @@ hdu = fits.PrimaryHDU(XS_cart_true_n)
 hdu.writeto('data/true_cart_n_{}.fits'.format(name), overwrite = True)
 
 
-# -------------------------------------------------------------------------------
-# plots
+# -------------------------------------------------------------------------------'''
+# load data
 # -------------------------------------------------------------------------------           
 
 hdu = fits.open('data/mean_cart_n_{}.fits'.format(name))
@@ -192,12 +192,15 @@ XS_cart_true_n = hdu[0].data
 hdu = fits.open('data/true_cyl_n_{}.fits'.format(name))
 XS_cyl_true_n = hdu[0].data
 
-# cuts in metallicity
-cut_high_feh = (labels['FE_H'] >= 0.0)
-cut_solar_feh = (labels['FE_H'] >= -0.2) * (labels['FE_H'] < 0.0)
-cut_low_feh = (labels['FE_H'] < -0.2) * (labels['FE_H'] >= -10) # remove [Fe/H] = -9999.0
-cut_names = list(['lowFEH', 'solarFEH', 'hiFEH'])
-cuts = list([cut_low_feh, cut_solar_feh, cut_high_feh])
+# -------------------------------------------------------------------------------
+# cuts in metallicity (NOT NEEDED AT THE MOMENT)
+# -------------------------------------------------------------------------------   
+
+#cut_high_feh = (labels['FE_H'] >= 0.0)
+#cut_solar_feh = (labels['FE_H'] >= -0.2) * (labels['FE_H'] < 0.0)
+#cut_low_feh = (labels['FE_H'] < -0.2) * (labels['FE_H'] >= -10) # remove [Fe/H] = -9999.0
[…]

### 2018-07-30 paper/sp.tex
note to self
notes to self; acks
fixing up a bit the data section

@@
 \shorttitle{spectrophotometric distance estimates}
-\shortauthors{eilers, hogg, rix}
+\shortauthors{hogg, eilers, rix}
 %\watermark{kittens}
@@
 
+HOGG: PUT WORDS HERE ABOUT GOING TWO-FOLD to pre-flash that we are going to do that
+in the next Section.
+
 \section{Data}
@@
 
-ACE: We match the \apogee\ stars making the $\logg$ cut also to
-the \zmass\ ACE WHICH CATALOG by ACE WHAT CRITERION?
-This gives, for each matching star $n$,
-near-infrared photometry $J_n$, $H_n$, and $K_n$.
+Because the \apogee\ target selection is based on \zmass, every \apogee\ star
+comes with \zmass photometry.
+That is, for each star $n$,
+we have near-infrared photometry $J_n$, $H_n$, and $K_n$.
 In detail we use from the catalog the entries ACE WHAT SPECIFIC CATALOG ENTRIES.
@@
 This gives, fore each matching star $n$,
-mid-infrared photometry $W_{1n}$ and $W_{2n}$ at ACE WHAT WAVELENGTHS?
+mid-infrared photometry $W_{1n}$ and $W_{2n}$ at 3.6 and 4.5\,$\mu$m
 In detail we use the ACE WHAT SPECIFIC CATALOG ENTRIES?
@@
 
-For each star $n$, we take from \gaia\ DR2 we the parallax $\varpi_n$
-and its uncertainty $\sigma_{\varpi n}$, which will be used in
-constructing the inverse-root-luminosity $Q_n$ and associated
-uncertainty $\sigma_n$.
-Importantly, to each \gaia\ parallax measurement we add a positive
-offset of $0.029$\,mas to adjust for the under-estimation of
-parallaxes reported by the \gaia\ Collaboration (HOGG CITE LINDEGREN).
-We also take from the \gaia\ data the photometric data $G_n$
+We take from the \gaia\ \acronym{DR2} Catalog the photometric data $G_n$
 (\code{phot_g_mean_mag}) and $G_{Bn}-G_{Rn}$ (\code{bp_rp}), which will
@@
 measurements are good.
-The training sets are shown in HOGG WHAT FIGURES?
-
 But---as we have emphasized above---we do not cut ever on parallax or
@@
 parallax signal-to-noise.
+
+In detail,
+for each star $n$ in the Training Sets, we take from \gaia\ DR2 we the parallax $\varpi_n$
+and its uncertainty $\sigma_{\varpi n}$, which will be used in
+constructing the inverse-root-luminosity $Q_n$ and associated
+uncertainty $\sigma_n$.
+Importantly, to each \gaia\ parallax measurement we add a positive
+offset of $0.029$\,mas to adjust for the under-estimation of
+parallaxes reported by the \gaia\ Collaboration (HOGG CITE LINDEGREN).
+The training sets are shown in HOGG WHAT FIGURES?
 
@@
 
+\wise\ Acknowledgement?
+
+\project{\acronym{SDSS-IV}} ack.
+
 \begin{thebibliography}{}

### 2018-07-30 paper/sp.tex
A/B split

@@
 
-HOGG: PUT WORDS HERE ABOUT GOING TWO-FOLD to pre-flash that we are going to do that
-in the next Section.
+One final detail: We perform all of the fitting in a two-fold train-and-test framework,
+in which the data are split into two disjoint subsets, A and B.
+The model trained on the A data is used to predict the $Q_m$ values (and hence
+the luminosities, distances, and parallaxes) of the B data,
+and the model trained on the B data is used to 
+predict the $Q_m$ values of the A data.
+This ensures that, in the estimate of any individal star's luminosity, distance,
+or parallax, none of the \gaia\ data pertaining to that star was used.
+This makes the parallax estimates from the spectrophotometric feature vectors
+$x_m$ statistically independent of the \gaia\ parallax measurements.
+They are not independent \emph{globally}, but they are independent on a
+star-by-star basis.
 

### 2018-07-31 paper/sp.tex
apogee data and cuts

@@
 $x_m$ statistically independent of the \gaia\ parallax measurements.
-They are not independent \emph{globally}, but they are independent on a
-star-by-star basis.
+They are not independent globally---\gaia\ data was used to train the model---but
+each star's spectrophotometric parallax estimate is independent
+of the astrometric estimate from \gaia\ on a star-by-star basis.
 
@@
 
-ACE: We take the \apogee\ data from what subset?
-We cut on $\logg$ how and why? Comment on the trustworthiness of the $\logg$ values.
+We take the \apogee\ (HOGG CITE) spectral data
+from \sdssiv\ (HOGG CITE) \acronym{DR14} (HOGG CITE).
+Because we want to make a purely linear model, we restrict our consideration
+to a small region in stellar parameter space.
+We cut the \apogee\ data down to the range $0<\logg<2.2$, which isolates
+stars that are more luminous than the red clump.
+The \apogee\ pipeline $\logg$ values (which we use) have uncertainties but
+this cut leads to a clean sample of luminous red giants, and it is a cut
+that is only on spectral properties of the stars (and not photometry nor astrometry).
 

### 2018-07-31 paper/sp.tex
much of the results section

@@
 
-HOGG: We do 2-fold train-and-test to make results independent of \gaia.
+We optimize two models, one for Training Set A and one for Training Set B.
+For optimization, we use \code{scipy} l-bfgs-b (HOGG CHECK), starting at
+an uninformed starting point in the $\theta$ space.
+The optimization is of the objective function given in (HOGG WHAT EQUATION) above,
+and although this optimization is not convex, it appears to converge to the
+same good solution quickly and reliably from multiple different starting points.
 
-We compute our precision how? Figures.
+The two models are applied to the two splits of the Parent Sample, Parent Sample
+A and Parent Sample B, using the A-trained model on Parent Sample B and
+\foreign{vice versa}.
+For the purposes of assessing the accuracy and precision of the model, this
+train-and-test framework constitutes a two-fold cross-validation.
+Results of this cross-validation is shown in HOGG WHAT FIGURE.
+For the stars with highest \gaia-measured parallax signal-to-noise,
+the spectrophotometric model is predicting the \gaia\ parallaxes with little bias
+and a scatter of better than 9~percent.
 
-We set the regularization parameters how? And why?
+We used this cross-validation framework to adjust the regularization parameters.
+We (arbitrarily) chose zero regularization on the photometric components of the
+$x_n$ vectors, and uniform, identical regularization strength for all of the
+spectral (\apogee-pixel) components.
+We did a coarse grid search in the value for the spectral-component value of the
+$\lambda$ vector, using the A/B split as the two-fold cross-validation.
+The 9-percent scatter is what we obtain at the optimal setting of the regularization
+strength.
+
+We adopt 9~percent as the method's precision.
+This estimate is conservative in that the 9-percent scatter includes contributions
+from both our spectrophotometric model and from the \gaia\ measurements themselves.
+However this estimate is not so conservative in that the test is performed with
+\gaia's best stars, which are bright, nearby, in low extinction regions, and
+(because of these things) near Solar metallicity.
+There could in principle be additional bias or scatter for stars that are in dustier
+regions or at lower metallicities.
+\figurename~HOGG WHAT FIGURE shows that
+there is no suggestion of any such trends when we color the residuals by metallicity
+or $H-W_2$ (which is a reddening proxy).
+
+HOGG: validation with clusters and Sag.
 

### 2018-07-31 rotation_curve_uncertainties.py
new

@@ -295,28 +295,57 @@ ax.set_aspect('equal')
 plt.savefig('plots/rotation/rotate_phi_wegde_{0}.pdf'.format(name), bbox_inches = 'tight')
 plt.close()  
 
+#dz = 1000. # pc
+#dr = 1. # kpc
+#bins_dr = np.arange(0, 40.01, dr/1.)
+#N_stars_annulus = np.zeros_like(bins_dr)
+#mean_XS_cyl_annulus = np.zeros((len(bins_dr), 6)) - np.inf
+##mean_XS_cart_annulus = np.zeros((len(bins_dr), 6)) - np.inf
+#var_XS_cyl_annulus = np.zeros((len(bins_dr), 3, 3)) - np.inf
+#error_var_XS_cyl_annulus = np.zeros((len(bins_dr), 3, 3)) - np.inf
+#vvT_cyl_annulus = np.zeros((len(bins_dr), 3, 3)) - np.inf
+#
+#for i, r_center in enumerate(bins_dr):
+#    cut_annulus = wedge * (abs(mean_XS_cyl_n[:, 2]) < dz/2.) * (abs(mean_XS_cyl_n[:, 0] - r_center) < dr/2.)
+#    N_stars_annulus[i] = np.sum(cut_annulus)
+#    #print(i, N_stars_annulus[i])
+#    if N_stars_annulus[i] > 0:
+#        mean_XS_cyl_annulus[i, :] = np.nanmean(mean_XS_cyl_n[cut_annulus], axis = 0)
+##        mean_XS_cart_annulus[i, :] = np.nanmean(mean_XS_cart_n[cut_annulus], axis = 0)
+##    if N_stars_annulus[i] > 7:
+##        dXS = mean_XS_cyl_n[cut_annulus] - mean_XS_cyl_annulus[i, :][None, :]
+##        var_XS_cyl_annulus[i, :, :] = np.dot(dXS[:, 3:].T, dXS[:, 3:]) / (N_stars_annulus[i] - 1.)
+#        error_var_XS_cyl_annulus[i, :, :] = np.nanmean(var_XS_cyl_n[cut_annulus], axis=0)
+#        vvT_cyl_annulus[i, :, :] = np.dot(XS_cyl_true_n[cut_annulus, 3:].T, XS_cyl_true_n[cut_annulus, 3:]) / (N_stars_annulus[i] - 1.)
+
 dz = 1000. # pc
 dr = 1. # kpc
[…]

### 2018-07-31 paper/sp.tex
spectral pixels and continuum

@@
 \newcommand{\zmass}{\project{\acronym{2MASS}}}
+\newcommand{\sdssiv}{\project{\acronym{SDSS-IV}}}
 \newcommand{\sdssv}{\project{\acronym{SDSS-V}}}
@@
 from \sdssiv\ (HOGG CITE) \acronym{DR14} (HOGG CITE).
-Because we want to make a purely linear model, we restrict our consideration
-to a small region in stellar parameter space.
+Because we want to make a purely linear model, which has very little capacity,
+we restrict our consideration to a small region in stellar parameter space.
 We cut the \apogee\ data down to the range $0<\logg<2.2$, which isolates
@@
 
+From the \apogee\ data on these low-$\logg$ stars, we take the spectral pixels,
+of which there are ACE HOW MANY, on a common rest-frame wavelength grid.
+That is, every \apogee\ star is extracted on (or interpolated to)
+the same wavelength scale.
+In detail, we obtain the
+the \apogee\ spectral pixels from the pipeline-generated \code{ACE whatever} files.
+We then pseudo-continuum-normalize the spectra according to the procedure developed
+in \project{The~Cannon} (HOGG CITE): That is, the pseudo-continuum is a spline fit to a set
+of pixels chosen to be insensitive to stellar parameters.
+We use as our spectral data the normalized spectral pixel values.
+
 Because the \apogee\ target selection is based on \zmass, every \apogee\ star
-comes with \zmass photometry.
+also comes with \zmass photometry.
 That is, for each star $n$,
@@
 
-Every Parent Sample object gets, in addition, a randomly assigned binary
+Every Parent-Sample star gets, in addition, a randomly assigned binary
 label (A or B).
@@
 For training, we need A and B Training Sets.
-We define the Training Set stars to be Parent Sample stars that also
+We define the Training Set stars to be Parent-Sample stars that also
 have a measured \gaia\ parallax.
@@
 
+HOGG: How to get the data and how to use it, probabilistically.
+
 \section{Discussion}

### 2018-07-31 paper/sp.tex
thank calj

@@
 It is a pleasure to thank
-  Andy L. Jones (HOGG),
+  Coryn Bailer-Jones (\acronym{MPIA}),
+  Andy L. Jones,
   Daniel Michalik (\acronym{ESTEC}), and

### 2018-07-31 paper/sp.tex
notes to self

@@
 % - address the non-convexity of the model (see twitter convo with @andy_l_jones).
+% - Coryn's point that we should carefully explain how our model is geometric.
 
@@
 without any cuts on parallax or parallax signal-to-noise ratio.
+There are no physical inputs to the model; it is a pure linear regression designed
+to accurately predict astrometric parallaxes.
 The model includes an L1 regularization that zeros out the contributions of
@@
 of data-driven models for these kinds of problems.
+How could it be that we are completely independent of stellar models?
 
@@
 (In addition: what did we assume about the \apogee\ data: Good $\logg$ values.
-
-What is this useful for and how do you use them safely?
+Okay pseudo-continuum normalization!)
 
@@
 
-Come back to SDSS-V and show the rotation curve of the Galaxy.
+What is this useful for and how do you use them safely?
+HOGG: Show a map of the kinematic rotation of the Galaxy as a function of
+x and y, colored by metallicity.
+We see the kinematic center of the Galaxy! We see the rotation curve.
+We see the chemical gradients.
+
+How do you build a proper probabilistic model using these? How is the answer to that
+related to the answer for \gaia\ itself?
 

### 2018-07-31 paper/sp.tex
validation tests

@@
 there is no suggestion of any such trends when we color the residuals by metallicity
-or $H-W_2$ (which is a reddening proxy).
+or $H-W_2$ color (which is a reddening proxy).
 
-HOGG: validation with clusters and Sag.
+Another validation of the results can be obtained by looking at known
+clusters or spatially compact objects in the data.
+In HOGG WHAT FIGURE, we show the parallax distribution from \gaia\ and
+the spectroscopic-parallax distribution from this work for stars in
+angular proximity to known stellar clusters and the Sagittarius dwarf
+galaxy.
+These clusters span some range in metallicity and age, and therefore test
+the accuracy of the method for different kinds of populations.
+They show that both \gaia\ and the spectroscopic parallaxes appear to be unbiased
+for these clusters and the range in abundances and ages they represent.
+HOGG: DO WE NEED TO ADDRESS whether why not these clusters show 9-percent spreads?
 
-Brief comment on the flexibility of the photometric model and interpretation of the photometric coefficients. Maybe plot some photometric properties vs metallicity or dust?
+HOGG: Brief comment on the flexibility of the photometric model and interpretation of the photometric coefficients. Maybe plot some photometric properties vs metallicity or dust?
 

### 2018-07-31 paper/sp.tex
interpretability

@@
 
+The linear model we use has many limitations: Linear models are inflexible.
+However, they have a great advantage, which is that it is easy to look inside
+the model and check whether the dependencies represented there make sense.
+One nice thing about the model is that the feature vector $x_n$ for star $n$
+contains a set of photometry in magnitudes.
+Linear combinations of these photometric measurements are like complex synthetic
+magnitudes or colors.
+Since the model was trained to deliver luminosities of stars given training data
+of different metallicities and extinctions, the synthetic photometry developed
+by the model should be, in some sense, metallicity- and extinction-independent
+photometry for luminous red-giant stars!
+HOGG: CONTINUE THAT THOUGHT.
+
 HOGG: Brief comment on the flexibility of the photometric model and interpretation of the photometric coefficients. Maybe plot some photometric properties vs metallicity or dust?

### 2018-08-01 paper/sp.tex
notes to self

@@
 % - draft introduction
-% - draft data+results section
 % - draft discussion
-% - make sure we address the cross-validation point
 % - make sure we address the sensibleness and generality of the linear model for the photometric space
@@
 % - Coryn's point that we should carefully explain how our model is geometric.
+% - related: HWR's point that this is standardizeable candles, like the Ia SNe results.
 
@@
 2018 \acronym{NYC} Gaia Sprint, hosted by the Center for Computational Astrophysics of
-the Flatiron Institute in New York City.
+the Flatiron Institute in New York City in 2018 June.
 

### 2018-08-01 paper/sp.tex
title tweaks; notes to self

@@
 % - related: HWR's point that this is standardizeable candles, like the Ia SNe results.
+% - related: Put in an explanation of WHY this works so well.
+% - related: Put in an explanation of how it is that we aren't really modeling luminosity or distance but really just generating parallaxes.
+% - related: Switch the model to be a model for parallax directly.
+% - does the model need a cute name, like The Swan?
 
@@
 
-\shorttitle{spectrophotometric distance estimates}
+\shorttitle{spectrophotometric parallax}
 \shortauthors{hogg, eilers, rix}
@@
 \title{\textbf{%
-Accurate spectrophotometric estimates of luminosity and distance\\
-for luminous red-giant stars%
+Spectrophotometric parallax:\\
+Accurate estimates of luminosity and distance\\
+for luminous red-giant stars
 }}
@@
 (plus \gaia, \zmass, and \wise\ photometry)
-that predicts absolute magnitude (logarithmic luminosity) for red stars with $0<\logg<2.2$.
+that predicts distance modulus (logarithmic parallax or distance)
+for red stars with $0<\logg<2.2$.
 The training makes use of a justified likelihood function for the \gaia\ parallaxes,
 so that the training can be performed with full use of uncertainties and
-without any cuts on parallax or parallax signal-to-noise ratio.
-There are no physical inputs to the model; it is a pure linear regression designed
+without any cuts on parallax or parallax signal-to-noise ratio (not even any non-negative
+requirements).
+There are no physical inputs to the model; it is a pure regression designed
 to accurately predict astrometric parallaxes.
@@
 with leave-out subsamples such that no star's astrometry is used even indirectly in its
-spectroscopic parallax estimate.
+spectrophotometric parallax estimate.
 The model is flexible enough---and has enough photometric information---to
@@
 We re-label each star in the sample
-with a new spectroscopic parallax; this new parallax has an empirical (1-sigma)
+with a new spectrophotometric parallax; this new parallax has an empirical (1-sigma)
 uncertainty of 9~percent, which is more precise than the \gaia\ parallax
@@
 are both accurate and precise.
-These distance and luminosity estimates open up great opportunities for
+These distance estimates open up great opportunities for
 mapping the Milky Way disk with the next
@@
 \gaia\ parallaxes. The disadvantages of this approach.
+
+\section{Why and how should we model red-giant stars?}
+
+HOGG: We need to, for \sdssv.
+
+HOGG: We know we can because of physical arguments.
+
+HOGG: What kinds of inputs do we need to predict parallaxes?
 
@@
 
-The linear model we use has many limitations: Linear models are inflexible.
-However, they have a great advantage, which is that it is easy to look inside
-the model and check whether the dependencies represented there make sense.
+The linear model we use has a big limitation: Linear models are inflexible.
+However, it has a great advantage over more general methods,
+which is that it is easy to look inside
+the linear model and check whether the dependencies represented there make sense.
 One nice thing about the model is that the feature vector $x_n$ for star $n$
@@
 photometry for luminous red-giant stars!
-HOGG: CONTINUE THAT THOUGHT.
-
-HOGG: Brief comment on the flexibility of the photometric model and interpretation of the photometric coefficients. Maybe plot some photometric properties vs metallicity or dust?
+In detail this synthetic magnitude is something like HOGG PUT EQUATION HERE AND
+SAY SOMETHING ABOUT IT. HOGG: Maybe plot this synthetic photometry against dust
+and metallicity.
 

### 2018-08-01 paper/sp.tex
switch over to parallax fitting

@@
 % math macros
+\newcommand{\T}{^{\mathsf{T}}}
+\DeclareMathOperator*{\argmin}{argmin}
 \newcommand{\logg}{\log g}
-\DeclareMathOperator*{\argmin}{argmin}
 
@@
 for the vast majority of targets.
-Validation with globular and open clusters shows that the spectroscopic parallaxes
+Validation with globular and open clusters shows that the spectrophotometric parallaxes
 are both accurate and precise.
@@
 \begin{description}
-\item[features] Perhaps our most fundamental assumption is that the intrinsic
-luminosity of a star can be predicted from the features we provide, which are
+\item[features] Perhaps our most fundamental assumption is that the parallax
+of a star can be predicted from the features we provide, which are
 the full set of (pseudo-continuum-normalized) pixels from the \apogee\ spectrum,
-plus the $G$ and $G_{BP}-G_{RP}$ photometry from \gaia,
+plus the $G$ and $({BP}-{RP})$ photometry from \gaia,
 plus the $J$, $H$, and $K_s$ photometry from \zmass,
@@
 That is, we assume that these spectrophotometric \emph{features} are sufficient
-to predict the intrinsic luminosity in the face of variations in stellar
-age, evolutionary phase, composition and other parameters, and also interstellar
+to predict the parallax in the face of variations in stellar
+age, evolutionary phase, composition, and other parameters, and also interstellar
 extinction.
@@
 \item[sparsity] We expect that only a small subset of the full complement of
-\apogee\ spectral pixels will be relevant to the prediction of stellar luminosity.
+\apogee\ spectral pixels will be relevant to the prediction of parallax.
 That is, we expect that many of the pixels will or ought to get no weight in the
@@
 \item[linearity] Perhaps the most restrictive assumption of this work is that
-the stellar logarithmic luminosity (or, equivalently, the absolute magnitude) in
-the \zmass\ $K_s$~band can be predicted as a \emph{completely linear function} of
+the logarithmic parallax (or, equivalently, the distance modulus) 
+can be predicted as a \emph{completely linear function} of
 the chosen features. We are only assuming this linearity in a small range of stellar
@@
 
-Since the photometry is (by the ``good features'' assumption above)
-better measured than the parallaxes, we
-presume that the parallax uncertainties dominate
-the uncertainties in the luminosity measurements.
-For each star $n$ in our training set, we define an inverse-square-root
-luminosity $Q_n$ estimate that will be subject to Gaussian noise under these assumptions.
+The model and log-likelihood function can be expressed heuristically as
 \begin{eqnarray}
-Q_n &\equiv& \varpi_n\,10^{0.2\,(K_n - 10)}
+\varpi_n &=& \exp(\theta\cdot x_n) + \mbox{noise}
 \\
-\sigma_n &\equiv& \sigma_{\varpi n}\,10^{0.2\,(K_n - 10)}
-\quad,
-\end{eqnarray}
-where
-$\varpi_n$ is the \gaia\ Catalog entry for the parallax (in mas) for star $n$,
-$K_n$ is the \zmass\ Catalog $K_s$-band magnitude for star $n$,
-the $Q_n$ have Gaussian errors of root-variance $\sigma_n$,
-and $\sigma_{\varpi n}$ is the \gaia\ Catalog entry for the uncertainty on parallax
-for star $n$.
-In what follows, the $Q_n$ are the training data.
-Naively, the $Q_n$ are related to the absolute magnitudes $M_{Kn}$ for the stars by
-\begin{eqnarray}
-M_{Kn} &\approx& 5\,\log_{10} Q_n
-\quad.
-\end{eqnarray}
-These would be the absolute magnitudes of the stars under the
-``the distance is the inverse of the
-measured parallax'' assumption, although we never, in this work, ever make that
-assumption, by God.
-We construct the $Q_n$ because they are related directly to the absolute magnitudes,
-but they (under assumptions) are subject to known Gaussian noise.
-
-The log-likelihood function can be expressed heuristically as
-\begin{eqnarray}
-Q_n &=& \exp(\theta\cdot x_n) + \mbox{noise}
-\\
-\chi^2(\theta) &\equiv& \sum_{n=1}^N \frac{[Q_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
+\chi^2(\theta) &\equiv& \sum_{n=1}^N \frac{[\varpi_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
 \quad ,
@@
 where
-the model is that the $\ln Q_n$ can be expressed as a linear combination of the components
+$\varpi_n$ is the \gaia\ measurement (adjusted; see below) of the parallax of star $n$,
+the model is that the logarithm of the true parallax
+can be expressed as a linear combination of the components
 of some $D$-dimensional feature vector $x_n$,
 $\theta$ is a $D$-vector of linear coefficients,
-and $\chi^2(\theta)$ is the negative-log-likelihood for the parameters $\theta$
+$\sigma_n$ is the \gaia\ estimate of the uncertainty on the parallax measurement,
+and $\chi^2(\theta)$ is (twice) the negative-log-likelihood for the parameters $\theta$
 under the assumption of known Gaussian noise and
 that there are $N$ independently measured stars $n$.
-This feature vector $x_n$ contains photometry in a few bands, and all 8000-ish pixels
+The feature vector $x_n$ contains photometry in a few bands, and all 8000-ish pixels
 in the pseudo-continuum-normalized \apogee\ spectrum, so $D$ is on the order of 8000.
@@
 
-Once the model is optimized, the prediction for a new stellar absolute magnitude
-$M_{Km}$ for a star $m$ in the validation or test set is
+Once the model is optimized, the prediction for a new stellar parallax
+$\varpi_m$ for a star $m$ in the validation or test set is
 \begin{eqnarray}
-Q_m &=& \exp(\theta\cdot x_m)
-\\
-M_{Km} &=& 5\,\log_{10} Q_m
+\varpi_m &=& \exp(\theta\cdot x_m)
 \quad ,
@@
 predictions for a validation set or a test set.
+The only requirements are that every test-set object $m$  must have a full feature
+vector $x_m$ just as every training-set object $n$ must have a full feature
+vector $x_n$.
 
@@
 in which the data are split into two disjoint subsets, A and B.
-The model trained on the A data is used to predict the $Q_m$ values (and hence
-the luminosities, distances, and parallaxes) of the B data,
+The model trained on the A data is used to predict the $\varpi_m$ values (and hence
+the distances and parallaxes) of the B data,
 and the model trained on the B data is used to 
-predict the $Q_m$ values of the A data.
-This ensures that, in the estimate of any individal star's luminosity, distance,
-or parallax, none of the \gaia\ data pertaining to that star was used.
+predict the $\varpi_m$ values of the A data.
+This ensures that, in the estimate of any individal star's
+parallax, none of the \gaia\ data pertaining to that star were used.
 This makes the parallax estimates from the spectrophotometric feature vectors
 $x_m$ statistically independent of the \gaia\ parallax measurements.
-They are not independent globally---\gaia\ data was used to train the model---but
+They are not independent globally---\gaia\ data were used to train the model---but
 each star's spectrophotometric parallax estimate is independent
@@
 We take from the \gaia\ \acronym{DR2} Catalog the photometric data $G_n$
-(\code{phot_g_mean_mag}) and $G_{Bn}-G_{Rn}$ (\code{bp_rp}), which will
+(\code{phot_g_mean_mag}) and $({BP}-{RP})_n$ (\code{bp_rp}), which will
 become part of the feature vector $x_n$.
@@
 meet the \apogee\ and \gaia\ cuts and also
-have the complete set of photometry: $G_n$, $G_{Bn}-G_{Rn}$, $J_n$,
+have the complete set of photometry: $G_n$, $({BP}-{RP})_n$, $J_n$,
 $H_n$, $K_n$, $W_{1n}$, and $W_{2n}$ and also an apogee spectrum.
@@
 for each star $n$ in the Training Sets, we take from \gaia\ DR2 we the parallax $\varpi_n$
-and its uncertainty $\sigma_{\varpi n}$, which will be used in
-constructing the inverse-root-luminosity $Q_n$ and associated
-uncertainty $\sigma_n$.
+and its uncertainty $\sigma_n$.
 Importantly, to each \gaia\ parallax measurement we add a positive
@@
 For every star $n$ in the full Parent Sample we construct the feature
-vector $x_n$ as HOGG PUT EQUATION HERE.
-The feature vectors live in a $D$-dimensional space where $D=$HOGG.
-For every star $n$ in either of the Training Sets we also construct a
-regression datum $Q_n$ and uncertainty $\sigma_n$ according to HOGG WHAT EQUATION ABOVE?
+vector $x_n$ as
+\begin{eqnarray}
+x_n\T &\equiv& [1, G_n, ({BP}-{RP})_n, J_n, H_n, K_n, W_{1n}, W_{2n}, f_{1n}, f_{2n}, \cdots, f_{Ln}]
+\quad ,
+\end{eqnarray}
+where the 1 permits a linear offset,
+the photometry is from \gaia, \zmass, and \wise, respectively,
+and
+the fluxes in the $L=$ACE \apogee\ spectral pixels for star $n$ are denoted
+$f_{1n}$, $f_{2n}$, and so on.
+These feature vectors live in a $D$-dimensional space where $D=$HOGG.
+For every star $n$ in either of the Training Sets we additionally require---in
+addition to these feature vectors---a parallax $\varpi_n$ and uncertainty $\sigma_n$.
 
@@
 In HOGG WHAT FIGURE, we show the parallax distribution from \gaia\ and
-the spectroscopic-parallax distribution from this work for stars in
+the spectrophotometric-parallax distribution from this work for stars in
 angular proximity to known stellar clusters and the Sagittarius dwarf
@@
 the accuracy of the method for different kinds of populations.
-They show that both \gaia\ and the spectroscopic parallaxes appear to be unbiased
+They show that both \gaia\ and the spectrophotometric parallaxes appear to be unbiased
 for these clusters and the range in abundances and ages they represent.

### 2018-08-01 paper/sp.tex
making the notation worse

@@
 \newcommand{\logg}{\log g}
+\newcommand{\gparallax}{\varpi^{(\mathrm{a})}}
+\newcommand{\sparallax}{\varpi^{(\mathrm{sp})}}
 
@@
 \item[sparsity] We expect that only a small subset of the full complement of
-\apogee\ spectral pixels will be relevant to the prediction of parallax.
+\apogee\ spectral pixels will be relevant to the prediction of parallax
+(the spectrophotometric parallax).
 That is, we expect that many of the pixels will or ought to get no weight in the
@@
 \begin{eqnarray}
-\varpi_n &=& \exp(\theta\cdot x_n) + \mbox{noise}
+\gparallax_n &=& \exp(\theta\cdot x_n) + \mbox{noise}
 \\
-\chi^2(\theta) &\equiv& \sum_{n=1}^N \frac{[\varpi_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
+\chi^2(\theta) &\equiv& \sum_{n=1}^N \frac{[\gparallax_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
 \quad ,
@@
 where
-$\varpi_n$ is the \gaia\ measurement (adjusted; see below) of the parallax of star $n$,
+$\gparallax_n$ is the \gaia\ measurement (adjusted; see below) of the parallax of star $n$,
 the model is that the logarithm of the true parallax
@@
 
-Once the model is optimized, the prediction for a new stellar parallax
-$\varpi_m$ for a star $m$ in the validation or test set is
+Once the model is optimized, the output of the model is a prediction of the parallax,
+or really what we will call the \emph{spectrophotometric parallax}.
+This spectrophotometric parallax
+$\sparallax_m$ for any star $m$ in the validation or test set is
+assigned according to
 \begin{eqnarray}
-\varpi_m &=& \exp(\theta\cdot x_m)
+\sparallax_m &\leftarrow& \exp(\hat{\theta}\cdot x_m)
 \quad ,
 \end{eqnarray}
-where $x_m$ is the feature $D$-vector for star $m$.
+where
+$\hat{\theta}$ is the optimal parameter vector according to (HOGG WHAT EQUATION),
+and
+$x_m$ is the feature $D$-vector for star $m$.
 Thus the model can be trained on a training set of stars and applied to make
@@
 in which the data are split into two disjoint subsets, A and B.
-The model trained on the A data is used to predict the $\varpi_m$ values (and hence
+The model trained on the A data is used to predict or produce
+the spectrophotometric parallaxes $\sparallax_m$ values (and hence
 the distances and parallaxes) of the B data,
 and the model trained on the B data is used to 
-predict the $\varpi_m$ values of the A data.
+produce the $\sparallax_m$ values of the A data.
 This ensures that, in the estimate of any individal star's
@@
 In detail,
-for each star $n$ in the Training Sets, we take from \gaia\ DR2 we the parallax $\varpi_n$
-and its uncertainty $\sigma_n$.
+for each star $n$ in the Training Sets, we take from \gaia\ DR2
+the parallax $\gparallax_n$ and its uncertainty $\sigma_n$.
 Importantly, to each \gaia\ parallax measurement we add a positive
@@
 For every star $n$ in either of the Training Sets we additionally require---in
-addition to these feature vectors---a parallax $\varpi_n$ and uncertainty $\sigma_n$.
+addition to these feature vectors---a \gaia-measured parallax $\gparallax_n$
+and uncertainty $\sigma_n$.
 

### 2018-08-01 paper/sp.tex
notes2self
thanks ACE

@@
 % - related: Put in an explanation of how it is that we aren't really modeling luminosity or distance but really just generating parallaxes.
-% - related: Switch the model to be a model for parallax directly.
 % - does the model need a cute name, like The Swan?
+% - we need a Facilities list / tags
+% - we need a Software list / tags
 
@@
 that there are $N$ independently measured stars $n$.
-The feature vector $x_n$ contains photometry in a few bands, and all 8000-ish pixels
-in the pseudo-continuum-normalized \apogee\ spectrum, so $D$ is on the order of 8000.
+The feature vector $x_n$ contains photometry in a few bands, and all 7400-ish pixels
+in the pseudo-continuum-normalized \apogee\ spectrum, so $D$ is on the order of 7400.
 
@@
 Because the \apogee\ target selection is based on \zmass, every \apogee\ star
-also comes with \zmass photometry.
+also comes with \zmass\ (HOGG CITE) photometry.
 That is, for each star $n$,
-we have near-infrared photometry $J_n$, $H_n$, and $K_n$.
-In detail we use from the catalog the entries ACE WHAT SPECIFIC CATALOG ENTRIES.
+we have \zmass\ near-infrared photometry $J_n$, $H_n$, and $K_n$.
 
@@
 
-We take from the \gaia\ \acronym{DR2} Catalog the photometric data $G_n$
-(\code{phot_g_mean_mag}) and $({BP}-{RP})_n$ (\code{bp_rp}), which will
-become part of the feature vector $x_n$.
+We take from the \gaia\ \acronym{DR2} Catalog the photometric data
+$G_n$ (\code{phot_g_mean_mag}),
+${BP}_n$ (\code{phot_bp_mean_mag}),
+and ${RP})_n$ (\code{phot_rp_mean_mag}),
+which will become part of the feature vector $x_n$.
 We do not use the entire \gaia\ \acronym{DR2} Catalog but instead the subset
@@
 meet the \apogee\ and \gaia\ cuts and also
-have the complete set of photometry: $G_n$, $({BP}-{RP})_n$, $J_n$,
+have the complete set of photometry: $G_n$, ${BP}_n$, ${RP}_n$, $J_n$,
 $H_n$, $K_n$, $W_{1n}$, and $W_{2n}$ and also an apogee spectrum.
@@
 \begin{eqnarray}
-x_n\T &\equiv& [1, G_n, ({BP}-{RP})_n, J_n, H_n, K_n, W_{1n}, W_{2n}, f_{1n}, f_{2n}, \cdots, f_{Ln}]
+x_n\T &\equiv& [1, G_n, {BP}_n, {RP}_n, J_n, H_n, K_n, W_{1n}, W_{2n}, f_{1n}, f_{2n}, \cdots, f_{Ln}]
 \quad ,
@@
 and
-the fluxes in the $L=$ACE \apogee\ spectral pixels for star $n$ are denoted
+the fluxes in the $L=7405$ \apogee\ spectral pixels (for which there are reliably
+and consistently data) for star $n$ are denoted
 $f_{1n}$, $f_{2n}$, and so on.
-These feature vectors live in a $D$-dimensional space where $D=$HOGG.
+These feature vectors live in a $D$-dimensional space where $D=7414$
 For every star $n$ in either of the Training Sets we additionally require---in
@@
   Andy L. Jones,
-  Daniel Michalik (\acronym{ESTEC}), and
-  Melissa Ness (Columbia)
+  Daniel Michalik (\acronym{ESTEC}),
+  Melissa Ness (Columbia),
+  and the participants in the \acronym{MPIA} Milky Way Group Meeting
 for help with this project.

### 2018-08-01 paper/sp.tex
affiliations

@@
 \title{\textbf{%
-Spectrophotometric parallax:\\
+Spectrophotometric parallaxes with \project{The Cygnet}:\\
 Accurate estimates of luminosity and distance\\
@@
 \author[0000-0003-2866-9403]{David W. Hogg}
-\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University}
-\affiliation{Center for Data Science, New York University}
-\affiliation{Max-Planck-Institut f\"ur Astronomie}
-\affiliation{Flatiron Institute}
+\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
+\affiliation{Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011}
+\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, 69117 Heidelberg, Germany}
+\affiliation{Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
 
 \author[0000-0003-2895-6218]{Anna-Christina Eilers}
-\affiliation{Max-Planck-Institut f\"ur Astronomie}
+\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, 69117 Heidelberg, Germany}
+\affiliation{International Max Planck Research School for Astronomy \& Cosmic Physics at the University of Heidelberg}
 
 \author[0000-0003-4996-9069]{Hans-Walter Rix}
-\affiliation{Max-Planck-Institut f\"ur Astronomie}
+\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, 69117 Heidelberg, Germany}
 

### 2018-08-01 paper/sp.tex
ACE inputs

@@
 % - we need a Software list / tags
+% - check with WISE experts that we have the correct WISE photometry
 
@@
 From the \apogee\ data on these low-$\logg$ stars, we take the spectral pixels,
-of which there are ACE HOW MANY, on a common rest-frame wavelength grid.
+of which there are 7405 (after cutting out pixels that rarely or never get data),
+on a common rest-frame wavelength grid.
 That is, every \apogee\ star is extracted on (or interpolated to)
@@
 In detail, we obtain the
-the \apogee\ spectral pixels from the pipeline-generated \code{ACE whatever} files.
+the \apogee\ spectral pixels from the pipeline-generated \code{aspcapStar} files.
 We then pseudo-continuum-normalize the spectra according to the procedure developed
-in \project{The~Cannon} (HOGG CITE): That is, the pseudo-continuum is a spline fit to a set
+in \project{The~Cannon} (HOGG CITE):
+That is, the pseudo-continuum is a spline fit to a set
 of pixels chosen to be insensitive to stellar parameters.
@@
 
-ACE: We match the \apogee+\zmass\ stars to he \wise\ ACE WHICH CATALOG by ACE WHAT CRITERION?
+We used the \gaia\ Archive (HOGG CITE) to match the \apogee+\zmass\ stars
+to the \wise\ Catalog (HOGG CITE),
+according to the \gaia\ Archive internal match criteria.
 This gives, fore each matching star $n$,
 mid-infrared photometry $W_{1n}$ and $W_{2n}$ at 3.6 and 4.5\,$\mu$m
-In detail we use the ACE WHAT SPECIFIC CATALOG ENTRIES?
+In detail we use the \code{w1mpro} and \code{w2mpro} Catalog entries.
 
-We match this full-match catalog to the \gaia\ \acronym{DR2} (HOGG CITE) by what criteria?
-
+We match this full-match catalog to the \gaia\ \acronym{DR2} (HOGG CITE)
+using the \gaia\ Archive official match (given the \zmass\ IDs).
 We take from the \gaia\ \acronym{DR2} Catalog the photometric data
@@
 ${BP}_n$ (\code{phot_bp_mean_mag}),
-and ${RP})_n$ (\code{phot_rp_mean_mag}),
+and ${RP}_n$ (\code{phot_rp_mean_mag}),
 which will become part of the feature vector $x_n$.
-We do not use the entire \gaia\ \acronym{DR2} Catalog but instead the subset
-that meets ACE WHAT ADDITIONAL PHOTOMETRIC CUTS?
 

### 2018-08-01 infer_magnitudes.py
new

@@ -29,10 +29,10 @@ fsize = 14
 # -------------------------------------------------------------------------------
 
 # make plots?
-prediction = False
+prediction = True
 
 print('loading labels...')
-hdu = fits.open('data/training_labels_parent_apogeedr15.fits')
+hdu = fits.open('data/training_labels_parent.fits')
 labels = Table(hdu[1].data)
 
 offset = 0.029 # mas as per Lindegren et al. 2018
@@ -42,7 +42,7 @@ if prediction:
     
     print('loading spectra...')
 
-    hdu = fits.open('data/all_flux_norm_parent_apogeedr15.fits')
+    hdu = fits.open('data/all_flux_norm_parent.fits')
     fluxes = hdu[0].data
                           
 # -------------------------------------------------------------------------------
@@ -115,7 +115,7 @@ def check_H_func(x, y, A, lams, ivar):
 
 Kfold = 2
 lam = 10000                      # hyperparameter -- needs to be tuned!
-name = 'N{0}_lam{1}_K{2}_dr15'.format(len(labels), lam, Kfold)
+name = 'N{0}_lam{1}_K{2}'.format(len(labels), lam, Kfold)
 
 if prediction:        
     

### 2018-08-01 map_making.py

@@ -38,8 +38,10 @@ fsize = 14
 
 N = 45787
 Kfold = 2
-lam = 100
-name = 'N{0}_lam{1}_K{2}'.format(N, lam, Kfold)
+lam = 30
+#name = 'N{0}_lam{1}_K{2}'.format(N, lam, Kfold)
+name = 'N{0}_lam{1}_K{2}_mag_allcolors_offset'.format(N, lam, Kfold)
+
 
 print('loading new labels...')   
 labels = Table.read('data/training_labels_new_{}_2.fits'.format(name), format = 'fits')    

### 2018-08-02 paper/sp.tex
thanks sven
small model issues

@@
 \begin{eqnarray}
-\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2(\theta) + ||\lambda\cdot\theta||_1^1\right]
+\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2(\theta) + ||\lambda\ast\theta||_1^1\right]
 \quad ,
@@
 different parts of the $x_n$ vectors differently),
-and $||x||_1^1$ is the L1-norm or sum of absolute values of the components of $x$.
+the ``$a\ast b$'' operator indicates element-wise multiplication of vectors $a$ and $b$,
+and $||y||_1^1$ is the L1-norm or sum of absolute values of the components of $y$.
 This kind of regularization adds a convex term to the optimization and leads to
@@
 \begin{eqnarray}
-x_n\T &\equiv& [1, G_n, {BP}_n, {RP}_n, J_n, H_n, K_n, W_{1n}, W_{2n}, f_{1n}, f_{2n}, \cdots, f_{Ln}]
+x_n\T &\equiv& [1, G_n, {BP}_n, {RP}_n, J_n, H_n, K_n, W_{1n}, W_{2n}, \ln f_{1n}, \ln f_{2n}, \cdots, \ln f_{Ln}]
 \quad ,
@@
 the photometry is from \gaia, \zmass, and \wise, respectively,
-and
 the fluxes in the $L=7405$ \apogee\ spectral pixels (for which there are reliably
 and consistently data) for star $n$ are denoted
-$f_{1n}$, $f_{2n}$, and so on.
+$f_{1n}$, $f_{2n}$, and so on,
+and we have taken logarithms of those fluxes.
 These feature vectors live in a $D$-dimensional space where $D=7414$
@@
   Coryn Bailer-Jones (\acronym{MPIA}),
+  Sven Buder (\acronym{MPIA}),
   Andy L. Jones,

### 2018-08-02 paper/sp.tex
intro paragraphs

@@
 Spectrophotometric parallaxes with \project{The Cygnet}:\\
-Accurate estimates of luminosity and distance\\
-for luminous red-giant stars
+Accurate distances for luminous red-giant stars
 }}
@@
 
+\section*{~}\clearpage
 \section{Introduction} \label{sec:intro}
@@
 
-In this contribution we develop data-driven techniques to predict
-stellar luminosities and distances with spectroscopy and photometry...
+Luminous red-giant stars are valuable for mapping the Milky Way disk for a number
+lf reasons, only one of which is their great luminosities (and hence brightnesses
+even at large distance).
+They are extremely common (frequent) stars, much more common than distance-indicating
+stars like Cepheids.
+They are produced by stellar populations at all metallicities and almost all ages,
+so they can be found in all parts of the Galaxy.
+They have temperatures and surface gravities such that it is
+straightforward to spectroscopically measure metallicities,
+detailed chamical abundances, and stellar parameters, including ages (CITE MARTIG, NESS).
+The red-giant branch is photometrically near-orthogonal to reddening vectors by dust,
+so they are easy to de-redden or dust-correct.
+And finally, in physical models of red giants, their predicted luminosities are simple
+functions of composition, surface gravity, temperature, and age.
+Thus, if we can measure these properties of red giants, we can (in principle) predict
+their luminosities and use them as standard candles.
+That is---like type Ia supernovae---red giants are standardizable candles.
 
-The history and value (and disadvantages)
-of data-driven models for these kinds of problems.
-How could it be that we are completely independent of stellar models?
+In this project, we develop purely data-driven techniques to predict
+red-giant parallaxes or distances with spectroscopy and photometry.
+Our method capitalizes on the point that red-giant stars are dust-correctable,
+standardizable candles, but it makes no use whatsoever of any stellar interior,
+stellar exterior, or dust-reddening models.
+We only use these physical expectations about red giants to recommend
+data features---inputs---for the data-driven model.
+The data-driven method uses patterns in the observed data themselves
+to discover the relationships between spectral features in the spectra of the stars,
+photometry (including colors), and parallax (or distance).
 
-The dangers of aggressive machine-learning methods. Why will we just
-look at linear models? The disadvantages of this approach.
+Data-driven models are very often more accurate than physical models of stars
+for this kind of work, because they have the flexibility to discover
+regularities or offsets in the data that are not captured by the physical
+models.
+They do better than physical models when there are lots of training data
+with properties that are good for performing regressions.
 
-The dangers of cutting on \gaia\ properties. Why we will not cut on
+Simple data-driven models (like the ones we will employ here)
+suffer from many limitations for this kind of work:
+Because they are constrained only by the data, and not physical law,
+they can learn relationships that we know
+to be physically unlikely, and (by the same token) they can't benefit from our
+physical knowledge.
+They are required, in some sense, to spend some of the information in the data on
+learning physical laws or relationships.
+Not all of the information in the data is being used
+optimally when we ignore our physical models.
+Relatedly, they are not (usually) interpretable or generalizable:
+A data-driven model trained with one kind of data can not usually be used in another
+context with very different input data.
+And the internal properties of the model are not useful, in general, for improving
+our understanding or intuition about the astrophysics in play.
+
+There is a huge range of complexity or capacity in data-driven
+models, from linear regressions (what we use here) to the most advanced machine-learning
+methods (like deep neural networks).
+In this work we stay on the low-complexity side of this.
+We do not want to face the (literally combinatoric) choices in model structure that
+advanced methods bring.
+And we want to control or understand the smoothness or flexibility of the model.
+Deep networks can model extremely complex function spaces, which is good in some
+contexts, but bad when you know that the functional dependencies you hope to
+model are smooth:
+In that case, much of the information in the data can get spent on learning the
+smoothness of the function.
+
+All that said, our goal here is to produce a precise mapping tool for surveys
+like \apogee\ and \sdssv.
+So we will take the limitations of the data-driven models in exchange for their
+performance.
+This work builds on the success we have seen in building other kinds of
+data-driven tools for  stellar spectroscopy, for example to measure stellar
+parameters (CITE NESS), abundances (CITE HO, NESS, CASEY), and masses and ages (CITE NESS).
+
+HOGG: The dangers of cutting on \gaia\ properties. Why we will not cut on
 \gaia\ parallaxes. The disadvantages of this approach.
-
-\section{Why and how should we model red-giant stars?}
-
-HOGG: We need to, for \sdssv.
-
-HOGG: We know we can because of physical arguments.
-
-HOGG: What kinds of inputs do we need to predict parallaxes?
 

### 2018-08-02 paper/sp.tex
censoring

@@
 
+All that said, our goal here is to produce a precise mapping tool for surveys
+like \apogee\ and \sdssv.
+So we will take the limitations of the data-driven models in exchange for their
+performance.
+This work builds on the success we have seen in building other kinds of
+data-driven tools for  stellar spectroscopy, for example to measure stellar
+parameters (CITE NESS), abundances (CITE HO, NESS, CASEY), and masses and ages (CITE NESS).
+
 There is a huge range of complexity or capacity in data-driven
@@
 
-All that said, our goal here is to produce a precise mapping tool for surveys
-like \apogee\ and \sdssv.
-So we will take the limitations of the data-driven models in exchange for their
-performance.
-This work builds on the success we have seen in building other kinds of
-data-driven tools for  stellar spectroscopy, for example to measure stellar
-parameters (CITE NESS), abundances (CITE HO, NESS, CASEY), and masses and ages (CITE NESS).
-
-HOGG: The dangers of cutting on \gaia\ properties. Why we will not cut on
-\gaia\ parallaxes. The disadvantages of this approach.
+What follows is a regression:
+We use \gaia\ measurements of parallax to learn relationships between 
+the spectroscopic and photometric properties of stars and their parallaxes.
+We then use that learned model to estimate improved parallaxes for stars with
+poor (or no) \gaia\ measurements.
+This regression is very sensitive to certain aspects of the data or experimental
+design..
+One is that we will do far better with the regression if we have an accurate
+model for noise process in the \gaia\ data.
+For this reason, we make use of a justifiable likelihood function for the
+\gaia\ astrometry (CITE HOGG).
+Another is that we can be strongly biased if we have a bad selection function
+or data censoring.
+HOGGGGGThis means that we cannot apply cuts to the \gaia\ astrometry or signal-to-noise
+unless those cuts are correctly accounted for in our generative model.HOGGG GGG
 

### 2018-08-02 paper/sp.tex
intro complete, but terrible

@@
 \gaia\ astrometry (CITE HOGG).
-Another is that we can be strongly biased if we have a bad selection function
+Another is that regressions can be strongly biased if we have a bad selection function
 or data censoring.
-HOGGGGGThis means that we cannot apply cuts to the \gaia\ astrometry or signal-to-noise
-unless those cuts are correctly accounted for in our generative model.HOGGG GGG
+This means that we cannot apply cuts to the \gaia\ astrometry or signal-to-noise
+unless those cuts are correctly accounted for in our generative model for the
+data sample.
+Below, we will make no such cuts. That means that our training step has the odd
+property that much of the training data is low in signal-to-noise and many of the
+training parallax measurements are even negative!
+The fact that we are using a justified likelihood function ensures that these
+data will only affect the model in sensible and justifiable ways.
+The alternative---cutting to high signal-to-noise data---is the more traditional
+approach, but it leads to biased inferences (unless the cuts are included correctly
+in the likelihood function).
+The method in which we make no cuts and keep an umodified, justified likelihood function
+is unbiased, and simple.
 

### 2018-08-02 paper/sp.tex
notes to self
citations and acks

@@
 % -----------
-% - draft introduction
 % - draft discussion
 % - make sure we address the sensibleness and generality of the linear model for the photometric space
-% - address the non-convexity of the model (see twitter convo with @andy_l_jones).
+% - address the non-convexity of the model (see twitter convo with @andy_l_jones). Plus optimization issues.
 % - Coryn's point that we should carefully explain how our model is geometric.
@@
 % text macros
+\newcommand{\equationname}{equation}
 \newcommand{\code}[1]{\texttt{\detokenize{#1}}}
 \newcommand{\foreign}[1]{\textsl{#1}}
+\newcommand{\etal}{\foreign{et al.}}
 \newcommand{\acronym}[1]{{\small{#1}}}
@@
 For this reason, we make use of a justifiable likelihood function for the
-\gaia\ astrometry (CITE HOGG).
+\gaia\ astrometry (\citealt{gaialf}).
 Another is that regressions can be strongly biased if we have a bad selection function
@@
 plus the $G$ and $({BP}-{RP})$ photometry from \gaia,
-plus the $J$, $H$, and $K_s$ photometry from \zmass,
+plus the $J$, $H$, and $K_s$ photometry from \zmass (\citealt{zmass}),
 plus the $W_1$ and $W_2$ photometry from \wise.
@@
 \hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2(\theta) + ||\lambda\ast\theta||_1^1\right]
-\quad ,
+\label{eq:objective}\quad ,
 \end{eqnarray}
@@
 where
-$\hat{\theta}$ is the optimal parameter vector according to (HOGG WHAT EQUATION),
+$\hat{\theta}$ is the optimal parameter vector according
+to \equationname~(\ref{eq:objective}),
 and
@@
 Because the \apogee\ target selection is based on \zmass, every \apogee\ star
-also comes with \zmass\ (HOGG CITE) photometry.
+also comes with \zmass\ (\citealt{zmass}) photometry.
 That is, for each star $n$,
@@
 
-\wise\ Acknowledgement?
+This publication makes use of data products from the Two Micron All Sky Survey, which is a joint project of the University of Massachusetts and the Infrared Processing and Analysis Center/California Institute of Technology, funded by the National Aeronautics and Space Administration and the National Science Foundation.
 
-\project{\acronym{SDSS-IV}} ack.
+This publication makes use of data products from the Wide-field Infrared Survey Explorer, which is a joint project of the University of California, Los Angeles, and the Jet Propulsion Laboratory/California Institute of Technology, funded by the National Aeronautics and Space Administration.
+
+Funding for the \project{Sloan Digital Sky Survey IV} has been provided by the Alfred P.
+Sloan Foundation, the U.S Department of Energy Office of Science, and the Participating Institutions.
+\sdssiv\ acknowledges
+support and resources from the Center for High-Performance Computing at
+the University of Utah. The \project{\acronym{SDSS}} web site is www.sdss.org.
+
+\sdssiv\ is managed by the Astrophysical Research Consortium for the 
+Participating Institutions of the \project{\acronym{SDSS}} Collaboration including the 
+Brazilian Participation Group, the Carnegie Institution for Science, 
+Carnegie Mellon University, the Chilean Participation Group, the French Participation Group, Harvard-Smithsonian Center for Astrophysics, 
+Instituto de Astrof\'isica de Canarias, The Johns Hopkins University, 
+Kavli Institute for the Physics and Mathematics of the Universe / 
+University of Tokyo, the Korean Participation Group, Lawrence Berkeley National Laboratory, 
+Leibniz Institut f\"ur Astrophysik Potsdam,  
+Max-Planck-Institut f\"ur Astronomie (Heidelberg), 
+Max-Planck-Institut f\"ur Astrophysik (Garching), 
+Max-Planck-Institut f\"ur Extraterrestrische Physik, 
+National Astronomical Observatories of China, New Mexico State University, 
+New York University, University of Notre Dame, 
+Observat\'ario Nacional~/~\acronym{MCTI}, The Ohio State University, 
+Pennsylvania State University, Shanghai Astronomical Observatory, 
+United Kingdom Participation Group,
+Universidad Nacional Aut\'onoma de M\'exico, University of Arizona, 
+University of Colorado Boulder, University of Oxford, University of Portsmouth, 
+University of Utah, University of Virginia, University of Washington, University of Wisconsin, 
+Vanderbilt University, and Yale University.
 
 \begin{thebibliography}{}
-\bibitem[Astropy Collaboration et al.(2013)]{2013A&A...558A..33A} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, \aap, 558, A33 
-\bibitem[Bertin \& Arnouts(1996)]{1996A&AS..117..393B} Bertin, E., \& Arnouts, S.\ 1996, \aaps, 117, 393 
-\bibitem[Corrales(2015)]{2015ApJ...805...23C} Corrales, L.\ 2015, \apj, 805, 23
-\bibitem[Ferland et al.(2013)]{2013RMxAA..49..137F} Ferland, G.~J., Porter, R.~L., van Hoof, P.~A.~M., et al.\ 2013, \rmxaa, 49, 137
-\bibitem[Hanisch \& Biemesderfer(1989)]{1989BAAS...21..780H} Hanisch, R.~J., \& Biemesderfer, C.~D.\ 1989, \baas, 21, 780 
-\bibitem[Lamport(1994)]{lamport94} Lamport, L. 1994, LaTeX: A Document Preparation System, 2nd Edition (Boston, Addison-Wesley Professional)
-\bibitem[Schwarz et al.(2011)]{2011ApJS..197...31S} Schwarz, G.~J., Ness, J.-U., Osborne, J.~P., et al.\ 2011, \apjs, 197, 31  
-\bibitem[Vogt et al.(2014)]{2014ApJ...793..127V} Vogt, F.~P.~A., Dopita, M.~A., Kewley, L.~J., et al.\ 2014, \apj, 793, 127  
+\bibitem[Hogg(2018)]{gaialf} Hogg, D.~W 2018, arXiv:1804.07766 
+\bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., et al.\ 2006, \aj, 131, 1163 
 \end{thebibliography}
 
-%% This command is needed to show the entire author+affilation list when
-%% the collaboration and author truncation commands are used.  It has to
-%% go at the end of the manuscript.
-%\allauthors
-
-%% Include this line if you are using the \added, \replaced, \deleted
-%% commands to see a summary list of all changes at the end of the article.
-%\listofchanges
-
 \end{document}
-
-% End of file `sample62.tex'.

### 2018-08-02 paper/sp.tex
lindegren

@@
 stochastic process, in which the difference between the \gaia-Catalog parallax (plus
-a small \gaia-recommended offset HOGG CITE) and the true parallax is effectively drawn from a
+a small \gaia-recommended offset; \citealt{lindegren}) and the true parallax is effectively drawn from a
 Gaussian with a width set by the \gaia-Catalog uncertainty on the parallax.
@@
 \bibitem[Hogg(2018)]{gaialf} Hogg, D.~W 2018, arXiv:1804.07766 
+\bibitem[Lindegren \etal(2018)]{lindegren} Lindegren, L., Hernandez, J., Bombrun, A., et al.\ 2018, arXiv:1804.09366 
 \bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., et al.\ 2006, \aj, 131, 1163 

### 2018-08-02 paper/sp.tex

@@
 We take the \apogee\ (HOGG CITE) spectral data
-from \sdssiv\ (HOGG CITE) \acronym{DR14} (HOGG CITE).
+from \sdssiv\ (\citealt{sdssiv}) \acronym{DR14} (HOGG CITE).
 Because we want to make a purely linear model, which has very little capacity,
@@
 support and resources from the Center for High-Performance Computing at
-the University of Utah. The \project{\acronym{SDSS}} web site is www.sdss.org.
+the University of Utah. The \project{\acronym{SDSS}} web site is \url{www.sdss.org}.
 
@@
 \begin{thebibliography}{}
+\bibitem[Blanton \etal(2017)]{sdssiv} Blanton, M.~R., Bershady, M.~A., Abolfathi, B., \etal\ 2017, \aj, 154, 28 
 \bibitem[Hogg(2018)]{gaialf} Hogg, D.~W 2018, arXiv:1804.07766 
-\bibitem[Lindegren \etal(2018)]{lindegren} Lindegren, L., Hernandez, J., Bombrun, A., et al.\ 2018, arXiv:1804.09366 
-\bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., et al.\ 2006, \aj, 131, 1163 
+\bibitem[Lindegren \etal(2018)]{lindegren} Lindegren, L., Hernandez, J., Bombrun, A., \etal\ 2018, arXiv:1804.09366 
+\bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., \etal\ 2006, \aj, 131, 1163 
 \end{thebibliography}

### 2018-08-02 infer_parallax.py
new

@@ -66,7 +66,7 @@ def H_func(x, y, A, lams, ivar):
 
 def check_H_func(x, y, A, lams, ivar):
     H0, dHdx0 = H_func(x, y, A, lams, ivar)
-    dx = 0.0001 # magic
+    dx = 1e-6 # magic
     for i in range(len(x)):
         x1 = 1. * x
         x1[i] += dx
@@ -83,6 +83,11 @@ Kfold = 2
 lam = 30                      # hyperparameter -- needs to be tuned!
 name = 'N{0}_lam{1}_K{2}_parallax'.format(len(labels), lam, Kfold)
 
+# optimization schedule
+# 1. photometry only
+# 2. go down with lambda?
+steps = 2
+
 if prediction:        
     
     # data
@@ -92,7 +97,8 @@ if prediction:
     
     # design matrix
     AT_0 = np.vstack([np.ones_like(y_all)])
-    AT_linear = np.vstack([labels['phot_g_mean_mag'], labels['phot_bp_mean_mag'], labels['phot_rp_mean_mag'], labels['J'], labels['H'], labels['K'], labels['w1mpro'], labels['w2mpro'], fluxes])
+    ln_fluxes = np.log(np.clip(fluxes, 0.01, 1.2))
+    AT_linear = np.vstack([labels['phot_g_mean_mag'], labels['phot_bp_mean_mag'], labels['phot_rp_mean_mag'], labels['J'], labels['H'], labels['K'], labels['w1mpro'], labels['w2mpro'], ln_fluxes])
     A_all = np.vstack([AT_0, AT_linear]).T
     
     # fucking brittle
@@ -123,30 +129,45 @@ if prediction:
         
         # cut in parallax_error
         cut_par = labels[train]['parallax_error'] < 0.1       # this cut is not strictly required!
-    
+        cut_burnin = labels[train]['parallax_over_error'] > 20.
+        
         # cut in astrometric_gof_al (should be around unity...?) *Daniel Michalik's advice*
         #cut_gof = labels[train]['astrometric_gof_al'] < 5  
         # Coryn's advice!
         cut_cal = (labels[train]['astrometric_chi2_al'] / np.sqrt(labels[train]['astrometric_n_good_obs_al']-5)) <= 35         
         
         # more cuts? e.g. astrometric_gof_al should be low!
-                          
-        cut_all = cut_parallax * cut_vis * cut_par * cut_cal      
[…]

### 2018-08-02 paper/sp.tex
refs

@@
 \newcommand{\etal}{\foreign{et al.}}
+\newcommand{\eg}{\foreign{e.g.}}
 \newcommand{\acronym}[1]{{\small{#1}}}
@@
 
-Finally, we never use Lutz--Kelker corrections (HOGG CITE) or distance
-posteriors (HOGG CITE). These both involve (implicit or explicit) priors on 
+Finally, we never use Lutz--Kelker corrections (\citealt{lk}) or distance
+posteriors (\eg, \citealt{calj}). These both involve (implicit or explicit) priors on 
 distance.
@@
 offset of $0.029$\,mas to adjust for the under-estimation of
-parallaxes reported by the \gaia\ Collaboration (HOGG CITE LINDEGREN).
+parallaxes reported by the \gaia\ Collaboration (\citealt{lindegren}).
 The training sets are shown in HOGG WHAT FIGURES?
@@
 \begin{thebibliography}{}
+\bibitem[Bailer-Jones et al.(2018)]{calj} Bailer-Jones, C.~A.~L., Rybizki, J., Fouesneau, M., Mantelet, G., \& Andrae, R.\ 2018, \aj, 156, 58 
 \bibitem[Blanton \etal(2017)]{sdssiv} Blanton, M.~R., Bershady, M.~A., Abolfathi, B., \etal\ 2017, \aj, 154, 28 
@@
 \bibitem[Lindegren \etal(2018)]{lindegren} Lindegren, L., Hernandez, J., Bombrun, A., \etal\ 2018, arXiv:1804.09366 
+\bibitem[Lutz \& Kelker(1973)]{lk} Lutz, T.~E., \& Kelker, D.~H.\ 1973, \pasp, 85, 573 
 \bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., \etal\ 2006, \aj, 131, 1163 

### 2018-08-02 paper/sp.tex
refs

@@
 We take the \apogee\ (HOGG CITE) spectral data
-from \sdssiv\ (\citealt{sdssiv}) \acronym{DR14} (HOGG CITE).
+from \sdssiv\ (\citealt{sdssiv}) \acronym{DR14} (\citealt{dr14}).
 Because we want to make a purely linear model, which has very little capacity,
@@
 
-Because the \apogee\ target selection is based on \zmass, every \apogee\ star
-also comes with \zmass\ (\citealt{zmass}) photometry.
+Because the \apogee\ target selection is based on \zmass (\citealt{zmass}),
+every \apogee\ star also comes with \zmass\ photometry.
 That is, for each star $n$,
@@
 \begin{thebibliography}{}
+\bibitem[Abolfathi \etal(2018)]{dr14} Abolfathi, B., Aguado, D.~S., Aguilar, G., \etal\ 2018, \apjs, 235, 42 
 \bibitem[Bailer-Jones et al.(2018)]{calj} Bailer-Jones, C.~A.~L., Rybizki, J., Fouesneau, M., Mantelet, G., \& Andrae, R.\ 2018, \aj, 156, 58 

### 2018-08-03 paper/sp.tex
regularization

@@
 \begin{eqnarray}
-\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2(\theta) + ||\lambda\ast\theta||_1^1\right]
+\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2(\theta) + \lambda\,||P\cdot\theta||_1^1\right]
 \label{eq:objective}\quad ,
@@
 where
-$\lambda$ is a $D$-vector of regularization parameters (such that we can regularize
-different parts of the $x_n$ vectors differently),
-the ``$a\ast b$'' operator indicates element-wise multiplication of vectors $a$ and $b$,
+$\lambda$ is a regularization parameter,
+$P$ is a projection operator that selects out from the $\theta$ vector only those components
+that pertain to the \apogee\ spectral pixels,
 and $||y||_1^1$ is the L1-norm or sum of absolute values of the components of $y$.
@@
 sparse optima.
-We will set the value $\lambda$ of the $D$-vector of regularization strengths by a
+The $P$ operator makes it such that we only regularize the components of $\theta$ that multiply
+the spectral pixels; we ask for the spectral model to be sparse but we don't ask for the photometric
+model to be sparse.
+We set the value of $\lambda$ by a 
 very coarse cross-validation.

### 2018-08-03 paper/sp.tex
notes to self

@@
 
+HOGG: Similarly, the spectral part of the model should also be
+interpretable. HOGG SHOW FIGURE THAT DEMONSTRATES THIS, qualitatively.
+
 HOGG: How to get the data and how to use it, probabilistically.
@@
 Why did we make the assumptions we made and what did they gain us and cost us?
+
+In that context: This model is only linear! We could have done much better with
+a more flexible model. But choices, execution.
+
+In that context: These features are bad! We could have used feature engineering
+or an autoencoder to do a much better job. Definitely should do this in the future.
 

### 2018-08-03 rotation_curve_uncertainties.py
new

@@ -39,10 +39,10 @@ fsize = 14
 N = 45787
 Kfold = 2
 lam = 30
-name = 'N{0}_lam{1}_K{2}_mag_allcolors_offset'.format(N, lam, Kfold)
+name = 'N{0}_lam{1}_K{2}_parallax'.format(N, lam, Kfold)
 
 print('loading new labels...')   
-labels = Table.read('data/training_labels_new_{}_2.fits'.format(name), format = 'fits')    
+labels = Table.read('data/training_labels_new_{}.fits'.format(name), format = 'fits')    
 labels.rename_column('ra_1', 'ra')
 labels.rename_column('dec_1', 'dec')
 
@@ -51,32 +51,23 @@ labels.rename_column('dec_1', 'dec')
 # -------------------------------------------------------------------------------           
 
 # Galactocentric position of the Sun:
-X_GC_sun_kpc = 8.3    #[kpc]
+X_GC_sun_kpc = 8.122 #[kpc] # (Gravity collaboration 2018)
 Z_GC_sun_kpc = 0.025 #[kpc] (e.g. Juric et al. 2008)
 
-# circular velocity of the Galactic potential at the radius of the Sun:
-vcirc_kms = 220. #[km/s] (e.g. Bovy 2015)
-
-# Velocity of the Sun w.r.t. the Local Standard of Rest (e.g. Schoenrich et al. 2009):
-U_LSR_kms = 11.1  # [km/s]
-V_LSR_kms = 12.24 # [km/s]
-W_LSR_kms = 7.25  # [km/s]
-
 # Galactocentric velocity of the Sun:
-vX_GC_sun_kms = -U_LSR_kms           # = -U              [km/s]
-vY_GC_sun_kms =  V_LSR_kms+vcirc_kms # = V+v_circ(R_Sun) [km/s]
-vZ_GC_sun_kms =  W_LSR_kms           # = W               [km/s]
-
-# keep proper motion of Sgr A* constant! 
-vY_GC_sun_kms = X_GC_sun_kpc * vY_GC_sun_kms / 8.
+vX_GC_sun_kms = -11.1 # [km/s]   (e.g. Schoenrich et al. 2009) 
+vY_GC_sun_kms =  245.8 # [km/s]  (combined with Sgr A* proper motions from Reid & Brunnthaler 2004)
+vZ_GC_sun_kms =  7.8 # [km/s]
 
 # -------------------------------------------------------------------------------
 # re-sample each star
[…]

### 2018-08-03 paper/sp.tex
symbols

@@
 \newcommand{\logg}{\log g}
+\newcommand{\BP}{B\!P}
+\newcommand{\RP}{R\!P}
 \newcommand{\gparallax}{\varpi^{(\mathrm{a})}}
@@
 the full set of (pseudo-continuum-normalized) pixels from the \apogee\ spectrum,
-plus the $G$ and $({BP}-{RP})$ photometry from \gaia,
+plus the $G$ and $(\BP-\RP)$ photometry from \gaia,
 plus the $J$, $H$, and $K_s$ photometry from \zmass (\citealt{zmass}),
@@
 $G_n$ (\code{phot_g_mean_mag}),
-${BP}_n$ (\code{phot_bp_mean_mag}),
-and ${RP}_n$ (\code{phot_rp_mean_mag}),
+$\BP_n$ (\code{phot_bp_mean_mag}),
+and $\RP_n$ (\code{phot_rp_mean_mag}),
 which will become part of the feature vector $x_n$.
@@
 meet the \apogee\ and \gaia\ cuts and also
-have the complete set of photometry: $G_n$, ${BP}_n$, ${RP}_n$, $J_n$,
+have the complete set of photometry: $G_n$, $\BP_n$, $\RP_n$, $J_n$,
 $H_n$, $K_n$, $W_{1n}$, and $W_{2n}$ and also an apogee spectrum.
@@
 \begin{eqnarray}
-x_n\T &\equiv& [1, G_n, {BP}_n, {RP}_n, J_n, H_n, K_n, W_{1n}, W_{2n}, \ln f_{1n}, \ln f_{2n}, \cdots, \ln f_{Ln}]
+x_n\T &\equiv& [1, G_n, \BP_n, \RP_n, J_n, H_n, K_n, W_{1n}, W_{2n}, \ln f_{1n}, \ln f_{2n}, \cdots, \ln f_{Ln}]
 \quad ,

### 2018-08-03 paper/sp.tex
scipy

@@
 We optimize two models, one for Training Set A and one for Training Set B.
-For optimization, we use \code{scipy} l-bfgs-b (HOGG CHECK), starting at
+For optimization, we use \code{scipy.optimize} \acronym{L-BFGS-B}, starting at
 an uninformed starting point in the $\theta$ space.
-The optimization is of the objective function given in (HOGG WHAT EQUATION) above,
+The optimization is of the objective function given in \equationname~(\ref{eq:objective}) above,
 and although this optimization is not convex, it appears to converge to the

### 2018-08-03 paper/sp.tex
sdssvv

@@
 the infrared.
-These arguments motivate the \apogee\ (CITE), \apogee-2 (CITE), and \sdssv\ (CITE)
+These arguments motivate the \apogee\ (CITE), \apogee-2 (CITE), and \sdssv\ (\citealt{sdssv})
 projects, which take spectra in the infrared, and deliver detailed abundances
@@
 
-This work has made use of data from the European Space Agency (ESA) mission
+This work has made use of data from the European Space Agency (\acronym{ESA}) mission
 \gaia\ (\url{https://www.cosmos.esa.int/gaia}), processed by the \gaia\ Data
@@
 \bibitem[Hogg(2018)]{gaialf} Hogg, D.~W 2018, arXiv:1804.07766 
+\bibitem[Kollmeier \etal(2017)]{sdssv} Kollmeier, J.~A., Zasowski, G., Rix, H.-W., \etal\ 2017, arXiv:1711.03234 
 \bibitem[Lindegren \etal(2018)]{lindegren} Lindegren, L., Hernandez, J., Bombrun, A., \etal\ 2018, arXiv:1804.09366 

### 2018-08-03 paper/sp.tex
figures

@@
 $H_n$, $K_n$, $W_{1n}$, and $W_{2n}$ and also an apogee spectrum.
-This Parent Sample contains ACE HOW MANY STARS? and is shown in HOGG WHAT FIGURES?
+This Parent Sample contains ACE HOW MANY STARS? and is shown in \figurename~\ref{fig:parent}.
+\begin{figure}
+\caption{A visualization of the Parent Sample stars. The left panel
+  shows all XXXX stars in the combined Parent Sample, colored by
+  the \apogee\ pipeline estimates of surface gravity. The right
+  panel shows small spectral segments of four example stars,
+  chosen to span metallicity and surface gravity.\label{fig:parent}}
+\end{figure}
 
@@
 parallaxes reported by the \gaia\ Collaboration (\citealt{lindegren}).
-The training sets are shown in HOGG WHAT FIGURES?
+The Training Sets are shown in \figurename~\ref{fig:training}.
+\begin{figure}
+\caption{A visualization of the Training Set stars.\label{fig:training}}
+\end{figure}
 

### 2018-08-03 paper/sp.tex
tweak

@@
 
-In addition, we assume that many entries in the $D$-vector will be zero
+In addition, we assume that many entries in the parameter $D$-vector $\theta$ will be zero
 (the ``sparsity'' assumption).
@@
 
-From the \apogee\ data on these low-$\logg$ stars, we take the spectral pixels,
+From the \apogee\ data on these low-gravity stars, we take the spectral pixels,
 of which there are 7405 (after cutting out pixels that rarely or never get data),

### 2018-08-03 paper/sp.tex
moar caption; parallaX

@@
 where
-$\gparallax_n$ is the \gaia\ measurement (adjusted; see below) of the parallax of star $n$,
+$\gparallax_n$ is the \gaia\ measurement (adjusted; see below) of the astrometric parallax of star $n$,
 the model is that the logarithm of the true parallax
@@
 \caption{A visualization of the Parent Sample stars. The left panel
-  shows all XXXX stars in the combined Parent Sample, colored by
+  shows a color--color diagram for all XXXX stars in the combined Parent Sample, colored by
   the \apogee\ pipeline estimates of surface gravity. The right
@@
 for each star $n$ in the Training Sets, we take from \gaia\ DR2
-the parallax $\gparallax_n$ and its uncertainty $\sigma_n$.
+the astrometric parallax $\gparallax_n$ and its uncertainty $\sigma_n$.
 Importantly, to each \gaia\ parallax measurement we add a positive
@@
 \begin{figure}
-\caption{A visualization of the Training Set stars.\label{fig:training}}
+\caption{A visualization of the Training Set stars. The left panel shows
+  same color--color diagram as in \figurename~\ref{fig:parent}, but for
+  the Training Set only. Note [HOGG SOME DIFFERENCES]. The right panel
+  shows HOGG WHAT?\label{fig:training}}
 \end{figure}
@@
 For every star $n$ in either of the Training Sets we additionally require---in
-addition to these feature vectors---a \gaia-measured parallax $\gparallax_n$
+addition to these feature vectors---a \gaia-measured astrometric parallax $\gparallax_n$
 and uncertainty $\sigma_n$.

### 2018-08-03 paper/sp.tex
typo

@@
 have the complete set of photometry: $G_n$, $\BP_n$, $\RP_n$, $J_n$,
-$H_n$, $K_n$, $W_{1n}$, and $W_{2n}$ and also an apogee spectrum.
+$H_n$, $K_n$, $W_{1n}$, and $W_{2n}$ and also an \apogee\ spectrum.
 This Parent Sample contains ACE HOW MANY STARS? and is shown in \figurename~\ref{fig:parent}.

### 2018-08-03 rotation_curve_uncertainties.py
new

@@ -71,7 +71,7 @@ N = len(labels) # 33044 stars!
 
 N_sample = 256
 np.random.seed(42)
-mean_XS_cart_n = np.zeros((N, 6))
+'''mean_XS_cart_n = np.zeros((N, 6))
 var_XS_cart_n = np.zeros((N, 3, 3))
 mean_XS_cyl_n = np.zeros((N, 6))
 var_XS_cyl_n = np.zeros((N, 3, 3))
@@ -93,7 +93,7 @@ for i in range(N):
     
     pmras = np.random.normal(labels['pmra'][i], scale = labels['pmra_error'][i], size = N_sample)
     pmdecs = np.random.normal(labels['pmdec'][i], scale = labels['pmdec_error'][i], size = N_sample)
-    vrs = np.random.normal(labels['VHELIO_AVG'][i], scale = np.sqrt(floor_rv**2 + labels['radial_velocity_error'][i]**2), size = N_sample)
+    vrs = np.random.normal(labels['VHELIO_AVG'][i], scale = np.sqrt(floor_rv**2), size = N_sample)
                              
     # -------------------------------------------------------------------------------
     # calculate cartesian coordinates
@@ -217,7 +217,7 @@ def overplot_ring_helio(r):
     thetas = np.arange(0., 2*np.pi + tiny, 0.001 * np.pi)
     xs = r * np.cos(thetas) - X_GC_sun_kpc
     ys = r * np.sin(thetas)
-    plt.plot(xs, ys, "k-", alpha=0.2, lw=1, zorder = -np.inf)
+    plt.plot(xs, ys, "k", alpha=0.2, lw=.8, linestyle = ':', zorder = np.inf)
     # plt.scatter(-X_GC_sun_kpc, 0, s = 10, color = 'k', alpha=0.2)
     return
 
@@ -233,17 +233,6 @@ def overplot_rings_helio():
 
 Xlimits = [[-30, 10], [-10, 30], [-20, 20], 
            [-200, 200], [-200, 200], [-200, 200]]
-
-# plot [FE/H] vs. radius
-cut_z = abs(mean_XS_cart_n[:, 2]) < 0.5
-fig, ax = plt.subplots(1, 1, figsize = (8, 8))        
[…]

### 2018-08-03 paper/sp.tex
optimization

@@
 % - make sure we address the sensibleness and generality of the linear model for the photometric space
-% - address the non-convexity of the model (see twitter convo with @andy_l_jones). Plus optimization issues.
 % - Coryn's point that we should carefully explain how our model is geometric.
@@
 
+This optimization is not convex---there are multiple optima in general.
+In particular, there is a large, degenerate, pathological optimum where
+the exponential function underflows, all parallaxes are predicted to vanish,
+and there is no gradient of anything with respect to the parameter $D$-vector $\theta$.
+This large, bad optimum (it is a local, not global, optimum) must be avoided in optimization.
+In practice we avoid it by optimizing first for the very highest signal-to-noise
+(in astrometric parallax) Training Set stars, and then use that first optimum as a
+good first guess or initialization for the full optimization.
+This method works because in the limit of high signal-to-noise, the problem asymptotically approaches
+the convex problem of L1-regularized linear least-square fitting.
+Optimization is performed with \code{scipy.optimize} \acronym{L-BFGS-B}.
+
 Once the model is optimized, the output of the model is a prediction of the parallax,
@@
 We optimize two models, one for Training Set A and one for Training Set B.
-For optimization, we use \code{scipy.optimize} \acronym{L-BFGS-B}, starting at
-an uninformed starting point in the $\theta$ space.
-The optimization is of the objective function given in \equationname~(\ref{eq:objective}) above,
-and although this optimization is not convex, it appears to converge to the
-same good solution quickly and reliably from multiple different starting points.
-
 The two models are applied to the two splits of the Parent Sample, Parent Sample

### 2018-08-03 paper/sp.tex
figure

@@
 train-and-test framework constitutes a two-fold cross-validation.
-Results of this cross-validation is shown in HOGG WHAT FIGURE.
+Results of this cross-validation is shown in \figurename~\ref{fig:xval}.
 For the stars with highest \gaia-measured parallax signal-to-noise,
@@
 and a scatter of better than 9~percent.
+\begin{figure}
+\caption{The predictive accuracy of the model, in the two-fold cross-validation.
+  Each panel shows \gaia\ astrometric parallaxes $\gparallax_n$ on the vertical axis
+  and our spectrophotometric parallaxes $\sparallax$ on the horizontal axis.
+  The left panel shows the full Parent Sample, the middle panel shows the Training Set
+  stars (which pass \gaia\ quality checks; see text), and the right panel shows a
+  subset of stars with very high astrometric parallax signal-to-noise. This latter
+  set plays no particular role in the method, but it can be used to demonstrate or
+  assess the prediction precision. Fractional precision of the prediction
+  is XXXX HOGG MEASURED HOW?\label{fig:xval}}
+\end{figure}
 

### 2018-08-03 paper/sp.tex
figures

@@
 regions or at lower metallicities.
-\figurename~HOGG WHAT FIGURE shows that
+\figurename~\ref{fig:residuals} shows that
 there is no suggestion of any such trends when we color the residuals by metallicity
 or $H-W_2$ color (which is a reddening proxy).
+\begin{figure}
+\caption{Prediction residuals colored by relevant features.
+  Each panel shows the prediction residuals (\gaia\ astrometric parallax minus
+  spectrophotometric parallax $\gparallax_n-\sparallax_n$) colored by a different
+  data feature that might be relevant to the residual.\label{fig:residuals}}
+\end{figure}
 
@@
 clusters or spatially compact objects in the data.
-In HOGG WHAT FIGURE, we show the parallax distribution from \gaia\ and
+In \figurename~\ref{fig:clusters}, we show the parallax distribution from \gaia\ and
 the spectrophotometric-parallax distribution from this work for stars in
@@
 HOGG: DO WE NEED TO ADDRESS whether why not these clusters show 9-percent spreads?
+\begin{figure}
+\caption{The histograms of \gaia\ astrometric parallaxes $\gparallax_n$
+  and spectrophotometric parallaxes $\sparallax_n$ for stars in angular proximity
+  to various stellar clusters in the \apogee\ observing footprint. In each panel,
+  the expected parallax of the cluster is shown as a vertical line.\label{fig:clusters}}
+\end{figure}
 
@@
 HOGG: Similarly, the spectral part of the model should also be
-interpretable. HOGG SHOW FIGURE THAT DEMONSTRATES THIS, qualitatively.
+interpretable. HOGG DESCRIBE \figurename~\ref{fig:spectral} THAT
+DEMONSTRATES THIS, qualitatively.
+\begin{figure}
+\caption{The coefficients of the spectral part of the linear model.\label{fig:spectral}}
+\end{figure}
 
@@
 HOGG: Show a map of the kinematic rotation of the Galaxy as a function of
-x and y, colored by metallicity.
+x and y, colored by metallicity in \figurename~\ref{fig:disk}.
 We see the kinematic center of the Galaxy! We see the rotation curve.
 We see the chemical gradients.
+\begin{figure}
+\caption{A map of the kinematics and element abundances in the Milky Way disk.
+  This map shows HOGG SOMETHING.\label{fig:disk}}
+\end{figure}
 

### 2018-08-03 infer_parallax.py
new

@@ -80,7 +80,7 @@ def check_H_func(x, y, A, lams, ivar):
 # -------------------------------------------------------------------------------
 
 Kfold = 2
-lam = 10                      # hyperparameter -- needs to be tuned!
+lam = 2n0                      # hyperparameter -- needs to be tuned!
 name = 'N{0}_lam{1}_K{2}_parallax'.format(len(labels), lam, Kfold)
 
 # optimization schedule

### 2018-08-03 rotation_curve_uncertainties.py

@@ -236,7 +236,11 @@ Xlimits = [[-30, 10], [-10, 30], [-20, 20],
 
 # -------------------------------------------------------------------------------
 # divide Milky Way into (x, y, z) patches
-# -------------------------------------------------------------------------------   
+# ------------------------------------------------------------------------------- 
+
+# cut in |b| instead of |z|! --> wedge in z! (not quite right?!)
+cut_b = abs(labels['b']) < 6.
+name = name + '_cutb6'
 
 box_size = .5               # that's just half of the box size
 all_x = np.arange(-30., 30.01, box_size)
@@ -254,7 +258,8 @@ mean_sigma_par = np.zeros((len(all_x), len(all_y))) - np.inf
 
 for i, box_center_x in enumerate(all_x):
     for j, box_center_y in enumerate(all_y):
-        cut_patch = (abs(mean_XS_cart_n[:, 2]) < box_size) * (abs(mean_XS_cart_n[:, 0] - box_center_x) < box_size) * (abs(mean_XS_cart_n[:, 1] - box_center_y) < box_size)
+        #cut_patch = (abs(mean_XS_cart_n[:, 2]) < box_size) * (abs(mean_XS_cart_n[:, 0] - box_center_x) < box_size) * (abs(mean_XS_cart_n[:, 1] - box_center_y) < box_size)
+        cut_patch = cut_b * (abs(mean_XS_cart_n[:, 0] - box_center_x) < box_size) * (abs(mean_XS_cart_n[:, 1] - box_center_y) < box_size)
         N_stars[i, j] = np.sum(cut_patch)
         if N_stars[i, j] > 0:
             mean_XS_cyl[i, j, :] = np.nanmean(mean_XS_cyl_n[cut_patch], axis = 0) # NEVER USE MEAN PHI -- DOESN'T MAKE SENSE!
@@ -279,27 +284,11 @@ deg_wedge = 30.
 wedge = np.abs(XS_cyl_true_n[:, 1]) < (deg_wedge/360. * 2. * np.pi)
 cut_z = abs(mean_XS_cyl_n[:, 2]) < 0.5
 
[…]

### 2018-08-05 paper/gc.tex
nothin

--- 2018-07-26 paper/gc.tex
@@
 
-Our position is that a methodological technique is correct inasmuch as
-it delivers correct or best results \emph{under its particular assumptions}.
-In that spirit, we present here the assumptions of the method
-explicitly.
-We will return to these assumptions in the Discussion section below,
-to address the costs and benefits of each in more depth.
-\begin{description}
-\item[features] Wassuh!
-
-\end{description}
-
 \section{Method}

### 2018-08-05 paper/sp.tex

@@
 
+\software{The code used in this project is available from
+\url{https://github.com/aceilers/spectroscopic_parallax}.
+It is released under the MIT (HOGG CHECK) open-source software license.
+This research additionally utilized:
+\code{Astropy} (\citealt{astropy, astropy2}),
+\code{IPython} (\citealt{ipython}),
+\code{matplotlib} (\citealt{matplotlib}),
+\code{numpy} (\citealt{numpy}), and
+\code{scipy} (\citealt{scipy})}
+
+\facility{\sdssiv, \apogee, \apogee-2, \gaia, \wise}
+
 \begin{thebibliography}{}
 \bibitem[Abolfathi \etal(2018)]{dr14} Abolfathi, B., Aguado, D.~S., Aguilar, G., \etal\ 2018, \apjs, 235, 42 
+\bibitem[Astropy Collaboration \etal(2013)]{astropy} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., \etal\ 2013, \aap, 558, A33 
+\bibitem[Astropy Collaboration \etal(2018)]{astropy2} The Astropy Collaboration, Price-Whelan, A.~M., Sip{\H o}cz, B.~M., \etal\ 2018, arXiv:1801.02634 
 \bibitem[Bailer-Jones et al.(2018)]{calj} Bailer-Jones, C.~A.~L., Rybizki, J., Fouesneau, M., Mantelet, G., \& Andrae, R.\ 2018, \aj, 156, 58 
 \bibitem[Blanton \etal(2017)]{sdssiv} Blanton, M.~R., Bershady, M.~A., Abolfathi, B., \etal\ 2017, \aj, 154, 28 
-\bibitem[Hogg(2018)]{gaialf} Hogg, D.~W 2018, arXiv:1804.07766 
+\bibitem[Hogg(2018)]{gaialf} Hogg, D.~W.\ 2018, arXiv:1804.07766
+\bibitem[Hunter(2007)]{matplotlib} Hunter, J.~D.\ 2007, CISE, 9(3), 90
+\bibitem[Jones \etal(2001)]{scipy} Jones, E., Oliphant, E., Peterson, P. \etal\ 2001 \url{http://www.scipy.org/}
 \bibitem[Kollmeier \etal(2017)]{sdssv} Kollmeier, J.~A., Zasowski, G., Rix, H.-W., \etal\ 2017, arXiv:1711.03234 
@@
 \bibitem[Lutz \& Kelker(1973)]{lk} Lutz, T.~E., \& Kelker, D.~H.\ 1973, \pasp, 85, 573 
+\bibitem[Oliphant(2006)]{numpy} Oliphant, T.~E.\ 2006 \emph{A guide to NumPy}, Trelgol Publishing
+\bibitem[P\'erez \& Granger(2007)]{ipython} P\'erez, F. \& Granger, B.~E.\ 2007, CISE, 9(3), 29
 \bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., \etal\ 2006, \aj, 131, 1163 

### 2018-08-05 paper/sp.tex
software and ref tweaks

@@
 
-\software{The code used in this project is available from
+\software{} % This command is wack
+The code used in this project is available from
 \url{https://github.com/aceilers/spectroscopic_parallax}.
@@
 \code{numpy} (\citealt{numpy}), and
-\code{scipy} (\citealt{scipy})}
+\code{scipy} (\citealt{scipy}).
 
-\facility{\sdssiv, \apogee, \apogee-2, \gaia, \wise}
+\facilities{\sdssiv, \apogee, \apogee-2, \gaia, \zmass, \wise}
 
@@
 \bibitem[Hunter(2007)]{matplotlib} Hunter, J.~D.\ 2007, CISE, 9(3), 90
-\bibitem[Jones \etal(2001)]{scipy} Jones, E., Oliphant, E., Peterson, P. \etal\ 2001 \url{http://www.scipy.org/}
+\bibitem[Jones \etal(2001)]{scipy} Jones, E., Oliphant, E., Peterson, P. \etal\ 2001, \url{http://www.scipy.org/}
 \bibitem[Kollmeier \etal(2017)]{sdssv} Kollmeier, J.~A., Zasowski, G., Rix, H.-W., \etal\ 2017, arXiv:1711.03234 
@@
 \bibitem[Lutz \& Kelker(1973)]{lk} Lutz, T.~E., \& Kelker, D.~H.\ 1973, \pasp, 85, 573 
-\bibitem[Oliphant(2006)]{numpy} Oliphant, T.~E.\ 2006 \emph{A guide to NumPy}, Trelgol Publishing
+\bibitem[Oliphant(2006)]{numpy} Oliphant, T.~E.\ 2006, \emph{A guide to NumPy}, Trelgol Publishing
 \bibitem[P\'erez \& Granger(2007)]{ipython} P\'erez, F. \& Granger, B.~E.\ 2007, CISE, 9(3), 29

### 2018-08-05 paper/sp.tex
tabular output and how to use it

@@
 the linear model and check whether the dependencies represented there make sense.
-One nice thing about the model is that the feature vector $x_n$ for star $n$
+The feature vector $x_n$ for star $n$
 contains a set of photometry in magnitudes.
@@
 
-HOGG: How to get the data and how to use it, probabilistically.
+In the end, this method produces a spectrophotometric parallax estimate $\sparallax_m$ for
+every star $m$.
+\tablename~\ref{tab:output} shows a snippet of this output, with the full output available
+as an electronic data table associated with this \documentname.
+In order to use this output with \gaia\ astrometry, we recommend joining the table to the
+\gaia\ Catalog on the \zmass\ ID (name), as per the instructions at the \Gaia\ Archive (CITE).
+
+In order to use these outputs in a probabilistic inference or forward model,
+an investigator needs a noise (uncertainty) model.
+The linear model permits straightforward propagation of uncertainty from the
+feature inputs.
+However, these propagated uncertainties clearly under-predict the scatter between
+the spectrophotometric parallaxes $\sparallax_m$ and the \gaia\ astrometric
+parallaxes $\gparallax_m$.
+Our recommendation for an uncertainty model is HOGG WAT?
 
@@
 
-\software{} % This command is wack
-The code used in this project is available from
-\url{https://github.com/aceilers/spectroscopic_parallax}.
-It is released under the MIT (HOGG CHECK) open-source software license.
-This research additionally utilized:
-\code{Astropy} (\citealt{astropy, astropy2}),
-\code{IPython} (\citealt{ipython}),
-\code{matplotlib} (\citealt{matplotlib}),
-\code{numpy} (\citealt{numpy}), and
-\code{scipy} (\citealt{scipy}).
+\facilities{
+\sdssiv(\apogee-2),
+\gaia,
+\zmass,
+\wise}
 
-\facilities{\sdssiv, \apogee, \apogee-2, \gaia, \zmass, \wise}
+\software{
+\code{Astropy} \citep{astropy, astropy2},
+\code{IPython} \citep{ipython},
+\code{matplotlib} \citep{matplotlib},
+\code{numpy} \citep{numpy},
+\code{scipy} \citep{scipy}}
 

### 2018-08-05 paper/sp.tex
notes to self
bug fixes

@@
 % - does the model need a cute name, like The Swan?
-% - we need a Facilities list / tags
-% - we need a Software list / tags
 % - check with WISE experts that we have the correct WISE photometry
+% - we need an uncertainty estimate or noise model for our outputs.
 
@@
 % text macros
+\newcommand{\documentname}{\textsl{Article}}
 \newcommand{\equationname}{equation}
@@
 In order to use this output with \gaia\ astrometry, we recommend joining the table to the
-\gaia\ Catalog on the \zmass\ ID (name), as per the instructions at the \Gaia\ Archive (CITE).
+\gaia\ Catalog on the \zmass\ ID (name), as per the instructions at the \gaia\ Archive (CITE).
+In order to use this output with \apogee\ radial velocities and element abundances, we recommend
+joining the table to the \sdssiv\ Catalog on the \zmass\ ID (name).
+\begin{table}
+\caption{The generated spectrophotometric parallaxes.\label{tab:output}}
+\end{table}
 

### 2018-08-06 paper/sp.tex
linearity assumption

@@
 
-In that context: This model is only linear! We could have done much better with
-a more flexible model. But choices, execution.
-
 In that context: These features are bad! We could have used feature engineering
 or an autoencoder to do a much better job. Definitely should do this in the future.
+
+As we said above, our most restrictive assumption is the linearity assumption.
+We required that the parallax prediction be constructed from a linear combination
+of the input data or feature vector components.
+This is absurd: Stars are not linear objects!
+There are two questions to ask about the linearity.
+One is: Why did it work? The other is: Could we have done better?
+The answer to the first is that we have taken such a small part of stellar parameter
+space with our surface gravity cut, that the linearized model of stellar
+spectrophotometry is not a bad approximation.
+Also, because our linear model included photometry in magnitudes (that is, logarithmic)
+and was exponentiated to make the parallax prediction, we knew in advance that much
+of the needed flexibility (for dust attenuation, for example) would be close to
+linear.
+That is, it was a combination of limited model scope and some cleverness in the
+feature engineering and model structure.
+
+The answer to the second question is that absolutely we could have done better!
+Since models like Gaussian processes and deep networks subsume the linear model
+(at least approximately), they could have delivered similar results.
+The issues with going to a much more complex model are manifold, however:
+We would have had many more decisions to make and many more hyper-parameters
+to set.
+We would have used some (or maybe much) of the information in the data to learn
+the simple fact that the problem is close to linear; that is, if you give a model
+a lot of freedom or capacity, it has to use a lot of the data to learn what part
+of that capacity it really needs.
+A more complex model might have had edge issues: Very flexible models don't extrapolate
+outside the convex hull of their training data, and our training data was probably not
+balanced in terms of parallax and dust attenuation.
+And finally, the complex model would have been harder to visualize and interpret;
+the linear model contains essentially only derivatives in the data space, which are
+relatively easy to understand.
 

### 2018-08-06 infer_parallax.py
plots

@@ -29,7 +29,7 @@ fsize = 14
 # -------------------------------------------------------------------------------
 
 # make plots?
-prediction = True
+prediction = False
 
 print('loading labels...')
 hdu = fits.open('data/training_labels_parent.fits')
@@ -80,7 +80,7 @@ def check_H_func(x, y, A, lams, ivar):
 # -------------------------------------------------------------------------------
 
 Kfold = 2
-lam = 2n0                      # hyperparameter -- needs to be tuned!
+lam = 70                      # hyperparameter -- needs to be tuned!
 name = 'N{0}_lam{1}_K{2}_parallax'.format(len(labels), lam, Kfold)
 
 # optimization schedule
@@ -107,6 +107,7 @@ if prediction:
        
     # split into training and validation set
     y_pred_all = np.zeros_like(y_all)
+    y_pred_all_err = np.zeros_like(y_all)
         
     for k in range(Kfold):    
         
@@ -163,7 +164,10 @@ if prediction:
                                
         # prediction
         y_pred = np.exp(np.dot(A_all[valid, :], x_new))
+#        foo = A_all[valid, :] * x_new_err
+#        y_pred_err = y_pred * np.sqrt(np.dot(foo, foo))
         y_pred_all[valid] = y_pred
+#        y_pred_all_err[valid] = y_pred_err
         
         plt.scatter(labels[valid]['parallax'], y_pred, alpha = .1)
         plt.xlim(-1, 3)
@@ -174,7 +178,9 @@ if prediction:
         f.close()   
     
     spec_parallax = y_pred_all
+    spec_parallax_err = y_pred_all_err
     labels.add_column(spec_parallax, name='spec_parallax')
+    labels.add_column(spec_parallax_err, name='spec_parallax_err')
     Table.write(labels, 'data/training_labels_new_{}.fits'.format(name), format = 'fits', overwrite = True)
     
 # -------------------------------------------------------------------------------
@@ -191,16 +197,15 @@ if not prediction:
     # visibility periods used
     cut_vis = labels['visibility_periods_used'] >= 8    
     # cut in parallax_error
-    cut_par = labels['parallax_error'] < 0.1           
-    # cut in b 
-    bcut = 0
-    cut_b = np.abs(labels['b']) >= bcut  
+    cut_par = labels['parallax_error'] < 0.1            
     # cut in astrometric_gof_al
-    cut_gof = labels['astrometric_gof_al'] < 5 
[…]

### 2018-08-06 make_figures_HER.py

@@ -1,0 +1,202 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Aug  6 13:13:17 2018
+
+@author: eilers
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import scipy.optimize as op
+import time
+import pickle
+from astropy.table import Column, Table, join, vstack, hstack
+import sys
+from astropy.io import fits
+from sklearn.decomposition import PCA
+from astropy import units as u
+from astropy.coordinates import SkyCoord
+import astropy.coordinates as coord
+from mpl_toolkits.mplot3d import Axes3D
+import corner
+
+
+# -------------------------------------------------------------------------------
+# plotting settings
+# -------------------------------------------------------------------------------
+
+matplotlib.rcParams['ytick.labelsize'] = 14
+matplotlib.rcParams['xtick.labelsize'] = 14
+matplotlib.rc('text', usetex=True)
+fsize = 14
+figsize = (8, 4.5)
+
+
+# -------------------------------------------------------------------------------
+# open inferred labels
+# -------------------------------------------------------------------------------
+
+N = 45787
+Kfold = 2
+lam = 30
+name = 'N{0}_lam{1}_K{2}_parallax'.format(N, lam, Kfold)
+
+print('loading new labels...')   
+labels = Table.read('data/training_labels_new_{}.fits'.format(name), format = 'fits')    
+labels.rename_column('ra_1', 'ra')
+labels.rename_column('dec_1', 'dec')
+
+# -------------------------------------------------------------------------------
+# Figure 1
+# -------------------------------------------------------------------------------
+
+fig, ax = plt.subplots(1, 2, figsize = figsize)
+ax[0].scatter(labels['bp_rp'], labels['J']-labels['K'], rasterized = True, s = 10, alpha = .1)
+ax[1].scatter(labels['bp_rp'], labels['H']-labels['w2mpro'], rasterized = True, s = 10, alpha = .1)
+ax[0].set_xlabel(r'$\rm B_P-R_p$', fontsize = fsize)
+ax[1].set_xlabel(r'$\rm B_P-R_p$', fontsize = fsize)
[…]

### 2018-08-06 test_distances.py

@@ -37,11 +37,11 @@ fsize = 14
 
 N = 45787
 Kfold = 2
-lam = 100
-name = 'N{0}_lam{1}_K{2}_mag_offset'.format(N, lam, Kfold)
+lam = 30
+name = 'N{0}_lam{1}_K{2}_parallax'.format(N, lam, Kfold)
 
 print('loading new labels...')   
-labels = Table.read('data/training_labels_new_{}_2.fits'.format(name), format = 'fits')    
+labels = Table.read('data/training_labels_new_{}.fits'.format(name), format = 'fits')    
 labels.rename_column('ra_1', 'ra')
 labels.rename_column('dec_1', 'dec')
 
@@ -59,23 +59,22 @@ cs = coord.ICRS(ra = labels['ra'] * u.degree,
                 pm_dec = labels['pmdec'] * u.mas/u.yr, 
                 radial_velocity = labels['VHELIO_AVG'] *u.km/u.s)
 
+# -------------------------------------------------------------------------------
+# position of the sun
+# -------------------------------------------------------------------------------           
 
-#Galactocentric position of the Sun:
-X_GC_sun_kpc = 8.    #[kpc]
+# Galactocentric position of the Sun:
+X_GC_sun_kpc = 8.122 #[kpc] # (Gravity collaboration 2018)
 Z_GC_sun_kpc = 0.025 #[kpc] (e.g. Juric et al. 2008)
 
-#circular velocity of the Galactic potential at the radius of the Sun:
-vcirc_kms = 220. #[km/s] (e.g. Bovy 2015)
+# Galactocentric velocity of the Sun:
+vX_GC_sun_kms = -11.1 # [km/s]   (e.g. Schoenrich et al. 2009) 
+vY_GC_sun_kms =  245.8 # [km/s]  (combined with Sgr A* proper motions from Reid & Brunnthaler 2004)
+vZ_GC_sun_kms =  7.8 # [km/s]
 
-#Velocity of the Sun w.r.t. the Local Standard of Rest (e.g. Schoenrich et al. 2009):
-U_LSR_kms = 11.1  # [km/s]
-V_LSR_kms = 12.24 # [km/s]
-W_LSR_kms = 7.25  # [km/s]
-
-#Galactocentric velocity of the Sun:
-vX_GC_sun_kms = -U_LSR_kms           # = -U              [km/s]
-vY_GC_sun_kms =  V_LSR_kms+vcirc_kms # = V+v_circ(R_Sun) [km/s]
[…]

### 2018-08-07 paper/sp.tex
notes2self
fig caption tweak
figures in
apogee logg

@@
 % - we need an uncertainty estimate or noise model for our outputs.
+% - we have set lambda = 30 by x-val.
 
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{./training_sample.pdf}
 \caption{A visualization of the Training Set stars. The left panel shows
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{residuals.pdf}
 \caption{The predictive accuracy of the model, in the two-fold cross-validation.
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{./residuals_training.pdf}
 \caption{Prediction residuals colored by relevant features.
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{clusters.pdf}
 \caption{The histograms of \gaia\ astrometric parallaxes $\gparallax_n$
@@
 \begin{figure}
-\caption{The coefficients of the spectral part of the linear model.\label{fig:spectral}}
+\includegraphics[width=\textwidth]{coefficients.pdf}
+\caption{The coefficients of the spectral part of the linear model.
+  Also shown are derivatives with respect to labels for a \project{Cannon}-like
+  linear regression for the Training Set data.\label{fig:spectral}}
 \end{figure}
@@
 
-(In addition: what did we assume about the \apogee\ data: Good $\logg$ values.
-Okay pseudo-continuum normalization!)
+Although the model, training, and prediction make no use whatsoever about stellar
+models, we did use stellar models indirectly to select the Parent Sample of stars:
+We used the \apogee\ pipeline surface-gravity $\logg$ measurements.
+This tiny use of stellar models could have been obviated by selecting stars not
+on the basis of the derived physical parameter, but instead selecting stars to
+be similar, observationally, in the spectral space.
+That would have lead to a clean sample and made the project independent completely
+of stellar models.
 

### 2018-08-07 rotation_curve_uncertainties.py
figures

@@ -208,7 +208,7 @@ def overplot_ring(r):
     thetas = np.arange(0., 2*np.pi + tiny, 0.001 * np.pi)
     xs = r * np.cos(thetas)
     ys = r * np.sin(thetas)
-    plt.plot(xs, ys, "k-", alpha=0.2, lw=1, zorder = -np.inf)
+    plt.plot(xs, ys, "k-", alpha=0.2, lw=1, zorder = np.inf)
     plt.scatter(0, 0, s = 10, color = 'k', alpha=0.2)
     return
 
@@ -217,7 +217,7 @@ def overplot_ring_helio(r):
     thetas = np.arange(0., 2*np.pi + tiny, 0.001 * np.pi)
     xs = r * np.cos(thetas) - X_GC_sun_kpc
     ys = r * np.sin(thetas)
-    plt.plot(xs, ys, "k", alpha=0.2, lw=.8, linestyle = ':', zorder = np.inf)
+    plt.plot(xs, ys, "k:", alpha=0.2, lw=1, zorder = -np.inf)
     # plt.scatter(-X_GC_sun_kpc, 0, s = 10, color = 'k', alpha=0.2)
     return
 
@@ -231,16 +231,17 @@ def overplot_rings_helio():
         overplot_ring_helio(r)
     return
 
-Xlimits = [[-30, 10], [-10, 30], [-20, 20], 
+Xlimits = [[-25, 10], [-12, 23], [-20, 20], 
            [-200, 200], [-200, 200], [-200, 200]]
 
 # -------------------------------------------------------------------------------
 # divide Milky Way into (x, y, z) patches
 # ------------------------------------------------------------------------------- 
 
-# cut in |b| instead of |z|! --> wedge in z! (not quite right?!)
-cut_b = abs(labels['b']) < 6.
-name = name + '_cutb6'
+# take wegde in z 
+deg_wedge_in_z = 6.
+wedge_z = (np.abs(XS_cyl_true_n[:, 2])/(XS_cyl_true_n[:, 0])) < np.tan(deg_wedge_in_z/360. * 2. * np.pi)
+cut_z = np.logical_or(abs(mean_XS_cyl_n[:, 2]) < 0.5, wedge_z)
 
[…]

### 2018-08-07 test_distances.py

@@ -322,7 +322,7 @@ for i, clus in enumerate(cluster_list):
     cut_clus = (abs(labels['ra'] - ra_clus) < .8) \
           * (abs(labels['dec'] - dec_clus) < .8) 
           
-    bins = np.linspace(-0.1, 1, 20)
+    bins = np.linspace(-0.1, .8, 20)
     ax[c, r].hist(labels['spec_parallax'][cut_clus], normed = True, bins = bins, histtype = 'step', lw = 3, color = 'k', label = r'$\varpi^{(sp)}$')
     ax[c, r].hist(labels['parallax'][cut_clus], normed = True, bins = bins, histtype = 'step', lw = 1, color = 'k', label = r'$\varpi^{(a)}$')
     ax[c, r].tick_params(axis=u'both', direction='in', which='both', right = 'on', top = 'on')
@@ -331,7 +331,7 @@ for i, clus in enumerate(cluster_list):
     ax[c, r].set_title(r'{0}, $\rm [Fe/H] = {1}$'.format(t['lat_name'][k][0], t['FE_H'][k][0]), fontsize = 13)
     ax[c, r].set_xlabel(r'$\varpi$ [mas]', fontsize = 14)
     ax[c, r].set_ylabel('normalized counts', fontsize = 14)
-    ax[c, r].set_xlim(-0.1, 1.)
+    ax[c, r].set_xlim(-0.1, .8)
     if r == 2: 
         c += 1
         r = 0    
@@ -342,7 +342,7 @@ ax[0, 0].legend(frameon = True)
 ax[3, 2].hist(labels_sgr['spec_parallax'], normed = True, bins = bins, histtype = 'step', lw=3, color='k', label = r'$\varpi^{(sp)}$')
 ax[3, 2].hist(labels_sgr['parallax'], normed = True, bins = bins, histtype = 'step', lw=1, color='k', label = r'$\varpi^{(a)}$')
 ax[3, 2].set_xlabel(r'$\varpi$ [mas]', fontsize = 14)
-ax[3, 2].set_xlim(-0.1, 1.)
+ax[3, 2].set_xlim(-0.1, .8)
 ax[3, 2].axvline(1./20, linestyle = '--', color = '#929591')
 ax[3, 2].set_title('Sagittarius')
 ax[3, 2].tick_params(axis=u'both', direction='in', which='both', right = 'on', top = 'on')

### 2018-08-07 make_figures_HER.py
figures

@@ -67,7 +67,7 @@ cbar_ax = fig.add_axes([1, 0.15, 0.03, 0.82])
 cb = fig.colorbar(sc, cax=cbar_ax)
 cb.set_label(r'$\log g$', fontsize = fsize)
 plt.tight_layout()
-plt.savefig('paper/parent_sample.pdf')
+plt.savefig('paper/parent_sample.pdf', pad_inches=.2, bbox_inches = 'tight')
 
 # -------------------------------------------------------------------------------
 # Figure 2
@@ -80,17 +80,20 @@ train = cut_vis * cut_par * cut_cal
 best = train * (labels['parallax_over_error'] >= 20)
 
 fig, ax = plt.subplots(1, 2, figsize = figsize)
-ax[0].scatter(labels[train]['bp_rp'], labels[train]['J']-labels[train]['K'], rasterized = True, s = 10, alpha = .1)
-ax[1].scatter(labels[train]['bp_rp'], labels[train]['H']-labels[train]['w2mpro'], rasterized = True, s = 10, alpha = .1)
+ax[0].scatter(labels[train]['bp_rp'], labels[train]['J']-labels[train]['K'], c = labels[train]['LOGG'], cmap = cm, vmin = 0, vmax = 2.2, rasterized = True, s = 10, alpha = .5)
+ax[1].scatter(labels[train]['bp_rp'], labels[train]['H']-labels[train]['w2mpro'], c = labels[train]['LOGG'], cmap = cm, vmin = 0, vmax = 2.2, rasterized = True, s = 10, alpha = .5)
 ax[0].set_xlabel(r'$\rm B_P-R_p$', fontsize = fsize)
 ax[1].set_xlabel(r'$\rm B_P-R_p$', fontsize = fsize)
 ax[0].set_ylabel(r'$\rm J-K$', fontsize = fsize)
 ax[1].set_ylabel(r'$\rm H-W_2$', fontsize = fsize)
 ax[0].tick_params(axis=u'both', direction='in', which='both', right = 'on', top = 'on')
 ax[1].tick_params(axis=u'both', direction='in', which='both', right = 'on', top = 'on')
-#plt.subplots_adjust(wspace = .1)
+fig.subplots_adjust(right = 0.8)
+cbar_ax = fig.add_axes([1, 0.15, 0.03, 0.82])
+cb = fig.colorbar(sc, cax=cbar_ax)
+cb.set_label(r'$\log g$', fontsize = fsize)
[…]

### 2018-08-07 paper/sp.tex
figure
features are limited

@@
 As we operate these spectroscopic surveys,
-the \gaia\ Mission has been revolutionizing our view of the disk.
+the \gaia\ Mission (CITE) has been revolutionizing our view of the disk.
 It delivers good proper motions everywhere,
@@
 plus the $G$ and $(\BP-\RP)$ photometry from \gaia,
-plus the $J$, $H$, and $K_s$ photometry from \zmass (\citealt{zmass}),
+plus the $J$, $H$, and $K_s$ photometry from \zmass\ (\citealt{zmass}),
 plus the $W_1$ and $W_2$ photometry from \wise.
@@
 
-Because the \apogee\ target selection is based on \zmass (\citealt{zmass}),
+Because the \apogee\ target selection is based on \zmass\ (\citealt{zmass}),
 every \apogee\ star also comes with \zmass\ photometry.
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{parent_sample.pdf}
 \caption{A visualization of the Parent Sample stars. The left panel
@@
 
-In that context: These features are bad! We could have used feature engineering
-or an autoencoder to do a much better job. Definitely should do this in the future.
+One strong assumption of this work is that our set of features is sufficient to
+build a predictive model for distance.
+Our arguments for our included features is that they ought to be sufficient to
+estimate a dust-independent apparent magnitude and the stellar parameters (especially
+age and surface gravity) that are sensitive to luminosity.
+The derived model parameters shown in \figurename~\ref{fig:spectral} do indicate that
+the model is capturing some or all of this information.
+However, our choice of features is very rigid:
+In principle the data themselves should tell us what features to include.
+In that direction, it would make sense to choose features using a more complex
+technique like deep learning or an auto-encoder (CITE).
+These methods generate good features automatically.
+However, they also involve an enormous set of hyper-parameter-like choices,
+and they involve sacrificing certain kinds of interpretability.
+However, we expect that a better set of features would do better; in this sense
+the project here is just a first step towards precise spectrophotometric distance
+estimates.
 
@@
 of stellar models.
+
+Complain about our lack of a noise model.
 

### 2018-08-07 paper/sp.tex

@@
 % - we need an uncertainty estimate or noise model for our outputs.
-% - we have set lambda = 30 by x-val.
+% - we have set lambda = 50 by x-val. We could go as high as 60 without hurting our results.
+% - We probably should cut in color-color space. Probably in J-K vs B-R and in H-W2.
+% - We will match axes on relevant plots.
+% - ...
 
@@
 \newcommand{\documentname}{\textsl{Article}}
+\newcommand{\sectionname}{Section}
 \newcommand{\equationname}{equation}
@@
 
-\section{Assumptions of the method}
+\section{Assumptions of the method}\label{sec:assumptions}
 
@@
 for building a model of the stars that is both justifiable probabilistically
-and consistent with the assumptions stated in the previous Section.
+and consistent with the assumptions stated in the previous \sectionname.
 Here we lay out the model and methodology.
-We apply the method to real data in the following Sections.
+We apply the method to real data in the following {\sectionname s}.
 
@@
 
-What did we do and what's cool?
+This project demonstrates that photometry and spectroscopy can be used to predict
+or estimate distances to stars.
+This isn't new!
+However, what's new is that it is possible to do this very well but with a purely
+linear model acting on magnitudes and spectral (logarithmic) fluxes, and entirely
+data-driven.
+That is, a linear model trained on noisy parallax data from \gaia.
 
-Why did we make the assumptions we made and what did they gain us and cost us?
+The model and method is based on a raft of assumptions, listed above
+in \sectionname~\ref{sec:assumptions}.
+We make some comments there, but we emphasize some particulars here:
 
@@
 age and surface gravity) that are sensitive to luminosity.
-The derived model parameters shown in \figurename~\ref{fig:spectral} do indicate that
-the model is capturing some or all of this information.
+In particular, since our model is linear in the logarithm of the parallax, it is
+also linear in any kind of apparent magnitude at fixed distance and also linear
+in distance modulus at fixed luminosity.
+Our feature decisions were highly motivated by these kinds of considerations.
+And the derived model parameters shown in \figurename~\ref{fig:spectral} do indicate that
+the model is capturing some or all of the expected information in the spectra, at least.
+
 However, our choice of features is very rigid:
@@
 
-Complain about our lack of a noise model.
+One significant limitation of the model is that it does not produce a sensible or
+useful noise estimate or error bar on each spectrophotometric parallax estimate.
+Although the linear model does permit simple noise propagation, that noise propagation
+seriously under-estimates the scatter between the spectrophotometric parallaxes
+and the best of the \gaia\ \acronym{DR2} astrometric parallaxes.
+We thus used an empirical noise estimate to get our HOGG DETAIL HERE.
+In principle it might have been better to build a noise model for our outputs,
+and learned that noise model along with the coefficients.
 
-(Mention independence of every result from the \gaia\ astrometry.)
+Because of our train-and-test framework, each star is given a spectrophotometric
+parallax based on coefficients derived from training on a complementary Training Set sample, of
+which that star is not a member.
+This framework ensures that the \gaia\ parallax $\gparallax_m$ of star $m$ is never used,
+even indirectly, in generating the spectrophotometric parallax $\sparallax_m$ of star $m$.
+That means that the parallax estimates generated here and given in \tablename~\ref{tab:output}
+is statistically independent of the \gaia\ parallaxes and can be combined with them in
+na\"ive ways, for example in inverse-variance-weighted averages.
 
@@
 We see the kinematic center of the Galaxy! We see the rotation curve.
-We see the chemical gradients.
+We see the chemical gradients. HOGG WHY CAN YOU NOT JUST READ OFF THINGS FROM THIS PLOT?
 \begin{figure}
 \caption{A map of the kinematics and element abundances in the Milky Way disk.
-  This map shows HOGG SOMETHING.\label{fig:disk}}
+  This map shows the rotation of the Galaxy out to large radius, and also
+  indicates the kinematic center of the disk rotation. Interpretation of
+  the quantitative information in this plot requires care (see text).\label{fig:disk}}
 \end{figure}
 
-How do you build a proper probabilistic model using these? How is the answer to that
-related to the answer for \gaia\ itself?
+\figurename~\ref{fig:disk} shows that this sample of spectrophotometric parallaxes is useful for
+mapping the Milky Way.
+How does an investigator use these spectrophotometric parallaxes wisely?
+How do you build a proper probabilistic model using these?
+The answer is similar to the answer to the same question for the \gaia\ \acronym{DR2} data
+themselves (\citealt{gaialf}): A likelihood function must be constructed for each of the
+spectroscopic parallaxes, given assumptions about the noise in those values.
+The sensible likelihood function for star $m$ would be HOGG PUT HERE.
+Model fitting can then be built on those likelihood functions.
 

### 2018-08-08 infer_parallax_rrl.py
new figures

@@ -15,6 +15,7 @@ from astropy.table import Column, Table, join, vstack, hstack
 from astropy.io import fits
 from scipy.stats import binned_statistic_2d
 import matplotlib.gridspec as gridspec
+from astropy import units as u
 
 # -------------------------------------------------------------------------------
 # plotting settings
@@ -28,8 +29,10 @@ fsize = 14
 # load spectra and labels
 # -------------------------------------------------------------------------------
 
+ps = True
 print('loading labels...')
-hdu = fits.open('data/RRL_GDR2_ALLWISE_full.fits')
+if ps: hdu = fits.open('data/RRL_PS1_Gaia_full.fits')
+else: hdu = fits.open('data/RRL_GDR2_ALLWISE_full.fits')
 labels = Table(hdu[1].data)
 
 offset = 0.029 # mas as per Lindegren et al. 2018
@@ -63,26 +66,45 @@ def check_H_func(x, y, A, lam1, lam2, ivar):
 # -------------------------------------------------------------------------------
 
 Kfold = 2
-name = 'N{0}_K{1}_rrl'.format(len(labels), Kfold)    
+name = 'N{0}_K{1}_rrl'.format(len(labels), Kfold)
+if ps: name += '_ps'   
 
-cut_finite = np.isfinite(labels['int_average_g']) * \
-             np.isfinite(labels['int_average_bp']) * \
-             np.isfinite(labels['int_average_rp']) * \
-             np.isfinite(labels['peak_to_peak_g']) * \
-             np.isfinite(labels['peak_to_peak_bp']) * \
-             np.isfinite(labels['peak_to_peak_rp']) * \
-             np.isfinite(labels['Jmag']) * \
-             np.isfinite(labels['Hmag']) * \
-             np.isfinite(labels['Kmag']) * \
-             np.isfinite(labels['W1mag']) * \
-             np.isfinite(labels['W2mag']) * \
-             np.isfinite(labels['pf']) * \
-             np.isfinite(labels['r21_g']) * \
-             np.isfinite(labels['r31_g']) 
-
-cut_clean = labels['num_clean_epochs_g'] > 40.
-cut_mag = labels['W1mag'] < 19
-cuts = cut_finite * cut_clean * cut_mag
+if ps:
+    cut_finite = np.isfinite(labels['phot_g_mean_mag']) * \
+                 np.isfinite(labels['phot_bp_mean_mag']) * \
[…]

### 2018-08-08 make_figures_HER.py

@@ -52,6 +52,15 @@ labels.rename_column('dec_1', 'dec')
 # Figure 1
 # -------------------------------------------------------------------------------
 
+# line parameters
+x = np.linspace(0, 8, 10)
+y1 = 0.4 + 0.45 * x
+y2 = -0.05 + 0. * x
+
+bprplim = (0.5, 6.0)
+jklim = (0.2, 2.7)
+hw2lim = (-0.2, 1.4)
+
 cm = 'viridis_r'
 fig, ax = plt.subplots(1, 2, figsize = figsize)
 ax[0].scatter(labels['bp_rp'], labels['J']-labels['K'], c = labels['LOGG'], cmap = cm, vmin = 0, vmax = 2.2, rasterized = True, s = 10, alpha = .5)
@@ -67,6 +76,12 @@ cbar_ax = fig.add_axes([1, 0.15, 0.03, 0.82])
 cb = fig.colorbar(sc, cax=cbar_ax)
 cb.set_label(r'$\log g$', fontsize = fsize)
 plt.tight_layout()
+ax[0].set_xlim(bprplim)
+ax[0].set_ylim(jklim)
+ax[1].set_xlim(bprplim)
+ax[1].set_ylim(hw2lim)
+ax[0].plot(x, y1, linestyle= '-', color = '0.6')
+ax[1].plot(x, y2, linestyle= '-', color = '0.6')
 plt.savefig('paper/parent_sample.pdf', pad_inches=.2, bbox_inches = 'tight')
 
 # -------------------------------------------------------------------------------
@@ -93,13 +108,19 @@ cbar_ax = fig.add_axes([1, 0.15, 0.03, 0.82])
 cb = fig.colorbar(sc, cax=cbar_ax)
 cb.set_label(r'$\log g$', fontsize = fsize)
 plt.tight_layout()
+ax[0].plot(x, y1, linestyle= '-', color = '0.6')
+ax[1].plot(x, y2, linestyle= '-', color = '0.6')
+ax[0].set_xlim(bprplim)
+ax[0].set_ylim(jklim)
+ax[1].set_xlim(bprplim)
+ax[1].set_ylim(hw2lim)
 plt.savefig('paper/training_sample.pdf', pad_inches=.2, bbox_inches = 'tight')
 
 # -------------------------------------------------------------------------------
 # Figure 3 (parallax vs. parallax for training and Gaia excellent, colored by SNR)
[…]

### 2018-08-08 rotation_curve_uncertainties.py

@@ -30,6 +30,7 @@ from plotting_helpers import histcont
 
 matplotlib.rcParams['ytick.labelsize'] = 14
 matplotlib.rcParams['xtick.labelsize'] = 14
+matplotlib.rc('text', usetex=True)
 fsize = 14
 
 # -------------------------------------------------------------------------------
@@ -256,6 +257,7 @@ mean_feh = np.zeros((len(all_x), len(all_y))) - np.inf
 cut_feh = labels['FE_H'] > -100
 mean_sigma_mu = np.zeros((len(all_x), len(all_y))) - np.inf
 mean_sigma_par = np.zeros((len(all_x), len(all_y))) - np.inf
+mean_rv = np.zeros((len(all_x), len(all_y))) - np.inf
 
 for i, box_center_x in enumerate(all_x):
     for j, box_center_y in enumerate(all_y):
@@ -267,6 +269,7 @@ for i, box_center_x in enumerate(all_x):
             mean_XS_cart[i, j, :] = np.nanmean(mean_XS_cart_n[cut_patch], axis = 0)
             mean_HW2[i, j] = np.nanmean(labels['H'][cut_patch] - labels['w2mpro'][cut_patch])
             mean_feh[i, j] = np.nanmean(labels['FE_H'][cut_patch * cut_feh])
+            mean_rv[i, j] = np.nanmean(labels['VHELIO_AVG'][cut_patch])
             mean_sigma_mu[i, j] = np.nanmean(np.sqrt(labels['pmra_error'][cut_patch] ** 2 + labels['pmdec_error'][cut_patch] ** 2))
             mean_sigma_par[i, j] = np.nanmean(0.09 * labels['spec_parallax'][cut_patch])
             error_var_XS_cyl[i, j, :, :] = np.nanmean(var_XS_cyl_n[cut_patch], axis=0)
@@ -487,7 +490,7 @@ plt.xlabel('$x$', fontsize = fsize)
 plt.ylabel('$y$', fontsize = fsize)
 ax.set_aspect('equal')
 plt.savefig('plots/rotation_curve/vrvp_{}.pdf'.format(name), bbox_inches = 'tight')
-'''plt.close()
+plt.close()
 
[…]

### 2018-08-08 paper/sp.tex
notes to self
color cuts; differences in samples

@@
 % -----------
-% - draft discussion
 % - make sure we address the sensibleness and generality of the linear model for the photometric space
@@
 % - we need an uncertainty estimate or noise model for our outputs.
-% - we have set lambda = 50 by x-val. We could go as high as 60 without hurting our results.
-% - We probably should cut in color-color space. Probably in J-K vs B-R and in H-W2.
-% - We will match axes on relevant plots.
-% - ...
+% - general HWR point that we aren't putting in to the model any idea about how apparent magnitude varies with distance; this means that we rely on data for EVERYTHING.
+% - HWR will deliver cluster memberships, and then we will improve the cluster plots.
+% - CODE: We will match axes on relevant plots.
+% - CODE: we need to look at how things change with the parallax offset (0.029). It is another hyper-par.
+% - CODE: we have set lambda = 50 by x-val -- final re-set?
 
@@
 \newcommand{\logg}{\log g}
-\newcommand{\BP}{B\!P}
-\newcommand{\RP}{R\!P}
+\newcommand{\BP}{{G_\mathrm{BP}}}
+\newcommand{\RP}{{G_\mathrm{RP}}}
 \newcommand{\gparallax}{\varpi^{(\mathrm{a})}}
@@
 $H_n$, $K_n$, $W_{1n}$, and $W_{2n}$ and also an \apogee\ spectrum.
-This Parent Sample contains ACE HOW MANY STARS? and is shown in \figurename~\ref{fig:parent}.
+In addition, we applied two light color cuts to remove stars with obviously
+contaminated or outlying photometry:
+\begin{eqnarray}
+(J_n - K_n) &<& (+0.4\,\mathrm{mag}) + 0.45 * (\BP_n - \RP_n)
+\\
+(H_n - W_{2n}) &>& (-0.05\,\mathrm{mag})
+\quad .
+\end{eqnarray}
+These cuts removed roughly 2~percent of the \apogee\ stars.
+This Parent Sample contains 44808 (HOGG CHECK) and is shown in \figurename~\ref{fig:parent}.
 \begin{figure}
 \includegraphics[width=\textwidth]{parent_sample.pdf}
-\caption{A visualization of the Parent Sample stars. The left panel
-  shows a color--color diagram for all XXXX stars in the combined Parent Sample, colored by
-  the \apogee\ pipeline estimates of surface gravity. The right
-  panel shows small spectral segments of four example stars,
-  chosen to span metallicity and surface gravity.\label{fig:parent}}
+\caption{Color--color diagrams for the Parent Sample stars.
+  The points are colored by
+  the \apogee\ pipeline estimates of surface gravity.
+  The color cuts used to trim outliers are shown as grey lines.\label{fig:parent}}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{./training_sample.pdf}
-\caption{A visualization of the Training Set stars. The left panel shows
-  same color--color diagram as in \figurename~\ref{fig:parent}, but for
-  the Training Set only. Note [HOGG SOME DIFFERENCES]. The right panel
-  shows HOGG WHAT?\label{fig:training}}
+\caption{Color--color diagrams for the Training Set stars.
+  Note that there is less color range in the Training Set than in the
+  Parent Sample.\label{fig:training}}
 \end{figure}
+
+Comparison of {\figurename s}~\ref{fig:parent} and \ref{fig:training} shows
+that there is more color range in the Parent Sample than in the Training Set.
+This challenges the representativeness assumption in \sectionname~\ref{sec:assumptions}.
+The principal reason for the difference is that the \gaia\ quality cuts
+exclude stars preferentially from crowded regions, which also tend to be the
+most dust-obscured.
+The hope of the model assumptions is that the Training Set will contain sufficient
+dust variation that the model will naturally learn the dust corrections and extrapolate
+acceptably.
 

### 2018-08-08 make_figures_HER.py
new

@@ -48,6 +48,10 @@ labels = Table.read('data/training_labels_new_{}.fits'.format(name), format = 'f
 labels.rename_column('ra_1', 'ra')
 labels.rename_column('dec_1', 'dec')
 
+cut_jk = (labels['J'] - labels['K']) < (0.4 + 0.45 * labels['bp_rp'])
+cut_hw2 = (labels['H'] - labels['w2mpro']) > -0.05
+labels = labels[cut_jk * cut_hw2]
+
 # -------------------------------------------------------------------------------
 # Figure 1
 # -------------------------------------------------------------------------------
@@ -59,7 +63,7 @@ y2 = -0.05 + 0. * x
 
 bprplim = (0.5, 6.0)
 jklim = (0.2, 2.7)
-hw2lim = (-0.2, 1.4)
+hw2lim = (-0.1, 1.3)
 
 cm = 'viridis_r'
 fig, ax = plt.subplots(1, 2, figsize = figsize)
@@ -120,7 +124,7 @@ plt.savefig('paper/training_sample.pdf', pad_inches=.2, bbox_inches = 'tight')
 # Figure 3 (parallax vs. parallax for training and Gaia excellent, colored by SNR)
 # -------------------------------------------------------------------------------
 
-'''fig, ax = plt.subplots(1, 2, figsize = figsize, sharey = True)
+fig, ax = plt.subplots(1, 2, figsize = figsize, sharey = True)
 sc = ax[0].scatter(labels['parallax'][train], labels['spec_parallax'][train], c = labels['SNR'][train], cmap = 'viridis_r', s = 10, vmin = 50, vmax = 1000, rasterized = True)
 ax[1].scatter(labels['parallax'][best], labels['spec_parallax'][best], c = labels['SNR'][best], cmap = 'viridis_r', s = 10, vmin = 50, vmax = 1000, rasterized = True)
 ax[0].set_xlabel(r'$\varpi^{\rm (a)}$', fontsize = fsize)
@@ -203,13 +207,24 @@ f1.close()
 f2 = open('optimization/opt_results_1_{}.pickle'.format(name), 'rb')
 res2 = pickle.load(f2)
 f2.close() 
-pixlist = np.loadtxt('data/pixtest8_dr13.txt', usecols = (0,), unpack = 1).astype(int)
 
-#wl = 
[…]

### 2018-08-08 test_distances.py

@@ -103,6 +103,8 @@ R = np.sqrt(XS[:, 0] ** 2 + XS[:, 1] ** 2) # in cylindrical coordinates! # + XS[
 
 t = Table(names = ('cluster', 'RA', 'DEC', 'FE_H', 'distance', 'lat_name'), dtype=('S8', 'S8', 'S8', 'S16', 'f8', 'S8'))
 
+# look up Melissa's paper: arxiv: 1701.07829
+
 # M67
 clus = 'm67'
 lat = 'M 67'

### 2018-08-08 make_figures_HER.py
new

@@ -221,12 +221,13 @@ hdu = fits.open('data/all_flux_norm_parent.fits')
 fluxes = hdu[0].data
 gaps = (np.sum(fluxes.T, axis = 0)) == float(fluxes.T.shape[0])
 
-fig, ax = plt.subplots(2, 1, figsize = figsize, sharex = True)
+fig, ax = plt.subplots(2, 1, figsize = figsize, sharex = True, sharey = True)
 ax[0].plot(np.array(wl)[~gaps], res1.x[9:], drawstyle = 'steps-mid', lw = .8, color = 'k')
 ax[1].plot(np.array(wl)[~gaps], res2.x[9:], drawstyle = 'steps-mid', lw = .8, color = 'k')
 ax[1].set_xlabel(r'$\lambda~\rm[{\AA}]$', fontsize = 14)
+ax[0].set_xlim(min(np.array(wl)[~gaps]), max(np.array(wl)[~gaps]))
+ax[0].set_ylim(-0.8, 0.8)
 plt.savefig('paper/coefficients.pdf')
-
 
 # -------------------------------------------------------------------------------
 # Figure 7 (map of kinematics & abundances)
@@ -236,8 +237,7 @@ plt.savefig('paper/coefficients.pdf')
 # patch colored by metallicty, opacity proportional to number of stars
 # superimposed arrow color coded by v_z, opacity also proportional to number of stars
 
-fig, ax = plt.subplots(1, 1, figsize = (12, 12))
-plt.savefig('paper/map.pdf')
+# run rotation_curve_uncertainties.py
 
 
 # -------------------------------------------------------------------------------'''

### 2018-08-08 map_making.py

@@ -145,7 +145,11 @@ def KeplerianRotation(R):
 
 # take only stars in mid-plane 
 vz_cut = (abs(labels['b']) < 2) # * (abs(XS[:, 5]) < 20)
-  
+
+# -------------------------------------------------------------------------------
+# other plots
+# -------------------------------------------------------------------------------
+ 
 ## test to rotate system
 #x = (np.random.random(size = 10000) -0.5) * 100
 #y = (np.random.random(size = 10000) -0.5) * 100
@@ -314,6 +318,22 @@ ax.set_aspect('equal')
 plt.savefig('plots/maps/xy_parallax_over_error.pdf', bbox_inches = 'tight', dpi = 120)
 plt.close()
 
+fig, ax = plt.subplots(1, 1, figsize = (12, 12))
+plt.quiver(XS[vz_cut, 0], XS[vz_cut, 1], XS[vz_cut, 3], XS[vz_cut, 4], 
+           np.clip(XS[vz_cut, 5], -10, 10), cmap = 'RdBu', scale_units='xy', 
+           scale=200, alpha =.8, headwidth = 3, headlength = 4, width = 0.002)
+cb = plt.colorbar(shrink = .85)
+cb.set_label(r'$v_z$', fontsize = 15)
+plt.xlim(Xlimits[0])
+plt.ylim(Xlimits[1])
+overplot_rings()
+plt.tick_params(axis=u'both', direction='in', which='both')
+plt.xlabel('$x$', fontsize = fsize)
+plt.ylabel('$y$', fontsize = fsize)
+ax.set_aspect('equal')
+plt.savefig('plots/maps/xy_arrow.pdf', bbox_inches = 'tight')
+plt.close()
+
 AKs = 0.918 * (labels['H'] - labels['w2mpro'] - 0.08)
 HW2 = labels['H'] - labels['w2mpro']
 fig, ax = plt.subplots(1, 1, figsize = (10, 10))
@@ -355,22 +375,6 @@ plt.xlabel('$x$', fontsize = fsize)
 plt.ylabel('$z$', fontsize = fsize)
 ax.set_aspect('equal')
 plt.savefig('plots/maps/xz_vtan.pdf', bbox_inches = 'tight', dpi = 120)
-plt.close()
-
-fig, ax = plt.subplots(1, 1, figsize = (12, 12))
[…]

### 2018-08-08 rotation_curve_uncertainties.py

@@ -316,11 +316,32 @@ for i, (r_start, r_end) in enumerate(zip(bin_start, bin_end)):
         cut_feh = labels['FE_H'] > -100
         mean_feh_annulus[i] = np.nanmean(labels['FE_H'][cut_annulus * cut_feh])
         
+        
+# -------------------------------------------------------------------------------
+# Figure 7 for HER18
+# -------------------------------------------------------------------------------           
+
+fig, ax = plt.subplots(1, 1, figsize = (12, 12))
+plt.quiver(XS_cart_true_n[cut_z, 0], XS_cart_true_n[cut_z, 1], XS_cart_true_n[cut_z, 3], XS_cart_true_n[cut_z, 4], 
+           np.clip(XS_cart_true_n[cut_z, 5], -10, 10), cmap = 'RdBu', scale_units='xy', 
+           scale=200, alpha =.8, headwidth = 3, headlength = 4, width = 0.002, rasterized = True)
+cb = plt.colorbar(shrink = .8)
+cb.set_label(r'$v_z$', fontsize = 15)
+plt.xlim(-25, 5)
+plt.ylim(-10, 20)
+overplot_rings()
+overplot_rings_helio()
+plt.tick_params(axis=u'both', direction='in', which='both')
+plt.xlabel('$x$', fontsize = fsize)
+plt.ylabel('$y$', fontsize = fsize)
+ax.set_aspect('equal')
+plt.savefig('paper/map.pdf', bbox_inches = 'tight')
+        
 # -------------------------------------------------------------------------------
 # calculate rotational velocity via Jeans equation (in patches)
 # -------------------------------------------------------------------------------        
 
-rho_R_exp = 3. # kpc
+'''rho_R_exp = 3. # kpc
 vrr_R_exp = 15. # kpc
 
 # velocity tensor!
@@ -562,7 +583,7 @@ plt.close()
 # plots with arrows!
 # -------------------------------------------------------------------------------
 
-'''fig, ax = plt.subplots(1, 1, figsize = (12, 12))        
+fig, ax = plt.subplots(1, 1, figsize = (12, 12))        
 plt.quiver(mean_XS_cart[:, :, 0].flatten(), mean_XS_cart[:, :, 1].flatten(), mean_XS_cart[:, :, 3].flatten(), mean_XS_cart[:, :, 4].flatten(), \
[…]

### 2018-08-09 infer_parallax_phot_only.py
new

@@ -35,7 +35,7 @@ print('loading labels...')
 hdu = fits.open('data/training_labels_parent.fits')
 labels = Table(hdu[1].data)
 
-offset = 0.009 # mas as per Lindegren et al. 2018
+offset = 0.029 # mas as per Lindegren et al. 2018
 labels['parallax'] += offset
 
 
@@ -51,21 +51,21 @@ cut_jk = (labels['J'] - labels['K']) < (0.4 + 0.45 * labels['bp_rp'])
 cut_hw2 = (labels['H'] - labels['w2mpro']) > -0.05
 labels = labels[cut_jk * cut_hw2]
 
-if prediction: 
-    
-    print('loading spectra...')
-
-    hdu = fits.open('data/all_flux_norm_parent.fits')
-    fluxes = hdu[0].data
-    fluxes = fluxes[:, cut_jk * cut_hw2]
-                          
-# -------------------------------------------------------------------------------
-# add pixel mask to remove gaps between chips! 
-# -------------------------------------------------------------------------------
-
-    print('removing chip gaps...')               
-    gaps = (np.sum(fluxes.T, axis = 0)) == float(fluxes.T.shape[0])
-    fluxes = fluxes[~gaps, :]
+#if prediction: 
+#    
+#    print('loading spectra...')
+#
+#    hdu = fits.open('data/all_flux_norm_parent.fits')
+#    fluxes = hdu[0].data
+#    fluxes = fluxes[:, cut_jk * cut_hw2]
+#                          
+## -------------------------------------------------------------------------------
+## add pixel mask to remove gaps between chips! 
+## -------------------------------------------------------------------------------
+#
+#    print('removing chip gaps...')               
+#    gaps = (np.sum(fluxes.T, axis = 0)) == float(fluxes.T.shape[0])
+#    fluxes = fluxes[~gaps, :]
 
 # -------------------------------------------------------------------------------
 # linear algebra
@@ -94,34 +94,52 @@ def check_H_func(x, y, A, lams, ivar):
 # -------------------------------------------------------------------------------
 
 Kfold = 2
-lam = 30                      # hyperparameter -- needs to be tuned!
-name = 'N{0}_lam{1}_K{2}_offset{3}_parallax'.format(len(labels), lam, Kfold, offset)
[…]

### 2018-08-09 linear_cannon.py

@@ -1,0 +1,107 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Aug  9 12:59:13 2018
+
+@author: eilers
+"""
+
+#!/usr/bin/env python2
+# -*- coding: utf-8 -*-
+"""
+Created on Tue Jun  5 15:15:21 2018
+
+@author: eilers
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import scipy.optimize as op
+import pickle
+from astropy.table import Column, Table, join, vstack, hstack
+from astropy.io import fits
+from scipy.stats import binned_statistic_2d
+import matplotlib.gridspec as gridspec
+
+# -------------------------------------------------------------------------------
+# plotting settings
+# -------------------------------------------------------------------------------
+
+matplotlib.rcParams['ytick.labelsize'] = 14
+matplotlib.rcParams['xtick.labelsize'] = 14
+fsize = 14
+
+# -------------------------------------------------------------------------------
+# load spectra and labels
+# -------------------------------------------------------------------------------
+
+print('loading labels...')
+hdu = fits.open('data/training_labels_parent.fits')
+labels = Table(hdu[1].data)
+
+
+hdulist = fits.open('./data/spectra/apStar-t9-2M00000002+7417074.fits')
+header = hdulist[1].header
+flux = hdulist[1].data[0]
+start_wl = header['CRVAL1']
+diff_wl = header['CDELT1']
+val = diff_wl * (len(flux)) + start_wl
+wl_full_log = np.arange(start_wl, val, diff_wl)
+wl = np.array([10**aval for aval in wl_full_log])
+
+# -------------------------------------------------------------------------------
+# color_cuts
+# -------------------------------------------------------------------------------
+
+cut_jk = (labels['J'] - labels['K']) < (0.4 + 0.45 * labels['bp_rp'])
+cut_hw2 = (labels['H'] - labels['w2mpro']) > -0.05
+cuts = (abs(labels['TEFF']) < 8000) * (abs(labels['LOGG']) < 10) * (abs(labels['FE_H']) < 10)
+labels = labels[cut_jk * cut_hw2 * cuts]
+    
+print('loading spectra...')
+
[…]

### 2018-08-09 normalize_all_spectra.py

@@ -49,9 +49,9 @@ def LoadAndNormalizeData(file_spectra, file_name, destinations, pca = False, X_m
     i=0
     no_data = []
     no_data_i = np.ones((len(file_spectra),), dtype = bool)
-    for entry, destination in zip(file_spectra, destinations):
+    for entry, destination_i in zip(file_spectra, destinations):
         print(i)
-        destination = './data/spectra/' #+ str(destination) + '/'
+        destination = './data/spectra/' + str(destination_i) + '/'
         try:
             hdulist = fits.open(destination + entry.strip())
             if len(hdulist[1].data) < 8575: 
@@ -71,7 +71,7 @@ def LoadAndNormalizeData(file_spectra, file_name, destinations, pca = False, X_m
             all_flux[i] = flux
             all_sigma[i] = sigma
         except:
-            entry = entry.replace('apStar', 'aspcapStar')
+            entry = entry.replace('apStar-r8', 'aspcapStar-r8-l31c.2')
             try:
                 hdulist = fits.open(destination + entry.strip())
                 if len(hdulist[1].data) < 8575: 
@@ -112,7 +112,7 @@ def LoadAndNormalizeData(file_spectra, file_name, destinations, pca = False, X_m
 #    f = open('data/' + file_name, 'wt')
 #    pickle.dump(data_norm, f)
 #    f.close()
-    fits.writeto('data/' + file_name, data_norm[:, :, 1], overwrite = True)
+    fits.writeto('data/' + file_name, data_norm[:, :, 1:], overwrite = True)
     
     return data_norm, continuum, no_data_i
 

### 2018-08-13 paper/sp.tex
map
notes to self
refs

@@
 % - general HWR point that we aren't putting in to the model any idea about how apparent magnitude varies with distance; this means that we rely on data for EVERYTHING.
-% - HWR will deliver cluster memberships, and then we will improve the cluster plots.
-% - CODE: We will match axes on relevant plots.
-% - CODE: we need to look at how things change with the parallax offset (0.029). It is another hyper-par.
-% - CODE: we have set lambda = 50 by x-val -- final re-set?
+% - write words about the setting of Lambda and the parallax offset (currently 0.0483 +/- 0.012 mas)
+% - we have shown that outliers in sp - gaia tend to have large H-W2 or excess flux like in crowded regions, but there are some exceptions.
+Our error estimates are consistent with no intrinsic scatter if we use a robust chi-squared estimate.
+That is, chi-squared is dominated by outliers.
+We clipped the bad apogee spectral pixels at an error of 0.05 in normalized flux.
+Our uncertainty estimates are sensitive to this.
+The reader is asked to suck it.
+% - write words about other linear models we have made or could make, probably in the discusssion.
 
@@
 straightforward to spectroscopically measure metallicities,
-detailed chamical abundances, and stellar parameters, including ages (CITE MARTIG, NESS).
+detailed chamical abundances, and stellar parameters, including ages (\citealt{martig, nessage}).
 The red-giant branch is photometrically near-orthogonal to reddening vectors by dust,
@@
 data-driven tools for  stellar spectroscopy, for example to measure stellar
-parameters (CITE NESS), abundances (CITE HO, NESS, CASEY), and masses and ages (CITE NESS).
+parameters (CITE NESS), abundances (CITE HO, NESS, CASEY), and masses and ages (\citealt{nessage}).
 
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{map.pdf}
 \caption{A map of the kinematics and element abundances in the Milky Way disk.
@@
 \bibitem[Lutz \& Kelker(1973)]{lk} Lutz, T.~E., \& Kelker, D.~H.\ 1973, \pasp, 85, 573 
+\bibitem[Martig \etal(2016)]{martig} Martig, M., Fouesneau, M., Rix, H.-W., \etal\ 2016, \mnras, 456, 3655 
+\bibitem[Ness \etal(2016)]{nessage} Ness, M., Hogg, D.~W., Rix, H.-W., \etal\ 2016, \apj, 823, 114 
 \bibitem[Oliphant(2006)]{numpy} Oliphant, T.~E.\ 2006, \emph{A guide to NumPy}, Trelgol Publishing

### 2018-08-13 make_figures_HER.py
new figures

@@ -217,16 +217,23 @@ val = diff_wl * (len(flux)) + start_wl
 wl_full_log = np.arange(start_wl, val, diff_wl)
 wl = [10**aval for aval in wl_full_log]
 
+hdu = fits.open('data/linear_cannon.fits')
+parameters = hdu[0].data
+
 hdu = fits.open('data/all_flux_norm_parent.fits')
 fluxes = hdu[0].data
 gaps = (np.sum(fluxes.T, axis = 0)) == float(fluxes.T.shape[0])
 
-fig, ax = plt.subplots(2, 1, figsize = figsize, sharex = True, sharey = True)
-ax[0].plot(np.array(wl)[~gaps], res1.x[9:], drawstyle = 'steps-mid', lw = .8, color = 'k')
-ax[1].plot(np.array(wl)[~gaps], res2.x[9:], drawstyle = 'steps-mid', lw = .8, color = 'k')
+fig, ax = plt.subplots(2, 1, figsize = (12, 12), sharex = True, sharey = True)
+ax[0].fill_between(np.array(wl)[~gaps], abs(res1.x[9:])) #, drawstyle = 'steps-mid', lw = .8, color = 'k')
+ax[1].fill_between(np.array(wl)[~gaps], abs(res2.x[9:])) #, drawstyle = 'steps-mid', lw = .8, color = 'k')
+for i in range(1, 2):
+    ax[0].fill_between(wl, abs(5 * parameters[:, i]), alpha = .5)
+    ax[1].fill_between(wl, abs(5 * parameters[:, i]), alpha = .5)
 ax[1].set_xlabel(r'$\lambda~\rm[{\AA}]$', fontsize = 14)
 ax[0].set_xlim(min(np.array(wl)[~gaps]), max(np.array(wl)[~gaps]))
-ax[0].set_ylim(-0.8, 0.8)
+ax[0].set_ylim(0., 0.5)
+ax[0].set_xlim(16700, 16900)
 plt.savefig('paper/coefficients.pdf')
 
 # -------------------------------------------------------------------------------
@@ -239,14 +246,33 @@ plt.savefig('paper/coefficients.pdf')
 
 # run rotation_curve_uncertainties.py
 
+# -------------------------------------------------------------------------------
+# Figure 8
+# -------------------------------------------------------------------------------
+
+apogee_table = fits.open('data/allStar-l31c.2.fits')
[…]

### 2018-08-13 test_distances.py

@@ -37,8 +37,8 @@ fsize = 14
 
 N = 44808 #45787
 Kfold = 2
-lam = 50
-name = 'N{0}_lam{1}_K{2}_parallax'.format(N, lam, Kfold)
+lam = 30
+name = 'N{0}_lam{1}_K{2}_offset0.049_parallax'.format(N, lam, Kfold)
 
 print('loading new labels...')   
 labels = Table.read('data/training_labels_new_{}.fits'.format(name), format = 'fits')    
@@ -324,12 +324,12 @@ for i, clus in enumerate(cluster_list):
     cut_clus = (abs(labels['ra'] - ra_clus) < .8) \
           * (abs(labels['dec'] - dec_clus) < .8) 
           
-    bins = np.linspace(-0.1, .8, 20)
+    bins = np.linspace(-0.1, .8, 40)
     ax[c, r].hist(labels['spec_parallax'][cut_clus], normed = True, bins = bins, histtype = 'step', lw = 3, color = 'k', label = r'$\varpi^{\rm (sp)}$')
     ax[c, r].hist(labels['parallax'][cut_clus], normed = True, bins = bins, histtype = 'step', lw = 1, color = 'k', label = r'$\varpi^{\rm (a)}$')
     ax[c, r].tick_params(axis=u'both', direction='in', which='both', right = 'on', top = 'on')
     
-    ax[c, r].axvline(1./t['distance'][k][0], linestyle = '--', color = '#929591', lw = 2)
+    ax[c, r].axvline(1./t['distance'][k][0], color = 'r', lw = 2)
     ax[c, r].set_title(r'{0}, $\rm [Fe/H] = {1}$'.format(t['lat_name'][k][0], t['FE_H'][k][0]), fontsize = 13)
     ax[c, r].set_xlabel(r'$\varpi$ [mas]', fontsize = 14)
     ax[c, r].set_ylabel('normalized counts', fontsize = 14)
@@ -345,11 +345,83 @@ ax[3, 2].hist(labels_sgr['spec_parallax'], normed = True, bins = bins, histtype 
 ax[3, 2].hist(labels_sgr['parallax'], normed = True, bins = bins, histtype = 'step', lw=1, color='k', label = r'$\varpi^{(a)}$')
 ax[3, 2].set_xlabel(r'$\varpi$ [mas]', fontsize = 14)
 ax[3, 2].set_xlim(-0.1, .8)
-ax[3, 2].axvline(1./20, linestyle = '--', color = '#929591')
[…]

### 2018-08-13 paper/sp.tex
refs

@@
 data-driven tools for  stellar spectroscopy, for example to measure stellar
-parameters (CITE NESS), abundances (CITE HO, NESS, CASEY), and masses and ages (\citealt{nessage}).
+parameters (\citealt{cannon}), abundances (\citealt{ho, casey, nessdopp}),
+and masses and ages (\citealt{nessage}).
 
@@
 \bibitem[Martig \etal(2016)]{martig} Martig, M., Fouesneau, M., Rix, H.-W., \etal\ 2016, \mnras, 456, 3655 
+\bibitem[Ness \etal(2015)]{cannon} Ness, M., Hogg, D.~W., Rix, H.-W., Ho, A.~Y.~Q., \& Zasowski, G.\ 2015, \apj, 808, 16 
 \bibitem[Ness \etal(2016)]{nessage} Ness, M., Hogg, D.~W., Rix, H.-W., \etal\ 2016, \apj, 823, 114 

### 2018-08-13 paper/sp.tex
refs

@@
 \bibitem[Astropy Collaboration \etal(2018)]{astropy2} The Astropy Collaboration, Price-Whelan, A.~M., Sip{\H o}cz, B.~M., \etal\ 2018, arXiv:1801.02634 
-\bibitem[Bailer-Jones et al.(2018)]{calj} Bailer-Jones, C.~A.~L., Rybizki, J., Fouesneau, M., Mantelet, G., \& Andrae, R.\ 2018, \aj, 156, 58 
+\bibitem[Bailer-Jones \etal(2018)]{calj} Bailer-Jones, C.~A.~L., Rybizki, J., Fouesneau, M., Mantelet, G., \& Andrae, R.\ 2018, \aj, 156, 58 
 \bibitem[Blanton \etal(2017)]{sdssiv} Blanton, M.~R., Bershady, M.~A., Abolfathi, B., \etal\ 2017, \aj, 154, 28 
+\bibitem[Casey \etal(2016)]{casey} Casey, A.~R., Hogg, D.~W., Ness, M., \etal\ 2016, arXiv:1603.03040 
+\bibitem[Ho \etal(2017)]{ho} Ho, A.~Y.~Q., Ness, M.~K., Hogg, D.~W., \etal\ 2017, \apj, 836, 5 
 \bibitem[Hogg(2018)]{gaialf} Hogg, D.~W.\ 2018, arXiv:1804.07766
@@
 \bibitem[Ness \etal(2016)]{nessage} Ness, M., Hogg, D.~W., Rix, H.-W., \etal\ 2016, \apj, 823, 114 
+\bibitem[Ness \etal(2018)]{nessdopp} Ness, M., Rix, H.-W., Hogg, D.~W., \etal\ 2018, \apj, 853, 198 
 \bibitem[Oliphant(2006)]{numpy} Oliphant, T.~E.\ 2006, \emph{A guide to NumPy}, Trelgol Publishing

### 2018-08-13 test_distances.py
new figures

@@ -390,7 +390,7 @@ plt.savefig('plots/open_clusters/ngc2862_new.pdf')
 
 cluster_list = ['M67', 'M71', 'M107', 'NGC2862']
 
-fig, ax = plt.subplots(2, 2, figsize = (12, 8), sharex = True, sharey = True)
+fig, ax = plt.subplots(2, 2, figsize = (12, 5), sharex = True, sharey = True)
 plt.subplots_adjust(wspace = 0.02, hspace = 0.02)
 c, r = 0, 0
 for i in range(4):
@@ -398,14 +398,14 @@ for i in range(4):
     mem = Table(hdu[1].data)
 
     xx = join(mem, labels, join_type = 'inner', keys = 'APOGEE_ID')
-    ax[c, r].errorbar(xx['spec_parallax_2'], xx['LOGG_2'], xerr = xx['spec_parallax_err'], fmt = 'o', color = "k", label = r'$\varpi^{\rm (sp)}$', zorder = 20)
-    ax[c, r].errorbar(xx['parallax_2'], xx['LOGG_2'], xerr = xx['parallax_error_2'], fmt = 'o', color = '#929591', label = r'$\varpi^{\rm (g)}$', zorder = 10)
+    ax[c, r].errorbar(xx['spec_parallax_2'], xx['LOGG_2'], xerr = xx['spec_parallax_err'], fmt = 'o', markersize = 5, color = "k", label = r'$\varpi^{\rm (sp)}$', zorder = 20)
+    ax[c, r].errorbar(xx['parallax_2'], xx['LOGG_2'], xerr = xx['parallax_error_2'], fmt = 'o', markersize = 5, color = '#929591', label = r'$\varpi^{\rm (g)}$', zorder = 10)
     avg = np.sum(xx['spec_parallax_2'] / xx['spec_parallax_err'] ** 2) / np.sum(1. / xx['spec_parallax_err'] **2)
     avg_gaia = np.sum(xx['parallax_2'] / xx['parallax_error_2'] ** 2) / np.sum(1. / xx['parallax_error_2'] **2)
     ax[c, r].axvline(avg, color="k", lw = 0.8, zorder = -10, linestyle = ':')
     ax[c, r].axvline(avg_gaia, color='#929591', lw = 0.8, zorder = -10, linestyle = ':')
     ax[c, r].tick_params(axis=u'both', direction='in', which='both', right = 'on', top = 'on')
[…]

### 2018-08-13 paper/sp.tex
refs

@@
 
-We take the \apogee\ (HOGG CITE) spectral data
+We take the \apogee\ (\citealt{aapogee, wapogee, apogee}) spectral data
 from \sdssiv\ (\citealt{sdssiv}) \acronym{DR14} (\citealt{dr14}).
@@
 models, we did use stellar models indirectly to select the Parent Sample of stars:
-We used the \apogee\ pipeline surface-gravity $\logg$ measurements.
+We used the \apogee\ pipeline (\citealt{aspcap}) surface-gravity $\logg$ measurements.
 This tiny use of stellar models could have been obviated by selecting stars not
@@
 \bibitem[Abolfathi \etal(2018)]{dr14} Abolfathi, B., Aguado, D.~S., Aguilar, G., \etal\ 2018, \apjs, 235, 42 
+\bibitem[Allende Prieto \etal(2008)]{aapogee} Allende Prieto, C., Majewski, S.~R., Schiavon, R., \etal\ 2008, Astronomische Nachrichten, 329, 1018 
 \bibitem[Astropy Collaboration \etal(2013)]{astropy} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., \etal\ 2013, \aap, 558, A33 
@@
 \bibitem[Casey \etal(2016)]{casey} Casey, A.~R., Hogg, D.~W., Ness, M., \etal\ 2016, arXiv:1603.03040 
+\bibitem[Garc{\'{\i}}a P{\'e}rez \etal(2016)]{aspcap} Garc{\'{\i}}a P{\'e}rez, A.~E., Allende Prieto, C., Holtzman, J.~A., \etal\ 2016, \aj, 151, 144 
 \bibitem[Ho \etal(2017)]{ho} Ho, A.~Y.~Q., Ness, M.~K., Hogg, D.~W., \etal\ 2017, \apj, 836, 5 
@@
 \bibitem[Lutz \& Kelker(1973)]{lk} Lutz, T.~E., \& Kelker, D.~H.\ 1973, \pasp, 85, 573 
+\bibitem[Majewski \etal(2017)]{apogee} Majewski, S.~R., Schiavon, R.~P., Frinchaboy, P.~M., \etal\ 2017, \aj, 154, 94 
 \bibitem[Martig \etal(2016)]{martig} Martig, M., Fouesneau, M., Rix, H.-W., \etal\ 2016, \mnras, 456, 3655 
@@
 \bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., \etal\ 2006, \aj, 131, 1163 
+\bibitem[Wilson \etal(2010)]{wapogee} Wilson, J.~C., Hearty, F., Skrutskie, M.~F., \etal\ 2010, \procspie, 7735, 77351C 
 \end{thebibliography}

### 2018-08-13 paper/sp.tex
cite

@@
 stars that are more luminous than the red clump.
-The \apogee\ pipeline $\logg$ values (which we use) have uncertainties but
+The \apogee\ pipeline (\citealt{aspcap})
+values of surface gravity $\logg$ (which we use) have uncertainties but
 this cut leads to a clean sample of luminous red giants, and it is a cut

### 2018-08-13 paper/sp.tex
refs

@@
 We then pseudo-continuum-normalize the spectra according to the procedure developed
-in \project{The~Cannon} (HOGG CITE):
+in \project{The~Cannon} (\citealt{cannon}):
 That is, the pseudo-continuum is a spline fit to a set
@@
 
-We used the \gaia\ Archive (HOGG CITE) to match the \apogee+\zmass\ stars
+We used the \gaia\ Archive (\citealt{gaiaarchive})
+official matches (\citealt{xmatch}) to match the \apogee+\zmass\ stars
 to the \wise\ Catalog (HOGG CITE),
@@
 \bibitem[Majewski \etal(2017)]{apogee} Majewski, S.~R., Schiavon, R.~P., Frinchaboy, P.~M., \etal\ 2017, \aj, 154, 94 
+\bibitem[Marrese \etal(2018)]{xmatch} Marrase, P.~M. \etal\ 2018, in preparation
 \bibitem[Martig \etal(2016)]{martig} Martig, M., Fouesneau, M., Rix, H.-W., \etal\ 2016, \mnras, 456, 3655 
+\bibitem[Mora \etal(2018)]{gaiaarchive} Mora, A., Gonz{\'a}lez-N{\'u}{\~n}ez, J., Baines, D., \etal\ 2018, IAU Symposium, 330, 35 
 \bibitem[Ness \etal(2015)]{cannon} Ness, M., Hogg, D.~W., Rix, H.-W., Ho, A.~Y.~Q., \& Zasowski, G.\ 2015, \apj, 808, 16 

### 2018-08-13 paper/sp.tex
gaia refs

@@
 % - write words about other linear models we have made or could make, probably in the discusssion.
+% - are we citing all the proper DR2 papers we need to cite?
 
@@
 
-We used the \gaia\ Archive (\citealt{gaiaarchive})
+We used the \gaia\ (\citealt{gaia}) \acronym{DR2} (\citealt{gaiadr2}) Data Archive (\citealt{gaiaarchive})
 official matches (\citealt{xmatch}) to match the \apogee+\zmass\ stars
-to the \wise\ Catalog (HOGG CITE),
+to the \wise\ Catalog (\citealt{wise}),
 according to the \gaia\ Archive internal match criteria.
@@
 
-We match this full-match catalog to the \gaia\ \acronym{DR2} (HOGG CITE)
-using the \gaia\ Archive official match (given the \zmass\ IDs).
+We match this full-match catalog to the \gaia\ \acronym{DR2}
+using the \gaia\ Archive official match (\citealt{xmatch}) to the the \zmass\ IDs.
 We take from the \gaia\ \acronym{DR2} Catalog the photometric data
@@
 \bibitem[Casey \etal(2016)]{casey} Casey, A.~R., Hogg, D.~W., Ness, M., \etal\ 2016, arXiv:1603.03040 
+\bibitem[Gaia Collaboration \etal(2016)]{gaia} Gaia Collaboration, Prusti, T., de Bruijne, J.~H.~J., \etal\ 2016, \aap, 595, A1 
+\bibitem[Gaia Collaboration \etal(2018)]{gaiadr2} Gaia Collaboration, Brown, A.~G.~A., Vallenari, A., \etal\ 2018, arXiv:1804.09365 
 \bibitem[Garc{\'{\i}}a P{\'e}rez \etal(2016)]{aspcap} Garc{\'{\i}}a P{\'e}rez, A.~E., Allende Prieto, C., Holtzman, J.~A., \etal\ 2016, \aj, 151, 144 
@@
 \bibitem[Wilson \etal(2010)]{wapogee} Wilson, J.~C., Hearty, F., Skrutskie, M.~F., \etal\ 2010, \procspie, 7735, 77351C 
+\bibitem[Wright \etal(2010)]{wise} Wright, E.~L., Eisenhardt, P.~R.~M., Mainzer, A.~K., \etal\ 2010, \aj, 140, 1868 
 \end{thebibliography}

### 2018-08-13 paper/sp.tex
noise model

@@
 % - check with WISE experts that we have the correct WISE photometry
-% - we need an uncertainty estimate or noise model for our outputs.
 % - general HWR point that we aren't putting in to the model any idea about how apparent magnitude varies with distance; this means that we rely on data for EVERYTHING.
@@
 These cuts removed roughly 2~percent of the \apogee\ stars.
-This Parent Sample contains 44808 (HOGG CHECK) and is shown in \figurename~\ref{fig:parent}.
+This Parent Sample contains 44808 stars and is shown in \figurename~\ref{fig:parent}.
 \begin{figure}
@@
 and distances to the B data and \foreign{vice versa}.
-The Parent Sample A has HOGG entries and Parent Sample B has HOGG entries.
 
@@
 In detail,
-for each star $n$ in the Training Sets, we take from \gaia\ DR2
+for each star $n$ in the Training Sets, we take from \gaia\ \acronym{DR2} (\citealt{gaiadr2})
 the astrometric parallax $\gparallax_n$ and its uncertainty $\sigma_n$.
 Importantly, to each \gaia\ parallax measurement we add a positive
-offset of $0.029$\,mas to adjust for the under-estimation of
+offset $\varpi_0$ to adjust for the under-estimation of
 parallaxes reported by the \gaia\ Collaboration (\citealt{lindegren}).
+The \gaia\ recommended value of $\varpi_0=0.029$\,mas (\citealt{lindegren}),
+but we adopt a value of $\varpi_0 = 0.048$\,mas because that value
+optimizes our cross-validation likelihood.
+The spectrophotometric parallaxes do not depend strongly on this choice, except
+for small differences at very small parallax (large distance).
 The Training Sets are shown in \figurename~\ref{fig:training}.
@@
 or $H-W_2$ color (which is a reddening proxy).
+However, that figure does show that stars with greater reddening do appear to have
+larger scatter against the \gaia\ parallaxes.
+Additional tests confirm that our spectroscopic parallaxes have largest scatter
+in the dustiest and most crowded regions, as expected given the relatively low angular resolution
+photometry.
 \begin{figure}
@@
 In \figurename~\ref{fig:clusters}, we show the parallax distribution from \gaia\ and
-the spectrophotometric-parallax distribution from this work for stars in
-angular proximity to known stellar clusters and the Sagittarius dwarf
-galaxy.
+the spectrophotometric-parallax distribution from this work for stars that are
+astrometrically confirmed members of known stellar clusters.
 These clusters span some range in metallicity and age, and therefore test
 the accuracy of the method for different kinds of populations.
-They show that both \gaia\ and the spectrophotometric parallaxes appear to be unbiased
+With the exception of \acronym{M71}, these figures 
+indicate that both \gaia\ and the spectrophotometric parallaxes appear to be unbiased
 for these clusters and the range in abundances and ages they represent.
-HOGG: DO WE NEED TO ADDRESS whether why not these clusters show 9-percent spreads?
+The case of cluster \acronym{M71} is troubling, but after searching for trends with
+housekeeping data (also see \figurename~\ref{fig:residuals}) we don't find any correlates
+with parallax offsets.
 \begin{figure}
@@
 \caption{The histograms of \gaia\ astrometric parallaxes $\gparallax_n$
-  and spectrophotometric parallaxes $\sparallax_n$ for stars in angular proximity
-  to various stellar clusters in the \apogee\ observing footprint. In each panel,
-  the expected parallax of the cluster is shown as a vertical line.\label{fig:clusters}}
+  and spectrophotometric parallaxes $\sparallax_n$ for member stars for
+  four stellar clusters in the \apogee\ observing footprint. The spectroscopic
+  parallaxes and their inverse-variance-weighted mean and the astrometric parallaxes
+  and their inverse-variance-weighted mean
+  the expected parallax of the cluster is shown as a vertical line.
+  Cluster members were identified to be close to the cluster in celestial coordinates
+  and comoving in proper motion.\label{fig:clusters}}
 \end{figure}
@@
 
-In order to use these outputs in a probabilistic inference or forward model,
-an investigator needs a noise (uncertainty) model.
 The linear model permits straightforward propagation of uncertainty from the
-feature inputs.
-However, these propagated uncertainties clearly under-predict the scatter between
-the spectrophotometric parallaxes $\sparallax_m$ and the \gaia\ astrometric
-parallaxes $\gparallax_m$.
-Our recommendation for an uncertainty model is HOGG WAT?
+feature inputs to the spectrophotometric parallaxes.
+Although the model fundamentally assumes that the prediction features are noise-free
+(see the ``good features'' assumption in \sectionname~\ref{sec:assumptions}),
+there are in fact small uncertainties on the features that we can propagate to 
+make uncertainty estimates on the spectroscopic parallaxes.
+We perform this propagation; these uncertainties are also given in
+\tablename~\ref{tab:output}.
 
@@
 
-One significant limitation of the model is that it does not produce a sensible or
-useful noise estimate or error bar on each spectrophotometric parallax estimate.
-Although the linear model does permit simple noise propagation, that noise propagation
-seriously under-estimates the scatter between the spectrophotometric parallaxes
-and the best of the \gaia\ \acronym{DR2} astrometric parallaxes.
-We thus used an empirical noise estimate to get our HOGG DETAIL HERE.
-In principle it might have been better to build a noise model for our outputs,
-and learned that noise model along with the coefficients.
+HOGG SOMETHING ABOUT THE CLUSTER M71 BEING CONCERNING AND OUR EXPLANATIONS.
 

### 2018-08-13 paper/sp.tex
words

@@
 % -----------
-% - make sure we address the sensibleness and generality of the linear model for the photometric space
 % - Coryn's point that we should carefully explain how our model is geometric.
-% - related: HWR's point that this is standardizeable candles, like the Ia SNe results.
-% - related: Put in an explanation of WHY this works so well.
-% - related: Put in an explanation of how it is that we aren't really modeling luminosity or distance but really just generating parallaxes.
-% - does the model need a cute name, like The Swan?
 % - check with WISE experts that we have the correct WISE photometry
-% - general HWR point that we aren't putting in to the model any idea about how apparent magnitude varies with distance; this means that we rely on data for EVERYTHING.
-% - write words about the setting of Lambda and the parallax offset (currently 0.0483 +/- 0.012 mas)
-% - we have shown that outliers in sp - gaia tend to have large H-W2 or excess flux like in crowded regions, but there are some exceptions.
-Our error estimates are consistent with no intrinsic scatter if we use a robust chi-squared estimate.
-That is, chi-squared is dominated by outliers.
-We clipped the bad apogee spectral pixels at an error of 0.05 in normalized flux.
-Our uncertainty estimates are sensitive to this.
-The reader is asked to suck it.
 % - write words about other linear models we have made or could make, probably in the discusssion.
@@
 \title{\textbf{%
-Spectrophotometric parallaxes with \project{The Cygnet}:\\
+Spectrophotometric parallaxes with linear models:\\
 Accurate distances for luminous red-giant stars
@@
 The training makes use of a justified likelihood function for the \gaia\ parallaxes,
-so that the training can be performed with full use of uncertainties and
-without any cuts on parallax or parallax signal-to-noise ratio (not even any non-negative
-requirements).
+so that we never take the inverse or logarithm of a parallax, and so that
+the training can be performed with full use of uncertainties and
+without any cuts on parallax or parallax signal-to-noise ratio.
 There are no physical inputs to the model; it is a pure regression designed
@@
 input of explicit extinction or reddening estimates.
-We re-label each star in the sample
-with a new spectrophotometric parallax; this new parallax has an empirical (1-sigma)
-uncertainty of 9~percent, which is more precise than the \gaia\ parallax
-for the vast majority of targets.
-Validation with globular and open clusters shows that the spectrophotometric parallaxes
-are both accurate and precise.
+We re-label each star in the sample with a new spectrophotometric parallax;
+these new parallaxes have uncertainties of a few to 15 percent, depending on data quality,
+which is more precise than the \gaia\ parallax for the vast majority of targets.
 These distance estimates open up great opportunities for
@@
 the infrared.
-These arguments motivate the \apogee\ (CITE), \apogee-2 (CITE), and \sdssv\ (\citealt{sdssv})
+These arguments motivate the \apogee\ (\citealt{apogee}),
+\apogee-2, and \sdssv\ (\citealt{sdssv})
 projects, which take spectra in the infrared, and deliver detailed abundances
@@
 As we operate these spectroscopic surveys,
-the \gaia\ Mission (CITE) has been revolutionizing our view of the disk.
+the \gaia\ Mission (\citealt{gaia}) has been revolutionizing our view of the disk.
 It delivers good proper motions everywhere,
@@
 straightforward to spectroscopically measure metallicities,
-detailed chamical abundances, and stellar parameters, including ages (\citealt{martig, nessage}).
+detailed chemical abundances, and stellar parameters, even including ages (\citealt{martig, nessage}).
 The red-giant branch is photometrically near-orthogonal to reddening vectors by dust,
@@
 We only use these physical expectations about red giants to recommend
-data features---inputs---for the data-driven model.
+data features---inputs---for a completely data-driven model.
 The data-driven method uses patterns in the observed data themselves
@@
 Simple data-driven models (like the ones we will employ here)
-suffer from many limitations for this kind of work:
+also suffer from many limitations for this kind of work:
 Because they are constrained only by the data, and not physical law,
@@
 This regression is very sensitive to certain aspects of the data or experimental
-design..
+design.
 One is that we will do far better with the regression if we have an accurate
@@
 Below, we will make no such cuts. That means that our training step has the odd
-property that much of the training data is low in signal-to-noise and many of the
+property that much of the training data is low in signal-to-noise and some of the
 training parallax measurements are even negative!
@@
 explicitly.
-We will return to these assumptions in the Discussion section below,
-to address the costs and benefits of each in more depth.
 \begin{description}
@@
 of a star can be predicted from the features we provide, which are
-the full set of (pseudo-continuum-normalized) pixels from the \apogee\ spectrum,
-plus the $G$ and $(\BP-\RP)$ photometry from \gaia,
+the full set of (pseudo-continuum-normalized) pixels from the \apogee\ (\citealt{apogee}) spectrum,
+plus the $G$, $\BP$, and $\RP$ photometry from \gaia\ (\citealt{gaia}),
 plus the $J$, $H$, and $K_s$ photometry from \zmass\ (\citealt{zmass}),
-plus the $W_1$ and $W_2$ photometry from \wise.
+plus the $W_1$ and $W_2$ photometry from \wise\ (\citealt{wise}).
 That is, we assume that these spectrophotometric \emph{features} are sufficient
@@
 the chosen features. We are only assuming this linearity in a small range of stellar
-surface gravity $\logg$, but this assumption is strong, and limits strongly the
+surface gravity, but this assumption is strong, and limits strongly the
 flexibility or capacity of the model.
@@
 to date, but it subsumes a number of related assumptions, like that the \gaia\ noise
-model is correct, that the stars are only covariant at negligible levels, and that
+model is correct, that the stars are only covariant with one another at negligible levels, and that
 there are no significant outliers or large-scale systematics.
@@
 and $||y||_1^1$ is the L1-norm or sum of absolute values of the components of $y$.
-This kind of regularization adds a convex term to the optimization and leads to
+This kind of regularization (L1) adds a convex term to the optimization and leads to
 sparse optima.
@@
 model to be sparse.
-We set the value of $\lambda$ by a 
-very coarse cross-validation.
+We set the value of $\lambda$ by cross-validation.
 
@@
 the convex problem of L1-regularized linear least-square fitting.
-Optimization is performed with \code{scipy.optimize} \acronym{L-BFGS-B}.
+Optimization is performed with \code{scipy.optimize} using the
+\acronym{L-BFGS-B} algorithm (HOGG CITE).
 
@@
 $x_m$ is the feature $D$-vector for star $m$.
-Thus the model can be trained on a training set of stars and applied to make
-predictions for a validation set or a test set.
+Thus the model can be trained on a training set of stars and applied to
+a validation set or a test set to make predictions.
 The only requirements are that every test-set object $m$  must have a full feature
@@
 This ensures that, in the estimate of any individal star's
-parallax, none of the \gaia\ data pertaining to that star were used.
+spectrophotometric parallax, none of the \gaia\ data pertaining to that star were used.
 This makes the parallax estimates from the spectrophotometric feature vectors
@@
 These cuts removed roughly 2~percent of the \apogee\ stars.
-This Parent Sample contains 44808 stars and is shown in \figurename~\ref{fig:parent}.
+This Parent Sample contains 44784 stars and is shown in \figurename~\ref{fig:parent}.
 \begin{figure}
 \includegraphics[width=\textwidth]{parent_sample.pdf}
-\caption{Color--color diagrams for the Parent Sample stars.
+\caption{Color--color diagrams for the 44784 Parent Sample stars.
   The points are colored by
@@
 This is used for two-fold validation and jackknife.
-In short, we will use the model trained on the A data to assign luminosities
-and distances to the B data and \foreign{vice versa}.
+In short, we will
+use the model trained on the Training Set A data to assign spectrophotometric parallaxes
+to the Parent Sample B data and
+use the model trained on the Training Set B data to assign spectrophotometric parallaxes
+to the Parent Sample A data.
 
@@
 measurements are good.
+These cuts leave 28226 stars total in the Training Sets.
 But---as we have emphasized above---we do not cut ever on parallax or
 parallax error or their ratio. For this reason, most of the training set
-do not have significantly measured parallaxes and HOGG percent even have
+do not have significantly measured parallaxes and more than 2~percent even have
 negative parallaxes!
@@
 parallaxes reported by the \gaia\ Collaboration (\citealt{lindegren}).
-The \gaia\ recommended value of $\varpi_0=0.029$\,mas (\citealt{lindegren}),
-but we adopt a value of $\varpi_0 = 0.048$\,mas because that value
-optimizes our cross-validation likelihood.
+The \gaia\ recommended value is $\varpi_0=0.029$\,mas (\citealt{lindegren}),
+but we adopt $\varpi_0 = 0.048$\,mas because that value
+optimizes our cross-validation objective.
 The spectrophotometric parallaxes do not depend strongly on this choice, except
@@
 \includegraphics[width=\textwidth]{./training_sample.pdf}
-\caption{Color--color diagrams for the Training Set stars.
+\caption{Color--color diagrams for the 28226 Training Set stars.
   Note that there is less color range in the Training Set than in the
@@
 and we have taken logarithms of those fluxes.
-These feature vectors live in a $D$-dimensional space where $D=7414$
+These feature vectors live in a $D$-dimensional space where $D=7414$.
 For every star $n$ in either of the Training Sets we additionally require---in
@@
 the spectrophotometric model is predicting the \gaia\ parallaxes with little bias
-and a scatter of better than 9~percent.
+and a scatter of less than 9~percent.
 \begin{figure}
@@
   set plays no particular role in the method, but it can be used to demonstrate or
-  assess the prediction precision. Fractional precision of the prediction
-  is XXXX HOGG MEASURED HOW?\label{fig:xval}}
+  assess the prediction precision. The fractional precision of the prediction in the
+  right panel is better than 9 percent.\label{fig:xval}}
 \end{figure}
 
-We used this cross-validation framework to adjust the regularization parameters.
-We (arbitrarily) chose zero regularization on the photometric components of the
-$x_n$ vectors, and uniform, identical regularization strength for all of the
-spectral (\apogee-pixel) components.
-We did a coarse grid search in the value for the spectral-component value of the
-$\lambda$ vector, using the A/B split as the two-fold cross-validation.
-The 9-percent scatter is what we obtain at the optimal setting of the regularization
-strength.
+We used this cross-validation framework to adjust the regularization parameter $\Lambda$.
+A coarse grid search in the value for the spectral-component value of the
+$\lambda$ vector, using the A/B split as the two-fold cross-validation, with the objective
+of the cross validation being prediction accuracy.
+We get a prediction accuracy of better than 9 percent 
+at the optimal setting of the regularization strength.
 
-We adopt 9~percent as the method's precision.
-This estimate is conservative in that the 9-percent scatter includes contributions
-from both our spectrophotometric model and from the \gaia\ measurements themselves.
-However this estimate is not so conservative in that the test is performed with
+This 9-percent precision estimate is conservative in the sense that it does not deconvolve
+the contributions from the \gaia\ noise model.
+However it is also not fully conservative, in that the test is performed with
 \gaia's best stars, which are bright, nearby, in low extinction regions, and
@@
 larger scatter against the \gaia\ parallaxes.
-Additional tests confirm that our spectroscopic parallaxes have largest scatter
+Additional tests confirm that our spectrophotometric parallaxes have largest scatter
 in the dustiest and most crowded regions, as expected given the relatively low angular resolution
@@
 In \figurename~\ref{fig:clusters}, we show the parallax distribution from \gaia\ and
-the spectrophotometric-parallax distribution from this work for stars that are
+the spectrophotometric-parallax distribution from our model for stars that are
 astrometrically confirmed members of known stellar clusters.
@@
   and spectrophotometric parallaxes $\sparallax_n$ for member stars for
-  four stellar clusters in the \apogee\ observing footprint. The spectroscopic
+  four stellar clusters in the \apogee\ observing footprint. The spectrophotometric
   parallaxes and their inverse-variance-weighted mean and the astrometric parallaxes
@@
 \includegraphics[width=\textwidth]{coefficients.pdf}
-\caption{The coefficients of the spectral part of the linear model.
+\caption{HOGG: THIS FIGURE ISN'T READY YET:
+  The coefficients of the spectral part of the linear model.
   Also shown are derivatives with respect to labels for a \project{Cannon}-like
@@
 In order to use this output with \gaia\ astrometry, we recommend joining the table to the
-\gaia\ Catalog on the \zmass\ ID (name), as per the instructions at the \gaia\ Archive (CITE).
+\gaia\ Catalog on the \zmass\ ID (name), as per the instructions at the \gaia\ Archive (\citealt{gaiaarchive}).
 In order to use this output with \apogee\ radial velocities and element abundances, we recommend
@@
 \begin{table}
-\caption{The generated spectrophotometric parallaxes.\label{tab:output}}
+\caption{HOGG TBD: The generated spectrophotometric parallaxes.\label{tab:output}}
 \end{table}
@@
 there are in fact small uncertainties on the features that we can propagate to 
-make uncertainty estimates on the spectroscopic parallaxes.
+make uncertainty estimates on the spectrophotometric parallaxes.
 We perform this propagation; these uncertainties are also given in
@@
 
-The model and method is based on a raft of assumptions, listed above
+The model and method are based on a raft of assumptions, listed above
 in \sectionname~\ref{sec:assumptions}.
@@
 In that direction, it would make sense to choose features using a more complex
-technique like deep learning or an auto-encoder (CITE).
+technique like deep learning or an auto-encoder.
 These methods generate good features automatically.
@@
 Since models like Gaussian processes and deep networks subsume the linear model
-(at least approximately), they could have delivered similar results.
+(at least approximately), they could have delivered better---or at least non-worse---results.
 The issues with going to a much more complex model are manifold, however:
@@
 A more complex model might have had edge issues: Very flexible models don't extrapolate
-outside the convex hull of their training data, and our training data was probably not
+outside the convex hull of their training data, and our training data was not
 balanced in terms of parallax and dust attenuation.
 And finally, the complex model would have been harder to visualize and interpret;
-the linear model contains essentially only derivatives in the data space, which are
+the linear model contains essentially only first derivatives in the data space, which are
 relatively easy to understand.
 
-Although the model, training, and prediction make no use whatsoever about stellar
+Although the model, training, and prediction make no use whatsoever of stellar
 models, we did use stellar models indirectly to select the Parent Sample of stars:
@@
 
+HOGG: CRITICISE OUR NOISE ESTIMATION, in two parts:
+HOGG: TURN THIS WORD SALAD INTO A PARAGRAPH: we have shown that outliers in sp - gaia tend to have large $H-W_2$ or excess flux like in crowded regions, but there are some exceptions.
+Our error estimates are consistent with no intrinsic scatter if we use a robust chi-squared estimate.
+That is, chi-squared is dominated by outliers.
+We clipped the bad apogee spectral pixels at an error of 0.05 in normalized flux.
+Our uncertainty estimates are sensitive to this.
+The reader is asked to suck it.
+Combine with M71 discussion and write a paragraph.
 HOGG SOMETHING ABOUT THE CLUSTER M71 BEING CONCERNING AND OUR EXPLANATIONS.
@@
 That means that the parallax estimates generated here and given in \tablename~\ref{tab:output}
-is statistically independent of the \gaia\ parallaxes and can be combined with them in
+are statistically independent of the \gaia\ parallaxes and can be combined with them in
 na\"ive ways, for example in inverse-variance-weighted averages.
+
+HOGG: ADD HERE a couple of sentences in the discussion about the setting of Lambda and the parallax offset (currently $0.0483\pm 0.012$\,mas).
 
@@
 themselves (\citealt{gaialf}): A likelihood function must be constructed for each of the
-spectroscopic parallaxes, given assumptions about the noise in those values.
+spectrophotometric parallaxes, given assumptions about the noise in those values.
 The sensible likelihood function for star $m$ would be HOGG PUT HERE.

### 2018-08-13 paper/sp.tex
HWR notes and new abstract

@@
 % - are we citing all the proper DR2 papers we need to cite?
+% - HWR: Footnote to explain the use of the word ``feature'' for astronomers.
+% - HWR: Figures 1 and 2 need to be combined to directly compare Parent and Training.
+% - HWR: Figure 3: Note that the y-axis is positive-definite!
+% - HWR: Figure 7: Caption needs to say what arrows represent.
 
@@
 \begin{abstract}\noindent
+% context
 With contemporary infrared spectroscopic surveys like \apogee,
 red-giant stars can be observed to distances and extinctions
-at which geometric parallaxes are not precisely measured by \gaia.
-Here we use the \apogee--\gaia\ overlap to train a purely linear model of
-continuum-normalized \apogee\ spectroscopy
-(plus \gaia, \zmass, and \wise\ photometry)
-that predicts distance modulus (logarithmic parallax or distance)
-for red stars with $0<\logg<2.2$.
-The training makes use of a justified likelihood function for the \gaia\ parallaxes,
-so that we never take the inverse or logarithm of a parallax, and so that
-the training can be performed with full use of uncertainties and
-without any cuts on parallax or parallax signal-to-noise ratio.
-There are no physical inputs to the model; it is a pure regression designed
-to accurately predict astrometric parallaxes.
-The model includes an L1 regularization that zeros out the contributions of
-most spectral pixels to the parallax estimates.
+at which \gaia\ parallaxes are not informative.
+Yet the combination of effective temperature, surface gravity, composition, and age---all
+accessible through spectroscopy---determines a giant's luminosity.
+Therefore spectroscopy plus photometry should enable 
+precise spectrophotometric distance estimates.
+% aims
+Here we use the \apogee--\gaia\ overlap to train a data-driven model
+to predict parallaxes for red-giant branch stars with $0<\logg<2.2$.
+% methods
+We employ (the exponentiation of)
+a linear function of \apogee\ spectral pixel intensities and multi-band photometry
+to predict a parallax spectrophotometrically.
+The data-driven-model training involves no logarithms or inverses of the \gaia\ parallaxes,
+and needs no cut on the \gaia\ parallax signal-to-noise ratio.
+It includes an L1 regularization to zero out the contributions of
+uninformative pixels.
 The training is performed
@@
 spectrophotometric parallax estimate.
-The model is flexible enough---and has enough photometric information---to
-correct for extinction by dust without any
-input of explicit extinction or reddening estimates.
-We re-label each star in the sample with a new spectrophotometric parallax;
-these new parallaxes have uncertainties of a few to 15 percent, depending on data quality,
+% results
+The model implicitly performs a reddening and extinction correction in its parallax prediction,
+without any explicit dust model.
+We assign to each star in the sample a new spectrophotometric parallax estimate;
+these parallaxes have uncertainties of a few to 15 percent, depending on data quality,
 which is more precise than the \gaia\ parallax for the vast majority of targets.
-These distance estimates open up great opportunities for
-mapping the Milky Way disk with the next
-generation of spectroscopic surveys, and especially \sdssv.
+We can now get 10-percent distance estimates out to heliocentric distances of 20\,kpc,
+allowing truly global maps of the Milky Way's disk.
 \end{abstract}

### 2018-09-30 paper/sp.tex
upadte paper

@@
 % - HWR: Figure 7: Caption needs to say what arrows represent.
+% - Make the comparison with Cannon as a function of L1 strength and maybe look at correlation coefficients between parameter vectors.
+% - Get model running on ASPCAP outputs and compare in Discussion.
+% - Make sure we incorporate / respond to Ness comments from Aug 13 email.
+% - Thank APOGEE telecon and add paragraph about AGB stars and the color-magnitude diagram in the discussion.
+% - Describe how we ran The Cannon; discussion of how we compare to The Cannon.
+% - Cite and describe Eilers et al.
 
@@
 Here we use the \apogee--\gaia\ overlap to train a data-driven model
-to predict parallaxes for red-giant branch stars with $0<\logg<2.2$.
+to predict parallaxes for red-giant branch stars with $0<\logg\leq2.2$.
 % methods
@@
 to predict a parallax spectrophotometrically.
-The data-driven-model training involves no logarithms or inverses of the \gaia\ parallaxes,
+The model training involves no logarithms or inverses of the \gaia\ parallaxes,
 and needs no cut on the \gaia\ parallax signal-to-noise ratio.
@@
 As we operate these spectroscopic surveys,
-the \gaia\ Mission (\citealt{gaia}) has been revolutionizing our view of the disk.
+the \gaia\ Mission (\citealt{gaia}) has also been revolutionizing our view of the disk.
 It delivers good proper motions everywhere,
-and extremely valuable parallax information
-locally.
+and extremely valuable parallax information locally.
 However, it does not deliver precise geometric parallaxes over a large fraction of
@@
 
+ACE: ADD HERE the Color-magnitude diagrams for high snr and training set.
+HOGG: DIscussion here of CMD and AGB stars.
+
 What is this useful for and how do you use them safely?

### 2018-09-30 fit_rot_curve.py

@@ -19,6 +19,10 @@ import matplotlib.gridspec as gridspec
 import astropy
 import astropy.units as u
 import astropy.cosmology as cosmo
+import emcee
+import scipy
+import matplotlib.patches as patches
+
 
 planck = cosmo.Planck13
 
@@ -36,12 +40,15 @@ figsize = (8, 4.5)
 # load rotation curve
 # -------------------------------------------------------------------------------
 
-f = open('data/rot_curve.txt', 'r')
+f = open('data/rot_curve_part2.txt', 'r')
 xx = np.loadtxt(f)
 f.close()
 
 bins_dr = xx[0, :]
 vc_annulus = xx[1, :]
+vc_annulus_err_m = xx[2, :]
+vc_annulus_err_p = xx[3, :]
+sigmas = 0.5 * (vc_annulus_err_m + vc_annulus_err_p)
 
 idx5 = sum(bins_dr < 5)
 
@@ -49,19 +56,72 @@ idx5 = sum(bins_dr < 5)
 # profiles
 # -------------------------------------------------------------------------------
 
-## Miyamoto - Nagai model (bulge, thin, thick disk)
-#def vc_MN(R, a, b, M, z = 0):  
-#    vc = R*u.kpc * np.sqrt(astropy.constants.G.to(u.kpc**3/u.M_sun/u.s**2) * M*u.M_sun / ((R*u.kpc)**2 + (a*u.kpc + np.sqrt((z*u.kpc)**2 + (b*u.kpc)**2))**2)**(3./2.))
-#    return vc.to(u.km/u.s)
-#
-#def vc_total(theta, R):    
-#    a1, M1, a2, M2, b3, M3 = theta
-#    vc_thin_disk = vc_MN(R, a1, 0.25, M1*1e7) # b = 0: infinitely thin disk
-#    vc_thick_disk = vc_MN(R, a2, 0.8, M2*1e7)
-#    vc_halo = vc_MN(R, 0, b3, M3*1e7) # a = 0: spherical symmetry (Plummer's sphere)
-#    vc_tot = np.sqrt(vc_thin_disk**2 + vc_thick_disk**2 + vc_halo**2)    
-#    return vc_tot.value
+# Miyamoto - Nagai model (bulge, thin, thick disk)
+def vc_MN(R, a, b, M, z = 0):  
[…]

### 2018-09-30 rotation_curve_uncertainties.py

@@ -23,6 +23,8 @@ from mpl_toolkits.mplot3d import Axes3D
 import corner
 from scipy.stats import binned_statistic_2d
 from plotting_helpers import histcont
+import matplotlib.gridspec as gridspec
+
 
 # -------------------------------------------------------------------------------
 # plotting settings
@@ -72,97 +74,118 @@ N = len(labels) # 33044 stars!
 
 N_sample = 256
 np.random.seed(42)
-'''mean_XS_cart_n = np.zeros((N, 6))
-var_XS_cart_n = np.zeros((N, 3, 3))
-mean_XS_cyl_n = np.zeros((N, 6))
-var_XS_cyl_n = np.zeros((N, 3, 3))
-XS_cart_true_n = np.zeros((N, 6))
-XS_cyl_true_n = np.zeros((N, 6))
 
-# RADIAL VELOCITY UNCERTAINTY?
-floor_rv = 0.1 # km/s
-
-for i in range(23697, N):
-    
-    if i%1000 == 0: print('working on star {0} out of {1}'.format(i, N))
-    spec_par = np.random.normal(labels['spec_parallax'][i], scale = labels['spec_parallax_err'][i], size = N_sample) * u.mas
-    #assert np.all(spec_par > 0)
-    distance = spec_par.to(u.parsec, equivalencies = u.parallax())
-    distance_true = (labels['spec_parallax'][i] * u.mas).to(u.parsec, equivalencies = u.parallax())
-    
-    pmras = np.random.normal(labels['pmra'][i], scale = labels['pmra_error'][i], size = N_sample)
-    pmdecs = np.random.normal(labels['pmdec'][i], scale = labels['pmdec_error'][i], size = N_sample)
-    vrs = np.random.normal(labels['VHELIO_AVG'][i], scale = np.sqrt(floor_rv**2), size = N_sample)
-                             
-    # -------------------------------------------------------------------------------
-    # calculate cartesian coordinates
-    # -------------------------------------------------------------------------------           
-    
-    cs = coord.ICRS(ra = np.ones((N_sample)) * labels['ra'][i] * u.degree, 
[…]

### 2018-10-12 fit_rot_curve.py
new plot

@@ -40,7 +40,7 @@ figsize = (8, 4.5)
 # load rotation curve
 # -------------------------------------------------------------------------------
 
-f = open('data/rot_curve_part2.txt', 'r')
+f = open('data/rot_curve.txt', 'r')
 xx = np.loadtxt(f)
 f.close()
 
@@ -51,6 +51,18 @@ vc_annulus_err_p = xx[3, :]
 sigmas = 0.5 * (vc_annulus_err_m + vc_annulus_err_p)
 
 idx5 = sum(bins_dr < 5)
+
+f = open('data/rot_curve_part1.txt', 'r')
+xx1 = np.loadtxt(f)
+f.close()
+idx5a = sum(xx1[0, :] < 5)
+
+
+f = open('data/rot_curve_part2.txt', 'r')
+xx2 = np.loadtxt(f)
+f.close()
+idx5b = sum(xx2[0, :] < 5)
+
 
 # -------------------------------------------------------------------------------
 # profiles
@@ -329,17 +341,19 @@ lopez = Table.read('data/data_lopezcorredoira2014_cleaned.txt', format = 'ascii'
 draws = 100
 np.random.seed(42)
 # LINES IN DARK GREY/BLACK!
-fig, ax = plt.subplots(1, 1, figsize = (9.5, 7.2), sharex = True)        
+fig, ax = plt.subplots(1, 1, figsize = (9.3, 7.2), sharex = True)        
 plt.errorbar(bins_dr[idx5:], vc_annulus[idx5:], yerr = [vc_annulus_err_m[idx5:], vc_annulus_err_p[idx5:]], fmt = 'o', markersize = 8, capsize=4, mfc='k', mec='k', ecolor = 'k', zorder = 300, label = r'Eilers et al.\ 2018 (this work)')
+#plt.plot(xx1[0, idx5a:], xx1[1, idx5a:], 'o', markersize = 8, zorder = -300, color = '#D7D7D7')
+#plt.plot(xx2[0, idx5b:], xx2[1, idx5b:], 'o', markersize = 8, zorder = -300, color = '#D7D7D7')
 plt.ylim(0, 400) #310
 plt.xlim(0, 25.2)
 plt.xticks([0, 5, 10, 15, 20, 25])
 plt.tick_params(axis=u'both', direction='in', which='both', right = 'on', top = 'on')
-plt.xlabel(r'$R_{\rm GC}\,\rm [kpc]$', fontsize = fsize)
[…]

### 2018-10-12 rotation_curve_uncertainties.py

@@ -75,6 +75,27 @@ N = len(labels) # 33044 stars!
 N_sample = 256
 np.random.seed(42)
 
+# -------------------------------------------------------------------------------
+# new plot
+# -------------------------------------------------------------------------------           
+distance = (labels['spec_parallax'] * u.mas).to(u.parsec, equivalencies = u.parallax())
+
+fig, ax = plt.subplots(1, 1, figsize = (7, 6))        
+cm = plt.cm.get_cmap('viridis')
+sc = ax.scatter(labels['phot_g_mean_mag'], distance/1000., c = labels['parallax_error']/labels['spec_parallax_err'], s=20, cmap=cm, alpha = .8, vmin = 0, vmax = 20)
+fig.subplots_adjust(right = 0.8)
+cbar_ax = fig.add_axes([1, 0.05, 0.02, 0.92])
+cb = fig.colorbar(sc, cax=cbar_ax)
+cb.set_label(r'$\sigma_{\varpi}^{(a)}/\sigma_{\varpi}^{(sp)}$', rotation=90, fontsize=18, labelpad=30)
+plt.tight_layout()
+ax.set_xlabel(r'$G$ [mag]', fontsize = fsize)
+ax.set_ylabel(r'distance [kpc]', fontsize = fsize)
+ax.tick_params(axis=u'both', direction='in', which='both')
+ax.set_ylim(0, 65)
+plt.savefig('paper/precicion.pdf', bbox_inches = 'tight', pad_inches=.2)
+
+
+
 ## -------------------------------------------------------------------------------
 ## HRD plot
 ## -------------------------------------------------------------------------------           
@@ -271,7 +292,7 @@ Xlimits = [[-25.2, 10], [-12, 23], [-20, 20],
 deg_wedge_in_z = 6.
 cut_vz = (XS_cart_true_n[:, 5] < 100)
 wedge_z = (np.abs(mean_XS_cyl_n[:, 2])/(mean_XS_cyl_n[:, 0])) < np.tan(deg_wedge_in_z/360. * 2. * np.pi)
[…]

### 2018-10-13 paper/sp.tex
notes to self
new plot
quality cuts

@@
 % - Cite and describe Eilers et al.
+% - comment on parallax offset being a function of position and color
+% - remove Cannon plot for the moment
+% - comment on new plot of new precision on parallax compared to Gaia's precision
 
@@
 The Training-Set stars, in addition to meeting all Parent-Sample cuts,
-also must meet ACE WHAT ADDITIONAL CUTS to ensure that the parallax
+also must meet $\sigma_{\varpi^{(\rm a)}} < 0.1$, it must be observed by \gaia\ for more than $8$ individual times, i.e.
+\code{visibility_periods_used}~$\geq~8$, as well as meet the additional quality criterion $\code{astrometric_chi2_al}/\sqrt{\code{astrometric_n_good_obs_al}-5} \leq 35$, which eliminates detections that are not consistent with a single PSF \citep[see][]{calj}, to ensure that the parallax
 measurements are good.
@@
   data feature that might be relevant to the residual.\label{fig:residuals}}
+\end{figure}
+
+% HOGG: NEW FIGURE -- put some words in the text about the figure. 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{./precision.pdf}
+\caption{some smart words :) \label{fig:precision}}
 \end{figure}

### 2018-10-14 paper/sp.tex
clean up gaia quality cuts

@@
 % - Get model running on ASPCAP outputs and compare in Discussion.
-% - Make sure we incorporate / respond to Ness comments from Aug 13 email.
+% - Make sure we incorporate / respond to Ness comments from Aug 13 email. And thank Ness.
 % - Thank APOGEE telecon and add paragraph about AGB stars and the color-magnitude diagram in the discussion.
@@
 % - comment on new plot of new precision on parallax compared to Gaia's precision
+% - Get figure showing how much better we are than Gaia at large distance and as a function of distance. And write a paragraph about said figure.
 
@@
 \chi^2(\theta) &\equiv& \sum_{n=1}^N \frac{[\gparallax_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
-\quad ,
 \end{eqnarray}
@@
 \hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2(\theta) + \lambda\,||P\cdot\theta||_1^1\right]
-\label{eq:objective}\quad ,
+\label{eq:objective}
 \end{eqnarray}
@@
 \sparallax_m &\leftarrow& \exp(\hat{\theta}\cdot x_m)
-\quad ,
 \end{eqnarray}
@@
 The Training-Set stars, in addition to meeting all Parent-Sample cuts,
-also must meet $\sigma_{\varpi^{(\rm a)}} < 0.1$, it must be observed by \gaia\ for more than $8$ individual times, i.e.
-\code{visibility_periods_used}~$\geq~8$, as well as meet the additional quality criterion $\code{astrometric_chi2_al}/\sqrt{\code{astrometric_n_good_obs_al}-5} \leq 35$, which eliminates detections that are not consistent with a single PSF \citep[see][]{calj}, to ensure that the parallax
-measurements are good.
+also must meet $\sigma_{\varpi^{(\rm a)}} < 0.1$, it must be observed by
+more than $8$ (widely separated) times
+\begin{equation}
+\code{visibility_periods_used >= 8}
+\end{equation}
+as well as meet the additional quality criterion
+\begin{equation}
+\code{astrometric_chi2_al / sqrt(astrometric_n_good_obs_al - 5) <= 35}
+\end{equation}
+which eliminates detections that are not consistent with a single PSF \citep[see][]{calj},
+to ensure that the parallax measurements are good.
 These cuts leave 28226 stars total in the Training Sets.
 But---as we have emphasized above---we do not cut ever on parallax or
-parallax error or their ratio. For this reason, most of the training set
+parallax error or their ratio. For this reason, most of the Training-Set stars
 do not have significantly measured parallaxes and more than 2~percent even have
@@
 x_n\T &\equiv& [1, G_n, \BP_n, \RP_n, J_n, H_n, K_n, W_{1n}, W_{2n}, \ln f_{1n}, \ln f_{2n}, \cdots, \ln f_{Ln}]
-\quad ,
 \end{eqnarray}

### 2018-10-14 paper/sp.tex
hogg edits

@@
 % - comment on new plot of new precision on parallax compared to Gaia's precision
-% - Get figure showing how much better we are than Gaia at large distance and as a function of distance. And write a paragraph about said figure.
+% - write caption and paragraph about the new figure showing how much better we are than Gaia at large distance
+% - need a snippet for Table 1.
 
@@
 \newcommand{\RP}{{G_\mathrm{RP}}}
-\newcommand{\gparallax}{\varpi^{(\mathrm{a})}}
-\newcommand{\sparallax}{\varpi^{(\mathrm{sp})}}
+\newcommand{\gparallax}{{\varpi^{(\mathrm{a})}}}
+\newcommand{\sparallax}{{\varpi^{(\mathrm{sp})}}}
+\newcommand{\gsigma}{{\sigma^{(\mathrm{a})}}}
+\newcommand{\ssigma}{{\sigma^{(\mathrm{sp})}}}
 
@@
 red-giant stars can be observed to distances and extinctions
-at which \gaia\ parallaxes are not informative.
+at which \gaia\ parallaxes are not highly informative.
 Yet the combination of effective temperature, surface gravity, composition, and age---all
@@
 % aims
-Here we use the \apogee--\gaia\ overlap to train a data-driven model
+Here we use the \apogee--\gaia--\zmass--\wise\ overlap to train a data-driven model
 to predict parallaxes for red-giant branch stars with $0<\logg\leq2.2$.
@@
 a linear function of \apogee\ spectral pixel intensities and multi-band photometry
-to predict a parallax spectrophotometrically.
+to predict parallax spectrophotometrically.
 The model training involves no logarithms or inverses of the \gaia\ parallaxes,
@@
 We can now get 10-percent distance estimates out to heliocentric distances of 20\,kpc,
-allowing truly global maps of the Milky Way's disk.
+and make global maps of the Milky Way's disk.
 \end{abstract}
@@
 The maps made with these data will reveal the dynamics of the Milky Way and its disk,
-and show us how and where stars form, and how they migrate around (and out of) the
+and show us how and where stars form, and how they migrate around the
 disk with time.
@@
 Luminous red-giant stars are valuable for mapping the Milky Way disk for a number
-lf reasons, only one of which is their great luminosities (and hence brightnesses
+of reasons, only one of which is their great luminosities (and hence brightnesses
 even at large distance).
@@
 their luminosities and use them as standard candles.
-That is---like type Ia supernovae---red giants are standardizable candles.
+That is, red giants are standardizable candles.
 
@@
 for this kind of work, because they have the flexibility to discover
-regularities or offsets in the data that are not captured by the physical
+regularities or offsets in the data that are not adequately captured by the physical
 models.
@@
 to interpret.
+It is also the case that linear models extrapolate better than higher-order or more
+flexible models; this is not extremely important to the present work, but it does
+protect our predictions at the edges (in terms of temperature, surface gravity,
+and chemical abundances) of our sample.
 
@@
 stochastic process, in which the difference between the \gaia-Catalog parallax (plus
-a small \gaia-recommended offset; \citealt{lindegren}) and the true parallax is effectively drawn from a
+a small offset, which we learn below) and the true parallax is effectively drawn from a
 Gaussian with a width set by the \gaia-Catalog uncertainty on the parallax.
@@
 \\
-\chi^2(\theta) &\equiv& \sum_{n=1}^N \frac{[\gparallax_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
+\chi^2(\theta) &\equiv& \sum_{n=1}^N \frac{[\gparallax_n - \exp(\theta\cdot x_n)]^2}{\gsigma_n^2}
 \end{eqnarray}
@@
 $\theta$ is a $D$-vector of linear coefficients,
-$\sigma_n$ is the \gaia\ estimate of the uncertainty on the parallax measurement,
+$\gsigma_n$ is the \gaia\ estimate of the uncertainty on the parallax measurement,
 and $\chi^2(\theta)$ is (twice) the negative-log-likelihood for the parameters $\theta$
@@
 (the ``sparsity'' assumption).
-In order to encourage this, we optimize not $\chi^2$ but a regularized objective function
+In order to represent this expectation,
+we optimize not $\chi^2$ but a regularized objective function
 \begin{eqnarray}
@@
 $x_m$ is the feature $D$-vector for star $m$.
-Thus the model can be trained on a training set of stars and applied to
+The model can be trained on a training set of stars and applied to
 a validation set or a test set to make predictions.
@@
 vector $x_n$.
+
+The linear model permits straightforward propagation of uncertainty from the
+feature inputs to the spectrophotometric parallaxes.
+Although the model fundamentally assumes that the prediction features are noise-free
+(see the ``good features'' assumption in \sectionname~\ref{sec:assumptions}),
+there are in fact small uncertainties on the features that we can propagate to 
+make uncertainty estimates on the spectrophotometric parallaxes.
+\begin{eqnarray}
+\ssigma_m^2 &\leftarrow& \sparallax_m^2\,\hat{\theta}^T\cdot C_m\cdot\hat{\theta}
+\end{eqnarray}
+where $\ssigma_m$ is the uncertainty on the inferred (or predicted)
+spectrophotometric parallax $\sparallax_m$,
+the left-hand side is a scalar product,
+and $C_m$ is the (presumed diagonal, actually) covariance matrix of the input
+features.
+That is, $C_m$ is a diagonal matrix with the uncertainty variances of the elements
+of $x_m$ along the diagonal.
 
@@
 according to the \gaia\ Archive internal match criteria.
-This gives, fore each matching star $n$,
+This gives, for each matching star $n$,
 mid-infrared photometry $W_{1n}$ and $W_{2n}$ at 3.6 and 4.5\,$\mu$m
@@
 The Training-Set stars, in addition to meeting all Parent-Sample cuts,
-also must meet $\sigma_{\varpi^{(\rm a)}} < 0.1$, it must be observed by
-more than $8$ (widely separated) times
+also must meet an uncertainty criterion $\gsigma_n < 0.1\,\mathrm{mas}$,
+\begin{equation}
+\code{parallax_error < 0.1}
+\end{equation}
+it must be observed by more than $8$ (widely separated) times
 \begin{equation}
@@
 But---as we have emphasized above---we do not cut ever on parallax or
-parallax error or their ratio. For this reason, most of the Training-Set stars
+or the ratio to its uncertainty. For this reason, most of the Training-Set stars
 do not have significantly measured parallaxes and more than 2~percent even have
@@
 for each star $n$ in the Training Sets, we take from \gaia\ \acronym{DR2} (\citealt{gaiadr2})
-the astrometric parallax $\gparallax_n$ and its uncertainty $\sigma_n$.
+the astrometric parallax $\gparallax_n$ and its uncertainty $\gsigma_n$.
 Importantly, to each \gaia\ parallax measurement we add a positive
@@
 optimizes our cross-validation objective.
-The spectrophotometric parallaxes do not depend strongly on this choice, except
+HOGG CITE STASSUN MAYBE AND OTHERS?
+The spectrophotometric parallaxes we generate
+do not depend strongly on this choice, except
 for small differences at very small parallax (large distance).
@@
 addition to these feature vectors---a \gaia-measured astrometric parallax $\gparallax_n$
-and uncertainty $\sigma_n$.
+and uncertainty $\gsigma_n$.
 
@@
 
-This 9-percent precision estimate is conservative in the sense that it does not deconvolve
-the contributions from the \gaia\ noise model.
+This 9-percent precision estimate is conservative in the sense that it does not
+deconvolve or correct for the contributions from the \gaia\ noise.
 However it is also not fully conservative, in that the test is performed with
@@
 \includegraphics[width=\textwidth]{./residuals_training.pdf}
-\caption{Prediction residuals colored by relevant features.
-  Each panel shows the prediction residuals (\gaia\ astrometric parallax minus
-  spectrophotometric parallax $\gparallax_n-\sparallax_n$) colored by a different
+\caption{Prediction validation colored by relevant features.
+  Each panel shows the predicted parallax \foreign{vs} the
+  spectrophotometric parallax colored by a different
   data feature that might be relevant to the residual.\label{fig:residuals}}
@@
 The case of cluster \acronym{M71} is troubling, but after searching for trends with
-housekeeping data (also see \figurename~\ref{fig:residuals}) we don't find any correlates
+housekeeping data (also see \figurename~\ref{fig:residuals}) we don't find any correlations
 with parallax offsets.
 \begin{figure}
-\includegraphics[width=\textwidth]{clusters.pdf}
-\caption{The histograms of \gaia\ astrometric parallaxes $\gparallax_n$
+\includegraphics[width=0.7\textwidth]{clusters.pdf}
+\caption{The \gaia\ astrometric parallaxes $\gparallax_n$
   and spectrophotometric parallaxes $\sparallax_n$ for member stars for
@@
   and their inverse-variance-weighted mean
-  the expected parallax of the cluster is shown as a vertical line.
+  are shown as vertical lines.
   Cluster members were identified to be close to the cluster in celestial coordinates
@@
 
-In the end, this method produces a spectrophotometric parallax estimate $\sparallax_m$ for
-every star $m$.
+In the end, this method produces a spectrophotometric parallax estimate $\sparallax_m$
+and uncertainty $\ssigma_m$ for every star $m$.
 \tablename~\ref{tab:output} shows a snippet of this output, with the full output available
@@
 \end{table}
-
-The linear model permits straightforward propagation of uncertainty from the
-feature inputs to the spectrophotometric parallaxes.
-Although the model fundamentally assumes that the prediction features are noise-free
-(see the ``good features'' assumption in \sectionname~\ref{sec:assumptions}),
-there are in fact small uncertainties on the features that we can propagate to 
-make uncertainty estimates on the spectrophotometric parallaxes.
-We perform this propagation; these uncertainties are also given in
-\tablename~\ref{tab:output}.
 
@@
 build a predictive model for distance.
-Our arguments for our included features is that they ought to be sufficient to
+Our argument for our included features is that they ought to be sufficient to
 estimate a dust-independent apparent magnitude and the stellar parameters (especially

### 2018-10-14 paper/sp.tex
remaining to-dos all in-text

@@
 % - HWR: Footnote to explain the use of the word ``feature'' for astronomers.
-% - HWR: Figures 1 and 2 need to be combined to directly compare Parent and Training.
 % - HWR: Figure 3: Note that the y-axis is positive-definite!
 % - HWR: Figure 7: Caption needs to say what arrows represent.
-% - Make the comparison with Cannon as a function of L1 strength and maybe look at correlation coefficients between parameter vectors.
-% - Get model running on ASPCAP outputs and compare in Discussion.
-% - Make sure we incorporate / respond to Ness comments from Aug 13 email. And thank Ness.
-% - Thank APOGEE telecon and add paragraph about AGB stars and the color-magnitude diagram in the discussion.
-% - Describe how we ran The Cannon; discussion of how we compare to The Cannon.
+% - Make sure we incorporate / respond to Ness comments from Aug 13 email.
+% - add paragraph about AGB stars and the color-magnitude diagram in the discussion.
 % - Cite and describe Eilers et al.
 % - comment on parallax offset being a function of position and color
-% - remove Cannon plot for the moment
+% - remove Cannon plot and all references for the moment
 % - comment on new plot of new precision on parallax compared to Gaia's precision
@@
 \ssigma_m^2 &\leftarrow& \sparallax_m^2\,\hat{\theta}^T\cdot C_m\cdot\hat{\theta}
+\label{eq:unc}
 \end{eqnarray}
@@
 These cuts removed roughly 2~percent of the \apogee\ stars.
-This Parent Sample contains 44784 stars and is shown in \figurename~\ref{fig:parent}.
+This Parent Sample contains 44784 stars and is shown in \figurename~\ref{fig:samples}.
 \begin{figure}
-\includegraphics[width=\textwidth]{parent_sample.pdf}
-\caption{Color--color diagrams for the 44784 Parent Sample stars.
+\includegraphics[width=\textwidth]{./parent_sample.pdf}\\
+\includegraphics[width=\textwidth]{./training_sample.pdf}
+\caption{\textsl{top:}
+  Color--color diagrams for the 44784 Parent Sample stars.
   The points are colored by
   the \apogee\ pipeline estimates of surface gravity.
-  The color cuts used to trim outliers are shown as grey lines.\label{fig:parent}}
+  The color cuts used to trim outliers are shown as grey lines.
+  \textsl{bottom:}
+  The same but for the 28226 Training Set stars.
+  Note that there is less color range in the Training Set than in the
+  Parent Sample.\label{fig:samples}}
 \end{figure}
@@
 for small differences at very small parallax (large distance).
-The Training Sets are shown in \figurename~\ref{fig:training}.
-\begin{figure}
-\includegraphics[width=\textwidth]{./training_sample.pdf}
-\caption{Color--color diagrams for the 28226 Training Set stars.
-  Note that there is less color range in the Training Set than in the
-  Parent Sample.\label{fig:training}}
-\end{figure}
+The Training Sets are shown in the bottom panels of \figurename~\ref{fig:samples}.
 
-Comparison of {\figurename s}~\ref{fig:parent} and \ref{fig:training} shows
+Comparison of the top and bottom panels of {\figurename}~\ref{fig:samples} shows
 that there is more color range in the Parent Sample than in the Training Set.
@@
 Our feature decisions were highly motivated by these kinds of considerations.
-And the derived model parameters shown in \figurename~\ref{fig:spectral} do indicate that
-the model is capturing some or all of the expected information in the spectra, at least.
 
@@
 The answer to the first is that we have taken such a small part of stellar parameter
-space with our surface gravity cut, that the linearized model of stellar
+space with our surface-gravity cut, that the linearized model of stellar
 spectrophotometry is not a bad approximation.
@@
 balanced in terms of parallax and dust attenuation.
-And finally, the complex model would have been harder to visualize and interpret;
+In the  linear model it is also---unlike in more flexible models---trivial to propagate
+uncertainties from the feature space to the prediction.
+We used that, above, to propagate uncertainty in \equationname~(\ref{eq:unc}).
+And finally---in principle---linear models are better for visualization and interpretation;
 the linear model contains essentially only first derivatives in the data space, which are

### 2018-10-15 paper/sp.tex
figure tweaks
references

@@
 Optimization is performed with \code{scipy.optimize} using the
-\acronym{L-BFGS-B} algorithm (HOGG CITE).
+\acronym{L-BFGS-B} algorithm (\citealt{lbfgsb}).
 
@@
 spectrophotometric parallax $\sparallax_m$,
-the left-hand side is a scalar product,
-and $C_m$ is the (presumed diagonal, actually) covariance matrix of the input
+the right-hand side is a scalar product,
+and $C_m$ is the covariance matrix of the input
 features.
-That is, $C_m$ is a diagonal matrix with the uncertainty variances of the elements
+In what follows we don't actually instantiate the full covariance matrix; we presume
+that input features are independently measured;
+that is, we make $C_m$ is a diagonal matrix with the uncertainty variances of the elements
 of $x_m$ along the diagonal.
@@
 optimizes our cross-validation objective.
-HOGG CITE STASSUN MAYBE AND OTHERS?
+It also happens to be in the range of other values found in the literature (\citealt{arenou, zinn}).
+Of course we really expect the offset to be a function of sky position and color (at least),
+so the offset we find is not recommended for further use as any kind of universal value.
 The spectrophotometric parallaxes we generate
@@
 \centering
-\includegraphics[width=\textwidth]{./precision.pdf}
+~\hfill\includegraphics[width=0.8\textwidth]{./precision.pdf}\hfill~
 \caption{some smart words :) \label{fig:precision}}
@@
 \begin{figure}
-\includegraphics[width=0.7\textwidth]{clusters.pdf}
+~\hfill\includegraphics[width=0.7\textwidth]{clusters.pdf}\hfill~
 \caption{The \gaia\ astrometric parallaxes $\gparallax_n$
@@
   Melissa Ness (Columbia),
+  Joel Zinn (OSU),
   and the participants in the \acronym{MPIA} Milky Way Group Meeting
@@
 \bibitem[Allende Prieto \etal(2008)]{aapogee} Allende Prieto, C., Majewski, S.~R., Schiavon, R., \etal\ 2008, Astronomische Nachrichten, 329, 1018 
+\bibitem[Arenou \etal(2018)]{arenou} Arenou, F., Luri, X., Babusiaux, C., et al.\ 2018, \aap, 616, A17 
 \bibitem[Astropy Collaboration \etal(2013)]{astropy} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., \etal\ 2013, \aap, 558, A33 
@@
 \bibitem[Blanton \etal(2017)]{sdssiv} Blanton, M.~R., Bershady, M.~A., Abolfathi, B., \etal\ 2017, \aj, 154, 28 
+\bibitem[Byrd \etal(1995)]{lbfgsb} Byrd, R.~H., Lu,~P., Nocedal,~J., \& Zhu,~C. 1995, SIAM J. Sci. Comput., 16, 1190
 \bibitem[Casey \etal(2016)]{casey} Casey, A.~R., Hogg, D.~W., Ness, M., \etal\ 2016, arXiv:1603.03040 
@@
 \bibitem[Wright \etal(2010)]{wise} Wright, E.~L., Eisenhardt, P.~R.~M., Mainzer, A.~K., \etal\ 2010, \aj, 140, 1868 
+\bibitem[Zinn \etal(2018)]{zinn} Zinn, J.~C., Pinsonneault, M.~H., Huber, D., \& Stello, D.\ 2018, arXiv:1805.02650 
 \end{thebibliography}

### 2018-10-15 paper/sp.tex
get rid of interpretation parts

@@
 
-The linear model we use has a big limitation: Linear models are inflexible.
-However, it has a great advantage over more general methods,
-which is that it is easy to look inside
-the linear model and check whether the dependencies represented there make sense.
-The feature vector $x_n$ for star $n$
-contains a set of photometry in magnitudes.
-Linear combinations of these photometric measurements are like complex synthetic
-magnitudes or colors.
-Since the model was trained to deliver luminosities of stars given training data
-of different metallicities and extinctions, the synthetic photometry developed
-by the model should be, in some sense, metallicity- and extinction-independent
-photometry for luminous red-giant stars!
-In detail this synthetic magnitude is something like HOGG PUT EQUATION HERE AND
-SAY SOMETHING ABOUT IT. HOGG: Maybe plot this synthetic photometry against dust
-and metallicity.
-
-HOGG: Similarly, the spectral part of the model should also be
-interpretable. HOGG DESCRIBE \figurename~\ref{fig:spectral} THAT
-DEMONSTRATES THIS, qualitatively.
-\begin{figure}
-\includegraphics[width=\textwidth]{coefficients.pdf}
-\caption{HOGG: THIS FIGURE ISN'T READY YET:
-  The coefficients of the spectral part of the linear model.
-  Also shown are derivatives with respect to labels for a \project{Cannon}-like
-  linear regression for the Training Set data.\label{fig:spectral}}
-\end{figure}
-
 In the end, this method produces a spectrophotometric parallax estimate $\sparallax_m$
@@
 relatively easy to understand.
+
+Another  advantage of a linear model over more general methods (like deep learning, say),
+is that it is possible to look inside
+the linear model and check whether the dependencies represented there make sense.
+For example, the feature vector $x_n$ for star $n$
+contains a set of photometry in magnitudes.
+Linear combinations of these photometric measurements are like complex synthetic
+magnitudes or colors.
+Similarly, linear combinations of spectral pixels are like a projection of the spectral space.
+Brief inspection of these linear combinations showed them to be sensible, but it is out of
+scope here to interpret them in detail.
 

### 2018-10-15 paper/sp.tex
table

@@
 \begin{table}
-\caption{HOGG TBD: The generated spectrophotometric parallaxes.\label{tab:output}}
+\caption{HOGG and ACE TBD: The generated spectrophotometric parallaxes.
+This table is just a snippet; the full table is available online (HOGG HOW?)\label{tab:output}}
 \end{table}
@@
   and the participants in the \acronym{MPIA} Milky Way Group Meeting
+  and the \apogee\ Science Telecon,
 for help with this project.

### 2018-10-15 paper/sp.tex
precision plot

@@
 
-% HOGG: NEW FIGURE -- put some words in the text about the figure. 
-\begin{figure}
-\centering
-~\hfill\includegraphics[width=0.8\textwidth]{./precision.pdf}\hfill~
-\caption{some smart words :) \label{fig:precision}}
-\end{figure}
-
 Another validation of the results can be obtained by looking at known
@@
   and comoving in proper motion.\label{fig:clusters}}
+\end{figure}
+
+It is interesting and valuable to compare the spectrophotometric parallax precision to
+the astrometric parallax precision.
+\figurename~\ref{fig:precision} compares uncertainties between the spectrophotometric
+outputs and the astrometric inputs
+(and recall that information is proportional to the inverse square of the uncertainty).
+The brief summary is that the spectrophotometric parallaxes are more precise (and even
+far more precise) than the astrometric parallaxes for stars further than a few kpc from
+the Sun.
+This is not unexpected; \gaia\ parallax uncertainties are in the 0.1-mas range, whereas
+our uncertainties are in the ten-percent range.
+\begin{figure}
+\centering
+~\hfill\includegraphics[width=0.8\textwidth]{./precision.pdf}\hfill~
+\caption{Comparison of precisions between the spectrophotometric parallaxes and
+the photometric parallaxes, as a function of magnitude and heliocentric distance.
+Note that the spectrophotometric estimates are more precise for most stars.
+The information content of a parallax estimate (or any measurement)
+is proportional to the inverse-square of the uncertainty.\label{fig:precision}}
 \end{figure}

### 2018-10-15 paper/sp.tex
more to-do items

@@
 % -----------
-% - Coryn's point that we should carefully explain how our model is geometric.
+% - need a snippet for Table 1.
 % - check with WISE experts that we have the correct WISE photometry
-% - write words about other linear models we have made or could make, probably in the discusssion.
-% - are we citing all the proper DR2 papers we need to cite?
-% - HWR: Footnote to explain the use of the word ``feature'' for astronomers.
-% - HWR: Figure 3: Note that the y-axis is positive-definite!
 % - HWR: Figure 7: Caption needs to say what arrows represent.
 % - Make sure we incorporate / respond to Ness comments from Aug 13 email.
-% - add paragraph about AGB stars and the color-magnitude diagram in the discussion.
 % - Cite and describe Eilers et al.
-% - comment on parallax offset being a function of position and color
-% - remove Cannon plot and all references for the moment
-% - comment on new plot of new precision on parallax compared to Gaia's precision
-% - write caption and paragraph about the new figure showing how much better we are than Gaia at large distance
-% - need a snippet for Table 1.
 
@@
 these parallaxes have uncertainties of a few to 15 percent, depending on data quality,
-which is more precise than the \gaia\ parallax for the vast majority of targets.
-We can now get 10-percent distance estimates out to heliocentric distances of 20\,kpc,
+which is more precise than the \gaia\ parallax for the vast majority of targets, and
+certainly any stars more than a few kpc distance.
+We obtain 10-percent distance estimates out to heliocentric distances of 20\,kpc,
 and make global maps of the Milky Way's disk.
@@
 \begin{description}
-\item[features] Perhaps our most fundamental assumption is that the parallax
+\item[features]\footnote{Here and everywhere in this \documentname, we
+  use the word ``feature'' to refer to an input or predictor or
+  independent variable for the regression.
+  That is, our parallax predictions will be linear combinations of features.
+  This is not common terminology in astrophysics, but it is standard in the machine-learning literature.}
+Perhaps our most fundamental assumption is that the parallax
 of a star can be predicted from the features we provide, which are
@@
   assess the prediction precision. The fractional precision of the prediction in the
-  right panel is better than 9 percent.\label{fig:xval}}
+  right panel is better than 9 percent. Note that by construction, there can be no negative
+  spectrophotometric parallaxes; that is, the quantity plotted on the vertical axis is
+  positive-definite.\label{fig:xval}}
 \end{figure}
@@
   spectrophotometric parallax colored by a different
-  data feature that might be relevant to the residual.\label{fig:residuals}}
+  data feature that might be relevant to the residual.
+  Again the quantity plotted on vertical axis is positive-definite.\label{fig:residuals}}
 \end{figure}

### 2018-10-15 paper/sp.tex
first draft

@@
 
-HOGG: CRITICISE OUR NOISE ESTIMATION, in two parts:
-HOGG: TURN THIS WORD SALAD INTO A PARAGRAPH: we have shown that outliers in sp - gaia tend to have large $H-W_2$ or excess flux like in crowded regions, but there are some exceptions.
-Our error estimates are consistent with no intrinsic scatter if we use a robust chi-squared estimate.
-That is, chi-squared is dominated by outliers.
-We clipped the bad apogee spectral pixels at an error of 0.05 in normalized flux.
-Our uncertainty estimates are sensitive to this.
-The reader is asked to suck it.
-Combine with M71 discussion and write a paragraph.
-HOGG SOMETHING ABOUT THE CLUSTER M71 BEING CONCERNING AND OUR EXPLANATIONS.
+Our uncertainty estimation is very naive; it presumes that the features are measured
+with no intrinsic correlations (probably not true for neighboring pixels in the \apogee\ spectra,
+for example) and it does not allow for any biases in the feature inputs.
+What's shown in \figurename~\ref{fig:residuals} is that the biggest deviations are for
+stars with large (red) $H-W_2$ colors.
+These are in dusty and crowded regions; we expect that is the crowding that's causing the
+greatest problem, because both \gaia\ and \wise\ (and probably also \zmass) photometry is
+affected by overlapping and blended sources.
+
+We validated the uncertainty estimates by looking at the chi-squared statistic and robust
+variations of it.
+We find that chi-squared is larger than expected if the errors are treated naively.
+However, robust estimates show that this is driven by outliers;
+median absolute differences are consistent with our expectations under this very naive noise model.
+We also find that the outliers that drive up the chi-squared statistic tend to be red
+in $H-W_2$.
+
+\figurename~\ref{fig:clusters} suggests that there may be some hidden biases in
+the results; the cluster M71 looks inconsistent between the \gaia\ and the spectrophotometric
+parallaxes.
+We have looked at colors, sky position, and cluster properties, and we don't have any
+simple explanation for the discrepancy.
+This is left as a caveat for the reader and our users.
 
@@
 
-HOGG: ADD HERE a couple of sentences in the discussion about the setting of Lambda and the parallax offset (currently $0.0483\pm 0.012$\,mas).
-
 ACE: ADD HERE the Color-magnitude diagrams for high snr and training set.
-HOGG: DIscussion here of CMD and AGB stars.
-
-What is this useful for and how do you use them safely?
-HOGG: Show a map of the kinematic rotation of the Galaxy as a function of
-x and y, colored by metallicity in \figurename~\ref{fig:disk}.
-We see the kinematic center of the Galaxy! We see the rotation curve.
-We see the chemical gradients. HOGG WHY CAN YOU NOT JUST READ OFF THINGS FROM THIS PLOT?
+HOGG: Discussion here of CMD and AGB stars.
+
+This \documentname\ and project delivers greatly improved distance estimates for
+luminous red-giant stars.
+The purpose of all this is to make global maps of the Milky Way.
+Just as a demonstration, we show kinematics and element abundances for stars as a function
+of position in the Milky Way disk in \figurename~\ref{fig:disk}.
+In this visualization, it is possible to see the rotation of the disk and the
+kinematic center of the Galaxy; that might even provide a distance estimate to the Galactic
+Center.
+It also shows extremely strong chemical gradients.
 \begin{figure}
@@
 \end{figure}
-
-\figurename~\ref{fig:disk} shows that this sample of spectrophotometric parallaxes is useful for
-mapping the Milky Way.
-How does an investigator use these spectrophotometric parallaxes wisely?
-How do you build a proper probabilistic model using these?
-The answer is similar to the answer to the same question for the \gaia\ \acronym{DR2} data
-themselves (\citealt{gaialf}): A likelihood function must be constructed for each of the
-spectrophotometric parallaxes, given assumptions about the noise in those values.
-The sensible likelihood function for star $m$ would be HOGG PUT HERE.
-Model fitting can then be built on those likelihood functions.
+In a companion paper (\citealt{eilers}), we are using this information to estimate
+the (asymmetric-drift-corrected) circular-velocity curve for the Milky Way.
+
+Finally, one comment on the use of these spectroscopic parallaxes.
+They are noisy.
+Precise use of these involves building a likelihood function and performing inference,
+just as the precise use of the \gaia\ data presented here required the same.
+We recommend making analogous inferences when using these outputs.
 
@@
 \bibitem[Casey \etal(2016)]{casey} Casey, A.~R., Hogg, D.~W., Ness, M., \etal\ 2016, arXiv:1603.03040 
+\bibitem[Eilers \etal(2018)]{eilers} Eilers,~A.~C., Hogg, D.~W., Rix, H.-W., \& Ness, M. 2018, in preparation
 \bibitem[Gaia Collaboration \etal(2016)]{gaia} Gaia Collaboration, Prusti, T., de Bruijne, J.~H.~J., \etal\ 2016, \aap, 595, A1 

### 2018-10-15 paper/sp.tex
todo

@@
 % - check with WISE experts that we have the correct WISE photometry
-% - HWR: Figure 7: Caption needs to say what arrows represent.
-% - Make sure we incorporate / respond to Ness comments from Aug 13 email.
-% - Cite and describe Eilers et al.
+% - ACE: What fraction of spectral pixels are (typically, or in each training set) zeroed out?
 
@@
 astrometrically confirmed members of known stellar clusters.
+The cluster members were found by hand, looking at sky and proper-motion distributions
+centered on literature values.
+In each case we are plotting only very securely identified cluster members; that
+is, membership is conservative.
 These clusters span some range in metallicity and age, and therefore test
@@
 \caption{A map of the kinematics and element abundances in the Milky Way disk.
+  Each arrow represents the three-space Galactocentric velocity of each star
+  in the sample, projected onto the Galactic disk plane.
+  To make these velocities, raw \gaia\ proper motions and \apogee\ radial velocities
+  were used (in a standard Galactocentric transformation), so they are
+  not in any way noise-deconvolved.
+  Each arrow has a color corresponding to the \apogee\ reported metallicity.
   This map shows the rotation of the Galaxy out to large radius, and also
   indicates the kinematic center of the disk rotation. Interpretation of
-  the quantitative information in this plot requires care (see text).\label{fig:disk}}
+  the quantitative information in this plot requires care.\label{fig:disk}}
 \end{figure}

### 2018-10-17 paper/sp.tex
CMD

@@
 The Training-Set stars, in addition to meeting all Parent-Sample cuts,
-also must meet an uncertainty criterion $\gsigma_n < 0.1\,\mathrm{mas}$,
+also must meet an uncertainty criterion $\gsigma_n < 0.1\,\mathrm{mas}$, i.e. 
 \begin{equation}
@@
 
-Finally, one comment on the use of these spectroscopic parallaxes.
+Finally, one comment on the use of these spectrophotometric parallaxes.
 They are noisy.

### 2018-10-17 make_figures_HER.py

@@ -160,17 +160,25 @@ plt.savefig('paper/residuals.pdf', pad_inches=.2, bbox_inches = 'tight')
 
 fig, ax = plt.subplots(1, 4, figsize = (16, 4.5))
 sc = ax[0].scatter(labels['parallax'][train], labels['spec_parallax'][train], c = labels['TEFF'][train], cmap = 'viridis_r', s = 10, vmin = 3600, vmax = 4800, rasterized = True)
-cb = fig.colorbar(sc, ax = ax[0], shrink = .8)
+cb = fig.colorbar(sc, ax = ax[0], shrink = .6)
 cb.set_label(r'$T_{\rm eff}$', fontsize = fsize)
 sc = ax[1].scatter(labels['parallax'][train], labels['spec_parallax'][train], c = labels['LOGG'][train], cmap = 'viridis_r', s = 10, vmin = 0.1, vmax = 2.2, rasterized = True)
-cb = fig.colorbar(sc, ax = ax[1], shrink = .8)
+cb = fig.colorbar(sc, ax = ax[1], shrink = .6)
 cb.set_label(r'$\log g$', fontsize = fsize)
 sc = ax[2].scatter(labels['parallax'][train], labels['spec_parallax'][train], c = labels['FE_H'][train], cmap = 'viridis_r', s = 10, vmin = -1, vmax = .5, rasterized = True)
-cb = fig.colorbar(sc, ax = ax[2], shrink = .8)
+cb = fig.colorbar(sc, ax = ax[2], shrink = .6)
 cb.set_label(r'$\rm [Fe/H]$', fontsize = fsize)
 sc = ax[3].scatter(labels['parallax'][train], labels['spec_parallax'][train], c = labels['H'][train] - labels['w2mpro'][train], cmap = 'viridis_r', s = 10, vmin = 0., vmax = 0.5, rasterized = True)
-cb = fig.colorbar(sc, ax = ax[3], shrink = .8)
+cb = fig.colorbar(sc, ax = ax[3], shrink = .6)
 cb.set_label(r'$\rm H-W_{2}$', fontsize = fsize)
+ax[0].set_xlabel(r'$\varpi^{\rm (a)}$', fontsize = fsize)
+ax[0].set_ylabel(r'$\varpi^{\rm (sp)}$', fontsize = fsize)
+ax[1].set_xlabel(r'$\varpi^{\rm (a)}$', fontsize = fsize)
[…]

### 2018-10-17 test_distances.py

@@ -35,10 +35,10 @@ fsize = 14
 # open inferred labels
 # -------------------------------------------------------------------------------
 
-N = 44808 #45787
+N = 44784
 Kfold = 2
 lam = 30
-name = 'N{0}_lam{1}_K{2}_offset0.049_parallax'.format(N, lam, Kfold)
+name = 'N{0}_lam{1}_K{2}_offset0.0483_parallax'.format(N, lam, Kfold)
 
 print('loading new labels...')   
 labels = Table.read('data/training_labels_new_{}.fits'.format(name), format = 'fits')    
@@ -309,7 +309,7 @@ labels_sgr = labels[xx]
 # plot for paper (3x2) histograms: Gaia vs. me (labeled by cluster name and metallicity)
 # -------------------------------------------------------------------------------
 
-cluster_list = ['m2', 'm3', 'm5', 'm13', 'm15', 'm53', 'm71', 'm92', 'm107', 'ngc6791', 'ngc6819']
+'''cluster_list = ['m2', 'm3', 'm5', 'm13', 'm15', 'm53', 'm71', 'm92', 'm107', 'ngc6791', 'ngc6819']
 
 fig, ax = plt.subplots(4, 3, figsize = (9, 12))
 plt.subplots_adjust(wspace = 0.08, hspace = 0.25)
@@ -427,7 +427,7 @@ plt.savefig('paper/clusters.pdf', bbox_inches = 'tight')
 
 # -------------------------------------------------------------------------------
 # take Sagittarius region...
-# -------------------------------------------------------------------------------
+# -------------------------------------------------------------------------------'''
 
 '''XS_sgr = np.vstack([labels_sgr['ra'], labels_sgr['dec'], labels_sgr['spec_parallax']]).T
 Xlabels_sgr = ['RA', 'DEC', r'$\varpi$']

### 2018-10-17 paper/sp.tex
CMD
table

@@
 joining the table to the \sdssiv\ Catalog on the \zmass\ ID (name).
-\begin{table}
-\caption{HOGG and ACE TBD: The generated spectrophotometric parallaxes.
+
+\newpage
+\begin{longrotatetable}
+\startlongtable
+\begin{deluxetable*}{ccccccc}
+\tablecaption{HOGG and ACE TBD: The generated spectrophotometric parallaxes.
 This table is just a snippet; the full table is available online (HOGG HOW?)\label{tab:output}}
-\end{table}
+\tablehead{\dcolhead{\texttt{2MASS\_ID}} & \dcolhead{\texttt{Gaia\_parallax}} & \dcolhead{\texttt{Gaia\_parallax\_err}} & \dcolhead{\texttt{spec\_parallax}} & \dcolhead{\texttt{spec\_parallax\_err}} & \dcolhead{\texttt{training\_set}} & \dcolhead{\texttt{sample}}}
+\startdata
+2M00000002+7417074 & 0.3191313429473075 & 0.032948037183877704 & 0.31849999096965226 & 0.01643285514291281 & 0 & B \\
+2M00000317+5821383 & 0.3643040248194837 & 0.05099089240834289 & 0.4428242906062965 & 0.01834130205824606 & 0 & A \\
+2M00000546+6152107 & 0.3154104939032093 & 0.035141087048246 & 0.3292112764247769 & 0.016388459680198957 & 0 & B \\
+2M00000797+6436119 & 0.07711986267105561 & 0.08408755179430512 & 0.19428078181594394 & 0.019370065406281896 & 1 & B \\
+2M00001719+6221324 & 0.13952084276830487 & 0.04389977390011321 & 0.1563998166245657 & 0.012831080487863935 & 0 & B \\
+2M00002005+5703467 & 0.3003205275388493 & 0.024656309709138968 & 0.2372046982030924 & 0.020327826647521956 & 0 & B \\
+2M00002036+6153416 & 0.06569746266321969 & 0.0636606119367412 & 0.1347406498933188 & 0.013122259531649945 & 1 & A \\
+2M00002118+6136420 & 0.19358517288879123 & 0.03229292271974642 & 0.2172989314297034 & 0.013278872163629496 & 1 & B \\
+2M00002227+6223341 & 0.1965229567661197 & 0.04503624629849183 & 0.1461267214260898 & 0.008928599127162944 & 0 & B \\
+2M00002472+5518473 & 0.09468551608472082 & 0.020679980798059373 & 0.14731238757486897 & 0.009570425952478014 & 1 & B \\
+... & ... & ... & ... & ... & ... & ...  \\
+ \enddata
+\tablecomments{Columns show...}
+\end{deluxetable*}
+\end{longrotatetable}
+\newpage
 
@@
 na\"ive ways, for example in inverse-variance-weighted averages.
+
+\begin{figure}
+\centering
+~\hfill\includegraphics[width=0.8\textwidth]{./CMD4.pdf}\hfill~
+\caption{color-magnitude-diagram.\label{fig:cmd}}
+\end{figure}
 

### 2018-10-17 make_figures_HER.py

@@ -21,6 +21,8 @@ from astropy.coordinates import SkyCoord
 import astropy.coordinates as coord
 from mpl_toolkits.mplot3d import Axes3D
 import corner
+from mpl_toolkits.axes_grid1.axes_divider import make_axes_locatable
+from mpl_toolkits.axes_grid1.colorbar import colorbar
 
 
 # -------------------------------------------------------------------------------
@@ -259,6 +261,7 @@ plt.savefig('paper/coefficients.pdf')
 
 # run rotation_curve_uncertainties.py
 
+
 # -------------------------------------------------------------------------------
 # Figure 8
 # -------------------------------------------------------------------------------
@@ -278,14 +281,40 @@ plt.savefig('paper/coefficients.pdf')
 #plt.savefig('paper/HRD.pdf')
 
 # color-color diagram?
-abs_mag_G = labels['phot_g_mean_mag'] + 5 + 5 * np.log10(labels['spec_parallax'])
+spec_par = labels['spec_parallax'] * u.mas
+distance = spec_par.to(u.parsec, equivalencies = u.parallax())
+abs_mag_G = labels['phot_g_mean_mag'] - 5 * np.log10(distance.value) + 5
 
 cut_vis = labels['visibility_periods_used'] >= 8    
 cut_par = labels['parallax_error'] < 0.1            
 cut_cal = (labels['astrometric_chi2_al'] / np.sqrt(labels['astrometric_n_good_obs_al']-5)) <= 35         
 train = cut_vis * cut_par * cut_cal  
 best = train * (labels['parallax_over_error'] >= 20)
+cut_feh = labels['FE_H'] > -100
 
+# -------------------------------------------------------------------------------
+# final catalog!
+# -------------------------------------------------------------------------------
+
+training_set = np.zeros((len(labels)), dtype = int)
+training_set[train] = 1
+sample = np.zeros((len(labels)), dtype = str)
+Kfold = 2
+sample_A = labels['random_index'] % Kfold == 0
[…]

### 2018-10-17 paper/sp.tex
table

--- 2018-10-17 paper/sp.tex
@@
 joining the table to the \sdssiv\ Catalog on the \zmass\ ID (name).
-\begin{table}
-\caption{HOGG and ACE TBD: The generated spectrophotometric parallaxes.
+
+\newpage
+\begin{longrotatetable}
+\startlongtable
+\begin{deluxetable*}{ccccccc}
+\tablecaption{HOGG and ACE TBD: The generated spectrophotometric parallaxes.
 This table is just a snippet; the full table is available online (HOGG HOW?)\label{tab:output}}
-\end{table}
+\tablehead{\dcolhead{\texttt{2MASS\_ID}} & \dcolhead{\texttt{Gaia\_parallax}} & \dcolhead{\texttt{Gaia\_parallax\_err}} & \dcolhead{\texttt{spec\_parallax}} & \dcolhead{\texttt{spec\_parallax\_err}} & \dcolhead{\texttt{training\_set}} & \dcolhead{\texttt{sample}}}
+\startdata
+2M00000002+7417074 & 0.3191313429473075 & 0.032948037183877704 & 0.31849999096965226 & 0.01643285514291281 & 0 & B \\
+2M00000317+5821383 & 0.3643040248194837 & 0.05099089240834289 & 0.4428242906062965 & 0.01834130205824606 & 0 & A \\
+2M00000546+6152107 & 0.3154104939032093 & 0.035141087048246 & 0.3292112764247769 & 0.016388459680198957 & 0 & B \\
+2M00000797+6436119 & 0.07711986267105561 & 0.08408755179430512 & 0.19428078181594394 & 0.019370065406281896 & 1 & B \\
+2M00001719+6221324 & 0.13952084276830487 & 0.04389977390011321 & 0.1563998166245657 & 0.012831080487863935 & 0 & B \\
+2M00002005+5703467 & 0.3003205275388493 & 0.024656309709138968 & 0.2372046982030924 & 0.020327826647521956 & 0 & B \\
+2M00002036+6153416 & 0.06569746266321969 & 0.0636606119367412 & 0.1347406498933188 & 0.013122259531649945 & 1 & A \\
+2M00002118+6136420 & 0.19358517288879123 & 0.03229292271974642 & 0.2172989314297034 & 0.013278872163629496 & 1 & B \\
+2M00002227+6223341 & 0.1965229567661197 & 0.04503624629849183 & 0.1461267214260898 & 0.008928599127162944 & 0 & B \\
+2M00002472+5518473 & 0.09468551608472082 & 0.020679980798059373 & 0.14731238757486897 & 0.009570425952478014 & 1 & B \\
+2M00002908+6140446 & 0.16751577398691722 & 0.04602771820716873 & 0.19135489809281317 & 0.01758606714554385 & 1 & A \\
+2M00003061+5820590 & 0.16692055579543114 & 0.03394660614876257 & 0.35023347779063707 & 0.029004068376645976 & 1 & A \\
+... & ... & ... & ... & ... & ... & ...  \\
+ \enddata
+\tablecomments{Columns show...}
+\end{deluxetable*}
+\end{longrotatetable}
+\newpage
 

### 2018-10-18 paper/sp.tex
table
notes to self

--- 2018-10-17 paper/sp.tex
@@
 joining the table to the \sdssiv\ Catalog on the \zmass\ ID (name).
-\begin{table}
-\caption{HOGG and ACE TBD: The generated spectrophotometric parallaxes.
+
+\newpage
+\begin{longrotatetable}
+\startlongtable
+\begin{deluxetable*}{lllllcc}
+\tablecaption{HOGG and ACE TBD: The generated spectrophotometric parallaxes.
 This table is just a snippet; the full table is available online (HOGG HOW?)\label{tab:output}}
-\end{table}
+\tablehead{\dcolhead{\texttt{2MASS\_ID}} & \dcolhead{\texttt{Gaia\_parallax}} & \dcolhead{\texttt{Gaia\_parallax\_err}} & \dcolhead{\texttt{spec\_parallax}} & \dcolhead{\texttt{spec\_parallax\_err}} & \dcolhead{\texttt{training\_set}} & \dcolhead{\texttt{sample}}}
+\startdata
+2M00000002+7417074 & 0.3191 & 0.0329 & 0.3185 & 0.0164 & 0 & B \\
+2M00000317+5821383 & 0.3643 & 0.051 & 0.4428 & 0.0183 & 0 & A \\
+2M00000546+6152107 & 0.3154 & 0.0351 & 0.3292 & 0.0164 & 0 & B \\
+2M00000797+6436119 & 0.0771 & 0.0841 & 0.1943 & 0.0194 & 1 & B \\
+2M00001719+6221324 & 0.1395 & 0.0439 & 0.1564 & 0.0128 & 0 & B \\
+2M00002005+5703467 & 0.3003 & 0.0247 & 0.2372 & 0.0203 & 0 & B \\
+2M00002036+6153416 & 0.0657 & 0.0637 & 0.1347 & 0.0131 & 1 & A \\
+2M00002118+6136420 & 0.1936 & 0.0323 & 0.2173 & 0.0133 & 1 & B \\
+2M00002227+6223341 & 0.1965 & 0.045 & 0.1461 & 0.0089 & 0 & B \\
+2M00002472+5518473 & 0.0947 & 0.0207 & 0.1473 & 0.0096 & 1 & B \\
+2M00002908+6140446 & 0.1675 & 0.046 & 0.1914 & 0.0176 & 1 & A \\
+2M00003061+5820590 & 0.1669 & 0.0339 & 0.3502 & 0.029 & 1 & A \\
+... & ... & ... & ... & ... & ... & ...  \\
+ \enddata
+\tablecomments{Columns show...}
+\end{deluxetable*}
+\end{longrotatetable}
+\newpage
 

### 2018-10-21 rotation_curve_uncertainties.py
new Fig4

@@ -403,11 +403,23 @@ wedge1d = abs(mean_XS_cyl_n[:, 1]) < (deg_wedge/360. * 2. * np.pi) * cut_vz
 #wedge1d = (mean_XS_cyl_n[:, 1] < (deg_wedge/360. * 2. * np.pi)) * (mean_XS_cyl_n[:, 1] > 0) * cut_vz
 #wedge1d = (-mean_XS_cyl_n[:, 1] < (deg_wedge/360. * 2. * np.pi)) * (mean_XS_cyl_n[:, 1] < 0) * cut_vz
 
+
+# -------------------------------------------------------------------------------
+# test of vertical gradient
+# -------------------------------------------------------------------------------   
+
+cut_z_plus = mean_XS_cyl_n[:, 2] > 0
+cut_z_minus = mean_XS_cyl_n[:, 2] < 0
+
+# -------------------------------------------------------------------------------
+# continue from above...
+# -------------------------------------------------------------------------------   
+
 dz = 1. # kpc
 bins_start = np.array([1.])
 bins_end = np.array([40.])
 
-cut_z_wedge = wedge1d * cut_z 
+cut_z_wedge = wedge1d * cut_z #* cut_z_minus
 foo = np.append(0., np.sort(mean_XS_cyl_n[cut_z_wedge, 0]))
 bar = np.append(np.sort(mean_XS_cyl_n[cut_z_wedge, 0]), 100.)
 min_stars_per_bin = 3
@@ -639,9 +651,9 @@ ax[0].set_ylabel(r'$y\,\rm [kpc]$', fontsize = fsize)
 ax[0].set_aspect('equal')
 ax[1].set_aspect('equal')
 ax[2].set_aspect('equal')
-ax[0].annotate(r'$\overline{v_{r}}$', (6.5, -10), fontsize = fsize, bbox=dict(boxstyle="square", fc="w"))
-ax[1].annotate(r'$\overline{v_{\varphi}}$', (6.5, -10), fontsize = fsize, bbox=dict(boxstyle="square", fc="w"))
-ax[2].annotate(r'$v_{\rm c}$', (6.5, -10), fontsize = fsize, bbox=dict(boxstyle="square", fc="w"))
[…]

### 2018-10-21 paper/sp.tex
table
adjusting cluster words

@@
 the spectrophotometric-parallax distribution from our model for stars that are
-astrometrically confirmed members of known stellar clusters.
+astrometrically confirmed members of three of the stellar clusters in which we
+validated our results.
 The cluster members were found by hand, looking at sky and proper-motion distributions
@@
 is, membership is conservative.
-These clusters span some range in metallicity and age, and therefore test
+We chose three clusters to span some range in metallicity and age, and therefore test
 the accuracy of the method for different kinds of populations.
-With the exception of \acronym{M71}, these figures 
+With the exception of \acronym{M71}, the clusters we tested
 indicate that both \gaia\ and the spectrophotometric parallaxes appear to be unbiased
@@
   and spectrophotometric parallaxes $\sparallax_n$ for member stars for
-  four stellar clusters in the \apogee\ observing footprint. The spectrophotometric
+  three of the stellar clusters in the \apogee\ observing footprint. The spectrophotometric
   parallaxes and their inverse-variance-weighted mean and the astrometric parallaxes
@@
 \startlongtable
-\begin{deluxetable*}{lllllcc}
+\begin{deluxetable*}{ccccccc}
 \tablecaption{HOGG and ACE TBD: The generated spectrophotometric parallaxes.
@@
 2M00000002+7417074 & 0.3191 & 0.0329 & 0.3185 & 0.0164 & 0 & B \\
-2M00000317+5821383 & 0.3643 & 0.051 & 0.4428 & 0.0183 & 0 & A \\
+2M00000317+5821383 & 0.3643 & 0.0510 & 0.4428 & 0.0183 & 0 & A \\
 2M00000546+6152107 & 0.3154 & 0.0351 & 0.3292 & 0.0164 & 0 & B \\
@@
 2M00002118+6136420 & 0.1936 & 0.0323 & 0.2173 & 0.0133 & 1 & B \\
-2M00002227+6223341 & 0.1965 & 0.045 & 0.1461 & 0.0089 & 0 & B \\
+2M00002227+6223341 & 0.1965 & 0.0450 & 0.1461 & 0.0089 & 0 & B \\
 2M00002472+5518473 & 0.0947 & 0.0207 & 0.1473 & 0.0096 & 1 & B \\
-2M00002908+6140446 & 0.1675 & 0.046 & 0.1914 & 0.0176 & 1 & A \\
-2M00003061+5820590 & 0.1669 & 0.0339 & 0.3502 & 0.029 & 1 & A \\
+2M00002908+6140446 & 0.1675 & 0.0460 & 0.1914 & 0.0176 & 1 & A \\
+2M00003061+5820590 & 0.1669 & 0.0339 & 0.3502 & 0.0290 & 1 & A \\
 ... & ... & ... & ... & ... & ... & ...  \\
@@
 
+HOGG: Discussion here of CMD and AGB stars.
 \begin{figure}
@@
 ~\hfill\includegraphics[width=0.8\textwidth]{./CMD4.pdf}\hfill~
-\caption{color-magnitude-diagram.\label{fig:cmd}}
+\caption{HOGG color-magnitude-diagram.\label{fig:cmd}}
 \end{figure}
-
-ACE: ADD HERE the Color-magnitude diagrams for high snr and training set.
-HOGG: Discussion here of CMD and AGB stars.
 

### 2018-10-21 paper/sp.tex
rix comments

@@
 % -----------
-% - need a snippet for Table 1.
 % - check with WISE experts that we have the correct WISE photometry
 % - ACE: What fraction of spectral pixels are (typically, or in each training set) zeroed out?
+% - are we clear that the map is just a demo and just shows the 10-percent-ish region of the MW disk?
 
@@
 Here we use the \apogee--\gaia--\zmass--\wise\ overlap to train a data-driven model
-to predict parallaxes for red-giant branch stars with $0<\logg\leq2.2$.
+to predict parallaxes for red-giant branch stars with $0<\logg\leq2.2$ (more luminous
+than the red clump).
 % methods
@@
 the \gaia\ Mission (\citealt{gaia}) has also been revolutionizing our view of the disk.
-It delivers good proper motions everywhere,
+It delivers good proper motions over most of the Galaxy,
 and extremely valuable parallax information locally.
-However, it does not deliver precise geometric parallaxes over a large fraction of
+However, \gaia\ does not deliver precise geometric parallaxes over a large fraction of
 the disk, and certainly not in the dusty and crowded regions.
@@
 even at large distance).
-They are extremely common (frequent) stars, much more common than distance-indicating
+They are very common stars, much more frequent than distance-indicating
 stars like Cepheids.
@@
 red-giant parallaxes or distances with spectroscopy and photometry.
-Our method capitalizes on the point that red-giant stars are dust-correctable,
-standardizable candles, but it makes no use whatsoever of any stellar interior,
+Our method is predicated on the expectation that red-giant stars are dust-correctable,
+standardizable candles with such data,
+but it makes no use whatsoever of any stellar interior,
 stellar exterior, or dust-reddening models.
@@
 optimally when we ignore our physical models.
-Relatedly, they are not (usually) interpretable or generalizable:
+Relatedly, they are not (usually) interpretable or generalizable or useful for
+extrapolations significantly outside the training set:
 A data-driven model trained with one kind of data can not usually be used in another
@@
 like \apogee\ and \sdssv.
-So we will take the limitations of the data-driven models in exchange for their
+So we accept the limitations of the data-driven models in exchange for their
 performance.
@@
 methods (like deep neural networks).
-In this work we stay on the low-complexity side of this.
-We do not want to face the (literally combinatoric) choices in model structure that
+In this work we stay on the low-complexity side of this, as
+we do not want to face the (literally combinatoric) choices in model structure that
 advanced methods bring.
@@
 Below, we will make no such cuts. That means that our training step has the odd
-property that much of the training data is low in signal-to-noise and some of the
-training parallax measurements are even negative!
+property that much of the training data is low in signal-to-noise; we include
+training parallax measurements even if they are negative!
 The fact that we are using a justified likelihood function ensures that these
@@
 In addition to making the above assumptions, we also avoid various practices
-with the data that are tempting but lead to strong biases.
+with the data that are tempting but lead to important biases.
 For example, we never cut on \gaia\ parallax or parallax signal-to-noise.
-The common practices of cutting to parallaxes that are good to 20~percent,
+The common practices of cutting to parallaxes that are good to 20~percent
+(see, for example, \citealt{trick, helmi}),
 or Catalog parallaxes that are positive, or parallaxes that are smaller or
@@
 
-The model and log-likelihood function can be expressed heuristically as
+The model for the parallax
+and the log-likelihood function can be expressed heuristically as
 \begin{eqnarray}
@@
 In practice we avoid it by optimizing first for the very highest signal-to-noise
-(in astrometric parallax) Training Set stars, and then use that first optimum as a
+(better than 5-percent measurements of astrometric parallax)
+Training Set stars, and then use that first optimum as a
 good first guess or initialization for the full optimization.
@@
 
-Once the model is optimized, the output of the model is a prediction of the parallax,
+Once the model is optimized---and therefore $\hat{\theta}$ is determined---the
+output of the model is a prediction of the parallax,
 or really what we will call the \emph{spectrophotometric parallax}.
@@
 
-One final detail: We perform all of the fitting in a two-fold train-and-test framework,
+We perform all of the fitting in a two-fold train-and-test framework,
 in which the data are split into two disjoint subsets, A and B.
@@
 
-Every Parent-Sample star gets, in addition, a randomly assigned binary
+Every Parent Sample star gets, in addition, a randomly assigned binary
 label (A or B).
@@
 For training, we need A and B Training Sets.
-We define the Training Set stars to be Parent-Sample stars that also
+We define the Training Set stars to be Parent Sample stars that also
 have a measured \gaia\ parallax.
-The Training-Set stars, in addition to meeting all Parent-Sample cuts,
+The Training Set stars, in addition to meeting all Parent Sample cuts,
 also must meet an uncertainty criterion $\gsigma_n < 0.1\,\mathrm{mas}$, i.e. 
@@
 But---as we have emphasized above---we do not cut ever on parallax or
-or the ratio to its uncertainty. For this reason, most of the Training-Set stars
+or the ratio to its uncertainty. For this reason, most of the Training Set stars
 do not have significantly measured parallaxes and more than 2~percent even have
@@
 parallaxes reported by the \gaia\ Collaboration (\citealt{lindegren}).
-The \gaia\ recommended value is $\varpi_0=0.029$\,mas (\citealt{lindegren}),
+The \gaia\ recommended all-sky-average value is $\varpi_0=0.029$\,mas (\citealt{lindegren}),
 but we adopt $\varpi_0 = 0.048$\,mas because that value
@@
 train-and-test framework constitutes a two-fold cross-validation.
-Results of this cross-validation is shown in \figurename~\ref{fig:xval}.
+Results of this cross-validation are shown in \figurename~\ref{fig:xval}.
 For the stars with highest \gaia-measured parallax signal-to-noise,
@@
   and our spectrophotometric parallaxes $\sparallax$ on the horizontal axis.
-  The left panel shows the full Parent Sample, the middle panel shows the Training Set
-  stars (which pass \gaia\ quality checks; see text), and the right panel shows a
+  The left panel shows the full Training Set (both A and B combined),
+  and the right panel shows a
   subset of stars with very high astrometric parallax signal-to-noise. This latter
@@
 deconvolve or correct for the contributions from the \gaia\ noise.
-However it is also not fully conservative, in that the test is performed with
+However, this test is performed with
 \gaia's best stars, which are bright, nearby, in low extinction regions, and
@@
 is, membership is conservative.
-We chose three clusters to span some range in metallicity and age, and therefore test
+We chose the three displayed clusters
+to span some range in metallicity and age and extinction, and therefore test
 the accuracy of the method for different kinds of populations.
@@
 housekeeping data (also see \figurename~\ref{fig:residuals}) we don't find any correlations
-with parallax offsets.
+with parallax offsets, although \acronym{M71} is a higher-extinction cluster.
 \begin{figure}
@@
 
-This project demonstrates that photometry and spectroscopy can be used to predict
+We have demonstrated that photometry and spectroscopy can be used to predict
 or estimate distances to stars.
 This isn't new!
-However, what's new is that it is possible to do this very well but with a purely
+However, what's new is that it is possible to do this very well for luminous giants
+with unknown ages, and with a purely
 linear model acting on magnitudes and spectral (logarithmic) fluxes, and entirely
@@
 These methods generate good features automatically.
-However, they also involve an enormous set of hyper-parameter-like choices,
+They also involve an enormous set of hyper-parameter-like choices,
 and they involve sacrificing certain kinds of interpretability.
-However, we expect that a better set of features would do better; in this sense
+All that said, we expect that a better set of features would do better; in this sense
 the project here is just a first step towards precise spectrophotometric distance
@@
 
-As we said above, our most restrictive assumption is the linearity assumption.
-We required that the parallax prediction be constructed from a linear combination
+As we said above,
+we required that the parallax prediction be constructed from a linear combination
 of the input data or feature vector components.
-This is absurd: Stars are not linear objects!
-There are two questions to ask about the linearity.
-One is: Why did it work? The other is: Could we have done better?
+This may seem like an absurd simplification; it begs the questions:
+Why did it work? And could we have done better?
 The answer to the first is that we have taken such a small part of stellar parameter
@@
 
-The answer to the second question is that absolutely we could have done better!
+The answer to the second question is that almost certainly we could have done better!
 Since models like Gaussian processes and deep networks subsume the linear model

### 2018-10-21 paper/sp.tex
arxiver, cmd, refs

@@
 %% Copyright 2018 the authors.
+
+%@arxiver{residuals.pdf,map.pdf}
 
@@
 % - ACE: What fraction of spectral pixels are (typically, or in each training set) zeroed out?
-% - are we clear that the map is just a demo and just shows the 10-percent-ish region of the MW disk?
 
@@
 
-%% Reintroduced the \received and \accepted commands from AASTeX v5.2
-%\received{TBD}
-%\revised{TBD}
-%\accepted{TBD}
-%% Command to document which AAS Journal the manuscript was submitted to.
-%% Adds "Submitted to " the arguement.
-%\submitjournal{AAS Journals}
-
 \shorttitle{spectrophotometric parallax}
 \shortauthors{hogg, eilers, rix}
-%\watermark{kittens}
 
@@
 
-\newpage
 \begin{longrotatetable}
@@
 \begin{deluxetable*}{ccccccc}
-\tablecaption{HOGG and ACE TBD: The generated spectrophotometric parallaxes.
-This table is just a snippet; the full table is available online (HOGG HOW?)\label{tab:output}}
+\tablecaption{The generated spectrophotometric parallaxes and their uncertainties..
+This \tablename\ is just a snippet; the full catalog of all 44784 objects in the combined Parent Samples
+is available online. In order to reproduce the {\figurename s} in this \documentname, this catalog
+should be joined to the \gaia, \zmass, \wise\, and \apogee\ Catalogs by \zmass\ ID.\label{tab:output}}
 \tablehead{\dcolhead{\texttt{2MASS\_ID}} & \dcolhead{\texttt{Gaia\_parallax}} & \dcolhead{\texttt{Gaia\_parallax\_err}} & \dcolhead{\texttt{spec\_parallax}} & \dcolhead{\texttt{spec\_parallax\_err}} & \dcolhead{\texttt{training\_set}} & \dcolhead{\texttt{sample}}}
@@
 
-HOGG: Discussion here of CMD and AGB stars.
-\begin{figure}
-\centering
-~\hfill\includegraphics[width=0.8\textwidth]{./CMD4.pdf}\hfill~
-\caption{HOGG color-magnitude-diagram.\label{fig:cmd}}
-\end{figure}
-
 This \documentname\ and project delivers greatly improved distance estimates for
-luminous red-giant stars.
+luminous red-giant stars, more luminous than the red clump.
 The purpose of all this is to make global maps of the Milky Way.
@@
   This map shows the rotation of the Galaxy out to large radius, and also
-  indicates the kinematic center of the disk rotation. Interpretation of
+  indicates the kinematic center of the disk rotation. The stars shown in this
+  map have distance uncertainties on the order of ten percent, so at significant
+  heliocentric distances, the interpretation of
   the quantitative information in this plot requires care.\label{fig:disk}}
@@
 \bibitem[Allende Prieto \etal(2008)]{aapogee} Allende Prieto, C., Majewski, S.~R., Schiavon, R., \etal\ 2008, Astronomische Nachrichten, 329, 1018 
-\bibitem[Arenou \etal(2018)]{arenou} Arenou, F., Luri, X., Babusiaux, C., et al.\ 2018, \aap, 616, A17 
+\bibitem[Arenou \etal(2018)]{arenou} Arenou, F., Luri, X., Babusiaux, C., \etal\ 2018, \aap, 616, A17 
 \bibitem[Astropy Collaboration \etal(2013)]{astropy} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., \etal\ 2013, \aap, 558, A33 
@@
 \bibitem[Garc{\'{\i}}a P{\'e}rez \etal(2016)]{aspcap} Garc{\'{\i}}a P{\'e}rez, A.~E., Allende Prieto, C., Holtzman, J.~A., \etal\ 2016, \aj, 151, 144 
+\bibitem[Helmi \etal(2018)]{helmi} Helmi, A., Babusiaux, C., Koppelman, H.~H., \etal\ 2018, arXiv:1806.06038 
 \bibitem[Ho \etal(2017)]{ho} Ho, A.~Y.~Q., Ness, M.~K., Hogg, D.~W., \etal\ 2017, \apj, 836, 5 
@@
 \bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., \etal\ 2006, \aj, 131, 1163 
+\bibitem[Trick \etal(2018)]{trick} Trick, W.~H., Coronado, J., \& Rix, H.-W.\ 2018, arXiv:1805.03653 
 \bibitem[Wilson \etal(2010)]{wapogee} Wilson, J.~C., Hearty, F., Skrutskie, M.~F., \etal\ 2010, \procspie, 7735, 77351C 

### 2018-10-21 paper/sp.tex
catalog

@@
 and uncertainty $\ssigma_m$ for every star $m$.
-\tablename~\ref{tab:output} shows a snippet of this output, with the full output available
+\tablename~\ref{tab:output} shows a snippet of this catalog of output, with the full catalog available
 as an electronic data table associated with this \documentname.
-In order to use this output with \gaia\ astrometry, we recommend joining the table to the
+In order to use this output with \gaia\ astrometry, we recommend joining our catalog to the
 \gaia\ Catalog on the \zmass\ ID (name), as per the instructions at the \gaia\ Archive (\citealt{gaiaarchive}).
 In order to use this output with \apogee\ radial velocities and element abundances, we recommend
-joining the table to the \sdssiv\ Catalog on the \zmass\ ID (name).
+joining our catalog to the \sdssiv\ Catalog on the \zmass\ ID (name).
 

### 2018-10-21 paper/sp.tex
code
50 percent

@@
 model to be sparse.
-We set the value of $\lambda$ by cross-validation.
+We set the value of $\lambda$ by cross-validation across the A/B split (see below).
+In the end, the L1 regularization zeros out about 50 percent of the model coefficients.
 
@@
 joining our catalog to the \sdssiv\ Catalog on the \zmass\ ID (name).
+In addition to this catalog, all the code used for this project is available online in an associated
+code repository.
 

### 2018-10-21 paper/sp.tex
75 percent
urls

--- 2018-10-21 paper/sp.tex
@@
 % - check with WISE experts that we have the correct WISE photometry
-% - ACE: What fraction of spectral pixels are (typically, or in each training set) zeroed out?
 
@@
 model to be sparse.
-We set the value of $\lambda$ by cross-validation.
+We set the value of $\lambda$ by cross-validation across the A/B split (see below).
+In the end, the L1 regularization zeros out about 75 percent of the model coefficients.
 
@@
 \tablename~\ref{tab:output} shows a snippet of this catalog of output, with the full catalog available
-as an electronic data table associated with this \documentname.
+as an electronic data table associated with this \documentname.\footnote{\url{https://zenodo.org/record/1468053}}
 In order to use this output with \gaia\ astrometry, we recommend joining our catalog to the
@@
 joining our catalog to the \sdssiv\ Catalog on the \zmass\ ID (name).
+In addition to this catalog, all the code used for this project is available online in an associated
+code repository.\footnote{\url{https://github.com/aceilers/spectroscopic_parallax}}
 

### 2018-11-23 fit_rot_curve.py
new

@@ -421,7 +421,7 @@ print('{:e} pm {:e}'.format(rho_0, sigma_rho_0))
 
 np.random.seed(42)
 # LINES IN DARK GREY/BLACK!
-fig, ax = plt.subplots(1, 1, figsize = (9.3, 7.2), sharex = True)        
+fig, ax = plt.subplots(1, 1, figsize = (9.3, 7.), sharex = True)        
 plt.errorbar(bins_dr[idx5:], vc_annulus[idx5:], yerr = [vc_annulus_err_m[idx5:], vc_annulus_err_p[idx5:]], fmt = 'o', markersize = 8, capsize=4, mfc='k', mec='k', ecolor = 'k', zorder = 300, label = r'Eilers et al.\ 2018 (this work)')
 #plt.plot(xx1[0, idx5a:], xx1[1, idx5a:], 'o', markersize = 8, zorder = -300, color = '#D7D7D7')
 #plt.plot(xx2[0, idx5b:], xx2[1, idx5b:], 'o', markersize = 8, zorder = -300, color = '#D7D7D7')
@@ -452,18 +452,61 @@ ax.add_patch(patches.Rectangle((0, 0), 5, 400, facecolor = '#EFEFEF', edgecolor 
 handles, labels = ax.get_legend_handles_labels()
 hand = [handles[7], handles[8], handles[10], handles[9], handles[0], handles[2], handles[1], handles[6], handles[3], handles[4], handles[5]]
 lab = [labels[7], labels[8], labels[10], labels[9], labels[0], labels[2], labels[1], labels[6], labels[3], labels[4], labels[5]]
-plt.errorbar(X_GC_sun_kpc, b_lin, yerr = b_lin * 0.05, markersize = 8, capsize=5, elinewidth=4, capthick = 4, mfc='w', mec='w', ecolor = 'w', zorder = 2000)
-plt.errorbar(X_GC_sun_kpc, b_lin, yerr = b_lin * 0.05, markersize = 8, capsize=4, elinewidth=2, capthick = 2, mfc='#047495', mec='#047495', ecolor = '#047495', zorder = 3000)
-#font = FontProperties()
-#font.set_weight('black')
-#font.set_size('x-large')
[…]

### 2018-11-23 rotation_curve_uncertainties.py

@@ -425,7 +425,7 @@ dz = 1. # kpc
 bins_start = np.array([1.])
 bins_end = np.array([40.])
 
-cut_z_wedge = wedge1d * cut_z #* cut_z_minus
+cut_z_wedge = wedge1d * cut_z 
 foo = np.append(0., np.sort(mean_XS_cyl_n[cut_z_wedge, 0]))
 bar = np.append(np.sort(mean_XS_cyl_n[cut_z_wedge, 0]), 100.)
 min_stars_per_bin = 3
@@ -486,7 +486,7 @@ indices = np.arange(mean_XS_cyl_n.shape[0])
 
 for i, (r_start, r_end) in enumerate(zip(bin_start, bin_end)):
     #cut_annulus = wedge * (abs(mean_XS_cyl_n[:, 2]) < dz/2.) * (mean_XS_cyl_n[:, 0] > r_start) * (mean_XS_cyl_n[:, 0] < r_end)
-    cut_annulus = cut_z_wedge * (mean_XS_cyl_n[:, 0] > r_start) * (mean_XS_cyl_n[:, 0] < r_end)
+    cut_annulus = cut_z_wedge * (mean_XS_cyl_n[:, 0] > r_start) * (mean_XS_cyl_n[:, 0] < r_end) #* cut_z_plus
     N_stars_annulus[i] = np.sum(cut_annulus)
     if N_stars_annulus[i] > 0:
         mean_XS_cyl_annulus[i, :] = np.nanmean(mean_XS_cyl_n[cut_annulus], axis = 0)
@@ -562,15 +562,48 @@ vrr_R_exp = theta_fit[1] # kpc
 print('vrr_R_exp = {}'.format(vrr_R_exp))
 
 # -------------------------------------------------------------------------------
+# calculate missing term in Jeans equation
+# -------------------------------------------------------------------------------
+
+vrvz_minus = np.array([-2.46887108e+03,  1.64845824e+03,  5.83353159e+01, -2.14045989e+03,
+        1.89241961e+01, -3.41014878e+02,  5.80391707e+01, -2.03978075e+02,
+       -4.14675591e+01, -3.54867829e+01,  4.53344803e+01, -7.01446949e+01,
+       -5.77536071e+01, -2.48209061e+01, -1.19106315e+02, -1.00391787e+02,
[…]

### 2019-03-24 paper/response_to_referee.txt
start

Dear Editor:

Enclosed please find the author response to the referee report on
AAS14152 "Spectrophotometric parallaxes with linear models: Accurate
distances for luminous red-giant stars" by Hogg, Eilers, & Rix.

Apologies for the tardiness of this response. The referee has
definitely made great suggestions and the paper is improved for us
having addressed them. Our response to the referee is below the line.

Hogg

-----

In what follows, the referee comments are indented, and our responses
are not. We have made changes to the paper in response to every point;
these comments were helpful and valuable.

    The paper presents a method to estimate spectrophotometric parallaxes
    based only on input data, without adoption of physically motivated
    models and priors. The paper is written in an way that is strongly
    focused on the method itself, but leaves several complementary aspects
    that are certainly important to assess the reliability of the inferred
    parallaxes and, most importantly, distances. I recommend its
    publication only after such major issues are addressed.

    Major issues

    1 - The data-driven method described in the paper details out how
    spectrophotometric parallaxes are inferred. But the paper does not
    give the reader a clue about how spectrophotometric distances are
    estimated. The authors make it clear in Section 2 that they never
    invert *observed* parallaxes to estimate distances in their
    method. But are the spectrophotometric distances they infer (shown in
    Figures 5 and 6) just the inverse of the *inferred* parallaxes? If so,
    this should be made explicit in the text, along with some discussion
    about the problems (or lack of) that may arise from this
    inversion. Even though the spectrophotometric parallaxes themselves
    are always positive, a reader would like to be convinced that the
    distances shown in Figures 5 and 6 do not suffer from biasses caused
    by the asymmetric nature of the distances distribution that results
    from a symmetric error distribution in parallaxes (as stated by the
    likelihood assumption of the method). How do these biasses simply go
    away when the method's w(sp) +/- sigma(sp) are used? In particular, a
    negative parallax for many stars may be consistent with their w(sp)
    +/- sigma(sp).

These issues are subtle! We are assuming that the GAIA PARALLAXES have
noise that is symmetric in parallax space. In our data-driven model,
we use a linear combination of features to predict or estimate a LOG
PARALLAX, so if anything we are claiming that our estimates are
symmetric in log parallax, not parallax. But we aren't even really
claiming that. The issues are subtle. However, this does deserve some
discussion and we have added that around Figures 5 and 6. The short answer
is that we just negative-logged our log-parallax estimates to make Figures
5 and 6.

There is an additional question of whether we are somehow biasing
things when we presume that the parallax must be positive? Our
position is that we are biasing things by presuming that the LOG
PARALLAX can be predicted as a linear combination of features. Of
course any other assumption would also introduce biases! There is no
way to fit that would be unbiased here, because by construction, our
model is not physically accurate. We added words about this into the
assumption list in Section 2.

    2 - The paper shows limited validation tests and comparisons to other
    work. There are several papers that provide spectrophotometric
    distances based on physical models for APOGEE stars (Queiroz et al
    2018, MNRAS, 476, 2556; Schultheis et al 2014, AJ, 148, 24 and
    others). The authors should compare their distances to those authors
    and discuss the results of this comparison. Moreover, asteroseismic
    distance estimates are also available for a fraction of the APOGEE
    sample of giants in common with the CoRoT and Kepler missions. This is
    an additional and important validation test.

This is a great point. We obtained these two data sets (and the
Queiroz sample required some emails!) and we made comparisons. We
added a short paragraph about the results of these comparisons in
Section XXX. We did not feel that the comparisons merited supporting
Figures, but we do have such figures if the referee disagrees.

During the long period in which we haven't responded to this report,
there was also a paper out by Leung & Bovy. They compare their results
to ours, and we also mention that in the newly added paragraph.

    3 - The authors apply colour cuts to eliminate outliers in the APOGEE
    data, eqs 6 and 7. They also apply a cut in parallax error, eq. 8.,
    along with other quality criteria, to define the training set. So, the
    training set parallax uncertainty distribution is truncated. How do
    these cuts affect the method and its assumptions? In particular, it is
    important to reconcile these selection criteria for the training set
    with the statements presented in the Introduction (page 3):
    "... regressions can be strongly biased if we have a bad selection
    function or data censoring. This means that we cannot apply cuts to
    the Gaia astrometry or signal-to-noise unless those cuts are correctly
    accounted for in our generative model for the data sample. Below, we
    will make no such cuts."  It is in fact clear from Figure 1 that the
    training set is well under represented of low log g stars with respect
    to the parent sample. This already jeopardizes the representativeness
    assumption somewhat.  One useful assessment of the effects of data
    censoring would be to show the spectral S/N or parallax distributions
    of the training data in comparison to that of the entire sample that
    the method is applied to.

This point is important! The Gaia parallax uncertainties are based on
a spacecraft model that is independent of the measurements themselves.
Therefore, cuts on parallax uncertainties are not cuts on parallax and
thus are not subject to the strong censure that we give to such cuts!
But the referee is correct that we have been very unclear about this
and we have attempted to fix this in Section YYYY.

The referee is completely correct that the detailed distribution of
stars in parallax, SNR, and log g is not identical in the training set
and the parent sample.  We don't have any reason to believe that this
biases our results. However, we do now show these distribution
differences in a newly added figure, Figure XXX. Certainly these
differences mean that we probably have uncertainties that, in detail,
vary with log g.

    4 - Finally, it is important to provide the reader with some
    understanding of how in practice the data are driving the resulting
    parallaxes. This may be achieved by discussing the behaviour of the
    linear coefficients theta. How do their values, as estimated by
    expr. (3), vary from training set A and to training set B? What
    spectral ranges are most relevant and survive the L1 regularization
    process? In other words, the authors could provide a discussion
    similar to the one around Figure 5 of Ness et al did on the Cannon,
    but this time comparing sets A and B across the spectrum or over the
    smaller photometric subspace.

[syntactical reference]

    Other issues:

    Introduction, 1st par.: could the authors say a few words about the
    importance of making "precise kinematic and element-abundance maps of
    the Milky Way disk out to large heliocentric radii..."?

    Section 2, 1st par. after the list of assumptions. The statement "We
    never cut on parallax or parallax signal-to-noise." is repeated.

    Section 4: "The hope of the model assumptions is that the Training Set
    will contain sufficient dust variation that the model will naturally
    learn the dust corrections and extrapolate acceptably.” Is this hope
    checked in any way?

    Section 4: The parallax offset w_0 is a contentious issue and it is
    important to quantify its effect on the final spectrophotometric
    parallaxes.

    The caption of Figure 2 is wrong: "...shows Gaia astrometric
    parallaxes w(a)n on the HORIZONTAL axis and our spectrophotometric
    parallaxes w(sp) on the VERTICAL axis."

    The panels and labels of Figure 3 are too small. Impossible to read
    the axes.

    Make an analogous of Figure 2 and 3 for all the stars with available
    parallaxes, regardless of whether criteria 8 and 9 are valid or not

    Section 5, 2nd paragraph: Is the parameter Lambda (capital letter) the
    same as lambda (lower case initial)? If so, what is meant by the value
    of the "lambda vector" which the authors mention in the same
    paragraph? Please clarify

    Typos:

    4th line after eq 5: "... is, we ASSUME THAT C_m is a diagonal matrix
    with the..." (instead of "... we make C_m is ...")

    Section 4, 2nd par., 4th line: please eliminate repeated "the" in "In
    detail, we obtain the the APOGEE..."

    Section 4, 4th par., 5th line: "...3.6 and 4.5 μmPERIOD In detail..."

    Section 4, 5th par., 2nd line: again, eliminate extra "the" in "...to
    the the 2MASS IDs"

    Section 4, 4th line below eq 10: eliminate extra "or" in "...cut ever
    on parallax or or the ratio..."

    Comments of the AAS Data and Software reviewer:

    - We recommend reproducibility material such as that hosted at the
      aceilers/spectroscopic parallax repository be archived to a
      repository, with the repository DOI added adjacent to the github URL
      in the final article. [This request is also true of submission
      AAS14147 for similar reasons].

    - Do the authors wish to see the full Table 1 hosted with the Journal?
      We'd gladly process the final table into the final article, which
      would speed ingest of that data into CDS/Vizier. Regardless the
      Zenodo URL in footnote # should be converted back to its DOI form.

### 2019-03-25 paper/response_to_referee.txt
more

@@
 is that we just negative-logged our log-parallax estimates to make Figures
-5 and 6.
+5 and 6. HOGG DO THIS
 
@@
 model is not physically accurate. We added words about this into the
-assumption list in Section 2.
+assumption list in Section 2. HOGG DO THIS
 
@@
 Section XXX. We did not feel that the comparisons merited supporting
-Figures, but we do have such figures if the referee disagrees.
+Figures, but we do have such figures if the referee disagrees. HOGG DO THIS
 
@@
 there was also a paper out by Leung & Bovy. They compare their results
-to ours, and we also mention that in the newly added paragraph.
+to ours, and we also mention that in the newly added paragraph. HOGG DO THIS
 
@@
 But the referee is correct that we have been very unclear about this
-and we have attempted to fix this in Section YYYY.
+and we have attempted to fix this in Section YYYY. HOGG DO THIS
 
@@
 differences mean that we probably have uncertainties that, in detail,
-vary with log g.
+vary with log g. HOGG DO THIS
 
@@
 
-[syntactical reference]
+Absolutely! Actually we had a section about this but dropped it just
+before submission. One of the reasons we dropped it is because the
+story is somewhat complicated. But you are right that this must be
+there, and we have added it back in, at the end of Section YYY. The
+short story (which you can read in the updated text) is that the
+photometric part of the model is interpretable and stable, but the
+spectroscopic part is less stable in its internals (while remaining
+very stable in its predictions in the data space). We don't fully
+understand this, but it connects to known issues in machine learning
+more generally, which we point out in the text. HOGG DO THIS
 
@@
 
+Good point! An expansion on this theme added. HOGG DO THIS.
+
     Section 2, 1st par. after the list of assumptions. The statement "We
     never cut on parallax or parallax signal-to-noise." is repeated.
+
+Yes, yes it is. That was supposed to be for emphasis, but we did HOGG
+WHAT NOW?
 
@@
 
+Hahaha no. Okay fine. We now comment on this point in Section YYY
+where we deliver the interpretation of the photometric part of the
+model. HOGG DO THIS.
+
     Section 4: The parallax offset w_0 is a contentious issue and it is
@@
     parallaxes.
+
+The effect is weak. We comment on this more now in Section XXX. HOGG
+DO THIS.
 
@@
 
+Whoops! Fixed. HOGG DO THIS.
+
     The panels and labels of Figure 3 are too small. Impossible to read
@@
 
+Fixed.
+
     Make an analogous of Figure 2 and 3 for all the stars with available
     parallaxes, regardless of whether criteria 8 and 9 are valid or not
+
+Done. HOGG: DO WE PUT THIS INTO THE PAPER?
 
@@
 
+Whoops. That's just a typo. Fixed.
+
     Typos:
@@
     with the..." (instead of "... we make C_m is ...")
+
+[syntactical reference]
 

### 2019-03-29 fit_rot_curve.py
non-axisymmetries

@@ -509,6 +509,28 @@ plt.tight_layout()
 plt.savefig('../applications/proposal_mw/vrot.pdf', bbox_inches = 'tight')
 
 # -------------------------------------------------------------------------------
+# plot talk
+# -------------------------------------------------------------------------------
+
+matplotlib.rc('axes',edgecolor='white')
+#colors = ["carolina blue", "amber", "greyish", "faded green", "dusty purple", "pale red", "orange"]
+#colors = sns.xkcd_palette(colors)
+
+R = np.linspace(0, 30, 1000)
+vc_bulge = vc_plummer(R, 0.5, 460*3e7)
+vc_halo = vc_NFW_old(R, 0, 2e12, 15.8, rho_c)
+vc_tot = np.sqrt(vc_bulge**2 + vc_halo**2)
+plt.plot(R, vc_bulge, '#feb308', lw = 3) #'#A8A8A8')
+plt.plot(R, vc_halo, '#d8dcd6', lw = 3, zorder = 40)
+#plt.plot(R, vc_tot, '#d9544d', lw = 3, zorder = 100)
+plt.xlim(-1, 30)
+plt.ylim(0, 250)
+plt.xlabel(r'$R~\rm [kpc]$', fontsize = 18, color = 'white')
+plt.ylabel(r'$v_{\rm c}~\rm [km/s]$', fontsize = 18, color = 'white')
+plt.tick_params(axis=u'both', direction='in', which='both', bottom = 'off', left = 'off', labelbottom = 'off', labelleft = 'off')
+plt.savefig('../talks_meetings/Galaxy_Coffee/Feb2019/vc2.pdf', transparent = True)
+
+# -------------------------------------------------------------------------------
 # plot webpage
 # -------------------------------------------------------------------------------
 

### 2019-03-29 infer_parallax.py

@@ -139,80 +139,91 @@ if prediction:
     # split into training and validation set
     y_pred_all = np.zeros_like(y_all)
     y_pred_all_err = np.zeros_like(y_all)
+    
+    # for referee
+    for run in range(1, 5):
+        np.random.seed(run)
+        random_vector = np.random.choice(2, (len(labels),))
+        try:
+            labels.remove_column('spec_parallax')
+            labels.remove_column('spec_parallax_err')
+        except (KeyError):
+            pass
         
-    for k in range(Kfold):    
-        
-        valid = labels['random_index'] % Kfold == k
-        train = np.logical_not(valid)
-        print('k = {0}: # of stars for prediction: {1}'.format(k, sum(valid)))
-        print('k = {0}: # of remaining of stars: {1}'.format(k, sum(train)))
+        for k in range(Kfold):    
             
-        # -------------------------------------------------------------------------------
-        # additional quality cuts for training set
-        # -------------------------------------------------------------------------------
+            #valid = labels['random_index'] % Kfold == k
+            valid = random_vector == k
+            train = np.logical_not(valid)
+            print('k = {0}: # of stars for prediction: {1}'.format(k, sum(valid)))
+            print('k = {0}: # of remaining of stars: {1}'.format(k, sum(train)))
+                
+            # -------------------------------------------------------------------------------
+            # additional quality cuts for training set
+            # -------------------------------------------------------------------------------
+            
+            print('more quality cuts for training sample...')
+            
+            # finite parallax required for training
+            cut_parallax = np.isfinite(labels[train]['parallax'])
+            
+            # visibility periods used
+            cut_vis = labels[train]['visibility_periods_used'] >= 8
+            
+            # cut in parallax_error
+            cut_par = labels[train]['parallax_error'] < 0.1       # this cut is not strictly required!
+            cut_burnin = labels[train]['parallax_over_error'] > 20.
+            
+            # cut in astrometric_gof_al (should be around unity...?) *Daniel Michalik's advice*
+            #cut_gof = labels[train]['astrometric_gof_al'] < 5  
+            # Coryn's advice!
[…]

### 2019-03-29 make_figures_HER.py

@@ -156,53 +156,129 @@ cb.set_label(r'$\rm S/N$', fontsize = fsize)
 plt.tight_layout()
 plt.savefig('paper/residuals.pdf', pad_inches=.2, bbox_inches = 'tight')
 
+# alternative version for referee
+fig, ax = plt.subplots(1, 3, figsize = (12, 4.5), sharey = True)
+sc = ax[0].scatter(labels['parallax'], labels['spec_parallax'], c = labels['SNR'], cmap = 'viridis_r', s = 10, vmin = 50, vmax = 1000, rasterized = True)
+ax[1].scatter(labels['parallax'][train], labels['spec_parallax'][train], c = labels['SNR'][train], cmap = 'viridis_r', s = 10, vmin = 50, vmax = 1000, rasterized = True)
+ax[2].scatter(labels['parallax'][best], labels['spec_parallax'][best], c = labels['SNR'][best], cmap = 'viridis_r', s = 10, vmin = 50, vmax = 1000, rasterized = True)
+ax[0].set_xlabel(r'$\varpi^{\rm (a)}$', fontsize = fsize)
+ax[1].set_xlabel(r'$\varpi^{\rm (a)}$', fontsize = fsize)
+ax[2].set_xlabel(r'$\varpi^{\rm (a)}$', fontsize = fsize)
+ax[0].set_ylabel(r'$\varpi^{\rm (sp)}$', fontsize = fsize)
+ax[0].tick_params(axis=u'both', direction='in', which='both', right = 'on', top = 'on')
+ax[1].tick_params(axis=u'both', direction='in', which='both', right = 'on', top = 'on')
+ax[2].tick_params(axis=u'both', direction='in', which='both', right = 'on', top = 'on')
+ax[0].plot(np.arange(-1, 3), np.arange(-1, 3), color = '0.6', linestyle = ':')
+ax[1].plot(np.arange(-1, 3), np.arange(-1, 3), color = '0.6', linestyle = ':')
+ax[2].plot(np.arange(-1, 3), np.arange(-1, 3), color = '0.6', linestyle = ':')
+ax[0].set_xlim(-.35, 2)
+ax[1].set_xlim(-.35, 2)
+ax[2].set_xlim(-.35, 2)
+ax[0].set_ylim(-.35, 2)
+ax[1].set_ylim(-.35, 2)
+ax[2].set_ylim(-.35, 2)
[…]

### 2019-03-29 non_axisymmetries.py

@@ -1,0 +1,833 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Feb 21 10:06:11 2019
+
+@author: eilers
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import scipy.optimize as op
+import time
+import pickle
+from astropy.table import Column, Table, join, vstack, hstack
+import sys
+from astropy.io import fits
+from sklearn.decomposition import PCA
+from astropy import units as u
+from astropy.coordinates import SkyCoord
+import astropy.coordinates as coord
+from mpl_toolkits.mplot3d import Axes3D
+from scipy.stats import binned_statistic_2d
+#from plotting_helpers import histcont
+import matplotlib.gridspec as gridspec
+import scipy.interpolate as interpol
+
+# -------------------------------------------------------------------------------
+# colormaps
+# -------------------------------------------------------------------------------
+
+import matplotlib.colors as mcolors
+
+c2 = plt.cm.Reds(np.linspace(0, 1, 32))
+c1 = plt.cm.Blues_r(np.linspace(0, 1, 32))
+c1 = np.vstack([c1, [1, 1, 1, 1]])
+mymap = mcolors.LinearSegmentedColormap.from_list('eddymap', np.vstack((c1, c2)))
+
+
+# -------------------------------------------------------------------------------
+# plotting settings
+# -------------------------------------------------------------------------------
+
+matplotlib.rcParams['ytick.labelsize'] = 18
+matplotlib.rcParams['xtick.labelsize'] = 18
+matplotlib.rc('text', usetex=True)
+fsize = 18
+
+# -------------------------------------------------------------------------------
+# open inferred labels
+# -------------------------------------------------------------------------------
+
+N = 44784 #66692
+Kfold = 2
+lam = 30
+name = 'N{0}_lam{1}_K{2}_offset0.0483_parallax'.format(N, lam, Kfold)
+
+print('loading new labels...')   
[…]

### 2019-03-29 normalize_all_spectra.py

@@ -14,7 +14,7 @@ from astropy.io import fits
 from astropy import units as u
 from astropy.coordinates import SkyCoord
 import pickle
-import matplotlib.pyplot as plt
+#import matplotlib.pyplot as plt
             
 # -------------------------------------------------------------------------------
 # download APOGEE catalogue
@@ -50,8 +50,9 @@ def LoadAndNormalizeData(file_spectra, file_name, destinations, pca = False, X_m
     no_data = []
     no_data_i = np.ones((len(file_spectra),), dtype = bool)
     for entry, destination_i in zip(file_spectra, destinations):
+        #entry = entry.replace('t9', 'r10')
         print(i)
-        destination = './data/spectra/' + str(destination_i) + '/'
+        destination = './data/spectra/' #+ str(destination_i).strip() + '/'
         try:
             hdulist = fits.open(destination + entry.strip())
             if len(hdulist[1].data) < 8575: 
@@ -71,7 +72,7 @@ def LoadAndNormalizeData(file_spectra, file_name, destinations, pca = False, X_m
             all_flux[i] = flux
             all_sigma[i] = sigma
         except:
-            entry = entry.replace('apStar-r8', 'aspcapStar-r8-l31c.2')
+            entry = entry.replace('apStar-t9', 'aspcapStar-t9-l31c.2')
             try:
                 hdulist = fits.open(destination + entry.strip())
                 if len(hdulist[1].data) < 8575: 

### 2019-03-29 parent_sample_dr15.py

@@ -100,13 +100,14 @@ subprocess.call(delete0, shell = True)
 
 for i, (fn, tel, field) in enumerate(zip(training_labels['FILE'], training_labels['TELESCOPE'], training_labels['FIELD'])):
     
+    fn = fn.replace('t9', 'r10')
     fn2 = fn.replace('apStar', 'aspcapStar')
     destination = './data/spectra/' + fn.strip() 
     destination2 = './data/spectra/' + fn2.strip()    
     #print(tel, destination, os.path.isfile(destination))
     
-    if not (os.path.isfile(destination) or os.path.isfile(destination2)):
-        urlbase = 'https://data.sdss.org/sas/apogeework/apogee/spectro/redux/t9/stars/' \
+    if True: #not (os.path.isfile(destination) or os.path.isfile(destination2)):
+        urlbase = 'https://data.sdss.org/sas/apogeework/apogee/spectro/redux/r10/stars/' \
         + str(tel).strip() + '/' + str(field).strip() + '/'
         url = urlbase + fn.strip()
         url2 = urlbase + fn2.strip()
@@ -146,22 +147,27 @@ for i in range(len(training_labels['FILE'])):
 training_labels = training_labels[found]
 print('spectra found for: {}'.format(len(training_labels)))
 
-# -------------------------------------------------------------------------------
+# -------------------------------------------------------------------------------'''
 # save training labels
 # -------------------------------------------------------------------------------
 
 print('save labels...')
 #fits.writeto('data/training_labels_parent_apogeedr15.fits', np.array(training_labels), overwrite = True)
-Table.write(training_labels, 'data/training_labels_parent_apogeedr15.fits', format = 'fits')
+Table.write(training_labels, 'data/training_labels_parent_apogeedr15_try3.fits', format = 'fits')
 
-# -------------------------------------------------------------------------------'''
+# -------------------------------------------------------------------------------
 # normalize spectra
 # -------------------------------------------------------------------------------
 
 print('load and normalize spectra...')
-file_name = 'all_flux_norm_parent_apogeedr15.fits'
-data_norm, continuum, not_found = LoadAndNormalizeData(training_labels['FILE'][:10], file_name, training_labels['FIELD'][:10])
-print('not found: {}'.format(np.sum(not_found)))       
+file_name = 'all_flux_norm_parent_apogeedr15_try3.fits'
[…]

### 2019-03-29 rotation_curve_uncertainties.py

@@ -41,10 +41,10 @@ fsize = 18
 # open inferred labels
 # -------------------------------------------------------------------------------
 
-N = 44784
+N = 66692 #44784
 Kfold = 2
 lam = 30
-name = 'N{0}_lam{1}_K{2}_offset0.0483_parallax'.format(N, lam, Kfold)
+name = 'N{0}_lam{1}_K{2}_offset0.0483_parallax_apogeedr15'.format(N, lam, Kfold)
 
 print('loading new labels...')   
 labels = Table.read('data/training_labels_new_{}.fits'.format(name), format = 'fits')    
@@ -109,7 +109,7 @@ np.random.seed(42)
 # -------------------------------------------------------------------------------
 # HRD plot
 # -------------------------------------------------------------------------------           
-#
+
 #distance = (labels['spec_parallax'] * u.mas).to(u.parsec, equivalencies = u.parallax())
 #MG = labels['phot_g_mean_mag'] - 5. * np.log10(distance.value) + 5
 #sn = 10
@@ -508,35 +508,35 @@ for i, (r_start, r_end) in enumerate(zip(bin_start, bin_end)):
 # Figure 7 for HER18
 # -------------------------------------------------------------------------------           
 
-#cut_feh = abs(labels['FE_H']) < 10
-#cuts = cut_z * cut_feh * cut_vz
-#fig, ax = plt.subplots(1, 1, figsize = (8, 8))
-#sc = plt.quiver(XS_cart_true_n[cuts, 0], XS_cart_true_n[cuts, 1], XS_cart_true_n[cuts, 3], XS_cart_true_n[cuts, 4], 
-#           np.clip(labels['FE_H'][cuts], -.7, .7), cmap = 'RdBu_r', scale_units='xy', 
-#           scale=200, alpha =.8, headwidth = 3, headlength = 5, width = 0.0015, rasterized = True)
-##cb = plt.colorbar(shrink = .8)
-##cb.set_label(r'$v_z$', fontsize = 15)
-#plt.xlim(-25, 5)
-#plt.ylim(-10, 20)
-#overplot_rings()
-#overplot_rings_helio()
-#plt.tick_params(axis=u'both', direction='in', which='both')
[…]

### 2019-03-29 paper/response_to_referee.txt
text fixes

@@
 
-Fixed.
+Fixed. This is a much better figure now.
 
@@
 
-Whoops. That's just a typo. Fixed.
+Whoops. That's just a typo. Fixed. We adjusted that whole paragraph,
+which was word salad, into something that is not wrong and also
+clearer.
 
@@
 
-[syntactical reference]
+fixed
 
@@
 
+fixed
+
     Section 4, 4th par., 5th line: "...3.6 and 4.5 μmPERIOD In detail..."
+
+fixed
 
@@
 
+fixed
+
     Section 4, 4th line below eq 10: eliminate extra "or" in "...cut ever
     on parallax or or the ratio..."
+
+fixed
 
@@
 
+Fixed the zenodo URL to a doi.org URL.
+
+We will do a final URL for the repo on acceptance; we can't do it
+before, because git hash! Happy to work to ensure reproducibility!
+
     - Do the authors wish to see the full Table 1 hosted with the Journal?
@@
       Zenodo URL in footnote # should be converted back to its DOI form.
+
+Yes we would like this. Let's discuss at acceptance.

### 2019-03-29 paper/sp.tex
table move
test commit
new
text fixes

@@
 \shortauthors{hogg, eilers, rix}
+\submitjournal{ApJ}
 
@@
 that input features are independently measured;
-that is, we make $C_m$ is a diagonal matrix with the uncertainty variances of the elements
+that is, we assume that $C_m$ is a diagonal matrix with the uncertainty variances of the elements
 of $x_m$ along the diagonal.
@@
 the same wavelength scale.
-In detail, we obtain the
+In detail, we obtain
 the \apogee\ spectral pixels from the pipeline-generated \code{aspcapStar} files.
@@
 This gives, for each matching star $n$,
-mid-infrared photometry $W_{1n}$ and $W_{2n}$ at 3.6 and 4.5\,$\mu$m
+mid-infrared photometry $W_{1n}$ and $W_{2n}$ at 3.6 and 4.5\,$\mu$m.
 In detail we use the \code{w1mpro} and \code{w2mpro} Catalog entries.
@@
 We match this full-match catalog to the \gaia\ \acronym{DR2}
-using the \gaia\ Archive official match (\citealt{xmatch}) to the the \zmass\ IDs.
+using the \gaia\ Archive official match (\citealt{xmatch}) to the \zmass\ IDs.
 We take from the \gaia\ \acronym{DR2} Catalog the photometric data
@@
 But---as we have emphasized above---we do not cut ever on parallax or
-or the ratio to its uncertainty. For this reason, most of the Training Set stars
+the ratio to its uncertainty. For this reason, most of the Training Set stars
 do not have significantly measured parallaxes and more than 2~percent even have
@@
 
-We used this cross-validation framework to adjust the regularization parameter $\Lambda$.
-A coarse grid search in the value for the spectral-component value of the
-$\lambda$ vector, using the A/B split as the two-fold cross-validation, with the objective
+We used this cross-validation framework to adjust the regularization parameter $\lambda$.
+We performed a coarse grid search in $\lambda$,
+using the A/B split as the two-fold cross-validation, with the objective
 of the cross validation being prediction accuracy.
 We get a prediction accuracy of better than 9 percent 
-at the optimal setting of the regularization strength.
+at the (coarsely) optimal setting of the regularization strength.
 
@@
 \tablename~\ref{tab:output} shows a snippet of this catalog of output, with the full catalog available
-as an electronic data table associated with this \documentname.\footnote{\url{https://zenodo.org/record/1468053}}
+as an electronic data table associated with this \documentname.\footnote{\url{https://doi.org/10.5281/zenodo.1468053}}
 In order to use this output with \gaia\ astrometry, we recommend joining our catalog to the
@@
 code repository.\footnote{\url{https://github.com/aceilers/spectroscopic_parallax}}
-
-\begin{longrotatetable}
-\startlongtable
-\begin{deluxetable*}{ccccccc}
-\tablecaption{The generated spectrophotometric parallaxes and their uncertainties..
-This \tablename\ is just a snippet; the full catalog of all 44784 objects in the combined Parent Samples
-is available online. In order to reproduce the {\figurename s} in this \documentname, this catalog
-should be joined to the \gaia, \zmass, \wise\, and \apogee\ Catalogs by \zmass\ ID.\label{tab:output}}
-\tablehead{\dcolhead{\texttt{2MASS\_ID}} & \dcolhead{\texttt{Gaia\_parallax}} & \dcolhead{\texttt{Gaia\_parallax\_err}} & \dcolhead{\texttt{spec\_parallax}} & \dcolhead{\texttt{spec\_parallax\_err}} & \dcolhead{\texttt{training\_set}} & \dcolhead{\texttt{sample}}}
-\startdata
-2M00000002+7417074 & 0.3191 & 0.0329 & 0.3185 & 0.0164 & 0 & B \\
-2M00000317+5821383 & 0.3643 & 0.0510 & 0.4428 & 0.0183 & 0 & A \\
-2M00000546+6152107 & 0.3154 & 0.0351 & 0.3292 & 0.0164 & 0 & B \\
-2M00000797+6436119 & 0.0771 & 0.0841 & 0.1943 & 0.0194 & 1 & B \\
-2M00001719+6221324 & 0.1395 & 0.0439 & 0.1564 & 0.0128 & 0 & B \\
-2M00002005+5703467 & 0.3003 & 0.0247 & 0.2372 & 0.0203 & 0 & B \\
-2M00002036+6153416 & 0.0657 & 0.0637 & 0.1347 & 0.0131 & 1 & A \\
-2M00002118+6136420 & 0.1936 & 0.0323 & 0.2173 & 0.0133 & 1 & B \\
-2M00002227+6223341 & 0.1965 & 0.0450 & 0.1461 & 0.0089 & 0 & B \\
-2M00002472+5518473 & 0.0947 & 0.0207 & 0.1473 & 0.0096 & 1 & B \\
-2M00002908+6140446 & 0.1675 & 0.0460 & 0.1914 & 0.0176 & 1 & A \\
-2M00003061+5820590 & 0.1669 & 0.0339 & 0.3502 & 0.0290 & 1 & A \\
-... & ... & ... & ... & ... & ... & ...  \\
- \enddata
-\tablecomments{Columns show...}
-\end{deluxetable*}
-\end{longrotatetable}
-\newpage
 
@@
 
+\begin{longrotatetable}
+\startlongtable
+\begin{deluxetable*}{ccccccc}
+\tablecaption{The generated spectrophotometric parallaxes and their uncertainties..
+This \tablename\ is just a snippet; the full catalog of all 44784 objects in the combined Parent Samples
+is available online. In order to reproduce the {\figurename s} in this \documentname, this catalog
+should be joined to the \gaia, \zmass, \wise\, and \apogee\ Catalogs by \zmass\ ID.\label{tab:output}}
+\tablehead{\dcolhead{\texttt{2MASS\_ID}} & \dcolhead{\texttt{Gaia\_parallax}} & \dcolhead{\texttt{Gaia\_parallax\_err}} & \dcolhead{\texttt{spec\_parallax}} & \dcolhead{\texttt{spec\_parallax\_err}} & \dcolhead{\texttt{training\_set}} & \dcolhead{\texttt{sample}}}
+\startdata
+2M00000002+7417074 & 0.3191 & 0.0329 & 0.3185 & 0.0164 & 0 & B \\
+2M00000317+5821383 & 0.3643 & 0.0510 & 0.4428 & 0.0183 & 0 & A \\
+2M00000546+6152107 & 0.3154 & 0.0351 & 0.3292 & 0.0164 & 0 & B \\
+2M00000797+6436119 & 0.0771 & 0.0841 & 0.1943 & 0.0194 & 1 & B \\
+2M00001719+6221324 & 0.1395 & 0.0439 & 0.1564 & 0.0128 & 0 & B \\
+2M00002005+5703467 & 0.3003 & 0.0247 & 0.2372 & 0.0203 & 0 & B \\
+2M00002036+6153416 & 0.0657 & 0.0637 & 0.1347 & 0.0131 & 1 & A \\
+2M00002118+6136420 & 0.1936 & 0.0323 & 0.2173 & 0.0133 & 1 & B \\
+2M00002227+6223341 & 0.1965 & 0.0450 & 0.1461 & 0.0089 & 0 & B \\
+2M00002472+5518473 & 0.0947 & 0.0207 & 0.1473 & 0.0096 & 1 & B \\
+2M00002908+6140446 & 0.1675 & 0.0460 & 0.1914 & 0.0176 & 1 & A \\
+2M00003061+5820590 & 0.1669 & 0.0339 & 0.3502 & 0.0290 & 1 & A \\
+... & ... & ... & ... & ... & ... & ...  \\
+ \enddata
+\tablecomments{All floating-point numbers are angles in mas.}
+\end{deluxetable*}
+\end{longrotatetable}
+
 \end{document}

### 2019-05-09 paper/sp.tex

@@
 %@arxiver{residuals.pdf,map.pdf}
-
-% To-do items
-% -----------
-% - check with WISE experts that we have the correct WISE photometry
 
@@
 \caption{The predictive accuracy of the model, in the two-fold cross-validation.
-  Each panel shows \gaia\ astrometric parallaxes $\gparallax_n$ on the vertical axis
-  and our spectrophotometric parallaxes $\sparallax$ on the horizontal axis.
+  Each panel shows \gaia\ astrometric parallaxes $\gparallax_n$ on the horizontal axis
+  and our spectrophotometric parallaxes $\sparallax$ on the vertical axis.
   The left panel shows the full Training Set (both A and B combined),

### 2019-05-09 paper/response_to_referee.txt
a few things done

@@
 we use a linear combination of features to predict or estimate a LOG
-PARALLAX, so if anything we are claiming that our estimates are
-symmetric in log parallax, not parallax. But we aren't even really
-claiming that. The issues are subtle. However, this does deserve some
-discussion and we have added that around Figures 5 and 6. The short answer
-is that we just negative-logged our log-parallax estimates to make Figures
-5 and 6. HOGG DO THIS
+PARALLAX, so if anything we are claiming that our estimates (not the
+data, but our estimates based on the data) are symmetric in log
+parallax, not parallax. But we aren't even really claiming that. The
+issues are subtle. However, this does deserve some discussion and we
+have added that to the captions of Figure 5 and 6, and in the last
+paragraph of the discussion, for good emphasis [HOGG DO THIS AT
+"PLACEHOLDER A"].  The short answer is that we just exponentiated the
+negative of our log-parallax estimates to make Figures 5 and 6.
 
@@
 model is not physically accurate. We added words about this into the
-assumption list in Section 2. HOGG DO THIS
+linearity assumption in the assumption list in Section 2.
 
@@
 
-Whoops! Fixed. HOGG DO THIS.
+Whoops! Fixed.
 
@@
 Yes we would like this. Let's discuss at acceptance.
+
+In addition to the above we made a few other small changes:
+
+- We added a reference to the Leung paper that followed ours in the
+relevant part of the discussion (where we discuss machine-learning
+alternatives).

### 2019-05-09 paper/sp.tex
a few things done

@@
   This is not common terminology in astrophysics, but it is standard in the machine-learning literature.}
-Perhaps our most fundamental assumption is that the parallax
+Perhaps our most fundamental assumption is that the parallax or distance
 of a star can be predicted from the features we provide, which are
@@
 \item[linearity] Perhaps the most restrictive assumption of this work is that
-the logarithmic parallax (or, equivalently, the distance modulus) 
+the logarithmic parallax (or, equivalently, the logarithmic distance, or,
+equivalently, the distance modulus) 
 can be predicted as a \emph{completely linear function} of
@@
 
+That said, the choice of this model (log parallax or log distance is a
+linear combination of input features) will certainly introduce
+particular and hard-to-learn biases into our results.
+Indeed, \emph{any} choice of model that is not strictly physically
+correct will introduce particular biases.
+We are effectively assuming that we can produce useful distance
+estimates nonetheless.
+
 \item[likelihood] We assume that the \gaia\ parallaxes are generated by a particular
@@
 Gaussian with a width set by the \gaia-Catalog uncertainty on the parallax.
-This is the standard assumption in all properly probabilistic \gaia-Mission inferences
-to date, but it subsumes a number of related assumptions, like that the \gaia\ noise
-model is correct, that the stars are only covariant with one another at negligible levels, and that
-there are no significant outliers or large-scale systematics.
+This is the standard assumption in all properly probabilistic
+\gaia-Mission inferences to date (as discussed in \citealt{gaialf}),
+but it subsumes a number of related assumptions, like that the
+\gaia\ noise model is correct, that the stars are only covariant with
+one another at negligible levels, and that there are no significant
+outliers or large-scale systematics.
 \end{description}
@@
 the photometric parallaxes, as a function of magnitude and heliocentric distance.
+For distances here are shown simply the inverses of the spectrophotometric parallaxes
+$\sparallax_n$.
 Note that the spectrophotometric estimates are more precise for most stars.
@@
 As we said above,
-we required that the parallax prediction be constructed from a linear combination
+we required that the log-parallax (or, equivalently, log-distance)
+prediction be constructed from a linear combination
 of the input data or feature vector components.
@@
 (at least approximately), they could have delivered better---or at least non-worse---results.
+And indeed, after the the appearance of this manuscript, such an investigation
+has appeared (\citealt{leung}) and it does produce very good results.
 The issues with going to a much more complex model are manifold, however:
@@
 of that capacity it really needs.
-A more complex model might have had edge issues: Very flexible models don't extrapolate
+A more complex model might also have edge issues: Very flexible models don't extrapolate
 outside the convex hull of their training data, and our training data was not
@@
   To make these velocities, raw \gaia\ proper motions and \apogee\ radial velocities
-  were used (in a standard Galactocentric transformation), so they are
+  and (for distance) inverses of spectrophotometric parallaxes were used
+  (in a standard Galactocentric transformation), so the map shown here is
   not in any way noise-deconvolved.
@@
 They are noisy.
+[HOGG PLACEHOLDER A]
 Precise use of these involves building a likelihood function and performing inference,
@@
 \bibitem[Kollmeier \etal(2017)]{sdssv} Kollmeier, J.~A., Zasowski, G., Rix, H.-W., \etal\ 2017, arXiv:1711.03234 
+\bibitem[Leung \& Bovy(2019)]{leung} Leung, H.~W., Bovy, J., 2019, arXiv:1902.08634
 \bibitem[Lindegren \etal(2018)]{lindegren} Lindegren, L., Hernandez, J., Bombrun, A., \etal\ 2018, arXiv:1804.09366 

### 2019-05-09 paper/response_to_referee.txt

@@
 issues are subtle. However, this does deserve some discussion and we
-have added that to the captions of Figure 5 and 6, and in the last
-paragraph of the discussion, for good emphasis [HOGG DO THIS AT
-"PLACEHOLDER A"].  The short answer is that we just exponentiated the
-negative of our log-parallax estimates to make Figures 5 and 6.
+have added that to the captions of Figure XXX and YYY (formerly Figure
+5 and 6), and in the last paragraph of the discussion, for good
+emphasis [HOGG DO THIS AT "PLACEHOLDER A"].  The short answer is that
+we just exponentiated the negative of our log-parallax estimates to
+make those figures.
 

### 2019-05-13 paper/response_to_referee.txt
more referee items

@@
 
-Hogg
+David W Hogg, for the authors
 
@@
 
-These issues are subtle! We are assuming that the GAIA PARALLAXES have
-noise that is symmetric in parallax space. In our data-driven model,
-we use a linear combination of features to predict or estimate a LOG
-PARALLAX, so if anything we are claiming that our estimates (not the
-data, but our estimates based on the data) are symmetric in log
-parallax, not parallax. But we aren't even really claiming that. The
-issues are subtle. However, this does deserve some discussion and we
-have added that to the captions of Figure XXX and YYY (formerly Figure
-5 and 6), and in the last paragraph of the discussion, for good
-emphasis [HOGG DO THIS AT "PLACEHOLDER A"].  The short answer is that
-we just exponentiated the negative of our log-parallax estimates to
-make those figures.
+These issues are subtle! We are assuming that the GAIA-MEASURED
+PARALLAXES have noise that is symmetric in parallax space. In our
+data-driven model, we use a linear combination of features to predict
+or estimate a LOG PARALLAX, so if anything we are claiming that our
+estimates (not the data, but our estimates based on the data) are
+symmetric in log parallax, not parallax. But we aren't even really
+claiming that. The issues are subtle. However, this does deserve some
+discussion and we have added that to the captions of Figure XXX and
+YYY (formerly Figure 5 and 6), and in the last paragraph of the
+discussion, for good emphasis [HOGG DO THIS AT "PLACEHOLDER A"].  The
+short answer is that we just exponentiated the negative of our
+log-parallax estimates to make those figures.
 
@@
 
-This point is important! The Gaia parallax uncertainties are based on
+This point is important. The Gaia parallax uncertainties are based on
 a spacecraft model that is independent of the measurements themselves.
@@
 But the referee is correct that we have been very unclear about this
-and we have attempted to fix this in Section YYYY. HOGG DO THIS
+and we have attempted to fix this in the relevant paragraph towards the
+end of Section 2. This has improved the paper.
 
@@
 
-Good point! An expansion on this theme added. HOGG DO THIS.
+Good point! The paper now opens with the big question and flows to the
+spectroscopy and precision. This change affects the first paragraph of
+the paper and changes how it starts.
 
@@
 
-Yes, yes it is. That was supposed to be for emphasis, but we did HOGG
-WHAT NOW?
+Yes, yes it is. Fixed and removed.
 
@@
 
-The effect is weak. We comment on this more now in Section XXX. HOGG
-DO THIS.
+The effect is very weak; we don't find that the change from 0.029 to 0.048
+makes much difference to our results except at very small parallaxes. And
+our 0.048 is not preferred over 0.029 at large confidence. The main reason
+it has little effect is that it effectively moves the training data only a
+tiny fraction of a sigma. We added two sentences to this effect in the relevant
+place.
 

### 2019-05-13 paper/sp.tex
more referee items

@@
 %% This file is part of the spectroscopic_parallax project
-%% Copyright 2018 the authors.
+%% Copyright 2018, 2019 the authors.
 
@@
 
-If we want to make precise kinematic and element-abundance maps
-of the Milky Way disk out to large heliocentric radii,
-and especially through the extinction to the other side of the Galactic Center, we will
-need to make use of luminous red giants, and we will need to observe in
-the infrared.
-These arguments motivate the \apogee\ (\citealt{apogee}),
+The Milky Way disk is a beautiful object of study for cosmology and galaxy evolution.
+Its evolution involves gas accretion and cooling, star formation, and orbital dynamics.
+To understand the details of this story, we need precise measurements of the
+stars and their properties over a large part of the disk, not just
+very near the Sun.
+Since we live in the middle of this disk, and the disk is large and
+dusty, these measurements require infrared spectroscopy of luminous
+giants.
+This motivates the \apogee\ (\citealt{apogee}),
 \apogee-2, and \sdssv\ (\citealt{sdssv})
@@
 in the likelihood function).
-The method in which we make no cuts and keep an umodified, justified likelihood function
+The method in which we make no cuts on astrometric measurements
+and we keep an umodified, justified likelihood function
 is unbiased, and simple.
@@
 with the data that are tempting but lead to important biases.
-For example, we never cut on \gaia\ parallax or parallax signal-to-noise.
 The common practices of cutting to parallaxes that are good to 20~percent
@@
 We never cut on parallax or parallax signal-to-noise.
+We will (below) cut on \gaia\ \emph{uncertainty estimates} on parallax.
+These uncertainty estimates are a function only of spacecraft orientation history
+(scanning law) and telemetry; they are not a function of the magnitudes of the
+parallaxes or their measurements.
+For these reasons, the uncertainty estimates on the parallaxes can be used to
+make sample cuts (and we do use them).
+These cuts will not bias our regressions.
 By using a justifiable likelihood function (the ``likelihood'' assumption above),
-we can use all of the \gaia\ parallaxes without the low signal-to-noise and
-negative parallaxes causing trouble for our regressions, and without the biases
+we can use all of the \gaia\ parallaxes without fearing that the low signal-to-noise and
+negative parallaxes causing trouble for our regressions, and without any of the biases
 that enter when cuts are made.
@@
 The spectrophotometric parallaxes we generate
-do not depend strongly on this choice, except
-for small differences at very small parallax (large distance).
+do not depend strongly on this choice.
+It does not have a large effect because it moves the training data by only a tiny
+fraction of the root-variance of the uncertainties.
+It is also the case (for the same reason) that we do not strongly prefer (in a statistical sense)
+the $\varpi_0 = 0.048$\,mas value over the \gaia\ value of 0.029\,mas.
+
 The Training Sets are shown in the bottom panels of \figurename~\ref{fig:samples}.
-
 Comparison of the top and bottom panels of {\figurename}~\ref{fig:samples} shows

### 2019-06-02 paper/response_to_referee.txt
one more item off

@@
 YYY (formerly Figure 5 and 6), and in the last paragraph of the
-discussion, for good emphasis [HOGG DO THIS AT "PLACEHOLDER A"].  The
-short answer is that we just exponentiated the negative of our
-log-parallax estimates to make those figures.
+discussion, for extra-special emphasis.  The short answer is that we
+just exponentiated the negative of our log-parallax estimates to make
+those figures.
 

### 2019-06-02 paper/sp.tex
one more item off

@@
 They are noisy.
-[HOGG PLACEHOLDER A]
+And they linearly combine a set of features into a logarithmic parallax or
+logarithmic distance.
+Because the input features have close-to-Gaussian noise (mainly
+because they are observed at high signal-to-noise), the linear
+combination will have close-to-Gaussian noise.
+That is, the noise in the estimated distances or parallaxes ought to be close
+to Gaussian in the logarithm.
+This is a consideration in the use of these outputs:
 Precise use of these involves building a likelihood function and performing inference,
 just as the precise use of the \gaia\ data presented here required the same.
-We recommend making analogous inferences when using these outputs.
+Justified likelihood functions are critical to making good inferences.
 

### 2019-06-02 paper/response_to_referee.txt
two more items

@@
 story is somewhat complicated. But you are right that this must be
-there, and we have added it back in, at the end of Section YYY. The
-short story (which you can read in the updated text) is that the
-photometric part of the model is interpretable and stable, but the
-spectroscopic part is less stable in its internals (while remaining
-very stable in its predictions in the data space). We don't fully
-understand this, but it connects to known issues in machine learning
-more generally, which we point out in the text. HOGG DO THIS
+there, and we have added it back in, in the middle of the Discussion
+section. The short story (which you can read in the updated text) is
+that the photometric part of the model is interpretable and stable,
+but the spectroscopic part is less stable in its internals (while
+remaining very stable in its predictions in the data space). We don't
+fully understand this, but it connects to known issues in machine
+learning more generally, which we point out in the text.
 
@@
 
-Hahaha no. Okay fine. We now comment on this point in Section YYY
-where we deliver the interpretation of the photometric part of the
-model. HOGG DO THIS.
+Hahaha no. Okay fine. We now comment on this point in the Discussion,
+where we now deliver an interpretation of the photometric part of the
+model, and comment on the fact that it *does* look dust-corrected.
 

### 2019-06-02 paper/sp.tex
two more items

@@
 \section*{~}\clearpage
-\section{Introduction} \label{sec:intro}
+\section{Introduction}\label{sec:intro}
 
@@
 
-\section{Method}
+\section{Method}\label{sec:method}
 
@@
 
-\section{Data}
+\section{Data}\label{sec:data}
 
@@
 
-\section{Results and validation}
+\section{Results and validation}\label{sec:results}
 
@@
 
-\section{Discussion}
+\section{Discussion}\label{sec:discussion}
 
@@
 Similarly, linear combinations of spectral pixels are like a projection of the spectral space.
-Brief inspection of these linear combinations showed them to be sensible, but it is out of
-scope here to interpret them in detail.
+
+For an example of how the model might be interpreted, we can look at
+the photometric part of the linear model and see if the combination of
+photometry does indeed look like a dust-corrected absolute magnitude,
+as we expected (see \sectionname~\ref{sec:data}).
+It does! Indeed the model automatically finds a photometric
+combination that includes a term close to $H-W2$, which is known as a
+good extinction estimator.
+And we find that on the two-fold A/B split, the photometric model is extremely stable.
+As for the spectroscopic-pixel part of the model, we find that interpretation is
+less simple, for two reasons:
+One is that the number of pixels is enormous, so the model has huge flexibility.
+Another is that although the two splits, A and B, both give similar-performing
+models, they have (in detail) different linear coefficients in the spectral space.
+That is, the model is not as stable in the spectroscopic domain as it is in the
+photometric domain.
+This is a generic feature of high-capacity machine-learning methods.
+And it means that we did not obtrain as much insight from the spectroscopic model
+as we had hoped up-front.
 

### 2019-06-02 paper/sp.tex
thank referee

@@
 It does! Indeed the model automatically finds a photometric
-combination that includes a term close to $H-W2$, which is known as a
+combination that includes a term close to $H-W_2$, which is known as a
 good extinction estimator.
@@
 for help with this project.
+The final version of this paper benefitted substantially from the comments
+  from the anonymous referee.
 This project was developed in part at the

### 2019-06-02 paper/response_to_referee.txt
another point down

@@
+To-do items:
+------------
+- find and kill all HOGG
+- find and kill all XXX
+
 Dear Editor:
@@
 discussion and we have added that to the captions of Figure XXX and
-YYY (formerly Figure 5 and 6), and in the last paragraph of the
+XXX (formerly Figure 5 and 6), and in the last paragraph of the
 discussion, for extra-special emphasis.  The short answer is that we
@@
 Queiroz sample required some emails!) and we made comparisons. We
-added a short paragraph about the results of these comparisons in
-Section XXX. We did not feel that the comparisons merited supporting
-Figures, but we do have such figures if the referee disagrees. HOGG DO THIS
+added a short paragraph about the results of these comparisons in the
+second-last paragraph of the Results Section. We did not feel that the
+comparisons merited supporting Figures, but we do have such figures if
+the referee disagrees.
 
@@
 there was also a paper out by Leung & Bovy. They compare their results
-to ours, and we also mention that in the newly added paragraph. HOGG DO THIS
+to ours, and we also mention that in the newly added paragraph.
 

### 2019-06-02 paper/sp.tex
another point down

@@
 
+It is also interesting to compare these estimates to other, previous
+spectroscopic distance efforts.
+We performed comparisons with two other efforts (\citealt{sch, que}).
+In both cases, the agreement is better at smaller distances (larger
+parallaxes) and worse at larger distances, with the distances
+(parallaxes) presented here being smaller (larger) on average than the other
+investigation results for the very most distant (and hence most difficult) stars.
+In detail, the comparison with the larger sample (\citealt{sch})
+are better:
+The scatter is at the 15-percent level for stars at distances
+$<8$\,kpc), and there is no visible bias between the results at these
+distances.
+The distances presented here also compare well with those from a recent
+machine-learning approach (\citealt{leung}), which appeared while this
+paper was under review.
+
 In the end, this method produces a spectrophotometric parallax estimate $\sparallax_m$
@@
 \bibitem[P\'erez \& Granger(2007)]{ipython} P\'erez, F. \& Granger, B.~E.\ 2007, CISE, 9(3), 29
+\bibitem[Queiroz \etal(2018)]{que} Queiroz, A.~B.~A., Anders, F., Santiago, B.~X., \etal\ 2018, \mnras, 476, 2556.
+\bibitem[Schultheis \etal(2014)]{sch} Schultheis, M., Zasowski, G., Allende Prieto, C., \etal\ 2014, \aj, 148, 24
 \bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., \etal\ 2006, \aj, 131, 1163 

### 2019-06-02 paper/response_to_referee.txt
last point I can do today

@@
 and the parent sample.  We don't have any reason to believe that this
-biases our results. However, we do now show these distribution
-differences in a newly added figure, Figure XXX. Certainly these
-differences mean that we probably have uncertainties that, in detail,
-vary with log g. HOGG DO THIS
+biases our results. Furthermore, we do have complete coverage. That
+is, there are no parallaxes, SNRs, or log g values that exist in the
+test set that do not exist in the training set: In each dimension, the
+test set is inside the hull of the training set. But in detail there
+are small distributional differences. We now comment on this in the
+text explicitly in two new sentences in the Data Section. We did not
+feel like this merited a figure, but we do have one if the referee
+would like to see it.
 
@@
 
-Done. HOGG: DO WE PUT THIS INTO THE PAPER?
+Done. EILERS DID WE MAKE THIS? SHOULD WE PUT THIS INTO THE PAPER?
 

### 2019-06-02 paper/response_to_referee.txt
cleaning

@@
 - find and kill all HOGG
-- find and kill all XXX
+- find and kill all EILERS
 
@@
 claiming that. The issues are subtle. However, this does deserve some
-discussion and we have added that to the captions of Figure XXX and
-XXX (formerly Figure 5 and 6), and in the last paragraph of the
-discussion, for extra-special emphasis.  The short answer is that we
-just exponentiated the negative of our log-parallax estimates to make
-those figures.
+discussion and we have added that to the captions of Figure 5 and 6,
+and in the last paragraph of the discussion, for extra-special
+emphasis.  The short answer is that we just exponentiated the negative
+of our log-parallax estimates to make those figures.
 

### 2019-06-03 paper/sp.tex

@@
 acceptably.
+In terms of the more mundane properties of \gaia\ parallax $\gparallax$,
+\apogee\ spectroscopic signal-to-noise ratio,
+or surface gravity $\logg$, there are small differences between the distributions
+in the Training Sets and the Parent Sample.
+However, in each of these dimensions, the Training Set do span the full range
+of the Parent Sample, so the Training Set does have full coverage in the
+space.
 

### 2019-06-03 make_figures_HER.py
figure

@@ -20,7 +20,6 @@ from astropy import units as u
 from astropy.coordinates import SkyCoord
 import astropy.coordinates as coord
 from mpl_toolkits.mplot3d import Axes3D
-import corner
 from mpl_toolkits.axes_grid1.axes_divider import make_axes_locatable
 from mpl_toolkits.axes_grid1.colorbar import colorbar
 
@@ -642,6 +641,11 @@ matches = join(schultheis, labels, keys='APOGEE_ID')
 spec_par_match = matches['spec_parallax'] * u.mas
 distance_match = spec_par_match.to(u.parsec, equivalencies = u.parallax())
 
+diff = np.log(matches['Dis']) - np.log(distance_match.value/1000.)
+offset = np.median(diff)
+scatter = np.percentile(diff, 84) - np.percentile(diff, 16)
+print(offset, scatter)
+
 chi2 = (distance_match.value/1000. - matches['Dis']) ** 2 / (matches['e_Dis']/100. * matches['Dis'])**2
 
 fsize = 16
@@ -680,6 +684,10 @@ plt.xlabel(r'$\ln d~$[kpc] (Queiroz et al. 2018)', fontsize = fsize)
 plt.ylabel(r'$\ln d~$[kpc] (Hogg, Eilers, Rix 2018)', fontsize = fsize)
 plt.savefig('paper/comparison_queiroz.pdf', pad_inches=.2, bbox_inches = 'tight')
 
+diff = np.log(matches['BPG_meandist']) - np.log(distance_match.value/1000.)
+offset = np.nanmedian(diff)
+scatter = np.nanpercentile(diff, 84) - np.nanpercentile(diff, 16)
+print(offset, scatter)
 
 
 

### 2019-06-03 non_axisymmetries.py

@@ -613,7 +613,7 @@ ax.tick_params(axis=u'both', direction='in', which='both')
 ax.set_aspect('equal')
 ax.set_xlabel(r'$x\rm~[kpc]$', fontsize=18)
 ax.set_ylabel(r'$y\rm~[kpc]$', fontsize=18)
-plt.savefig('non_axisymmetries/mesh_vz_bc.pdf'.format(name_plot, name)) 
+plt.savefig('non_axisymmetries/{}.pdf'.format(name_plot)) 
 #plt.close()
 
 

### 2019-06-03 paper/sp.tex
caption update

@@
 \begin{figure}
-\includegraphics[width=\textwidth]{residuals.pdf}
+\includegraphics[width=\textwidth]{residuals_inclduing_parent_sample.pdf}
 \caption{The predictive accuracy of the model, in the two-fold cross-validation.
@@
   and our spectrophotometric parallaxes $\sparallax$ on the vertical axis.
-  The left panel shows the full Training Set (both A and B combined),
+  The points are colored by spectroscopic (\apogee) signal-to-noise ratio.
+  The left panel shows the full Parent Sample,
+  the middle panel shows the Training Set (both A and B combined),
   and the right panel shows a
@@
 investigation results for the very most distant (and hence most difficult) stars.
-In detail, the comparison with the larger sample (\citealt{sch})
+In detail, the comparison with the more recent sample (\citealt{que})
 are better:

### 2019-07-29 paper/sp.tex
reference
response 2

@@
 contexts, but bad when you know that the functional dependencies you hope to
-model are smooth:
+model are smooth.
 In that case, much of the information in the data can get spent on learning the
@@
 We chose the three displayed clusters
-to span some range in metallicity and age and extinction, and therefore test
+to span some range in metallicity, age, and extinction, and therefore test
 the accuracy of the method for different kinds of populations.
@@
 In detail, the comparison with the more recent sample (\citealt{que})
-are better:
+is better:
 The scatter is at the 15-percent level for stars at distances
-$<8$\,kpc), and there is no visible bias between the results at these
+$<8$\,kpc, and there is no visible bias between the results at these
 distances.
@@
 
-However, our choice of features is very rigid:
+However, our choice of features is very rigid.
 In principle the data themselves should tell us what features to include.
@@
 Since models like Gaussian processes and deep networks subsume the linear model
-(at least approximately), they could have delivered better---or at least non-worse---results.
+(at least approximately), they could have delivered better---or at least similar---results.
 And indeed, after the the appearance of this manuscript, such an investigation
 has appeared (\citealt{leung}) and it does produce very good results.
-The issues with going to a much more complex model are manifold, however:
+The issues with going to a much more complex model are manifold, however.
 We would have had many more decisions to make and many more hyper-parameters
@@
 We would have used some (or maybe much) of the information in the data to learn
-the simple fact that the problem is close to linear; that is, if you give a model
-a lot of freedom or capacity, it has to use a lot of the data to learn what part
-of that capacity it really needs.
-A more complex model might also have edge issues: Very flexible models don't extrapolate
-outside the convex hull of their training data, and our training data was not
-balanced in terms of parallax and dust attenuation.
+the simple fact that the problem is \emph{close} to linear.
+A more complex model might also have edge issues, because flexible models don't extrapolate
+outside the convex hull of their training data.
 In the  linear model it is also---unlike in more flexible models---trivial to propagate
@@
 And finally---in principle---linear models are better for visualization and interpretation;
-the linear model contains essentially only first derivatives in the data space, which are
-relatively easy to understand.
+the linear model contains essentially only first derivatives in the data space.
 
@@
 This is a generic feature of high-capacity machine-learning methods.
-And it means that we did not obtrain as much insight from the spectroscopic model
+And it means that we did not obtain as much insight from the spectroscopic model
 as we had hoped up-front.
@@
 Although the model, training, and prediction make no use whatsoever of stellar
-models, we did use stellar models indirectly to select the Parent Sample of stars:
+models, we did use stellar models indirectly to select the Parent Sample of stars.
 We used the \apogee\ pipeline (\citealt{aspcap}) surface-gravity $\logg$ measurements.
@@
 which that star is not a member.
-This framework ensures that the \gaia\ parallax $\gparallax_m$ of star $m$ is never used,
-even indirectly, in generating the spectrophotometric parallax $\sparallax_m$ of star $m$.
-That means that the parallax estimates generated here and given in \tablename~\ref{tab:output}
-are statistically independent of the \gaia\ parallaxes and can be combined with them in
-na\"ive ways, for example in inverse-variance-weighted averages.
+As we emphasize above, this also ensures that the \gaia\ measurement $\gparallax_m$ of star $m$ is never used,
+even indirectly, in generating the spectrophotometric estimate $\sparallax_m$ for that star $m$.
+This, in turn, ensures statistical independence of the
+parallax estimates generated here from the \gaia\ parallaxes.
+The two could even be combined na\"ively in an inference.
 

### 2019-07-30 paper/sp.tex

@@
 which that star is not a member.
-As we emphasize above, this also ensures that the \gaia\ measurement $\gparallax_m$ of star $m$ is never used,
+As we emphasize above, this ensures that the \gaia\ measurement $\gparallax_m$ of star $m$ is never used,
 even indirectly, in generating the spectrophotometric estimate $\sparallax_m$ for that star $m$.
-This, in turn, ensures statistical independence of the
+This leads to true statistical independence of the
 parallax estimates generated here from the \gaia\ parallaxes.
@@
 \bibitem[Casey \etal(2016)]{casey} Casey, A.~R., Hogg, D.~W., Ness, M., \etal\ 2016, arXiv:1603.03040 
-\bibitem[Eilers \etal(2018)]{eilers} Eilers,~A.~C., Hogg, D.~W., Rix, H.-W., \& Ness, M. 2018, in preparation
+\bibitem[Eilers \etal(2018)]{eilers} Eilers,~A.~C., Hogg, D.~W., Rix, H.-W., \& Ness, M. 2019, \apj, 871, 120
 \bibitem[Gaia Collaboration \etal(2016)]{gaia} Gaia Collaboration, Prusti, T., de Bruijne, J.~H.~J., \etal\ 2016, \aap, 595, A1 

### 2019-07-31 paper/response_to_referee2.txt
response2

Dear Editor:

Enclosed please find the author response to the referee report on
AAS14152R1 "Spectrophotometric parallaxes with linear models: Accurate
distances for luminous red-giant stars" by Hogg, Eilers, & Rix.

Apologies for the tardiness of this response. Our response to the
referee is below the line.

David W Hogg, for the authors

-----

In what follows, the referee comments are indented, and our responses
are not. These comments improved the paper; we appreciate the close
read!

   page 3: "... hope to model are smoothPERIOD In that case, much of
   the information in the data can ..."

Done.

   page 11: Section 5, 2nd paragraph: please use the same verbal tense
   in the sentences.

We disagree here. The first two sentences say what we did, and the last
says what we find, having done that.

   page 12, 2nd par: "... metallicity, age and extinction,..." (I
   suggest to replace the first "and" by a comma)

Done.

   page 13, par. that compares to other distance estimates: "the
   comparison with the more recent sample (Queiroz et al. 2018) IS
   better:"

Done.

   page 13, par. that compares to other distance estimates: remove ")"
   after "at distances < 8 kpc"

Done.

   page 15, section 6, 3rd par.: "...our choice of features is very
   rigidPERIOD In principle..."

Done.

   In fact, in this last section there several other examples where
   colon was used where a period would better separate the sentences,
   in my humble opinion!

Hahaha! Understood. Removed some.

   page 15, last par.: Perhaps replace "non-worse", which sounds
   weird, by "similar quality"?

Done.

   top of page 16: "...are manifold, howeverPERIOD We would have had..."

Done.

   page 16, end of 3rd par.: "...that we did not OBTAIN as much insight.."

Done.

   Section 6: perhaps the authors could shorten a bit the comparison
   with more complex models based on machine learning and the
   discussion on the advantages and disadvantages of the linear model
   adopted.

Shortened. That improves this paragraph.

   In this last section, there are also a few sentences that are
   repeated from previous sections, such as: "This framework ensures
   that the Gaia parallax ... in generating the spectrophotometric
   parallax ... That means that the parallax estimates generated here
   and given in Table 1 are statistically independent of the Gaia
   parallaxes" This sentence is basically repeated from the end of
   Section 3.

Rephrased these; we do want to emphasize these points by repetition,
so we didn't delete them, just rephrased them to make it clear that
it is a repetition for emphasis.
